---
title: "H3K18la Analysis - May 28th, 2024"
output: html_notebook

---

### Project Summary
This project investigates the H3K18la during during aging. The samples are for P4,
P10, and P28. One replicate of P28 was run on 240410 and the rest of the samples
were run on 240522. The CUT&TAG was performed using the Active Motif kit with the 
Henikoff pA-Tn5 enzyme. The primary analysis was perfomed using the nf-core cutandrun 
pipeline...details in src/data. The no antibody control was not used for background 
subtraction since the read numbers were too low for the P4 and the P28 rep1. The 
MACS consensus peaks were filtered for peaks coming from at least two replicates.

**DESIGN**:  

**NOTES**: The data were generated from nf-core cutandrun pipeline.

**UPDATE**: The definition of promoter (-1000 to +500) was used for peak annotation.
            Filter the peaks for FDR of 10E-6 prior to generating the concensus peaks.



#### Get custom scripts and functions and outside data used for analysis
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

script_dir <- "https://raw.githubusercontent.com/MBrooks313/R_codes/master"
dest_tools_dir <- "../src/tools"
dest_vis_dir <- "../src/visualization"

download.file(url = file.path(script_dir, "GO_leaf.R"),
              destfile = file.path(dest_tools_dir, "GO_leaf.R"),
              method = "curl")


# system('cp ~/github/R_codes/DFave.R ../src/tools')
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/et_edgeR_v0.1.0.R ../src/tools")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/FuncGeneEnrich_v0.1.0.R ../src/tools")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/GO_leaf_v1.1.R ../src/tools")

#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/APk_cluster_BF_v0.1.0.R ../src/tools')
#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/HM_PCAsort_BF_v0.1.0.R ../src/visualization')
#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/parPlot_hc_v0.2.0.R ../src/visualization')
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/PCA_simple_v0.1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/VolcanoPlot_v0.1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_v1.7.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Dotplot_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Barplot_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_v1.7.R ../src/visualization")
# 
# system("cp /Volumes/NGS/RNA-seq/ENS_v98/Mouse/Counts/by_project/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv ../data/external")
# system("cp /Volumes/NGS/RNA-seq/ENS_v98/Mouse/Counts/by_project/Mouse_FSPR_WT_Development_Gene_nCPM_ENSv98.tsv ../data/external")

# Function to convert ENS ids to gene symbol 
# ens2sym <- function(x){paste(sort(unique(AnnotationDbi::mapIds(org.Mm.eg.db, keys = x, keytype = "ENSEMBL", column = "SYMBOL"))), sep = "", collapse = ",")}
library(tidyverse)
options(java.parameters = "-Xmx8000m")
print("Completed...")
```

# Get Ensembl annotation database
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

## Query for all available EnsDb databases
ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

ensDb <- ahdb[[1]]

# supportedFilters(ensDb)

```




################################################################################
## All H3K18la samples
################################################################################

## Get metadata for samples
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
# ord <- c(1:4)

# Macs2 beds
macs_beds <- list.files(path = "../data/interim/bed/macs2/", pattern = "narrowPeak", recursive = T)

# Meta sheet
meta <- data.frame(Beds = macs_beds[1:7],
                   Samp = gsub("H3K18La_(P.+)\\.macs.+", "\\1", macs_beds[1:7]),
                   Group = gsub("H3K18La_(P\\d+)_R.+", "\\1", macs_beds[1:7]))
meta <- meta[c(6,7,1:5),]
# meta$Beds <- factor(meta$Beds, levels = levels(meta$Beds))
meta$Samp <- factor(meta$Samp, levels = unique(meta$Samp))
meta$Group <- factor(meta$Group, levels = unique(meta$Group))

print("Meta data acquired...")
```


# Import the peaks into a merged list object and plot for number of peaks at each timepoint
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
library(ChIPpeakAnno)

macs_dir <- "../data/interim/bed/macs2/"

# Import the cons peaks
peakList <- sapply(file.path(macs_dir, meta$Beds), toGRanges, format="narrowPeak")
names(peakList) <- meta$Samp

peakList_1E6 <- GRangesList()
# Sort the cons peaks
for (i in 1:length(peakList)){
  
  # seqlevelsStyle(peakList[[i]]) <- "NCBI"
  tmp_peak <- sortSeqlevels(peakList[[i]])
  tmp_peak <- GenomicRanges::sort(tmp_peak)
  tmp_peak <- tmp_peak[(elementMetadata(tmp_peak)[,"qValue"] > 6) & (seqnames(tmp_peak) != "MT")]
  peakList_1E6[[i]] <- tmp_peak
}
names(peakList_1E6) <- meta$Samp

# ####################
# # FDR = 1E-6
# 
# # Filter for 0.0001
# peakList_1E6 <- GRangesList()
# for(i in 1:length(peakList)){
#   tmp_peak <- peakList[[i]][(elementMetadata(peakList[[i]])[,"qValue"] > 6) & (seqnames(peakList[[i]]) != "MT")]
#   peakList_1E6[[i]] <- tmp_peak
# }
# names(peakList_1E6) <- meta$Samp

# Overlap peaks per timepoint
# Get overlaps
ol_P4 <- findOverlapsOfPeaks(c(peakList_1E6[1],
                             peakList_1E6[2]), connectedPeaks="keepAll", ignore.strand = T)

ol_P10 <- findOverlapsOfPeaks(c(peakList_1E6[3],
                             peakList_1E6[4]), connectedPeaks="keepAll", ignore.strand = T)

ol_P28 <- findOverlapsOfPeaks(c(peakList_1E6[5],
                             peakList_1E6[6],
                             peakList_1E6[7]), connectedPeaks="keepAll", ignore.strand = T)

peakList_cons <- GRangesList(ol_P4$mergedPeaks, ol_P10$mergedPeaks, ol_P28$mergedPeaks)
names(peakList_cons) <- c("P4", "P10", "P28")

```










<!-- ## Get metadata for samples -->
<!-- ```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE} -->
<!-- # ord <- c(1:4) -->

<!-- # Macs2 beds -->
<!-- macs_beds <- list.files(path = "../data/interim/bed/consensus_nf-core", pattern = ".+bed", recursive = T) -->

<!-- # Meta sheet -->
<!-- meta <- data.frame(Beds = macs_beds, -->
<!--                    Samp = gsub("H3K18La_(.+)\\.consensus.+", "\\1", macs_beds)) -->
<!-- meta <- meta[c(3,1,2),] -->
<!-- # meta$Beds <- factor(meta$Beds, levels = levels(meta$Beds)) -->
<!-- meta$Samp <- factor(meta$Samp, levels = unique(meta$Samp)) -->


<!-- print("Meta data acquired...") -->
<!-- ``` -->


# Import the peaks into a merged list object and plot for number of peaks at each timepoint
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
library(ChIPpeakAnno)

# cons_dir <- "../data/interim/bed/consensus_nf-core"
# 
# # Import the cons peaks
# peakList_cons <- sapply(file.path(cons_dir, meta$Beds), toGRanges, format="BED")
# names(peakList_cons) <- meta$Samp
# 
# # Sort the cons peaks
# for (i in 1:length(peakList_cons)){
#   
#   # seqlevelsStyle(peakList_cons[[i]]) <- "NCBI"
#   peakList_cons[i] <- sortSeqlevels(peakList_cons[[i]])
#   peakList_cons[i] <- GenomicRanges::sort(peakList_cons[[i]])
# }

# Generate plot for peak number
df <- lapply(peakList_cons, length) %>% as.data.frame() %>% gather("Sample", "Peaks")
df$Sample <- gsub("X", "", df$Sample)
df$Sample <- factor(df$Sample, levels = df$Sample)

b <- ggplot(df, aes(Sample, Peaks))+
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 0.8, vjust = 0.8)) +
  labs(title = "Peaks") + xlab(NULL)

cowplot::ggsave2(plot = b, filename = "../data/processed/Peak_number_allPeaks.pdf", height = 4, width = 3)

print("Peak data acquired...")
```


## Annotate peaks with ensembl db
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


peakAnnoList_ens <- list()

for (i in 1:length(peakList_cons)){
  
  peak_tmp <- peakList_cons[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  # peakAnnoList_ens[[i]] <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
  #                        TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T,
  #                        overlap = "all")
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  anno_tmp@anno$GENENAME <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "GENENAME", 
                                         keytype = "GENEID")
  peakAnnoList_ens[[i]] <- anno_tmp

}

names(peakAnnoList_ens) <- names(peakList_cons)

```


# Evaluate location of peaks relative to annotation features
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
#library(clusterProfiler)
promoter <- getPromoters(TxDb = ensDb, upstream = 3000, downstream = 3000)
seqlevelsStyle(promoter) <- "NCBI"

tagMatrixList <- list()
for (i in names(peakList_cons)){
  tagMatrixList[[i]] <- getTagMatrix(peakList_cons[[i]], windows=promoter)
}


# # tagMatrix <- getTagMatrix(peakList_cons, windows=promoter)
# tagHeatmap(tagMatrixList, xlim=c(-3000, 3000))

# Plot peak histogram around TSS
p <- plotAvgProf(tagMatrixList, xlim=c(-3000, 3000), resmple = 100,
            xlab="Genomic Region (5'->3')", ylab = "Read Count Frequency")
print(p)

pdf("../data/processed/01_Peak_TSS_hist_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


# Plot annotation bargraph
p <- plotAnnoBar(peakAnnoList_ens, title = "H3K18la peak annotation")
print(p)

pdf("../data/processed/01_Peak_annot_bar_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


# Plot disrtobution of peaks realtive to TSS
p <- plotDistToTSS(peakAnnoList_ens,
              title="Distribution of H3K18la peak to TSS")
print(p)

pdf("../data/processed/01_Peak_TSSdist_bar_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


```

## Export peak annotation as bed and xlsx files
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(rtracklayer)
library(xlsx)


for (i in 1:length(peakAnnoList_ens)){
  
  gr <- peakAnnoList_ens[[i]]@anno
  name_gr <- names(peakAnnoList_ens[i])
  
  df <- data.frame(seqnames=seqnames(gr),
    starts=start(gr)-1,
    ends=end(gr),
    SYMBOL=gr$SYMBOL,
    # names=names(gr),
    #scores=gr$score,
    scores=".",
    strands=strand(gr),
    #signalValue=gr$signalValue,
    #qValue=gr$qValue,
    annotation=gr$annotation,
    # names=names(gr),
    # SYMBOL=gr$SYMBOL,
    # ENTREZID=gr$ENTREZID,
    geneId=gr$geneId,
    transcriptId=gr$transcriptId,
    distanceToTSS=gr$distanceToTSS,
    GENENAME=gr$GENENAME,
    biotype=gr$transcriptBiotype)
  
  write.table(df, file=paste0("../data/processed/01_", name_gr, "_annotated.bed"), 
              quote=F, sep="\t", row.names=F, col.names=F)

  # write.xlsx(df, file=paste0("../data/processed/01_", name_gr, "_annotated.xlsx"),
  #            row.names=F, col.names=T)
  
}


```



# Export the annotated peak file
```{r}

saveRDS(peakList_cons, file = "../data/interim/peakList_cons_240911.RDS")
saveRDS(peakAnnoList_ens, file = "../data/interim/peakAnnoList_cons_240911.RDS")

```



################################################################################
# Prepare plot for figures - Oct 30th, 2023 
################################################################################

####################
# Timecourse K18La
####################

<!-- # Peak enrichment region plots -->
<!-- ```{r} -->
<!-- dir_export <- "../data/processed" -->

<!-- ##################### -->
<!-- # Prep tag matrix -->
<!-- ##################### -->

<!-- promoter <- getPromoters(TxDb = ensDb, upstream = 3000, downstream = 3000) -->
<!-- seqlevelsStyle(promoter) <- "UCSC" -->

<!-- tagMatrixList <- list() -->
<!-- for (i in names(peakList_cons)){ -->
<!--   tagMatrixList[[i]] <- getTagMatrix(peakList_cons[[i]], windows=promoter) -->
<!-- } -->


<!-- ##################### -->
<!-- # Make plots -->
<!-- ##################### -->

<!-- # Generate plot for peak number -->
<!-- df <- lapply(peakList_cons[c(1:3)], length) %>% as.data.frame() %>% gather("Sample", "Peaks") -->
<!-- df$Sample <- gsub("X", "", df$Sample) -->
<!-- df$Sample <- factor(df$Sample, levels = df$Sample) -->

<!-- b <- ggplot(df, aes(Sample, Peaks))+ -->
<!--   geom_bar(stat = "identity") + -->
<!--   theme_minimal() + -->
<!--   theme(axis.text.x = element_text(angle = 30, hjust = 0.8, vjust = 0.8)) + -->
<!--   labs(title = "Peaks") + xlab(NULL) -->

<!-- cowplot::ggsave2(plot = b, filename = "../data/processed/Peak_number_TP.pdf", height = 3, width = 2) -->


<!-- # Plot peak histogram around TSS -->
<!-- p <- plotAvgProf(tagMatrixList[c(1:3)], xlim=c(-3000, 3000), resmple = 100, -->
<!--             xlab="Genomic Region (5'->3')", ylab = "Read Count Frequency") -->
<!-- print(p) -->

<!-- cowplot::ggsave2(p, filename = file.path(dir_export, "01_Peak_TSS_hist_TP.pdf"),  -->
<!--                 device = "pdf",  -->
<!--                  height = 4.5, width = 6) -->


<!-- # Plot annotation bargraph -->
<!-- pp <- plotAnnoBar(peakAnnoList_ens[c(1:3)], title = "H3K18la peak annotation") -->
<!-- print(pp) -->

<!-- cowplot::ggsave2(pp, filename = file.path(dir_export, "01_Peak_annot_bar_TP.pdf"),  -->
<!--                 device = "pdf",  -->
<!--                  height = 2, width = 6) -->


<!-- # Plot disrtibution of peaks realtive to TSS -->
<!-- ppp <- plotDistToTSS(peakAnnoList_ens[c(1:3)], -->
<!--               title="Distribution of H3K18la peak to TSS") -->
<!-- print(ppp) -->

<!-- cowplot::ggsave2(ppp, filename = file.path(dir_export, "01_Peak_TSSdist_bar_TP.pdf"),  -->
<!--                 device = "pdf",  -->
<!--                  height = 2, width = 6) -->

<!-- # Heatmap 1000 -->
<!-- promoter <- getPromoters(TxDb = ensDb, upstream = 1000, downstream = 1000) -->
<!-- seqlevelsStyle(promoter) <- "UCSC" -->

<!-- tagMatrixList2 <- list() -->
<!-- for (i in names(peakList_cons)[1:3]){ -->
<!--   tagMatrixList2[[i]] <- getTagMatrix(peakList_cons[[i]], windows=promoter) -->
<!-- } -->

<!-- # Heatmap -->
<!-- pdf(file = file.path(dir_export, "01_Peak_TSSdist_heatmap_TP.pdf"), height = 4, width = 5) -->
<!-- tagHeatmap(tagMatrixList2[c(1:3)], xlim=c(-1000, 1000), color = "grey25") -->
<!-- dev.off() -->

<!-- # print(h) -->
<!-- # cowplot::ggsave2(h, filename = file.path(dir_export, "01_Peak_TSSdist_heatmap_TP.pdf"),  -->
<!-- #                 device = "pdf",  -->
<!-- #                  height = 4, width = 6) -->
<!-- ``` -->


# FGE of annotated genes
```{r}

library(clusterProfiler)
library(enrichplot)
library(org.Mm.eg.db)
source(file.path(dest_tools_dir, "GO_leaf.R"))
# source("../src/tools/Dotplot_GO.R")

dir_export <- "../data/processed"

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_enrich <- file.path(dir_export, paste0("01_FuncGeneEnrich_", enrich_db, "_", GOont))
dir.create(dir_enrich)


# Get list of genes for enrichment analysis

# goi.list <- list(P4 = peakAnnoList_ens[[1]]@anno$SYMBOL %>% unique(),
#                  P10 = peakAnnoList_ens[[2]]@anno$SYMBOL %>% unique(),
#                  P28 = peakAnnoList_ens[[3]]@anno$SYMBOL %>% unique())

goi.list <- list(P4 = peakAnnoList_ens[[1]]@anno$SYMBOL %>% unique(),
                 P10 = peakAnnoList_ens[[2]]@anno$SYMBOL %>% unique(),
                 P28 = peakAnnoList_ens[[3]]@anno$SYMBOL %>% unique())

# # Get universe
# univ <- row.names(seu_rodbp@assays$decontXcounts@data)

# Get enrichment
enrich_results <- clusterProfiler::compareCluster(geneClusters = goi.list, fun = "enrichGO",
                                            keyType = "SYMBOL", ont = GOont, OrgDb = org)





###########################
# Look at GO Level 6
###########################

go_level <- 8
go_ids <- GOxploreR::Level2GOTermBP(go_level, organism = "mouse")
top <- 10

# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.0001, Count > 2)

enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Prep the enrichment DF
enrich <- enrich_results_filt_fix %>% 
  as.data.frame() %>% 
  dplyr::filter(ID %in% go_ids)
enrich$Description <- factor(enrich$Description, levels = rev(unique(enrich$Description)))


# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich %>% 
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich %>% 
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>% 
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
p <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1))


#Open the pdf device
cowplot::ggsave2(plot = p, filename = file.path(dir_enrich, paste0("Timepoint_FGE_top_", top, "_level_", go_level, "_0001_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)



###########################
# Look at leaf terms
###########################


# # Filter enrichment for significance
# enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.0001, Count > 2)

# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# # Fix the Ratio
# enrich_results_leaf_fix <- enrich_results_leaf %>%
#   as.data.frame() %>%
#   mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>%
#   rowwise() %>%
#   mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>% 
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>% 
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("Timepoint_FGE_top_", top, "_leaf_0001_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)


# 
# # Make dotplot of results
# dot_go(enrich_results_leaf, top = 10, grp = "Cluster", desc = "Description", p.val = "p.adjust", ratio = "Ratio", filenm = file.path(dir_enrich, "plots", paste0(i, "_cluster_dotplot_05")))
# 
# dot_go(enrich_results_leaf, top = 5, grp = "Cluster", desc = "Description", p.val = "p.adjust", ratio = "Ratio", filenm = file.path(dir_enrich, "plots", paste0(i, "_cluster_dotplot_05_5")))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", "timepoint", "_filt.tsv")),
              quote = F, row.names = F, sep = "\t")

# Export the FGE results table
# tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", "timepoint", "_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")


```


#
### Session Information
```{r, echo=FALSE}
sessionInfo()
```

#
## Explanation of Output Files

#### File structure of the data

* data
    + external
    + interim
    + processed   
    + raw
        - bams
        - beds
        - bigwig
* notebooks

#
#### Output File Description

Output files from analysis contain both tab-separated values table (.tsv), as well as, plots in pdf form. The tsv tables can be opened in Excel.

#
##### 'data/external'
* DevRet_Gene38v84_nCPM_MSTR.tsv
    + Gene expression values from deveolping mouse retina. Values are normalized CPM values. File is tab-separated value file.

#
##### 'data/interim'
Output from combining replicate peaks for the concensus bed file used for further analysis

* beds
    + Consensus bed files for each timepoint grouping using overlap method
    + cons
        - Consensus bed files for each timepoint using the cons method (*narrowPeak)
* bigwig
    + Merged bigwig files using samples from cons samples
        
#
##### 'data/processed'

* 01_PeakGeneAnno_*.txt
    + Peak annotation at each tf-timepoint
* gProf_AllPeaks
    + Directory containing the export table and visualization results from the functional gene enrichment analysis from genes containing peaks from the consensus bed files for the timepoint groupings
        - gProfileR_all.tsv - output from gProfileR
        - Heatmaps : heatmaps of all the top enriched categories using log2 CPM values 
* gProf_AllPeaks_Nrl_PN
    + Replot of the dotplots using all timepoints except Nrl E14
* gProf_[Crx|Nrl]
    + Replot of the dotplots separarting Crx and Nrl
* gProf_Nrl_PN
    + Replot of the dotplots separating Nrl and excluding the E14 timepoint

#
##### 'data/raw

* bams
    + Bam files generated from the pipeline for each sample.
* beds
    + The 1% FDR Macs peak bed files for each sample used in the analysis
* bigwig
    + Bigwig file exports for each sample. The bigwig file is a bin per million normalization. 