---
title: "H3K18la Alluvial Analysis - Sept 18th, 2024"
output: html_notebook

---

### Project Summary
This project investigates the H3K18la during development. 

**DESIGN**:  

**NOTES**: The data were generated from the ENCODE histone pipeline with duplicate 
removal with no controls. Alluvial plot of histone overlaps being generated by 
quantitation of reads over mstr_features.bed then filtering for binding prior
to generation of alluvial plot.



	

#### Get custom scripts and functions and outside data used for analysis
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

script_dir <- "https://raw.githubusercontent.com/MBrooks313/R_codes/master"
dest_tools_dir <- "../src/tools"
dest_vis_dir <- "../src/visualization"

# download.file(url = file.path(script_dir, "GO_leaf.R"),
#               destfile = file.path(dest_tools_dir, "GO_leaf.R"),
#               method = "curl")
# 
download.file(url = file.path(script_dir, "DFave.R"),
              destfile = file.path(dest_tools_dir, "DFave.R"),
              method = "curl")
# 
# download.file(url = file.path(script_dir, "Heatmap.R"),
#               destfile = file.path(dest_tools_dir, "Heatmap.R"),
#               method = "curl")

download.file(url = file.path(script_dir, "Expr_Filter.R"),
              destfile = file.path(dest_tools_dir, "Expr_Filter.R"),
              method = "curl")

download.file(url = file.path(script_dir, "Barplot_GO.R"),
              destfile = file.path(dest_tools_dir, "Barplot_GO.R"),
              method = "curl")

# system('cp ~/github/R_codes/DFave.R ../src/tools')
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/et_edgeR_v0.1.0.R ../src/tools")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/FuncGeneEnrich_v0.1.0.R ../src/tools")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/GO_leaf_v1.1.R ../src/tools")

#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/APk_cluster_BF_v0.1.0.R ../src/tools')
#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/HM_PCAsort_BF_v0.1.0.R ../src/visualization')
#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/parPlot_hc_v0.2.0.R ../src/visualization')
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/PCA_simple_v0.1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/VolcanoPlot_v0.1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_v1.7.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Dotplot_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Barplot_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_v1.7.R ../src/visualization")
# 
# system("cp /Volumes/NGS/RNA-seq/ENS_v98/Mouse/Counts/by_project/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv ../data/external")
# system("cp /Volumes/NGS/RNA-seq/ENS_v98/Mouse/Counts/by_project/Mouse_FSPR_WT_Development_Gene_nCPM_ENSv98.tsv ../data/external")

# Function to convert ENS ids to gene symbol 
# ens2sym <- function(x){paste(sort(unique(AnnotationDbi::mapIds(org.Mm.eg.db, keys = x, keytype = "ENSEMBL", column = "SYMBOL"))), sep = "", collapse = ",")}
library(tidyverse)
# options(java.parameters = "-Xmx8000m")
print("Completed...")
```

<!-- # Get Ensembl annotation database -->
<!-- ```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE} -->
<!-- library(AnnotationHub) -->
<!-- library(ensembldb) -->

<!-- ah <- AnnotationHub() -->

<!-- ## Query for all available EnsDb databases -->
<!-- ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102)) -->

<!-- ensDb <- ahdb[[1]] -->

<!-- # supportedFilters(ensDb) -->

<!-- ``` -->


# Get Ensembl annotation database
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

## Query for all available EnsDb databases
ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

ensDb <- ahdb[[1]]

# supportedFilters(ensDb)

```


################################################################################
## Import outside datasets
################################################################################
```{r}

source(file.path(dest_tools_dir, "DFave.R"))

#############################################
# Import expression data for developing retina
retina.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)

# Get group info for averaging
grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.dwnld)[10:31])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
retina_cpm <- data.frame(merge(retina.dwnld[,1:9],
                               log2(DFave(retina.dwnld[,10:31], exp_grp) + 1),
                               by = 0),
                         row.names = 1)


#############################################
# Import the H3K27ac from Aldiri
k27_aldiri_beds <- list.files(path = "/Volumes/data/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bed/", pattern = "*H3K27Ac*", recursive = T, full.names = T)
names(k27_aldiri_beds) <- gsub(".+\\/(.+)_H3K27.+", "\\1", k27_aldiri_beds)

# Import the beds 
k27_aldiri <- sapply(k27_aldiri_beds, toGRanges, format="narrowPeak")

for (i in 1:length(k27_aldiri)){
  k27_aldiri[[i]] <- GenomicRanges::reduce(k27_aldiri[[i]])
  seqlevelsStyle(k27_aldiri[[i]]) <- "NCBI"
}


#############################################
# Import the H3K4me3 from Aldiri
k4me3_aldiri_beds <- list.files(path = "/Volumes/data/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bed/", pattern = "*H3K4me3_*", recursive = T, full.names = T)
names(k4me3_aldiri_beds) <- gsub(".+\\/(.+)_H3K4me3.+", "\\1", k4me3_aldiri_beds)

# Import the beds 
k4me3_aldiri <- sapply(k4me3_aldiri_beds, toGRanges, format="narrowPeak")

for (i in 1:length(k4me3_aldiri)){
  k4me3_aldiri[[i]] <- GenomicRanges::reduce(k4me3_aldiri[[i]])
  seqlevelsStyle(k4me3_aldiri[[i]]) <- "NCBI"
}



#############################################
# Import the time course sig DB regions
tc_prom_pc_sig <- readRDS(file = "../data/interim/03_prom_pc_sig_240913.RDS")
posi_cor <- readRDS(file = "../data/interim/03_h3k18la_db_dge_positive_corr_241001.RDS")
anti_cor <- readRDS(file = "../data/interim/03_h3k18la_db_dge_anti_corr_241001.RDS")

tc_peakAnno_cons <- readRDS("../data/interim/peakAnnoList_cons_240911.RDS")
# retina_z <- data.frame(merge(retina_cpm[,1:9],
#                                t(scale(t(retina_cpm[,10:19]), scale = T, center = T)),
#                                by = 0),
#                          row.names = 1)
                         
```                   
                         

################################################################################
# Annotate outside data
################################################################################
```{r}

# Aldiri k27ac 
k27AnnoList_ens <- list()

for (i in 1:length(k27_aldiri)){
  
  peak_tmp <- k27_aldiri[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  # peakAnnoList_ens[[i]] <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
  #                        TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T,
  #                        overlap = "all")
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  # anno_tmp@anno$GENENAME <- mapIds(x = ensDb, 
  #                                        keys = anno_tmp@anno$geneId, 
  #                                        column = "GENENAME", 
  #                                        keytype = "GENEID")
  k27AnnoList_ens[[i]] <- anno_tmp

}

names(k27AnnoList_ens) <- names(k27_aldiri)


# Aldiri k4me3 
k4me3AnnoList_ens <- list()

for (i in 1:length(k4me3_aldiri)){
  
  peak_tmp <- k4me3_aldiri[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  # peakAnnoList_ens[[i]] <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
  #                        TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T,
  #                        overlap = "all")
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  # anno_tmp@anno$GENENAME <- mapIds(x = ensDb, 
  #                                        keys = anno_tmp@anno$geneId, 
  #                                        column = "GENENAME", 
  #                                        keytype = "GENEID")
  k4me3AnnoList_ens[[i]] <- anno_tmp

}

names(k4me3AnnoList_ens) <- names(k4me3_aldiri)


```



################################################################################
## Get the counts over annotated feature regions
################################################################################
```{r}
library(tidyverse)
library(ChIPpeakAnno)
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)
library(csaw)

############################
# Master features
mstr_features_anno <- readRDS(file = "../data/interim/03_mstr_features_anno_240913.RDS")
mstr_features_anno_ucsc <- mstr_features_anno
seqlevels(mstr_features_anno_ucsc@anno) <- paste0("chr", seqlevels(mstr_features_anno@anno))

# Params
pe.param <- readParam(max.frag=600, pe="both", minq = 20, dedup = T)
se.param <- readParam(max.frag=600, pe="none", minq = 20, dedup = T)

############################
# H3K18la

# Get bam files
bam_dir <- "../data/raw/bams/"
bams <- list.files(path = bam_dir, pattern = "H3K18La_P.+.bam$")
bams <- bams[c(6:7, 1:2, 3:5)]
bam.files <- paste0(bam_dir, bams)

# Annotated region counts
h3k18la_counts <- regionCounts(bam.files, mstr_features_anno@anno, ext=150, param=pe.param)

# Normalization
binned <- windowCounts(bam.files, bin=TRUE, width=10000, param=se.param)
h3k18la_se <- normFactors(binned, se.out=T)


############################
# H3K27ac

# Get bam files
bam_dir <- "/Volumes/data/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bam/"
bams <- list.files(path = bam_dir, pattern = ".+H3K27Ac.+.bam$")
bams <- bams[c(16:17, 10:11, 1:2)]
bam.files <- paste0(bam_dir, bams)

# Annotated region counts
h3k27ac_counts <- regionCounts(bam.files, mstr_features_anno_ucsc@anno, ext=150, param=se.param)

# Normalization
binned <- windowCounts(bam.files, bin=TRUE, width=10000, param=se.param)
h3k27ac_se <- normFactors(binned, se.out=T)


############################
# H3K4me3

# Get bam files
bam_dir <- "/Volumes/data/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bam/"
bams <- list.files(path = bam_dir, pattern = ".+H3K4me3.+.bam$")
bams <- bams[c(16:17, 9:10, 1:2)]
bam.files <- paste0(bam_dir, bams)

# Annotated region counts
h3k4me3_counts <- regionCounts(bam.files, mstr_features_anno_ucsc@anno, ext=150, param=se.param)

# Normalization
binned <- windowCounts(bam.files, bin=TRUE, width=10000, param=se.param)
h3k4me3_se <- normFactors(binned, se.out=T)



```



################################################################################
#### Prep the DGE list objects
################################################################################
```{r}
library(edgeR)

# source(file.path(dest_tools_dir, "DFave.R"))
source(file.path(dest_tools_dir, "Expr_Filter.R"))


##########
## h3k18la
samp_nm <- gsub(".+(H3K18La_.+)\\.target.+", "\\1", h3k18la_counts@colData[[1]])

h3k18la.dge <- asDGEList(h3k18la_counts,
               norm.factors = h3k18la_se$norm.factors,
               group = gsub("H3K18La_(.+)_.+", "\\1", samp_nm),
               genes = as.data.frame(mstr_features_anno))
colnames(h3k18la.dge) <- samp_nm
h3k18la.dge$samples$group <- factor(h3k18la.dge$samples$group, levels = unique(h3k18la.dge$samples$group))

h3k18la.dge$rpkm <- rpkm(h3k18la.dge, gene.length = "width")

h3k18la.dge$rpkm_mean <- h3k18la.dge$rpkm %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(P4 = mean(c(H3K18La_P4_R1, H3K18La_P4_R2)),
         P10 = mean(c(H3K18La_P10_R1, H3K18La_P10_R2)),
         P28 = mean(c(H3K18La_P28_R1, H3K18La_P28_R2, H3K18La_P28_R3)))

# ExpFilt(h3k18la.dge$rpkm, exp = 0.0, group = h3k18la.dge$samples$group)
# 
# h3k18la.dge.filt <- h3k18la.dge[idx.filt, keep.lib.sizes = T]
# h3k18la.dge.filt$rpkm_mean <- h3k18la.dge.filt$rpkm_mean[idx.filt,]

h3k18la.pc.prom <- cbind(h3k18la.dge$genes, h3k18la.dge$rpkm_mean) %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")




##########
## h3k27ac
samp_nm <- gsub(".+([Pa].+H3K27Ac).+", "\\1", h3k27ac_counts@colData[[1]])
samp_nm <- paste(samp_nm, rep(c("R1", "R2"), 3), sep = "_")
h3k27ac.dge <- asDGEList(h3k27ac_counts,
               norm.factors = h3k27ac_se$norm.factors,
               group = gsub("_.+", "", samp_nm),
               genes = as.data.frame(mstr_features_anno))
colnames(h3k27ac.dge) <- samp_nm
h3k27ac.dge$samples$group <- factor(h3k27ac.dge$samples$group, levels = unique(h3k27ac.dge$samples$group))

h3k27ac.dge$rpkm <- rpkm(h3k27ac.dge, gene.length = "width")

h3k27ac.dge$rpkm_mean <- h3k27ac.dge$rpkm %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(P4 = mean(c(P3_H3K27Ac_R1, P3_H3K27Ac_R2)),
         P10 = mean(c(P10_H3K27Ac_R1, P10_H3K27Ac_R2)),
         P28 = mean(c(adult_H3K27Ac_R1, adult_H3K27Ac_R2)))

h3k27ac.pc.prom <- cbind(h3k27ac.dge$genes, h3k27ac.dge$rpkm_mean) %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")


##########
## h3k4me3
samp_nm <- gsub(".+([Pa].+H3K4me3).+", "\\1", h3k4me3_counts@colData[[1]])
samp_nm <- paste(samp_nm, rep(c("R1", "R2"), 3), sep = "_")
h3k4me3.dge <- asDGEList(h3k4me3_counts,
               norm.factors = h3k4me3_se$norm.factors,
               group = gsub("_.+", "", samp_nm),
               genes = as.data.frame(mstr_features_anno))
colnames(h3k4me3.dge) <- samp_nm
h3k4me3.dge$samples$group <- factor(h3k4me3.dge$samples$group, levels = unique(h3k4me3.dge$samples$group))

h3k4me3.dge$rpkm <- rpkm(h3k4me3.dge, gene.length = "width")

h3k4me3.dge$rpkm_mean <- h3k4me3.dge$rpkm %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(P4 = mean(c(P3_H3K4me3_R1, P3_H3K4me3_R2)),
         P10 = mean(c(P10_H3K4me3_R1, P10_H3K4me3_R2)),
         P28 = mean(c(adult_H3K4me3_R1, adult_H3K4me3_R2)))

h3k4me3.pc.prom <- cbind(h3k4me3.dge$genes, h3k4me3.dge$rpkm_mean) %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")

```


################################################################################
# Make alluvial plots and associated plots based on quantitation of histone marks
# against the K18la mstr_features
################################################################################


##################
# Alluvial plot of promoter peak histone presence
##################
```{r}
library(ggalluvial)


thresh = 2
# Prep the gene lists for the FGE loop
gene_list_P4 <- list(H3K27ac = h3k27ac.pc.prom %>% 
                       dplyr::filter(P4 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K4me3 = h3k4me3.pc.prom %>% 
                       dplyr::filter(P4 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K18la = h3k18la.pc.prom %>% 
                       dplyr::filter(P4 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F))

gene_list_P10 <- list(H3K27ac = h3k27ac.pc.prom %>% 
                       dplyr::filter(P10 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K4me3 = h3k4me3.pc.prom %>% 
                       dplyr::filter(P10 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K18la = h3k18la.pc.prom %>% 
                       dplyr::filter(P10 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F))

gene_list_P28 <- list(H3K27ac = h3k27ac.pc.prom %>% 
                       dplyr::filter(P28 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K4me3 = h3k4me3.pc.prom %>% 
                       dplyr::filter(P28 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K18la = h3k18la.pc.prom %>% 
                       dplyr::filter(P28 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F))




goi_list <- list(gene_list_P4, gene_list_P10, gene_list_P28)
names(goi_list) <- c("P4", "P10", "P28")

total_gois <- c(gene_list_P4 %>% unlist(use.names = F) %>% unique(),
                gene_list_P10 %>% unlist(use.names = F) %>% unique(),
                gene_list_P28 %>% unlist(use.names = F) %>% unique()) %>% unique()


# Prep DF for alluvial plot
df_alluv <- data.frame()

# 
for (i in names(goi_list)){
  
  tmp_venn <- ggVennDiagram::Venn(goi_list[[i]])
  tmp_data <- ggVennDiagram::process_data(tmp_venn)
  tmp_regions <- tmp_data@region$item
  names(tmp_regions) <- tmp_data@region$name

  for (ii in names(tmp_regions)){
    
    tmp_genes <- tmp_regions[[ii]]
    tmp_df <- data.frame(Symbol = tmp_genes,
                         TP = i,
                         Region = ii)
   
    
    df_alluv <- rbind(df_alluv, tmp_df)
      
  }
  
   tmp_df_missing <- data.frame(Symbol = total_gois[!(total_gois %in% (tmp_regions %>% unlist(use.names = F) %>% unique()))],
                                TP = i,
                                Region = NA)
   df_alluv <- rbind(df_alluv, tmp_df_missing)
}

df_alluv$TP <- factor(df_alluv$TP, levels = c("P4", "P10", "P28"))
df_alluv <- df_alluv %>% as.tibble() %>% replace_na(list(Region = "no_mark"))

hist_barplot_levels <- df_alluv$Region %>% unique() %>% sort()
hist_barplot_levels <- hist_barplot_levels[c(8,1,6,2,7,3,4,5)]
# hist_barplot_levels[1] <- NA
df_alluv$Region <- factor(df_alluv$Region, levels = hist_barplot_levels)

gga <- ggplot(df_alluv, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) +
  # scale_fill_brewer(type = "qual", palette = "Set2") +
  # scale_fill_viridis_b() +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) +
  geom_stratum(alpha = 0.75) +
  theme(legend.position = "bottom") +
  ggtitle("Histone occupancy in promoter region during development") 
cowplot::ggsave2(gga, filename = "../data/processed/08_Alluvial_Prom_18la-27ac-4me3_241008_filt2.pdf",
                 height = 8, width = 12)

saveRDS(object = df_alluv, file = "../data/interim/08_df_alluvial_241008_filt2.RDS")

```


###################################
## Perform FGE on Alluvial segments
###################################
```{r}

source(file.path(dest_tools_dir, "GO_Leaf.R"))

df_alluv_spread <- df_alluv %>% spread(TP, Region)


goi_list_alluv <- list(all3_allTP = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3..H3K18la", 
                                       P10 == "H3K27ac..H3K4me3..H3K18la", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P4.all3_P10.K4..K27_P28.all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3..H3K18la", 
                                       P10 == "H3K27ac..H3K4me3", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P4.K4..K27_P10.all3_P28.all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3", 
                                       P10 == "H3K27ac..H3K4me3..H3K18la", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P4.K4..K27_P10.K4..K27_P28.all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3", 
                                       P10 == "H3K27ac..H3K4me3", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       K4..K27_allTP = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3", 
                                       P10 == "H3K27ac..H3K4me3", 
                                       P28 == "H3K27ac..H3K4me3") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28.K27..K18 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K27ac..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28.K4..K18 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28.K18 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F))


# Perform FGE on the GOIs from each subset of peaks

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 12

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_export <- "../data/processed"
dir_enrich <- file.path(dir_export, paste0("08_Alluvial_Regions_241009_", enrich_db, "_", GOont))
dir.create(dir_enrich)

# univ <- df_alluv_spread$Symbol
univ <- mstr_features_anno@anno$SYMBOL %>% unique() 
  
enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list_alluv, fun = "enrichGO",
                                          keyType = "SYMBOL", ont = GOont, OrgDb = org, univ = univ)
# enrich_results2 <- clusterProfiler::compareCluster(geneClusters = goi_list_alluv, fun = "enrichGO",
                                          # keyType = "SYMBOL", ont = GOont, OrgDb = org, univ = univ2)


# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>% 
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>% 
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("H3K18la Peaks in ", i))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0(i, "_Alluvial-Region_FGE_top_", top, "_leaf_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)
    
```


###################################
## Perform FGE on Alluvial segments - redo based on discussions from Sep 25, 2024 Mtg
###################################
```{r}

source(file.path(dest_tools_dir, "GO_Leaf.R"))

df_alluv_spread <- df_alluv %>% spread(TP, Region)


goi_list_alluv <- list(P4_all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P10_acquired_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(!(P4 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P10 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_acquired_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(!(P10 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la_H3K4me3 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la_H3K27ac = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K27ac..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       H3K18la_acquired = df_alluv_spread %>% 
                         dplyr::filter(!(P4 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P10 == "H3K27ac..H3K4me3..H3K18la",
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F))


# Perform FGE on the GOIs from each subset of peaks

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 12

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_export <- "../data/processed"
dir_enrich <- file.path(dir_export, paste0("08_Alluvial_Regions_241009v2_", enrich_db, "_", GOont))
dir.create(dir_enrich)

# univ <- df_alluv_spread$Symbol
univ <- mstr_features_anno@anno$SYMBOL %>% unique() 
  
enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list_alluv, fun = "enrichGO",
                                          keyType = "SYMBOL", ont = GOont, OrgDb = org, univ = univ)


# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_full.tsv")),
              quote = F, row.names = F, sep = "\t")



# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Simplify the ernichment
simp <- clusterProfiler::simplify(enrich_results)
write.table(simp@compareClusterResult, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_simp.tsv")),
              quote = F, row.names = F, sep = "\t")


# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>% 
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>% 
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("H3K18la Peaks in ", i))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0(i, "_Alluvial-Region_FGE_top_", top, "_leaf_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)
    

# Make a bargraph of top 25 from simplified enriched results
source("../src/tools/Barplot_GO.R")

bar_plot_list <- list()
top = 50
for (i in simp@compareClusterResult$Cluster %>% unique()){
  tmp_res <- simp@compareClusterResult %>% dplyr::filter(Cluster == i)
  tmp_res <- tmp_res %>% slice_head(n = top)
  tmp_res$Description <- factor(tmp_res$Description, levels = rev(tmp_res$Description))
  bp <- ggplot(tmp_res, aes(x = Description, y = Count, fill = -log10(p.adjust))) +
    geom_bar(stat = "identity") +
    coord_flip() +
   # scale_fill_gradient(low="#3B9AB2", high="#F21A00", name="P Value \n(-log10)",limits=c(0,10), oob = scales::squish) +
     scale_fill_gradient2(low="#225ea8", mid="#7fcdbb", high="#ffffd9", midpoint = 5, name="P Value \n(-log10)",limits=c(0,10), oob = scales::squish) +
    labs(y="Number of Significant Genes", x=NULL) +
    theme_bw()
  cowplot::ggsave2(plot = bp, filename = file.path(dir_enrich, paste0(i, "_Alluvial-Region_FGEsimp_top_", top, "_barplotv2.pdf")),
                 height = 12, 
                 width = 10)
  # bar_go(tmp_res, top = 25, desc = "Description", p.val = "p.adjust", genenum = "Count", filenm = file.path(dir_enrich, paste0(i, "_FGE_simp_", i,"_top_", top, "_barplat.pdf")))
}
# bar_go(simp@compareClusterResult, top = 25, desc = "Description", p.val = "p.adjust", genenum = "Count")




```



###########################################################################
# Figure 4 - Euler/Boxplot of promoter genes from H3K18la, H3K27ac, H3K4me3
###########################################################################
```{r}
library(ggVennDiagram)
library(eulerr)

# P28
venn <- Venn(gene_list_P28)
data <- process_data(venn)
 
data_item <- data@region$item
names(data_item) <- data@region$name


df <- data.frame()
tp = "P28"

for (i in names(data_item)){
  tmp_data <- retina_cpm %>% dplyr::filter(external_gene_name %in% data_item[[i]])
  tmp_tidy <- data.frame(Section = i, 
                         Gene = tmp_data$external_gene_name, 
                         CPM = tmp_data[tp], row.names = NULL)
  df <- rbind(df, tmp_tidy)
}

# Add no_mark gene quant to dataframe
no_mark <- data_item %>% unlist(use.names = F)
tmp_data <- retina_cpm %>% dplyr::filter(!(external_gene_name %in% no_mark) & gene_biotype == "protein_coding")
tmp_tidy <-  data.frame(Section = "no_mark", 
                         Gene = tmp_data$external_gene_name, 
                         CPM = tmp_data[tp], row.names = NULL)

df <- rbind(df, tmp_tidy)

df$Section <- factor(df$Section, levels = c(names(data_item)[c(7,4:6,1:3)], "no_mark") %>% rev())

tmp_count <- df %>% count(Section) %>% dplyr::select(n)
tmp_count <- paste0("n=", tmp_count$n)

gb <- ggplot(df, aes(x = Section, y = P28, color = Section)) +
  geom_boxplot() + coord_flip() + theme_classic() + theme(legend.position = "none") +
  annotate("text", label = tmp_count, x = seq(1:8), y = 16) +
  guides(color = guide_legend(reverse=T)) + ylab(label = "CPM (log2)") + ggtitle(paste0(tp, " gene expression with presence of histone marks in promoter"))

cowplot::ggsave2(gb, filename = file.path(dir_export, paste0("08_Box_P28_Prom_18la-27ac-4me3.pdf")),
                 height = 6, width = 6)



# P28
tmp <- lapply(gene_list_P28, function(x){x[!duplicated(x)]})
eu <- euler(tmp)
eup <- plot(eu, quantities = list(type = "percent"))

cowplot::ggsave2(eup, filename = file.path(dir_export, paste0("08_Euler_P28_Prom_18la-27ac-4me3.pdf")),
                 height = 4, width = 4)

```



################################################################################
## RNAseq correlated Alluvial plot
################################################################################
```{r}
library(ggalluvial)

# Load alluvial data.frame from 02
# df_alluv <- readRDS("../data/interim/02_df_alluvial.RDS")

# Correlated genes
df_alluv_corr <- df_alluv %>% 
  dplyr::filter(Symbol %in% posi_cor$SYMBOL)

gga <- ggplot(df_alluv_corr, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) +
  # scale_fill_brewer(type = "qual", palette = "Set2") +
  # scale_fill_viridis_b() +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) +
  geom_stratum(alpha = 0.75) +
  theme(legend.position = "bottom") +
  ggtitle("Histone occupancy in correlated promoter region during development") 

cowplot::ggsave2(gga, filename = file.path(dir_export, paste0("08_Alluvial_Corr_Prom_18la-27ac-4me3.pdf")),
                 height = 8, width = 12)


# Anti-Correlated genes
df_alluv_anticorr <- df_alluv %>% 
  dplyr::filter(Symbol %in% anti_cor$SYMBOL)

gga <- ggplot(df_alluv_anticorr, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) +
  # scale_fill_brewer(type = "qual", palette = "Set2") +
  # scale_fill_viridis_b() +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) +
  geom_stratum(alpha = 0.75) +
  theme(legend.position = "bottom") +
  ggtitle("Histone occupancy in anti-correlated promoter region during development") 

cowplot::ggsave2(gga, filename = file.path(dir_export, paste0("08_Alluvial_AntiCorr_Prom_18la-27ac-4me3.pdf")),
                 height = 8, width = 12)


# Un-Correlated genes
df_alluv_uncorr <- df_alluv %>% 
  dplyr::filter(!(Symbol %in% c(posi_cor$SYMBOL, anti_cor$SYMBOL)))

gga <- ggplot(df_alluv_uncorr, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) +
  # scale_fill_brewer(type = "qual", palette = "Set2") +
  # scale_fill_viridis_b() +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) +
  geom_stratum(alpha = 0.75) +
  theme(legend.position = "bottom") +
  ggtitle("Histone occupancy in uncorrelated promoter region during development") 

cowplot::ggsave2(gga, filename = file.path(dir_export, paste0("08_Alluvial_UnCorr_Prom_18la-27ac-4me3.pdf")),
                 height = 8, width = 12)


```


################################################################################
## Perform FGE on Alluvial segments
################################################################################
```{r}
source("../src/tools/GO_leaf.R")

df_alluv_spread <- df_alluv_corr %>% spread(TP, Region)


goi_list_alluv <- list(P4_all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P10_acquired_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(!(P4 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P10 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_acquired_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(!(P10 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la_H3K4me3 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la_H3K27ac = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K27ac..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       H3K18la_acquired = df_alluv_spread %>% 
                         dplyr::filter(!(P4 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P10 == "H3K27ac..H3K4me3..H3K18la",
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F))


# Perform FGE on the GOIs from each subset of peaks

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 10

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_enrich <- file.path(dir_export, paste0("08_Alluvial_Regions_Corr_", enrich_db, "_", GOont))
dir.create(dir_enrich)

# univ <- df_alluv_spread$Symbol
univ <- mstr_features_anno@anno$SYMBOL %>% unique()
  
enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list_alluv, fun = "enrichGO",
                                          keyType = "SYMBOL", ont = GOont, OrgDb = org)


# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Simplify the ernichment
simp <- clusterProfiler::simplify(enrich_results)
write.table(simp@compareClusterResult, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_simp.tsv")),
              quote = F, row.names = F, sep = "\t")


# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get top 5 pathways per cell type to get union set of pathways
# top_union <- enrich_results_leaf %>%
top_union <- simp@compareClusterResult %>%
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top_union$ID %>% unique()))

# Prep the df for plotting
top_enrich <- top_enrich %>% 
  arrange(Cluster, p.adjust)
top_enrich$Description <- substring(top_enrich$Description, 1, 60)
top_enrich$Description <- factor(top_enrich$Description, levels = rev(unique(top_enrich$Description)))
top_enrich$p.adjust <- -log10(top_enrich$p.adjust)

# Make plot
pp <- ggplot(top_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("GO:BP in Histone Occupancy in Correlated peaks"))

#Export plot
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0(i, "_Alluvial-Region_FGE_top_", top, "_", i, "_simp_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)

# Make barplots for each section

for (i in top_union$Cluster){
  tmp_dat <- top_union %>% dplyr::filter(Cluster == i)
  tmp_dat$Description <- factor(tmp_dat$Description, levels = rev(tmp_dat$Description))
  
  bp <- ggplot(tmp_dat) +
      geom_bar(stat = 'identity', aes(x=Description, y = Count, fill = -log10(p.adjust))) +
      coord_flip() +
      scale_fill_gradient(low="#3B9AB2", high="#F21A00", name="P Value \n(-log10)",limits=c(0,10), oob = scales::squish) +
      #scale_fill_gradientn(colors = c("#3B9AB2", "#F21A00"), name="P Value (log10)", trans = 'reverse', limits=c(-50,0), oob=squish) +
      labs(y="Number of Significant Genes", x=NULL) +
      ggtitle(paste0("GO:BP - ", i)) +
      theme_bw()
  cowplot::ggsave2(plot = bp, filename = file.path(dir_enrich, paste0("Alluvial-Region_FGE_top_", top, "_", i, "_leaf_01_barplot.pdf")),
                 height = 4, 
                 width = 8)
}

top <- 10
for (i in simp@compareClusterResult$Cluster){
  tmp_dat <- simp@compareClusterResult %>% as.data.frame() %>% dplyr::filter(Cluster == i)
  tmp_dat <- tmp_dat %>% slice_head(n = top)
  tmp_dat$Description <- factor(tmp_dat$Description, levels = rev(tmp_dat$Description))
  
  bp <- ggplot(tmp_dat) +
      geom_bar(stat = 'identity', aes(x=Description, y = Count, fill = -log10(p.adjust))) +
      coord_flip() +
      # scale_fill_gradient(low="#3B9AB2", high="#F21A00", name="P Value \n(-log10)",limits=c(0,10), oob = scales::squish) +
      scale_fill_gradient2(low="#225ea8", mid="#7fcdbb", high="#ffffd9", midpoint = 5, name="P Value \n(-log10)",limits=c(0,10), oob = scales::squish) +
      #scale_fill_gradientn(colors = c("#3B9AB2", "#F21A00"), name="P Value (log10)", trans = 'reverse', limits=c(-50,0), oob=squish) +
      labs(y="Number of Significant Genes", x=NULL) +
      ggtitle(paste0("GO:BP - ", i)) +
      theme_bw()
  cowplot::ggsave2(plot = bp, filename = file.path(dir_enrich, paste0("Alluvial-Region_FGE_top_", top, "_", i, "_simp_01_barplot_v2.pdf")),
                 height = 4, 
                 width = 8)
}

tmp_res <- simp@compareClusterResult %>% dplyr::filter(Cluster == i)
  tmp_res <- tmp_res %>% slice_head(n = top)
  tmp_res$Description <- factor(tmp_res$Description, levels = rev(tmp_res$Description))

```











################################################################################
################################################################################
################################################################################
# Make alluvial plots and associated plots based on quantitation of K18la
# mstr_features and K27ac and K4me MACS peaks bed annotation
################################################################################

################################################################################
## Perform alluvial of Overlap the K27ac, K4me3 and K18la at adult/P28  
################################################################################
```{r}

################
# P4
idx_k18 <- h3k18la.pc.prom %>% 
  dplyr::filter(P4 > 1) %>% 
  row.names() %>% as.numeric()
idx_k27 <- k27AnnoList_ens$P3@anno %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
  dplyr::select(rowname) %>% 
  as.vector() %>% unlist(use.names = F) %>% as.numeric()
idx_k4 <- k4me3AnnoList_ens$P3@anno %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
  dplyr::select(rowname) %>% 
  as.vector() %>% unlist(use.names = F) %>% as.numeric()
       
k18_p4_pc_prom <- mstr_features_anno@anno[idx_k18,]        
k27_p4_pc_prom <- k27AnnoList_ens$P3@anno[idx_k27,]
k4_p4_pc_prom <- k4me3AnnoList_ens$P3@anno[idx_k4,]
      
k27ac_k4me3_k18la_ol_p4 <- findOverlapsOfPeaks(k18_p4_pc_prom, 
                                               k27_p4_pc_prom, 
                                               k4_p4_pc_prom, 
                                               connectedPeaks = "keepAll", maxgap = 100L)


################
# P10
idx_k18 <- h3k18la.pc.prom %>% 
  dplyr::filter(P10 > 1) %>% 
  row.names() %>% as.numeric()
idx_k27 <- k27AnnoList_ens$P10@anno %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
  dplyr::select(rowname) %>% 
  as.vector() %>% unlist(use.names = F) %>% as.numeric()
idx_k4 <- k4me3AnnoList_ens$P10@anno %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
  dplyr::select(rowname) %>% 
  as.vector() %>% unlist(use.names = F) %>% as.numeric()

k18_p10_pc_prom <- mstr_features_anno@anno[idx_k18,]        
k27_p10_pc_prom <- k27AnnoList_ens$P10@anno[idx_k27,]
k4_p10_pc_prom <- k4me3AnnoList_ens$P10@anno[idx_k4,]
      
k27ac_k4me3_k18la_ol_p10 <- findOverlapsOfPeaks(k18_p10_pc_prom, 
                                               k27_p10_pc_prom, 
                                               k4_p10_pc_prom, 
                                               connectedPeaks = "keepAll", maxgap = 100L)

################
# P28
idx_k18 <- h3k18la.pc.prom %>% 
  dplyr::filter(P28 > 1) %>% 
  row.names() %>% as.numeric()
idx_k27 <- k27AnnoList_ens$adult@anno %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
  dplyr::select(rowname) %>% 
  as.vector() %>% unlist(use.names = F) %>% as.numeric()
idx_k4 <- k4me3AnnoList_ens$adult@anno %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
  dplyr::select(rowname) %>% 
  as.vector() %>% unlist(use.names = F) %>% as.numeric()
                      
k18_p28_pc_prom <- mstr_features_anno@anno[idx_k18,]        
k27_p28_pc_prom <- k27AnnoList_ens$adult@anno[idx_k27,]
k4_p28_pc_prom <- k4me3AnnoList_ens$adult@anno[idx_k4,]
      
k27ac_k4me3_k18la_ol_p28 <- findOverlapsOfPeaks(k18_p28_pc_prom, 
                                               k27_p28_pc_prom, 
                                               k4_p28_pc_prom, 
                                               connectedPeaks = "keepAll", maxgap = 100L)

```


##################
# Alluvial plot of promoter peak histone presence
##################
```{r}
library(ggalluvial)

#############
# P4
tmp_ol <- k27ac_k4me3_k18la_ol_p4$overlappingPeaks
tmp_k18_k27 <- tmp_ol$`k18_p4_pc_prom///k27_p4_pc_prom` %>% data.frame() %>% dplyr::filter(!duplicated(SYMBOL))
tmp_k18_k4 <- tmp_ol$`k18_p4_pc_prom///k4_p4_pc_prom` %>% data.frame() %>% dplyr::filter(!duplicated(SYMBOL))
tmp_k27_k4 <- tmp_ol$`k27_p4_pc_prom///k4_p4_pc_prom` %>% data.frame() %>% dplyr::filter(!duplicated(SYMBOL))
tmp_uniq <- k27ac_k4me3_k18la_ol_p4$uniquePeaks %>% as.data.frame() %>% rownames_to_column() %>% rowwise() %>% mutate(source = str_split(rowname, "__")[[1]][1])
tp = "P4"

df_p4 <- rbind(data.frame(Symbol = inner_join(tmp_k18_k4, 
                                               tmp_k18_k27, 
                                               by = "SYMBOL") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "all3"),
                data.frame(Symbol = tmp_k18_k27 %>% 
                             dplyr::filter(! SYMBOL %in% tmp_k18_k4$SYMBOL) %>%
                             dplyr::select(SYMBOL) %>%
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K18la..H3K27ac"),
                data.frame(Symbol = tmp_k18_k4 %>% 
                             dplyr::filter(! SYMBOL %in% tmp_k18_k27$SYMBOL) %>%
                             dplyr::select(SYMBOL) %>%
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K18la..H3K4me3"),
                data.frame(Symbol = tmp_k27_k4 %>% 
                             dplyr::filter(! SYMBOL %in% tmp_k18_k27$SYMBOL) %>%
                             dplyr::select(SYMBOL) %>%
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K4me3..H3K27ac"),
                data.frame(Symbol = tmp_uniq %>% 
                             dplyr::filter(source == "k18_p4_pc_prom") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K18la"),
                data.frame(Symbol = tmp_uniq %>% 
                             dplyr::filter(source == "k27_p4_pc_prom") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K27ac"),
                data.frame(Symbol = tmp_uniq %>% 
                             dplyr::filter(source == "k4_p4_pc_prom") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K4me3")) 

#############
# P10
tmp_ol <- k27ac_k4me3_k18la_ol_p10$overlappingPeaks
tmp_k18_k27 <- tmp_ol$`k18_p10_pc_prom///k27_p10_pc_prom` %>% data.frame() %>% dplyr::filter(!duplicated(SYMBOL))
tmp_k18_k4 <- tmp_ol$`k18_p10_pc_prom///k4_p10_pc_prom` %>% data.frame() %>% dplyr::filter(!duplicated(SYMBOL))
tmp_k27_k4 <- tmp_ol$`k27_p10_pc_prom///k4_p10_pc_prom` %>% data.frame() %>% dplyr::filter(!duplicated(SYMBOL))
tmp_uniq <- k27ac_k4me3_k18la_ol_p10$uniquePeaks %>% as.data.frame() %>% rownames_to_column() %>% rowwise() %>% mutate(source = str_split(rowname, "__")[[1]][1])
tp = "P10"

df_p10 <- rbind(data.frame(Symbol = inner_join(tmp_k18_k4, 
                                               tmp_k18_k27, 
                                               by = "SYMBOL") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "all3"),
                data.frame(Symbol = tmp_k18_k27 %>% 
                             dplyr::filter(! SYMBOL %in% tmp_k18_k4$SYMBOL) %>%
                             dplyr::select(SYMBOL) %>%
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K18la..H3K27ac"),
                data.frame(Symbol = tmp_k18_k4 %>% 
                             dplyr::filter(! SYMBOL %in% tmp_k18_k27$SYMBOL) %>%
                             dplyr::select(SYMBOL) %>%
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K18la..H3K4me3"),
                data.frame(Symbol = tmp_k27_k4 %>% 
                             dplyr::filter(! SYMBOL %in% tmp_k18_k27$SYMBOL) %>%
                             dplyr::select(SYMBOL) %>%
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K4me3..H3K27ac"),
                data.frame(Symbol = tmp_uniq %>% 
                             dplyr::filter(source == "k18_p10_pc_prom") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K18la"),
                data.frame(Symbol = tmp_uniq %>% 
                             dplyr::filter(source == "k27_p10_pc_prom") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K27ac"),
                data.frame(Symbol = tmp_uniq %>% 
                             dplyr::filter(source == "k4_p10_pc_prom") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K4me3")) 


#############
# P28
tmp_ol <- k27ac_k4me3_k18la_ol_p28$overlappingPeaks
tmp_k18_k27 <- tmp_ol$`k18_p28_pc_prom///k27_p28_pc_prom` %>% data.frame() %>% dplyr::filter(!duplicated(SYMBOL))
tmp_k18_k4 <- tmp_ol$`k18_p28_pc_prom///k4_p28_pc_prom` %>% data.frame() %>% dplyr::filter(!duplicated(SYMBOL))
tmp_k27_k4 <- tmp_ol$`k27_p28_pc_prom///k4_p28_pc_prom` %>% data.frame() %>% dplyr::filter(!duplicated(SYMBOL))
tmp_uniq <- k27ac_k4me3_k18la_ol_p28$uniquePeaks %>% as.data.frame() %>% rownames_to_column() %>% rowwise() %>% mutate(source = str_split(rowname, "__")[[1]][1])
tp = "P28"

df_p28 <- rbind(data.frame(Symbol = inner_join(tmp_k18_k4, 
                                               tmp_k18_k27, 
                                               by = "SYMBOL") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "all3"),
                data.frame(Symbol = tmp_k18_k27 %>% 
                             dplyr::filter(! SYMBOL %in% tmp_k18_k4$SYMBOL) %>%
                             dplyr::select(SYMBOL) %>%
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K18la..H3K27ac"),
                data.frame(Symbol = tmp_k18_k4 %>% 
                             dplyr::filter(! SYMBOL %in% tmp_k18_k27$SYMBOL) %>%
                             dplyr::select(SYMBOL) %>%
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K18la..H3K4me3"),
                data.frame(Symbol = tmp_k27_k4 %>% 
                             dplyr::filter(! SYMBOL %in% tmp_k18_k27$SYMBOL) %>%
                             dplyr::select(SYMBOL) %>%
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K4me3..H3K27ac"),
                data.frame(Symbol = tmp_uniq %>% 
                             dplyr::filter(source == "k18_p28_pc_prom") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K18la"),
                data.frame(Symbol = tmp_uniq %>% 
                             dplyr::filter(source == "k27_p28_pc_prom") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K27ac"),
                data.frame(Symbol = tmp_uniq %>% 
                             dplyr::filter(source == "k4_p28_pc_prom") %>% 
                             dplyr::select(SYMBOL) %>% 
                             as.vector() %>% unlist(use.names = F),
                           TP = tp,
                           Region = "H3K4me3")) 


##################
# get df of NA per timepoint
goi_all <- c(df_p4$Symbol, df_p10$Symbol, df_p28$Symbol) %>% unique()

na_p4 <- data.frame(Symbol = goi_all[! goi_all %in% df_p4$Symbol],
                    TP = "P4",
                    Region = "no_mark")
na_p10 <- data.frame(Symbol = goi_all[! goi_all %in% df_p10$Symbol],
                    TP = "P10",
                    Region = "no_mark")
na_p28 <- data.frame(Symbol = goi_all[! goi_all %in% df_p28$Symbol],
                    TP = "P28",
                    Region = "no_mark")

df_alluv2 <- rbind(df_p4, df_p10, df_p28,
                   na_p4, na_p10, na_p28)

df_alluv2$TP <- factor(df_alluv2$TP, levels = c("P4", "P10", "P28"))

hist_barplot_levels <- df_alluv2$Region %>% unique() %>% sort()
hist_barplot_levels <- hist_barplot_levels[c(8,1,6,2,7,3,4,5)]
df_alluv2$Region <- factor(df_alluv2$Region, levels = hist_barplot_levels)

gga <- ggplot(df_alluv %>% as_tibble(), aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) +
  # scale_fill_brewer(type = "qual", palette = "Set2") +
  # scale_fill_viridis_b() +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) +
  geom_stratum(alpha = 0.75) +
  theme(legend.position = "bottom") +
  ggtitle("Histone occupancy in promoter region during development") 
cowplot::ggsave2(gga, filename = "../data/processed/08_Alluvial_Prom_18la-27ac-4me3_v3.pdf",
                 height = 8, width = 12)
```




##################
# Alluvial plot of promoter peak histone presence
##################
```{r}
library(ggalluvial)

# Prep the gene lists for the FGE loop
gene_list_P4 <- list(H3K27ac = k27_p4_pc_prom$SYMBOL %>% unique(),
                     H3K4me3 = k4_p4_pc_prom$SYMBOL %>% unique(),
                     H3K18la = k18_p4_pc_prom$SYMBOL %>% unique())

gene_list_P10 <- list(H3K27ac = k27_p10_pc_prom$SYMBOL %>% unique(),
                     H3K4me3 = k4_p10_pc_prom$SYMBOL %>% unique(),
                     H3K18la = k18_p10_pc_prom$SYMBOL %>% unique())

gene_list_P28 <- list(H3K27ac = k27_p28_pc_prom$SYMBOL %>% unique(),
                     H3K4me3 = k4_p28_pc_prom$SYMBOL %>% unique(),
                     H3K18la = k18_p28_pc_prom$SYMBOL %>% unique())


goi_list <- list(gene_list_P4, gene_list_P10, gene_list_P28)
names(goi_list) <- c("P4", "P10", "P28")

total_gois <- c(gene_list_P4 %>% unlist(use.names = F) %>% unique(),
                gene_list_P10 %>% unlist(use.names = F) %>% unique(),
                gene_list_P28 %>% unlist(use.names = F) %>% unique()) %>% unique()


# Prep DF for alluvial plot
df_alluv <- data.frame()

# 
for (i in names(goi_list)){
  
  tmp_venn <- ggVennDiagram::Venn(goi_list[[i]])
  tmp_data <- ggVennDiagram::process_data(tmp_venn)
  tmp_regions <- tmp_data@region$item
  names(tmp_regions) <- tmp_data@region$name

  for (ii in names(tmp_regions)){
    
    tmp_genes <- tmp_regions[[ii]]
    tmp_df <- data.frame(Symbol = tmp_genes,
                         TP = i,
                         Region = ii)
   
    
    df_alluv <- rbind(df_alluv, tmp_df)
      
  }
  
   tmp_df_missing <- data.frame(Symbol = total_gois[!(total_gois %in% (tmp_regions %>% unlist(use.names = F) %>% unique()))],
                                TP = i,
                                Region = NA)
   df_alluv <- rbind(df_alluv, tmp_df_missing)
}

df_alluv$TP <- factor(df_alluv$TP, levels = c("P4", "P10", "P28"))
df_alluv <- df_alluv %>% as.tibble() %>% replace_na(list(Region = "no_mark"))

hist_barplot_levels <- df_alluv$Region %>% unique() %>% sort()
hist_barplot_levels <- hist_barplot_levels[c(8,1,6,2,7,3,4,5)]
# hist_barplot_levels[1] <- NA
df_alluv$Region <- factor(df_alluv$Region, levels = hist_barplot_levels)

gga <- ggplot(df_alluv, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) +
  # scale_fill_brewer(type = "qual", palette = "Set2") +
  # scale_fill_viridis_b() +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) +
  geom_stratum(alpha = 0.75) +
  theme(legend.position = "bottom") +
  ggtitle("Histone occupancy in promoter region during development") 
cowplot::ggsave2(gga, filename = "../data/processed/08_Alluvial_Prom_18la-27ac-4me3_v2.pdf",
                 height = 8, width = 12)

saveRDS(object = df_alluv, file = "../data/interim/08_df_alluvial.RDS")

```



###################################
## Perform FGE on Alluvial segments
###################################
```{r}

source(file.path(dest_tools_dir, "GO_Leaf.R"))

df_alluv_spread <- df_alluv %>% spread(TP, Region)


goi_list_alluv <- list(all3_allTP = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3..H3K18la", 
                                       P10 == "H3K27ac..H3K4me3..H3K18la", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P4.all3_P10.K4..K27_P28.all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3..H3K18la", 
                                       P10 == "H3K27ac..H3K4me3", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P4.K4..K27_P10.all3_P28.all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3", 
                                       P10 == "H3K27ac..H3K4me3..H3K18la", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P4.K4..K27_P10.K4..K27_P28.all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3", 
                                       P10 == "H3K27ac..H3K4me3", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       K4..K27_allTP = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3", 
                                       P10 == "H3K27ac..H3K4me3", 
                                       P28 == "H3K27ac..H3K4me3") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28.K27..K18 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K27ac..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28.K4..K18 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28.K18 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F))


# Perform FGE on the GOIs from each subset of peaks

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 12

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_export <- "../data/processed"
dir_enrich <- file.path(dir_export, paste0("08_Alluvial_Regions_", enrich_db, "_", GOont))
dir.create(dir_enrich)

univ <- df_alluv_spread$Symbol

  
enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list_alluv, fun = "enrichGO",
                                          keyType = "SYMBOL", ont = GOont, OrgDb = org, univ = univ)


# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>% 
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>% 
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("H3K18la Peaks in ", i))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0(i, "_Alluvial-Region_FGE_top_", top, "_leaf_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)
    
```


<!-- ################################################################################ -->
<!-- ## Figure 3 Correlate peaks with RNA-seq -->
<!-- ################################################################################ -->
<!-- ```{r} -->
<!-- library(csaw) -->
<!-- library(limma) -->
<!-- library(edgeR) -->

<!-- # Make master peak domain -->
<!-- masterPeak <- GenomicRanges::reduce(c(peakList_cons[[1]], -->
<!--                                  peakList_cons[[2]], -->
<!--                                  peakList_cons[[3]])) -->

<!-- tmp_gr_list <- list() -->
<!-- for (i in 1:3){ -->
<!--   tmp_gr <- peakList_cons[[i]] -->
<!--   seqlevelsStyle(tmp_gr) <- "NCBI" -->
<!--   tmp_gr_list[[i]] <- tmp_gr  -->
<!-- } -->

<!-- masterPeak_anno <- GenomicRanges::reduce(c(tmp_gr_list[[1]], -->
<!--                                  tmp_gr_list[[2]], -->
<!--                                  tmp_gr_list[[3]])) -->

<!-- masterPeak_anno <- annotatePeak(masterPeak_anno, tssRegion=c(-1000, 500), -->
<!--                          TxDb=ensDb, annoDb=NULL, addFlankGeneInfo = T, -->
<!--                          overlap = "all") -->

<!-- # Get fragment counts for domains -->
<!-- bams <- list.files(path = "../data/interim/", pattern = "*bam$", recursive = T, full.names = T) -->
<!-- bams <- bams[c(23:26, 13:17, 18:22)] -->
<!-- names(bams) <- paste(gsub(".+\\/K18La-(.+)_cr.+", "\\1", bams), c(seq(1:4), seq(1:5), seq(1:5)), sep = "_" ) -->

<!-- pe.param <- readParam(max.frag=400, pe="both", minq = 20, dedup = F) -->

<!-- k18la.counts <- regionCounts(bams, masterPeak, ext=150, param=pe.param) -->
<!-- colnames(k18la.counts) <- paste(gsub(".+\\/K18La-(.+)_cr.+", "\\1", bams), c(seq(1:4), seq(1:5), seq(1:5)), sep = "_" ) -->


<!-- # Seq depth normalization -->
<!-- k18la.dge <- asDGEList(k18la.counts, -->
<!--                # norm.factors = filtered.data$norm.factors, -->
<!--                group = gsub(".+\\/K18La-(.+)_cr.+", "\\1", bams), -->
<!--                genes = as.data.frame(masterPeak_anno@anno)) -->
<!-- colnames(k18la.dge) <- names(bams) -->
<!-- k18la.dge <- edgeR::calcNormFactors(k18la.dge) -->
<!-- k18la.dge$rpkm <- edgeR::rpkm(k18la.dge, gene.length = "width") -->

<!-- k18la.dge$logRPKM <- log2(k18la.dge$rpkm + 1) -->

<!-- color.grp.pal <- wesanderson::wes_palette("Darjeeling1", 5, type = "discrete")[c(1,3,5)] -->

<!-- #Prep the PCA data -->
<!-- pca.k18la <- PCAtools::pca(k18la.dge$logRPKM, metadata = data.frame(k18la.dge$samples)) -->
<!-- pca.k18la$metadata$group <- factor(pca.k18la$metadata$group, levels = c("P4", "P10", "P28")) -->
<!-- sampkey <- color.grp.pal -->
<!-- names(sampkey) <- unique(pca.k18la$metadata$group) -->

<!-- #Plot all the samples -->
<!-- X='PC1'; Y='PC2' -->
<!-- p <- PCAtools::biplot(pca.k18la, x = X, y = Y, -->
<!--        colby = "group", colkey = sampkey, -->
<!--        #lab = "replicate", labSize = 2, -->
<!--        legendPosition = 'right',  -->
<!--        title = 'PCA All Samples', -->
<!--        subtitle = paste0(X, ' versus ', Y)) -->

<!-- cowplot::ggsave2(p, filename = file.path(dir_export, paste0("02_PCA_H3K18la_PC1vPC2.pdf")), -->
<!--                  height = 5, width = 6) -->


<!-- # Correlation with RNA -->
<!-- ``` -->



<!-- # Test EnrichedHeatmaps -->
<!-- https://bioconductor.org/packages/release/bioc/vignettes/EnrichedHeatmap/inst/doc/EnrichedHeatmap.html -->

<!-- ```{r} -->
<!-- library(EnrichedHeatmap) -->

<!-- load(system.file("extdata", "chr21_test_data.RData", package = "EnrichedHeatmap")) -->
<!-- ls() -->

<!-- tss = promoters(genes, upstream = 0, downstream = 1) -->
<!-- mat1 = normalizeToMatrix(H3K4me3, tss, value_column = "coverage",  -->
<!--     extend = 5000, mean_mode = "w0", w = 50) -->

<!-- EnrichedHeatmap(mat1, name = "H3K4me3") -->

<!-- tss2 <- getPromoters(TxDb = ensDb, upstream = 0, downstream = 1) -->
<!-- seqlevelsStyle(tss2) <- "UCSC" -->

<!-- mat_P28 <- normalizeToMatrix(peakList_cons$`K18La-P28`, tss2, value_column = "score",  -->
<!--     extend = 1000, mean_mode = "w0", w = 50) -->

<!-- EnrichedHeatmap(mat_P28, name = "P28-H3K18la", ylim = c(0,100)) -->

<!-- ``` -->






<!-- ## Get metadata for samples -->
<!-- ```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE} -->
<!-- # ord <- c(1:4) -->

<!-- # Macs2 beds -->
<!-- macs_beds <- list.files(path = "../data/interim/bed/", pattern = "optimal.+narrowPeak.gz", recursive = T) -->

<!-- # Meta sheet -->
<!-- meta <- data.frame(Beds = macs_beds, -->
<!--                    Samp = gsub("(.+)_overlap.+", "\\1", macs_beds)) -->
<!-- meta <- meta[c(5,3,4,2,1,8,6,7),] -->
<!-- # meta$Beds <- factor(meta$Beds, levels = levels(meta$Beds)) -->
<!-- meta$Samp <- factor(meta$Samp, levels = unique(meta$Samp)) -->


<!-- print("Meta data acquired...") -->
<!-- ``` -->


<!-- # Import the peaks into a merged list object and plot for number of peaks at each timepoint -->
<!-- ```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE} -->
<!-- library(ChIPpeakAnno) -->

<!-- cons_dir <- "../data/interim/bed/" -->

<!-- # Import the cons peaks -->
<!-- peakList_cons <- sapply(file.path(cons_dir, meta$Beds), toGRanges, format="narrowPeak") -->
<!-- names(peakList_cons) <- meta$Samp -->

<!-- # Sort the cons peaks -->
<!-- for (i in 1:length(peakList_cons)){ -->

<!--   # seqlevelsStyle(peakList_cons[[i]]) <- "NCBI" -->
<!--   peakList_cons[i] <- sortSeqlevels(peakList_cons[[i]]) -->
<!--   peakList_cons[i] <- GenomicRanges::sort(peakList_cons[[i]]) -->
<!-- } -->

<!-- # Generate plot for peak number -->
<!-- df <- lapply(peakList_cons, length) %>% as.data.frame() %>% gather("Sample", "Peaks") -->
<!-- df$Sample <- gsub("X", "", df$Sample) -->
<!-- df$Sample <- factor(df$Sample, levels = df$Sample) -->

<!-- b <- ggplot(df, aes(Sample, Peaks))+ -->
<!--   geom_bar(stat = "identity") + -->
<!--   theme_minimal() + -->
<!--   theme(axis.text.x = element_text(angle = 30, hjust = 0.8, vjust = 0.8)) + -->
<!--   labs(title = "Peaks") + xlab(NULL) -->

<!-- cowplot::ggsave2(plot = b, filename = "../data/processed/Peak_number_allPeaks.pdf", height = 4, width = 3) -->

<!-- print("Peak data acquired...") -->
<!-- ``` -->


<!-- ## Annotate peaks with ensembl db -->
<!-- ```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE} -->
<!-- library(ChIPseeker) -->
<!-- library(org.Mm.eg.db) -->
<!-- library(biomaRt) -->
<!-- library(GenomicFeatures) -->


<!-- peakAnnoList_ens <- list() -->

<!-- for (i in 1:length(peakList_cons)){ -->

<!--   peak_tmp <- peakList_cons[[i]] -->
<!--   seqlevelsStyle(peak_tmp) <- "NCBI" -->

<!--   peakAnnoList_ens[[i]] <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500), -->
<!--                          TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T, -->
<!--                          overlap = "all") -->

<!-- } -->

<!-- names(peakAnnoList_ens) <- levels(meta$Samp) -->

<!-- ``` -->




<!-- ################################################################################ -->
<!-- # Prepare plot for figures - Oct 30th, 2023  -->
<!-- ################################################################################ -->

<!-- #################### -->
<!-- # Timecourse K18La -->
<!-- #################### -->



<!-- # FGE of annotated genes -->
<!-- ```{r} -->

<!-- library(clusterProfiler) -->
<!-- library(enrichplot) -->
<!-- library(org.Mm.eg.db) -->
<!-- source(file.path(dest_tools_dir, "GO_leaf.R")) -->
<!-- # source("../src/tools/Dotplot_GO.R") -->

<!-- org = "org.Mm.eg.db" -->

<!-- enrich_db <- "GO_220310" -->
<!-- GOont <- "BP" -->

<!-- # dir_enrich <- file.path(dir_res, "FuncGeneEnrich") -->
<!-- dir_enrich <- file.path(dir_export, paste0("FuncGeneEnrich_", enrich_db, "_", GOont)) -->
<!-- dir.create(dir_enrich) -->


<!-- # Get list of genes for enrichment analysis -->

<!-- goi.list <- list(P4 = peakAnnoList_ens[[1]]@anno$SYMBOL %>% unique(), -->
<!--                  P10 = peakAnnoList_ens[[2]]@anno$SYMBOL %>% unique(), -->
<!--                  P28 = peakAnnoList_ens[[3]]@anno$SYMBOL %>% unique()) -->

<!-- # # Get universe -->
<!-- # univ <- row.names(seu_rodbp@assays$decontXcounts@data) -->

<!-- # Get enrichment -->
<!-- enrich_results <- clusterProfiler::compareCluster(geneClusters = goi.list, fun = "enrichGO", -->
<!--                                             keyType = "SYMBOL", ont = GOont, OrgDb = org) -->





<!-- ########################### -->
<!-- # Look at GO Level 6 -->
<!-- ########################### -->

<!-- go_level <- 8 -->
<!-- go_ids <- GOxploreR::Level2GOTermBP(go_level, organism = "mouse") -->
<!-- top <- 10 -->

<!-- # Filter enrichment for significance -->
<!-- enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.0001, Count > 2) -->

<!-- enrich_results_filt_fix <- enrich_results_filt %>%  -->
<!--   as.data.frame() %>%  -->
<!--   mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>%  -->
<!--   rowwise() %>%  -->
<!--   mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2])) -->

<!-- # Prep the enrichment DF -->
<!-- enrich <- enrich_results_filt_fix %>%  -->
<!--   as.data.frame() %>%  -->
<!--   dplyr::filter(ID %in% go_ids) -->
<!-- enrich$Description <- factor(enrich$Description, levels = rev(unique(enrich$Description))) -->


<!-- # Get top 5 pathways per cell type to get union set of pathways -->
<!-- top5_union <- enrich %>%  -->
<!--   group_by(Cluster) %>%  -->
<!--   arrange(p.adjust) %>%  -->
<!--   slice_head(n = top) -->


<!-- # Filter the full table for the union of top5 pathways -->
<!-- top5_enrich <- enrich %>%  -->
<!--   dplyr::filter(ID %in% (top5_union$ID %>% unique())) -->

<!-- # Prep the df for plotting -->
<!-- top5_enrich <- top5_enrich %>%  -->
<!--   arrange(Cluster, p.adjust) -->
<!-- top5_enrich$Description <- substring(top5_enrich$Description, 1, 60) -->
<!-- top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description))) -->
<!-- top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust) -->

<!-- # Make plot -->
<!-- p <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) +  -->
<!--     geom_point(aes(size = Ratio)) +  -->
<!--     theme_bw() + -->
<!--     scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) + -->
<!--     guides(size=guide_legend(title="Ratio")) + -->
<!--     theme(axis.title.x = element_blank(), axis.title.y = element_blank(), -->
<!--           axis.text.x = element_text(angle = 30, hjust = 1)) -->


<!-- #Open the pdf device -->
<!-- cowplot::ggsave2(plot = p, filename = file.path(dir_enrich, paste0("Timepoint_FGE_top_", top, "_level_", go_level, "_0001_dotplot.pdf")), -->
<!--                  height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5,  -->
<!--                  width = 6) -->



<!-- ########################### -->
<!-- # Look at leaf terms -->
<!-- ########################### -->


<!-- # # Filter enrichment for significance -->
<!-- # enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.0001, Count > 2) -->

<!-- # Get leaf terms -->
<!-- enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont) -->

<!-- # # Fix the Ratio -->
<!-- # enrich_results_leaf_fix <- enrich_results_leaf %>% -->
<!-- #   as.data.frame() %>% -->
<!-- #   mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% -->
<!-- #   rowwise() %>% -->
<!-- #   mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2])) -->

<!-- # Get top 5 pathways per cell type to get union set of pathways -->
<!-- top5_union <- enrich_results_leaf %>%  -->
<!--   group_by(Cluster) %>%  -->
<!--   arrange(p.adjust) %>%  -->
<!--   slice_head(n = top) -->


<!-- # Filter the full table for the union of top5 pathways -->
<!-- top5_enrich <- enrich_results_filt_fix %>%  -->
<!--   dplyr::filter(ID %in% (top5_union$ID %>% unique())) -->

<!-- # Prep the df for plotting -->
<!-- top5_enrich <- top5_enrich %>%  -->
<!--   arrange(Cluster, p.adjust) -->
<!-- top5_enrich$Description <- substring(top5_enrich$Description, 1, 60) -->
<!-- top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description))) -->
<!-- top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust) -->

<!-- # Make plot -->
<!-- pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) +  -->
<!--     geom_point(aes(size = Ratio)) +  -->
<!--     theme_bw() + -->
<!--     scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) + -->
<!--     guides(size=guide_legend(title="Ratio")) + -->
<!--     theme(axis.title.x = element_blank(), axis.title.y = element_blank(), -->
<!--           axis.text.x = element_text(angle = 30, hjust = 1)) -->

<!-- #Open the pdf device -->
<!-- cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("Timepoint_FGE_top_", top, "_leaf_0001_dotplot.pdf")), -->
<!--                  height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5,  -->
<!--                  width = 6) -->


<!-- #  -->
<!-- # # Make dotplot of results -->
<!-- # dot_go(enrich_results_leaf, top = 10, grp = "Cluster", desc = "Description", p.val = "p.adjust", ratio = "Ratio", filenm = file.path(dir_enrich, "plots", paste0(i, "_cluster_dotplot_05"))) -->
<!-- #  -->
<!-- # dot_go(enrich_results_leaf, top = 5, grp = "Cluster", desc = "Description", p.val = "p.adjust", ratio = "Ratio", filenm = file.path(dir_enrich, "plots", paste0(i, "_cluster_dotplot_05_5"))) -->

<!-- # Export the FGE results table -->
<!-- tmp <- enrich_results_filt_fix[,-11] -->
<!-- # Export the enrichment analysis -->
<!-- write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", "timepoint", "_filt.tsv")), -->
<!--               quote = F, row.names = F, sep = "\t") -->

<!-- # Export the FGE results table -->
<!-- tmp <- enrich_results_leaf_fix[,-11] -->
<!-- # Export the enrichment analysis -->
<!-- write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", "timepoint", "_leaf.tsv")), -->
<!--               quote = F, row.names = F, sep = "\t") -->


<!-- ``` -->


#
### Session Information
```{r, echo=FALSE}
sessionInfo()
```

#
## Explanation of Output Files

#### File structure of the data

* data
    + external
    + interim
    + processed   
    + raw
        - bams
        - beds
        - bigwig
* notebooks

#
#### Output File Description

Output files from analysis contain both tab-separated values table (.tsv), as well as, plots in pdf form. The tsv tables can be opened in Excel.

#
##### 'data/external'
* DevRet_Gene38v84_nCPM_MSTR.tsv
    + Gene expression values from deveolping mouse retina. Values are normalized CPM values. File is tab-separated value file.

#
##### 'data/interim'
Output from combining replicate peaks for the concensus bed file used for further analysis

* beds
    + Consensus bed files for each timepoint grouping using overlap method
    + cons
        - Consensus bed files for each timepoint using the cons method (*narrowPeak)
* bigwig
    + Merged bigwig files using samples from cons samples
        
#
##### 'data/processed'

* 01_PeakGeneAnno_*.txt
    + Peak annotation at each tf-timepoint
* gProf_AllPeaks
    + Directory containing the export table and visualization results from the functional gene enrichment analysis from genes containing peaks from the consensus bed files for the timepoint groupings
        - gProfileR_all.tsv - output from gProfileR
        - Heatmaps : heatmaps of all the top enriched categories using log2 CPM values 
* gProf_AllPeaks_Nrl_PN
    + Replot of the dotplots using all timepoints except Nrl E14
* gProf_[Crx|Nrl]
    + Replot of the dotplots separarting Crx and Nrl
* gProf_Nrl_PN
    + Replot of the dotplots separating Nrl and excluding the E14 timepoint

#
##### 'data/raw

* bams
    + Bam files generated from the pipeline for each sample.
* beds
    + The 1% FDR Macs peak bed files for each sample used in the analysis
* bigwig
    + Bigwig file exports for each sample. The bigwig file is a bin per million normalization. 