---
title: "H3K18la Tissue Comparison Analysis - November 19th, 2024"
output: html_notebook

---

### Project Summary
This project investigates the H3K18la during development. 

**DESIGN**:  

**NOTES**: The data were generated from the ENCODE histone pipeline with duplicate 
removal with no controls. Alluvial plot of histone overlaps being generated by 
quantitation of reads over mstr_features.bed then filtering for binding prior
to generation of alluvial plot.



	

#### Get custom scripts and functions and outside data used for analysis
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

script_dir <- "https://raw.githubusercontent.com/MBrooks313/R_codes/master"
dest_tools_dir <- "../src/tools"
dest_vis_dir <- "../src/visualization"

# download.file(url = file.path(script_dir, "GO_leaf.R"),
#               destfile = file.path(dest_tools_dir, "GO_leaf.R"),
#               method = "curl")
# 

# Function to convert ENS ids to gene symbol 
# ens2sym <- function(x){paste(sort(unique(AnnotationDbi::mapIds(org.Mm.eg.db, keys = x, keytype = "ENSEMBL", column = "SYMBOL"))), sep = "", collapse = ",")}

library(tidyverse)
# options(java.parameters = "-Xmx8000m")
print("Completed...")
```


# Get Ensembl annotation database
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

## Query for all available EnsDb databases
ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

ensDb <- ahdb[[1]]



ahdb_hs <- query(ah, pattern = c("EnsDb", "Homo sapiens", 109))

ensDb_hs <- ahdb_hs[[1]]


# supportedFilters(ensDb)

```


################################################################################
## Import outside datasets
################################################################################
```{r}
library(GenomicRanges)
library(ChIPpeakAnno)
# source(file.path(dest_tools_dir, "DFave.R"))

#############################################
# Import the H3K18La from Galle
k18la_galle_beds <- list.files(path = "/Volumes/data/People/brooksma/PEGASUS/Projects/Gaur/GSE195856/", pattern = "*H3K18la*", recursive = T, full.names = T)
names(k18la_galle_beds) <- gsub(".+_(.+)_H3K18la.+", "\\1", k18la_galle_beds)

# Import the beds
k18la_galle <- sapply(k18la_galle_beds, toGRanges, format="narrowPeak")


#############################################
# Import the time course sig DB regions
# tc_prom_pc_sig <- readRDS(file = "../data/interim/03_prom_pc_sig_240913.RDS")
# posi_cor <- readRDS(file = "../data/interim/03_h3k18la_db_dge_positive_corr_241001.RDS")
# anti_cor <- readRDS(file = "../data/interim/03_h3k18la_db_dge_anti_corr_241001.RDS")

tc_peakAnno_cons <- readRDS("../data/interim/peakAnnoList_cons_240911.RDS")

                         
```        

## Annotate outside peaks with ensembl db
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


peakAnnoList_galle <- list()

for (i in 1:length(k18la_galle)){
  
  peak_tmp <- k18la_galle[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  # peakAnnoList_ens[[i]] <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
  #                        TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T,
  #                        overlap = "all")
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  anno_tmp@anno$GENENAME <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "GENENAME", 
                                         keytype = "GENEID")
  peakAnnoList_galle[[i]] <- anno_tmp

}

names(peakAnnoList_galle) <- names(k18la_galle)

```

## Export peak annotation as bed and xlsx files
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(rtracklayer)
library(xlsx)


for (i in 1:length(peakAnnoList_galle)){
  
  gr <- peakAnnoList_galle[[i]]@anno
  name_gr <- names(peakAnnoList_galle[i])
  
  df <- data.frame(seqnames=seqnames(gr),
    starts=start(gr)-1,
    ends=end(gr),
    SYMBOL=gr$SYMBOL,
    # names=names(gr),
    #scores=gr$score,
    scores=".",
    strands=strand(gr),
    #signalValue=gr$signalValue,
    #qValue=gr$qValue,
    annotation=gr$annotation,
    # names=names(gr),
    # SYMBOL=gr$SYMBOL,
    # ENTREZID=gr$ENTREZID,
    geneId=gr$geneId,
    transcriptId=gr$transcriptId,
    distanceToTSS=gr$distanceToTSS,
    GENENAME=gr$GENENAME,
    biotype=gr$transcriptBiotype)
  
  write.table(df, file=paste0("../data/processed/01_", name_gr, "_annotated.bed"), 
              quote=F, sep="\t", row.names=F, col.names=F)

  # write.xlsx(df, file=paste0("../data/processed/01_", name_gr, "_annotated.xlsx"),
  #            row.names=F, col.names=T)
  
}


```
          
          
################################################################################
## Import outside datasets
################################################################################
```{r}               
ol_galle <- findOverlapsOfPeaks(tc_peakAnno_cons$P28, 
                                k18la_galle$ADIPO,
                                k18la_galle$BMDM,
                                k18la_galle$GAS,
                                k18la_galle$MB,
                                k18la_galle$MT,
                                k18la_galle$PIM,
                                connectedPeaks = "keepAll", maxgap = 500L)




```


#
# Import venn counts from Homer overlap of Galle and Gaur H3K18la peaks
```{r}
library(UpSetR)

###################
# All H3K18La peaks
venn_all <- read.csv("/Volumes/data-2/Gaur/homer/mergePeaks/all_peaks/venn", sep = "\t")

# Prep input
input_all <- venn_all$Total

# Prep input names
venn_nm <- gsub("data\\/", "", venn_all$Name)
venn_nm <- gsub("\\.txt", "", venn_nm)
venn_nm <- gsub("\\|", "\\&", venn_nm)

names(input_all) <- venn_nm

# Plot
nint <- (input_all>500) %>% sum()
u <- upset(fromExpression(input_all), 
      nintersects = nint, 
      nsets = 7, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
      )

pdf("../data/processed/09_allPeak_upset.pdf", 
    width = 9, height = 6, onefile = F)
u
dev.off()


###################
# P28 peaks
venn_P28 <- read.csv("/Volumes/data-2/Gaur/homer/mergePeaks/P28_peaks/venn", sep = "\t")

# Prep input
input_p28 <- venn_P28$Total

# Prep input names
venn_nm <- gsub("data\\/", "", venn_P28$Name)
venn_nm <- gsub("\\.txt", "", venn_nm)
venn_nm <- gsub("\\|", "\\&", venn_nm)

names(input_p28) <- venn_nm

# Plot
nint <- (input_p28>500) %>% sum()
u <- upset(fromExpression(input_p28), 
      nintersects = nint, 
      nsets = 7, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
      )

pdf("../data/processed/09_P28Peak_upset2.pdf", 
    width = 9, height = 6, onefile = F)
u
dev.off()


###################
# P28 Promoter, proteinc oding peaks
venn_P28 <- read.csv("/Volumes/data-2/Gaur/homer/mergePeaks/P28_Prom-PC_peaks/venn", sep = "\t")

# Prep input
input_p28 <- venn_P28$Total

# Prep input names
venn_nm <- gsub("data\\/", "", venn_P28$Name)
venn_nm <- gsub("\\.txt", "", venn_nm)
venn_nm <- gsub("\\|", "\\&", venn_nm)

names(input_p28) <- venn_nm

# Plot
nint <- (input_p28>500) %>% sum()
u <- upset(fromExpression(input_p28), 
      nintersects = nint, 
      nsets = 7, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
      )

pdf("../data/processed/09_P28_Prom-PCPeak_upset2.pdf", 
    width = 9, height = 6, onefile = F)
u
dev.off()



###################
# P28 Promoter, protein coding peaks with Galle promoter, protein coding 
venn_P28 <- read.csv("/Volumes/data-2/Gaur/homer/mergePeaks/P28_Prom-PC_peaks_v2/venn", sep = "\t")

# Prep input
input_p28 <- venn_P28$Total

# Prep input names
venn_nm <- gsub("data\\/", "", venn_P28$Name)
venn_nm <- gsub("\\.txt", "", venn_nm)
venn_nm <- gsub("\\|", "\\&", venn_nm)

names(input_p28) <- venn_nm

# Plot
nint <- (input_p28>200) %>% sum()
u <- upset(fromExpression(input_p28), 
      nintersects = nint, 
      nsets = 7, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
      )

updf("../data/processed/09_P28-Galle_Prom-PC_peak_upset.pdf", 
    width = 9, height = 6, onefile = F)
u
dev.off()

```
#



################################################################################
# Import the Galle peaks from nf-core pipeline, process then the same way as retina and then compare
################################################################################

# Import the peaks into a merged list object and plot for number of peaks at each timepoint
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
library(ChIPpeakAnno)

# # Get 
# # Macs2 beds
# macs_beds <- list.files(path = "../data/interim/bed/macs2/", pattern = "narrowPeak", recursive = T)
# 
# # Meta sheet
# meta <- data.frame(Beds = macs_beds[1:7],
#                    Samp = gsub("H3K18La_(P.+)\\.macs.+", "\\1", macs_beds[1:7]),
#                    Group = gsub("H3K18La_(P\\d+)_R.+", "\\1", macs_beds[1:7]))
# meta <- meta[c(6,7,1:5),]
# # meta$Beds <- factor(meta$Beds, levels = levels(meta$Beds))
# meta$Samp <- factor(meta$Samp, levels = unique(meta$Samp))
# meta$Group <- factor(meta$Group, levels = unique(meta$Group))



macs_dir <- "../data/external/Galle_nf-core/04_called_peaks/macs2/"
macs_beds <- list.files(path = macs_dir, pattern = "H3K18.+narrowPeak", recursive = T)

# Import the cons peaks
peakList <- sapply(file.path(macs_dir, macs_beds), toGRanges, format="narrowPeak")
names(peakList) <- gsub("(.+).macs.+", "\\1", macs_beds)

peakList_1E6 <- GRangesList()
# Sort the cons peaks
for (i in 1:length(peakList)){
  
  # seqlevelsStyle(peakList[[i]]) <- "NCBI"
  tmp_peak <- sortSeqlevels(peakList[[i]])
  tmp_peak <- GenomicRanges::sort(tmp_peak)
  tmp_peak <- tmp_peak[(elementMetadata(tmp_peak)[,"qValue"] > 6) & (seqnames(tmp_peak) != "MT")]
  peakList_1E6[[i]] <- tmp_peak
}
names(peakList_1E6) <- gsub("(.+).macs.+", "\\1", macs_beds)

# ####################
# # FDR = 1E-6
# 
# # Filter for 0.0001
# peakList_1E6 <- GRangesList()
# for(i in 1:length(peakList)){
#   tmp_peak <- peakList[[i]][(elementMetadata(peakList[[i]])[,"qValue"] > 6) & (seqnames(peakList[[i]]) != "MT")]
#   peakList_1E6[[i]] <- tmp_peak
# }
# names(peakList_1E6) <- meta$Samp

# Overlap peaks per timepoint
# Get overlaps
ol_ADIPO <- findOverlapsOfPeaks(c(peakList_1E6[1],
                             peakList_1E6[2]), connectedPeaks="keepAll", ignore.strand = T)

ol_ESC_2i <- findOverlapsOfPeaks(c(peakList_1E6[4],
                             peakList_1E6[5]), connectedPeaks="keepAll", ignore.strand = T)

ol_ESC_ser <- findOverlapsOfPeaks(c(peakList_1E6[6],
                             peakList_1E6[7]), connectedPeaks="keepAll", ignore.strand = T)

ol_GAS <- findOverlapsOfPeaks(c(peakList_1E6[8],
                             peakList_1E6[9]), connectedPeaks="keepAll", ignore.strand = T)

ol_MT <- findOverlapsOfPeaks(c(peakList_1E6[11],
                             peakList_1E6[12]), connectedPeaks="keepAll", ignore.strand = T)

ol_PIM <- findOverlapsOfPeaks(c(peakList_1E6[13],
                             peakList_1E6[14]), connectedPeaks="keepAll", ignore.strand = T)

peakList_cons <- GRangesList(ol_ADIPO$mergedPeaks, peakList_1E6$BMDM_H3K18la_R1, 
                             ol_ESC_2i$mergedPeaks, ol_ESC_ser$mergedPeaks,
                             ol_GAS$mergedPeaks, peakList_1E6$MB_H3K18la_R1, 
                             ol_MT$mergedPeaks, ol_PIM$mergedPeaks)
nm <- gsub("_H3K18la.+", "", names(peakList_1E6)) %>% unique()
nm <- gsub("-", "_", nm)
names(peakList_cons) <- nm

```


## Annotate outside peaks with ensembl db
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


peakAnnoList_galle_nf <- list()

for (i in 1:length(peakList_cons)){
  
  peak_tmp <- peakList_cons[[i]]
  names(peak_tmp) <- paste0("X", 1:length(peak_tmp))
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  # peakAnnoList_ens[[i]] <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
  #                        TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T,
  #                        overlap = "all")
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  anno_tmp@anno$GENENAME <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "GENENAME", 
                                         keytype = "GENEID")
  peakAnnoList_galle_nf[[i]] <- anno_tmp

}

names(peakAnnoList_galle_nf) <- names(peakList_cons)

```

## Export peak annotation as bed and xlsx files
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(rtracklayer)
library(xlsx)


for (i in 1:length(peakAnnoList_galle_nf)){
  
  gr <- peakAnnoList_galle_nf[[i]]@anno
  name_gr <- names(peakAnnoList_galle_nf[i])
  
  df <- data.frame(seqnames=seqnames(gr),
    starts=start(gr)-1,
    ends=end(gr),
    SYMBOL=gr$SYMBOL,
    # names=names(gr),
    #scores=gr$score,
    scores=".",
    strands=strand(gr),
    #signalValue=gr$signalValue,
    #qValue=gr$qValue,
    annotation=gr$annotation,
    # names=names(gr),
    # SYMBOL=gr$SYMBOL,
    # ENTREZID=gr$ENTREZID,
    geneId=gr$geneId,
    transcriptId=gr$transcriptId,
    distanceToTSS=gr$distanceToTSS,
    GENENAME=gr$GENENAME,
    biotype=gr$transcriptBiotype)
  
  write.table(df, file=paste0("../data/processed/01_", name_gr, "_nf_annotated.bed"), 
              quote=F, sep="\t", row.names=F, col.names=F)

  # write.xlsx(df, file=paste0("../data/processed/01_", name_gr, "_annotated.xlsx"),
  #            row.names=F, col.names=T)
  
}


```



################################################################################
# Import the Yu 2021 peaks from nf-core pipeline, process then the same way as retina and then compare
################################################################################
 *** The 2nd replicate is not good...266 macs peaks ***
# Import the peaks into a merged list object and plot for number of peaks at each timepoint
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
library(ChIPpeakAnno)


macs_dir <- "../data/external/Yu_nf-core//04_called_peaks/macs2/"
macs_beds <- list.files(path = macs_dir, pattern = "H3K18.+narrowPeak", recursive = T)

seacr_dir <- "../data/external/Yu_nf-core//04_called_peaks/seacr//"
seacr_beds <- list.files(path = seacr_dir, pattern = "H3K18.+bed", recursive = T)

# Import the cons peaks
peakList <- sapply(file.path(macs_dir, macs_beds), toGRanges, format="narrowPeak")
names(peakList) <- gsub("(.+).macs.+", "\\1", macs_beds)

peakList_seacr <- sapply(file.path(seacr_dir, seacr_beds), toGRanges, format="BED")
names(peakList) <- gsub("(.+).macs.+", "\\1", macs_beds)

peakList_1E6 <- GRangesList()
# Sort the cons peaks
for (i in 1:length(peakList)){
  
  # seqlevelsStyle(peakList[[i]]) <- "NCBI"
  tmp_peak <- sortSeqlevels(peakList[[i]])
  tmp_peak <- GenomicRanges::sort(tmp_peak)
  tmp_peak <- tmp_peak[(elementMetadata(tmp_peak)[,"qValue"] > 6) & (seqnames(tmp_peak) != "MT")]
  peakList_1E6[[i]] <- tmp_peak
}
names(peakList_1E6) <- gsub("(.+).macs.+", "\\1", macs_beds)

# ####################
# # FDR = 1E-6
# 
# # Filter for 0.0001
# peakList_1E6 <- GRangesList()
# for(i in 1:length(peakList)){
#   tmp_peak <- peakList[[i]][(elementMetadata(peakList[[i]])[,"qValue"] > 6) & (seqnames(peakList[[i]]) != "MT")]
#   peakList_1E6[[i]] <- tmp_peak
# }
# names(peakList_1E6) <- meta$Samp

# Overlap peaks per timepoint
# Get overlaps
ol_OCM1 <- findOverlapsOfPeaks(c(peakList_1E6[1],
                             peakList_1E6[2]), connectedPeaks="keepAll", ignore.strand = T)


```


## Annotate outside peaks with ensembl db
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


peakAnnoList_galle_nf <- list()

for (i in 1:length(peakList_cons)){
  
  peak_tmp <- peakList_cons[[i]]
  names(peak_tmp) <- paste0("X", 1:length(peak_tmp))
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  # peakAnnoList_ens[[i]] <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
  #                        TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T,
  #                        overlap = "all")
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  anno_tmp@anno$GENENAME <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "GENENAME", 
                                         keytype = "GENEID")
  peakAnnoList_galle_nf[[i]] <- anno_tmp

}

names(peakAnnoList_galle_nf) <- names(peakList_cons)

```

## Export peak annotation as bed and xlsx files
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(rtracklayer)
library(xlsx)


for (i in 1:length(peakAnnoList_galle_nf)){
  
  gr <- peakAnnoList_galle_nf[[i]]@anno
  name_gr <- names(peakAnnoList_galle_nf[i])
  
  df <- data.frame(seqnames=seqnames(gr),
    starts=start(gr)-1,
    ends=end(gr),
    SYMBOL=gr$SYMBOL,
    # names=names(gr),
    #scores=gr$score,
    scores=".",
    strands=strand(gr),
    #signalValue=gr$signalValue,
    #qValue=gr$qValue,
    annotation=gr$annotation,
    # names=names(gr),
    # SYMBOL=gr$SYMBOL,
    # ENTREZID=gr$ENTREZID,
    geneId=gr$geneId,
    transcriptId=gr$transcriptId,
    distanceToTSS=gr$distanceToTSS,
    GENENAME=gr$GENENAME,
    biotype=gr$transcriptBiotype)
  
  write.table(df, file=paste0("../data/processed/01_", name_gr, "_nf_annotated.bed"), 
              quote=F, sep="\t", row.names=F, col.names=F)

  # write.xlsx(df, file=paste0("../data/processed/01_", name_gr, "_annotated.xlsx"),
  #            row.names=F, col.names=T)
  
}


```



```{r}

###################
# P28 Promoter, protein coding peaks with Galle nf-core pipeline promoter, protein coding 
venn_P28 <- read.csv("/Volumes/data-2/Gaur/homer/mergePeaks/P28_Prom-PC_nf_peaks/venn", sep = "\t")

# Prep input
input_p28 <- venn_P28$Total

# Prep input names
venn_nm <- gsub("data\\/", "", venn_P28$Name)
venn_nm <- gsub("\\.txt", "", venn_nm)
venn_nm <- gsub("\\|", "\\&", venn_nm)

names(input_p28) <- venn_nm

# Plot
nint <- (input_p28>200) %>% sum()
u <- upset(fromExpression(input_p28), 
      nintersects = nint, 
      nsets = 7, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
      )

pdf("../data/processed/09_P28-Galle_Prom-PC_nf_peak_upset.pdf", 
    width = 9, height = 6, onefile = F)
u
dev.off()



```


#
# Make distance plot of promoter genes for the different tissues
```{r}

# Make upset plot based on presence of peak in promoter of gene
genes_list <- list(RETINA = (tc_peakAnno_cons$P28@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        ADIPO = (peakAnnoList_galle_nf$ADIPO@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        BMDM = (peakAnnoList_galle_nf$BMDM@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        GAS = (peakAnnoList_galle_nf$GAS@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        MB = (peakAnnoList_galle_nf$MB@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        MT = (peakAnnoList_galle_nf$MT@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        PIM = (peakAnnoList_galle_nf$PIM@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()))

u2 <- upset(fromList(genes_list), 
      nintersects = 14, 
      nsets = 7, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
      )

pdf("../data/processed/09_P28-Galle_Prom-PC_nf_peak_upset_GeneBased.pdf", 
    width = 9, height = 6, onefile = F)
u2
dev.off()


# Make correlation and dendro plots based on presence of peak in promoter of gene
library(rafalib)
library(Hmisc)
library(corrplot)

# Get genes for matrix
genes_galle <- lapply(peakAnnoList_galle_nf, function(x){x@anno %>% 
    as.data.frame() %>% 
    dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
    dplyr::select(SYMBOL)}) %>% 
  unlist() %>% sort() %>% unique()

genes_P28 <- tc_peakAnno_cons$P28@anno %>% 
  as.data.frame() %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
  dplyr::select(SYMBOL) %>% 
  unlist() %>% sort() %>% unique()

genes_all <- c(genes_galle, genes_P28) %>% sort() %>% unique()

# Make matrix


genes_mat <- data.frame(RETINA = ifelse(genes_all %in% (tc_peakAnno_cons$P28@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        ADIPO = ifelse(genes_all %in% (peakAnnoList_galle_nf$ADIPO@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        BMDM = ifelse(genes_all %in% (peakAnnoList_galle_nf$BMDM@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        GAS = ifelse(genes_all %in% (peakAnnoList_galle_nf$GAS@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        MB = ifelse(genes_all %in% (peakAnnoList_galle_nf$MB@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        MT = ifelse(genes_all %in% (peakAnnoList_galle_nf$MT@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        PIM = ifelse(genes_all %in% (peakAnnoList_galle_nf$PIM@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0), row.names = genes_all)



#####################################
# Make phi correlation of full matrix

# Correlation plot
c <- cor(genes_mat, use = "pair")

pdf(file="../data/processed/09_CorrPlot_allGenes.pdf", width=8, height=8, useDingbats = F)
corrplot(c, method="color", cl.cex=1.25)
dev.off()

# Dendrogram of correlation
d2 <- dist(c)
hc2 <- hclust(d2, method="ward.D2")

pdf(file="../data/processed/09_CorrDendro_allGenes.pdf", width=8, height=4, useDingbats = F)
par(mar=c(4,4,1,1))
myplclust(hclust = hc2)
dev.off()

#####################################
# Make phi correlation of only shared genes

# Drop similar genes
genes_nonsim <- genes_mat[rowSums(genes_mat) %in% c(2:6),]

# Correlation plot
c2 <- cor(genes_nonsim, use = "pair")

pdf(file="../data/processed/09_CorrPlot_sharedGenes.pdf", width=8, height=8, useDingbats = F)
corrplot(c2, method="color", cl.cex=1.25)
dev.off()

# Dendrogram of correlation
d2 <- dist(c2)
hc2 <- hclust(d2, method="ward.D2")

pdf(file="../data/processed/09_CorrDendro_sharedGenes.pdf", width=8, height=4, useDingbats = F)
par(mar=c(4,4,1,1))
myplclust(hclust = hc2)
dev.off()

```



# Perform FGE on Retina only, Ret-Adipo, and common genes
```{r}

library(clusterProfiler)
library(enrichplot)
library(org.Mm.eg.db)
source("~/Github/R_codes/GO_leaf.R")

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"

univ <- row.names(genes_mat)

gene_list <- list(RETINA = genes_mat %>% dplyr::filter(RETINA == 1, rowSums(genes_mat) == 1) %>% row.names(),
                  RETINA_ADIPO = genes_mat %>% dplyr::filter(RETINA == 1, ADIPO == 1, rowSums(genes_mat) == 2) %>% row.names(),
                  RETINA_BMDM = genes_mat %>% dplyr::filter(RETINA == 1, BMDM == 1, rowSums(genes_mat) == 2) %>% row.names(),
                  RETINA_BMDM_ADIPO = genes_mat %>% dplyr::filter(RETINA == 1, BMDM == 1, ADIPO == 1, rowSums(genes_mat) == 3) %>% row.names(),
                  Common = genes_mat %>% dplyr::filter(rowSums(genes_mat) == 7) %>% row.names())
# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
# dir_res <- "../data/processed/"
# dir_enrich <- file.path(dir_res, paste0("FuncGeneEnrich_", id1, "v", id2, ""), enrich_db)
# dir.create(file.path(dir_enrich, "plots", "IndiPlots"), recursive = T)

# perform enrichment
enrich_results <- clusterProfiler::compareCluster(geneClusters = gene_list, fun = "enrichGO",
                                                keyType = "SYMBOL", ont = GOont, OrgDb = org, universe = univ)
  
# Fiter enrichment for significance
enrich_results <- enrich_results %>% dplyr::filter(p.adjust < 0.05)


# Get leaf term
enrich_leaf <- GO_leaf(enrich_results@compareClusterResult, query = "Cluster", id.col = "ID")

# Get simplified output
enrich_simp <- clusterProfiler::simplify(enrich_results)

tmp_res <- enrich_results@compareClusterResult


write.table(enrich_simp@compareClusterResult, file = "../data/processed/09_GO_220310_BP_upset_simp.tsv", 
            sep = "\t", quote = F, col.names = T, row.names = F)

```


#
### Session Information
```{r, echo=FALSE}
sessionInfo()
```

#
## Explanation of Output Files

#### File structure of the data

* data
    + external
    + interim
    + processed   
    + raw
        - bams
        - beds
        - bigwig
* notebooks

#
#### Output File Description

Output files from analysis contain both tab-separated values table (.tsv), as well as, plots in pdf form. The tsv tables can be opened in Excel.

#
##### 'data/external'
* DevRet_Gene38v84_nCPM_MSTR.tsv
    + Gene expression values from deveolping mouse retina. Values are normalized CPM values. File is tab-separated value file.

#
##### 'data/interim'
Output from combining replicate peaks for the concensus bed file used for further analysis

* beds
    + Consensus bed files for each timepoint grouping using overlap method
    + cons
        - Consensus bed files for each timepoint using the cons method (*narrowPeak)
* bigwig
    + Merged bigwig files using samples from cons samples
        
#
##### 'data/processed'

* 01_PeakGeneAnno_*.txt
    + Peak annotation at each tf-timepoint
* gProf_AllPeaks
    + Directory containing the export table and visualization results from the functional gene enrichment analysis from genes containing peaks from the consensus bed files for the timepoint groupings
        - gProfileR_all.tsv - output from gProfileR
        - Heatmaps : heatmaps of all the top enriched categories using log2 CPM values 
* gProf_AllPeaks_Nrl_PN
    + Replot of the dotplots using all timepoints except Nrl E14
* gProf_[Crx|Nrl]
    + Replot of the dotplots separarting Crx and Nrl
* gProf_Nrl_PN
    + Replot of the dotplots separating Nrl and excluding the E14 timepoint

#
##### 'data/raw

* bams
    + Bam files generated from the pipeline for each sample.
* beds
    + The 1% FDR Macs peak bed files for each sample used in the analysis
* bigwig
    + Bigwig file exports for each sample. The bigwig file is a bin per million normalization. 