---
title: "H3K18la Alluvial Analysis"
output: html_notebook

---

### Get custom scripts and functions and outside data used for analysis
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

script_dir <- "https://raw.githubusercontent.com/MBrooks313/R_codes/master"
dest_tools_dir <- "../src/tools"
dest_vis_dir <- "../src/visualization"

download.file(url = file.path(script_dir, "DFave.R"),
              destfile = file.path(dest_tools_dir, "DFave.R"),
              method = "curl")

download.file(url = file.path(script_dir, "Expr_Filter.R"),
              destfile = file.path(dest_tools_dir, "Expr_Filter.R"),
              method = "curl")

download.file(url = file.path(script_dir, "Barplot_GO.R"),
              destfile = file.path(dest_tools_dir, "Barplot_GO.R"),
              method = "curl")


library(tidyverse)

```


### Get Ensembl annotation database
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

## Query for all available EnsDb databases
ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

ensDb <- ahdb[[1]]

```


################################################################################
### Import outside datasets
################################################################################
```{r}

source(file.path(dest_tools_dir, "DFave.R"))

#############################################
# Import expression data for developing retina
retina.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)

# Get group info for averaging
grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.dwnld)[10:31])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
retina_cpm <- data.frame(merge(retina.dwnld[,1:9],
                               log2(DFave(retina.dwnld[,10:31], exp_grp) + 1),
                               by = 0),
                         row.names = 1)


#############################################
# Import the H3K27ac from Aldiri
k27_aldiri_beds <- list.files(path = "/Volumes/data/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bed/", pattern = "*H3K27Ac*", recursive = T, full.names = T)
names(k27_aldiri_beds) <- gsub(".+\\/(.+)_H3K27.+", "\\1", k27_aldiri_beds)

# Import the beds 
k27_aldiri <- sapply(k27_aldiri_beds, toGRanges, format="narrowPeak")

for (i in 1:length(k27_aldiri)){
  k27_aldiri[[i]] <- GenomicRanges::reduce(k27_aldiri[[i]])
  seqlevelsStyle(k27_aldiri[[i]]) <- "NCBI"
}


#############################################
# Import the H3K4me3 from Aldiri
k4me3_aldiri_beds <- list.files(path = "/Volumes/data/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bed/", pattern = "*H3K4me3_*", recursive = T, full.names = T)
names(k4me3_aldiri_beds) <- gsub(".+\\/(.+)_H3K4me3.+", "\\1", k4me3_aldiri_beds)

# Import the beds 
k4me3_aldiri <- sapply(k4me3_aldiri_beds, toGRanges, format="narrowPeak")

for (i in 1:length(k4me3_aldiri)){
  k4me3_aldiri[[i]] <- GenomicRanges::reduce(k4me3_aldiri[[i]])
  seqlevelsStyle(k4me3_aldiri[[i]]) <- "NCBI"
}



#############################################
# Import the time course sig DB regions
tc_prom_pc_sig <- readRDS(file = "../data/interim/03_prom_pc_sig_240913.RDS")
posi_cor <- readRDS(file = "../data/interim/03_h3k18la_db_dge_positive_corr_241001.RDS")
anti_cor <- readRDS(file = "../data/interim/03_h3k18la_db_dge_anti_corr_241001.RDS")

tc_peakAnno_cons <- readRDS("../data/interim/peakAnnoList_cons_240911.RDS")

```                   
                         

### Annotate outside data
```{r}

# Aldiri k27ac 
k27AnnoList_ens <- list()

for (i in 1:length(k27_aldiri)){
  
  peak_tmp <- k27_aldiri[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")

  k27AnnoList_ens[[i]] <- anno_tmp

}

names(k27AnnoList_ens) <- names(k27_aldiri)


# Aldiri k4me3 
k4me3AnnoList_ens <- list()

for (i in 1:length(k4me3_aldiri)){
  
  peak_tmp <- k4me3_aldiri[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")

  k4me3AnnoList_ens[[i]] <- anno_tmp

}

names(k4me3AnnoList_ens) <- names(k4me3_aldiri)


```



################################################################################
### Get the counts over annotated feature regions
################################################################################
```{r}
library(tidyverse)
library(ChIPpeakAnno)
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)
library(csaw)

############################
# Master features
mstr_features_anno <- readRDS(file = "../data/interim/03_mstr_features_anno_240913.RDS")
mstr_features_anno_ucsc <- mstr_features_anno
seqlevels(mstr_features_anno_ucsc@anno) <- paste0("chr", seqlevels(mstr_features_anno@anno))

# Params
pe.param <- readParam(max.frag=600, pe="both", minq = 20, dedup = T)
se.param <- readParam(max.frag=600, pe="none", minq = 20, dedup = T)

############################
# H3K18la

# Get bam files
bam_dir <- "../data/raw/bams/"
bams <- list.files(path = bam_dir, pattern = "H3K18La_P.+.bam$")
bams <- bams[c(6:7, 1:2, 3:5)]
bam.files <- paste0(bam_dir, bams)

# Annotated region counts
h3k18la_counts <- regionCounts(bam.files, mstr_features_anno@anno, ext=150, param=pe.param)

# Normalization
binned <- windowCounts(bam.files, bin=TRUE, width=10000, param=se.param)
h3k18la_se <- normFactors(binned, se.out=T)


############################
# H3K27ac

# Get bam files
bam_dir <- "/Volumes/data/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bam/"
bams <- list.files(path = bam_dir, pattern = ".+H3K27Ac.+.bam$")
bams <- bams[c(16:17, 10:11, 1:2)]
bam.files <- paste0(bam_dir, bams)

# Annotated region counts
h3k27ac_counts <- regionCounts(bam.files, mstr_features_anno_ucsc@anno, ext=150, param=se.param)

# Normalization
binned <- windowCounts(bam.files, bin=TRUE, width=10000, param=se.param)
h3k27ac_se <- normFactors(binned, se.out=T)


############################
# H3K4me3

# Get bam files
bam_dir <- "/Volumes/data/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bam/"
bams <- list.files(path = bam_dir, pattern = ".+H3K4me3.+.bam$")
bams <- bams[c(16:17, 9:10, 1:2)]
bam.files <- paste0(bam_dir, bams)

# Annotated region counts
h3k4me3_counts <- regionCounts(bam.files, mstr_features_anno_ucsc@anno, ext=150, param=se.param)

# Normalization
binned <- windowCounts(bam.files, bin=TRUE, width=10000, param=se.param)
h3k4me3_se <- normFactors(binned, se.out=T)



```



################################################################################
### Prep the DGE list objects
################################################################################
```{r}
library(edgeR)

source(file.path(dest_tools_dir, "Expr_Filter.R"))


##########
## h3k18la
samp_nm <- gsub(".+(H3K18La_.+)\\.target.+", "\\1", h3k18la_counts@colData[[1]])

h3k18la.dge <- asDGEList(h3k18la_counts,
               norm.factors = h3k18la_se$norm.factors,
               group = gsub("H3K18La_(.+)_.+", "\\1", samp_nm),
               genes = as.data.frame(mstr_features_anno))
colnames(h3k18la.dge) <- samp_nm
h3k18la.dge$samples$group <- factor(h3k18la.dge$samples$group, levels = unique(h3k18la.dge$samples$group))

h3k18la.dge$rpkm <- rpkm(h3k18la.dge, gene.length = "width")

h3k18la.dge$rpkm_mean <- h3k18la.dge$rpkm %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(P4 = mean(c(H3K18La_P4_R1, H3K18La_P4_R2)),
         P10 = mean(c(H3K18La_P10_R1, H3K18La_P10_R2)),
         P28 = mean(c(H3K18La_P28_R1, H3K18La_P28_R2, H3K18La_P28_R3)))

h3k18la.pc.prom <- cbind(h3k18la.dge$genes, h3k18la.dge$rpkm_mean) %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")




##########
## h3k27ac
samp_nm <- gsub(".+([Pa].+H3K27Ac).+", "\\1", h3k27ac_counts@colData[[1]])
samp_nm <- paste(samp_nm, rep(c("R1", "R2"), 3), sep = "_")
h3k27ac.dge <- asDGEList(h3k27ac_counts,
               norm.factors = h3k27ac_se$norm.factors,
               group = gsub("_.+", "", samp_nm),
               genes = as.data.frame(mstr_features_anno))
colnames(h3k27ac.dge) <- samp_nm
h3k27ac.dge$samples$group <- factor(h3k27ac.dge$samples$group, levels = unique(h3k27ac.dge$samples$group))

h3k27ac.dge$rpkm <- rpkm(h3k27ac.dge, gene.length = "width")

h3k27ac.dge$rpkm_mean <- h3k27ac.dge$rpkm %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(P4 = mean(c(P3_H3K27Ac_R1, P3_H3K27Ac_R2)),
         P10 = mean(c(P10_H3K27Ac_R1, P10_H3K27Ac_R2)),
         P28 = mean(c(adult_H3K27Ac_R1, adult_H3K27Ac_R2)))

h3k27ac.pc.prom <- cbind(h3k27ac.dge$genes, h3k27ac.dge$rpkm_mean) %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")


##########
## h3k4me3
samp_nm <- gsub(".+([Pa].+H3K4me3).+", "\\1", h3k4me3_counts@colData[[1]])
samp_nm <- paste(samp_nm, rep(c("R1", "R2"), 3), sep = "_")
h3k4me3.dge <- asDGEList(h3k4me3_counts,
               norm.factors = h3k4me3_se$norm.factors,
               group = gsub("_.+", "", samp_nm),
               genes = as.data.frame(mstr_features_anno))
colnames(h3k4me3.dge) <- samp_nm
h3k4me3.dge$samples$group <- factor(h3k4me3.dge$samples$group, levels = unique(h3k4me3.dge$samples$group))

h3k4me3.dge$rpkm <- rpkm(h3k4me3.dge, gene.length = "width")

h3k4me3.dge$rpkm_mean <- h3k4me3.dge$rpkm %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(P4 = mean(c(P3_H3K4me3_R1, P3_H3K4me3_R2)),
         P10 = mean(c(P10_H3K4me3_R1, P10_H3K4me3_R2)),
         P28 = mean(c(adult_H3K4me3_R1, adult_H3K4me3_R2)))

h3k4me3.pc.prom <- cbind(h3k4me3.dge$genes, h3k4me3.dge$rpkm_mean) %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")

```



################################################################################
### Alluvial plot of promoter peak histone presence - Figure 3
################################################################################
```{r}
library(ggalluvial)


thresh = 2
# Prep the gene lists for the FGE loop
gene_list_P4 <- list(H3K27ac = h3k27ac.pc.prom %>% 
                       dplyr::filter(P4 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K4me3 = h3k4me3.pc.prom %>% 
                       dplyr::filter(P4 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K18la = h3k18la.pc.prom %>% 
                       dplyr::filter(P4 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F))

gene_list_P10 <- list(H3K27ac = h3k27ac.pc.prom %>% 
                       dplyr::filter(P10 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K4me3 = h3k4me3.pc.prom %>% 
                       dplyr::filter(P10 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K18la = h3k18la.pc.prom %>% 
                       dplyr::filter(P10 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F))

gene_list_P28 <- list(H3K27ac = h3k27ac.pc.prom %>% 
                       dplyr::filter(P28 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K4me3 = h3k4me3.pc.prom %>% 
                       dplyr::filter(P28 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F),
                     H3K18la = h3k18la.pc.prom %>% 
                       dplyr::filter(P28 > thresh) %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F))




goi_list <- list(gene_list_P4, gene_list_P10, gene_list_P28)
names(goi_list) <- c("P4", "P10", "P28")

total_gois <- c(gene_list_P4 %>% unlist(use.names = F) %>% unique(),
                gene_list_P10 %>% unlist(use.names = F) %>% unique(),
                gene_list_P28 %>% unlist(use.names = F) %>% unique()) %>% unique()


# Prep DF for alluvial plot
df_alluv <- data.frame()

# 
for (i in names(goi_list)){
  
  tmp_venn <- ggVennDiagram::Venn(goi_list[[i]])
  tmp_data <- ggVennDiagram::process_data(tmp_venn)
  tmp_regions <- tmp_data@region$item
  names(tmp_regions) <- tmp_data@region$name

  for (ii in names(tmp_regions)){
    
    tmp_genes <- tmp_regions[[ii]]
    tmp_df <- data.frame(Symbol = tmp_genes,
                         TP = i,
                         Region = ii)
   
    
    df_alluv <- rbind(df_alluv, tmp_df)
      
  }
  
   tmp_df_missing <- data.frame(Symbol = total_gois[!(total_gois %in% (tmp_regions %>% unlist(use.names = F) %>% unique()))],
                                TP = i,
                                Region = NA)
   df_alluv <- rbind(df_alluv, tmp_df_missing)
}

df_alluv$TP <- factor(df_alluv$TP, levels = c("P4", "P10", "P28"))
df_alluv <- df_alluv %>% as.tibble() %>% replace_na(list(Region = "no_mark"))

hist_barplot_levels <- df_alluv$Region %>% unique() %>% sort()
hist_barplot_levels <- hist_barplot_levels[c(8,1,6,2,7,3,4,5)]
# hist_barplot_levels[1] <- NA
df_alluv$Region <- factor(df_alluv$Region, levels = hist_barplot_levels)

gga <- ggplot(df_alluv, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) +
  # scale_fill_brewer(type = "qual", palette = "Set2") +
  # scale_fill_viridis_b() +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) +
  geom_stratum(alpha = 0.75) +
  theme(legend.position = "bottom") +
  ggtitle("Histone occupancy in promoter region during development") 
cowplot::ggsave2(gga, filename = "../data/processed/08_Alluvial_Prom_18la-27ac-4me3_241008_filt2.pdf",
                 height = 8, width = 12)

saveRDS(object = df_alluv, file = "../data/interim/08_df_alluvial_241008_filt2.RDS")

```


### Perform FGE on Alluvial segments - Figure 3
```{r}

source(file.path(dest_tools_dir, "GO_Leaf.R"))

df_alluv_spread <- df_alluv %>% spread(TP, Region)


goi_list_alluv <- list(P4_all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P10_acquired_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(!(P4 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P10 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_acquired_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(!(P10 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la_H3K4me3 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la_H3K27ac = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K27ac..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       H3K18la_acquired = df_alluv_spread %>% 
                         dplyr::filter(!(P4 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P10 == "H3K27ac..H3K4me3..H3K18la",
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F))


# Perform FGE on the GOIs from each subset of peaks

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 12

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_export <- "../data/processed"
dir_enrich <- file.path(dir_export, paste0("08_Alluvial_Regions_241009v2_", enrich_db, "_", GOont))
dir.create(dir_enrich)

# univ <- df_alluv_spread$Symbol
univ <- mstr_features_anno@anno$SYMBOL %>% unique() 
  
enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list_alluv, fun = "enrichGO",
                                          keyType = "SYMBOL", ont = GOont, OrgDb = org, univ = univ)


# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_full.tsv")),
              quote = F, row.names = F, sep = "\t")



# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Simplify the ernichment
simp <- clusterProfiler::simplify(enrich_results)
write.table(simp@compareClusterResult, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_simp.tsv")),
              quote = F, row.names = F, sep = "\t")


# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>% 
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>% 
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("H3K18la Peaks in ", i))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0(i, "_Alluvial-Region_FGE_top_", top, "_leaf_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)
    

# Make a bargraph of top 25 from simplified enriched results
source("../src/tools/Barplot_GO.R")

bar_plot_list <- list()
top = 50
for (i in simp@compareClusterResult$Cluster %>% unique()){
  tmp_res <- simp@compareClusterResult %>% dplyr::filter(Cluster == i)
  tmp_res <- tmp_res %>% slice_head(n = top)
  tmp_res$Description <- factor(tmp_res$Description, levels = rev(tmp_res$Description))
  bp <- ggplot(tmp_res, aes(x = Description, y = Count, fill = -log10(p.adjust))) +
    geom_bar(stat = "identity") +
    coord_flip() +
   # scale_fill_gradient(low="#3B9AB2", high="#F21A00", name="P Value \n(-log10)",limits=c(0,10), oob = scales::squish) +
     scale_fill_gradient2(low="#225ea8", mid="#7fcdbb", high="#ffffd9", midpoint = 5, name="P Value \n(-log10)",limits=c(0,10), oob = scales::squish) +
    labs(y="Number of Significant Genes", x=NULL) +
    theme_bw()
  cowplot::ggsave2(plot = bp, filename = file.path(dir_enrich, paste0(i, "_Alluvial-Region_FGEsimp_top_", top, "_barplotv2.pdf")),
                 height = 12, 
                 width = 10)
  # bar_go(tmp_res, top = 25, desc = "Description", p.val = "p.adjust", genenum = "Count", filenm = file.path(dir_enrich, paste0(i, "_FGE_simp_", i,"_top_", top, "_barplat.pdf")))
}
# bar_go(simp@compareClusterResult, top = 25, desc = "Description", p.val = "p.adjust", genenum = "Count")




```




################################################################################
### RNAseq correlated Alluvial plot - Figure 4
################################################################################
```{r}
library(ggalluvial)

# Load alluvial data.frame from 02
# df_alluv <- readRDS("../data/interim/02_df_alluvial.RDS")

# Correlated genes
df_alluv_corr <- df_alluv %>% 
  dplyr::filter(Symbol %in% posi_cor$SYMBOL)

gga <- ggplot(df_alluv_corr, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) +
  # scale_fill_brewer(type = "qual", palette = "Set2") +
  # scale_fill_viridis_b() +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) +
  geom_stratum(alpha = 0.75) +
  theme(legend.position = "bottom") +
  ggtitle("Histone occupancy in correlated promoter region during development") 

cowplot::ggsave2(gga, filename = file.path(dir_export, paste0("08_Alluvial_Corr_Prom_18la-27ac-4me3.pdf")),
                 height = 8, width = 12)


# Anti-Correlated genes
df_alluv_anticorr <- df_alluv %>% 
  dplyr::filter(Symbol %in% anti_cor$SYMBOL)

gga <- ggplot(df_alluv_anticorr, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) +
  # scale_fill_brewer(type = "qual", palette = "Set2") +
  # scale_fill_viridis_b() +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) +
  geom_stratum(alpha = 0.75) +
  theme(legend.position = "bottom") +
  ggtitle("Histone occupancy in anti-correlated promoter region during development") 

cowplot::ggsave2(gga, filename = file.path(dir_export, paste0("08_Alluvial_AntiCorr_Prom_18la-27ac-4me3.pdf")),
                 height = 8, width = 12)


# Un-Correlated genes
df_alluv_uncorr <- df_alluv %>% 
  dplyr::filter(!(Symbol %in% c(posi_cor$SYMBOL, anti_cor$SYMBOL)))

gga <- ggplot(df_alluv_uncorr, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) +
  # scale_fill_brewer(type = "qual", palette = "Set2") +
  # scale_fill_viridis_b() +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) +
  geom_stratum(alpha = 0.75) +
  theme(legend.position = "bottom") +
  ggtitle("Histone occupancy in uncorrelated promoter region during development") 

cowplot::ggsave2(gga, filename = file.path(dir_export, paste0("08_Alluvial_UnCorr_Prom_18la-27ac-4me3.pdf")),
                 height = 8, width = 12)


```



### Perform FGE on Alluvial segments - Figure 4
```{r}
source("../src/tools/GO_leaf.R")

df_alluv_spread <- df_alluv_corr %>% spread(TP, Region)


goi_list_alluv <- list(P4_all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P10_acquired_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(!(P4 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P10 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_acquired_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(!(P10 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la_H3K4me3 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28_H3K18la_H3K27ac = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K27ac..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       H3K18la_acquired = df_alluv_spread %>% 
                         dplyr::filter(!(P4 == "H3K27ac..H3K4me3..H3K18la"), 
                                       P10 == "H3K27ac..H3K4me3..H3K18la",
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         dplyr::select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F))


# Perform FGE on the GOIs from each subset of peaks

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 10

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_enrich <- file.path(dir_export, paste0("08_Alluvial_Regions_Corr_", enrich_db, "_", GOont))
dir.create(dir_enrich)

# univ <- df_alluv_spread$Symbol
univ <- mstr_features_anno@anno$SYMBOL %>% unique()
  
enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list_alluv, fun = "enrichGO",
                                          keyType = "SYMBOL", ont = GOont, OrgDb = org)


# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Simplify the ernichment
simp <- clusterProfiler::simplify(enrich_results)
write.table(simp@compareClusterResult, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_simp.tsv")),
              quote = F, row.names = F, sep = "\t")


# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get top 5 pathways per cell type to get union set of pathways
# top_union <- enrich_results_leaf %>%
top_union <- simp@compareClusterResult %>%
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top_union$ID %>% unique()))

# Prep the df for plotting
top_enrich <- top_enrich %>% 
  arrange(Cluster, p.adjust)
top_enrich$Description <- substring(top_enrich$Description, 1, 60)
top_enrich$Description <- factor(top_enrich$Description, levels = rev(unique(top_enrich$Description)))
top_enrich$p.adjust <- -log10(top_enrich$p.adjust)

# Make plot
pp <- ggplot(top_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("GO:BP in Histone Occupancy in Correlated peaks"))

#Export plot
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0(i, "_Alluvial-Region_FGE_top_", top, "_", i, "_simp_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)

# Make barplots for each section

for (i in top_union$Cluster){
  tmp_dat <- top_union %>% dplyr::filter(Cluster == i)
  tmp_dat$Description <- factor(tmp_dat$Description, levels = rev(tmp_dat$Description))
  
  bp <- ggplot(tmp_dat) +
      geom_bar(stat = 'identity', aes(x=Description, y = Count, fill = -log10(p.adjust))) +
      coord_flip() +
      scale_fill_gradient(low="#3B9AB2", high="#F21A00", name="P Value \n(-log10)",limits=c(0,10), oob = scales::squish) +
      #scale_fill_gradientn(colors = c("#3B9AB2", "#F21A00"), name="P Value (log10)", trans = 'reverse', limits=c(-50,0), oob=squish) +
      labs(y="Number of Significant Genes", x=NULL) +
      ggtitle(paste0("GO:BP - ", i)) +
      theme_bw()
  cowplot::ggsave2(plot = bp, filename = file.path(dir_enrich, paste0("Alluvial-Region_FGE_top_", top, "_", i, "_leaf_01_barplot.pdf")),
                 height = 4, 
                 width = 8)
}

top <- 10
for (i in simp@compareClusterResult$Cluster){
  tmp_dat <- simp@compareClusterResult %>% as.data.frame() %>% dplyr::filter(Cluster == i)
  tmp_dat <- tmp_dat %>% slice_head(n = top)
  tmp_dat$Description <- factor(tmp_dat$Description, levels = rev(tmp_dat$Description))
  
  bp <- ggplot(tmp_dat) +
      geom_bar(stat = 'identity', aes(x=Description, y = Count, fill = -log10(p.adjust))) +
      coord_flip() +
      # scale_fill_gradient(low="#3B9AB2", high="#F21A00", name="P Value \n(-log10)",limits=c(0,10), oob = scales::squish) +
      scale_fill_gradient2(low="#225ea8", mid="#7fcdbb", high="#ffffd9", midpoint = 5, name="P Value \n(-log10)",limits=c(0,10), oob = scales::squish) +
      #scale_fill_gradientn(colors = c("#3B9AB2", "#F21A00"), name="P Value (log10)", trans = 'reverse', limits=c(-50,0), oob=squish) +
      labs(y="Number of Significant Genes", x=NULL) +
      ggtitle(paste0("GO:BP - ", i)) +
      theme_bw()
  cowplot::ggsave2(plot = bp, filename = file.path(dir_enrich, paste0("Alluvial-Region_FGE_top_", top, "_", i, "_simp_01_barplot_v2.pdf")),
                 height = 4, 
                 width = 8)
}

tmp_res <- simp@compareClusterResult %>% dplyr::filter(Cluster == i)
  tmp_res <- tmp_res %>% slice_head(n = top)
  tmp_res$Description <- factor(tmp_res$Description, levels = rev(tmp_res$Description))

```


