---
title: "03 - H3K18la DB Peak Analysis"
output: html_notebook
---
	
	
################################################################################
#### Get custom scripts and functions and outside data used for analysis
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

script_dir <- "https://raw.githubusercontent.com/MBrooks313/R_codes/master"
dest_tools_dir <- "../src/tools"
dest_vis_dir <- "../src/visualization"

download.file(url = file.path(script_dir, "GO_leaf.R"),
              destfile = file.path(dest_tools_dir, "GO_leaf.R"),
              method = "curl")

download.file(url = file.path(script_dir, "DFave.R"),
              destfile = file.path(dest_tools_dir, "DFave.R"),
              method = "curl")

download.file(url = file.path(script_dir, "Heatmap.R"),
              destfile = file.path(dest_tools_dir, "Heatmap.R"),
              method = "curl")

download.file(url = file.path(script_dir, "Expr_Filter.R"),
              destfile = file.path(dest_tools_dir, "Expr_Filter.R"),
              method = "curl")

library(tidyverse)

dir_export <- "../data/processed"

```


################################################################################
#### Import outside RNAseq data
################################################################################
```{r}
source("../src/tools/DFave.R")

#############################################
# Import expression data for flow-sorted rods
fsprwt.dwnld <- read.csv("../data/external/Mouse_FSPR_WT_Development_Gene_nCPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)
# Get group info for averaging
grp <- gsub("(.+)\\.FS.+", "\\1", colnames(fsprwt.dwnld)[10:25])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
fsprwt_cpm <- data.frame(merge(fsprwt.dwnld[,1:9],
                               log2(DFave(fsprwt.dwnld[,10:25], exp_grp) + 1),
                               by = 0),
                         row.names = 1)

fsprwt_z <- data.frame(merge(fsprwt_cpm[,1:9],
                               t(scale(t(fsprwt_cpm[,10:15]), scale = T, center = T)),
                               by = 0),
                         row.names = 1)



#############################################
# Import expression data for developing retina
retina.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)

# Get group info for averaging
grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.dwnld)[10:31])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
retina_cpm <- data.frame(merge(retina.dwnld[,1:9],
                               log2(DFave(retina.dwnld[,10:31], exp_grp) + 1),
                               by = 0),
                         row.names = 1)

retina_z <- data.frame(merge(retina_cpm[,1:9],
                               t(scale(t(retina_cpm[,10:19]), scale = T, center = T)),
                               by = 0),
                         row.names = 1)


#############################################
# Import transcript expression data for developing retina
retina.trans.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Trans_nRKPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)

# Get group info for averaging
grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.trans.dwnld)[12:33])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
retina_trans_cpm <- data.frame(merge(retina.trans.dwnld[,1:11],
                               log2(DFave(retina.trans.dwnld[,12:33], exp_grp) + 1),
                               by = 0),
                         row.names = 1)

retina_trans_z <- data.frame(merge(retina_trans_cpm[,1:11],
                               t(scale(t(retina_trans_cpm[,12:21]), scale = T, center = T)),
                               by = 0),
                         row.names = 1)


```


################################################################################
### Get Ensembl annotation database
################################################################################
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

## Query for all available EnsDb databases
ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

ensDb <- ahdb[[1]]
ensDb_pc <- addFilter(ensDb, TxBiotypeFilter("protein_coding", condition = "=="))
ensDb_pc_hard <- genes(ensDb_pc)
ensDb_pc_hard <- ensembldb::filter(ensDb, filter = ~ tx_biotype == "protein_coding")

```


################################################################################
### Get list of bam files and make master bed feature for quantitation
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(tidyverse)
library(ChIPpeakAnno)
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


# Macs2 beds
bam_dir <- "../data/raw/bams/"
macs_bams <- list.files(path = bam_dir, pattern = "H3K18La_P.+.bam$")
macs_bams <- macs_bams[c(6:7, 1:2, 3:5)]

# Import consensus peaks
peakList_cons <- readRDS(file = "../data/interim/peakList_cons_240911.RDS")
peakList_cons <- peakList_cons[c(1:3)]

mstr_features <- IRanges::reduce(c(peakList_cons[[1]],
                         peakList_cons[[2]],
                         peakList_cons[[3]]))

# Annotate the features
mstr_features_anno <- annotatePeak(mstr_features, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
mstr_features_anno@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = mstr_features_anno@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")

# Export bed file of the regions
mstr_feature_bed <- data.frame(chrom = seqnames(mstr_features_anno@anno),
                             start = start(mstr_features_anno@anno),
                             end = end(mstr_features_anno@anno),
                             name = mstr_features_anno@anno$SYMBOL,
                             score = width(mstr_features_anno@anno),
                             strand = strand(mstr_features_anno@anno))

write.table(mstr_feature_bed, file = "../data/processed/03_mstr_feature.bed", sep = "\t", quote = F, col.names = F, row.names = F)

```


################################################################################
### Get bam counts over mstr_feature
################################################################################
```{r}
library(csaw)

# Perform region counts
pe.param <- readParam(max.frag=600, pe="both", minq = 20, dedup = T)
bam.files <- paste0(bam_dir, macs_bams)

h3k18la.counts <- regionCounts(bam.files, mstr_features_anno@anno, ext=150, param=pe.param)


# Normalization
binned <- windowCounts(bam.files, bin=TRUE, width=10000, param=pe.param)
h3k18la_se <- normFactors(binned, se.out=T)

```


################################################################################
### Prep the DGE list objects
################################################################################
```{r}
library(edgeR)

# source(file.path(dest_tools_dir, "DFave.R"))
source(file.path(dest_tools_dir, "Expr_Filter.R"))


##########
## h3k18la
samp_nm <- gsub(".+H3K18La_(.+)\\.target.+", "\\1", h3k18la.counts@colData[[1]])

h3k18la.dge <- asDGEList(h3k18la.counts,
               norm.factors = h3k18la_se$norm.factors,
               group = gsub("_.+", "", samp_nm),
               genes = as.data.frame(mstr_features_anno))
colnames(h3k18la.dge) <- samp_nm
h3k18la.dge$samples$group <- factor(h3k18la.dge$samples$group, levels = unique(h3k18la.dge$samples$group))

h3k18la.dge$rpkm <- rpkm(h3k18la.dge, gene.length = "width")
ExpFilt(h3k18la.dge$rpkm, exp = 1, group = h3k18la.dge$samples$group)

h3k18la.dge.filt <- h3k18la.dge[idx.filt, keep.lib.sizes = T]

h3k18la.dge.filt$logCPM <- log2(cpm(h3k18la.dge.filt) + 1)
h3k18la.dge.filt$logFPKM <- rpkm(h3k18la.dge.filt, gene.length = "width", normalized.lib.sizes = T, log = T, prior.count = 1)
h3k18la.dge.filt$FPKM <- rpkm(h3k18la.dge.filt, gene.length = "width", normalized.lib.sizes = T, log = F)
h3k18la.dge.filt$FPKM_mean <- h3k18la.dge.filt$FPKM %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(P4 = mean(c(P4_R1, P4_R2)),
         P10 = mean(c(P10_R1, P10_R2)),
         P28 = mean(c(P28_R1, P28_R2, P28_R3)))

# Export RPKM quant
tmp <- cbind(as.data.frame(mstr_features_anno), h3k18la.dge$rpkm)
write.table(tmp, file = "../data/processed/03_tc_quant_rpkm.tsv", sep = "\t", quote = F, col.names = F, row.names = F)


```


################################################################################
### PCA analysis - Figure 4
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(plot3D)
library(PCAtools)
library(wesanderson)

#Prep the PCA data
pca.h3k18la <- PCAtools::pca(h3k18la.dge.filt$logFPKM, metadata = data.frame(h3k18la.dge$samples))


X='PC1'; Y='PC2'
pca1 <- biplot(pca.h3k18la, x = X, y = Y,
       colby = "group", 
       legendPosition = 'right', 
       title = 'Peak Quant PCA',
       subtitle = paste0(X, ' versus ', Y))

cowplot::ggsave2(filename = "../data/processed/03_PCA_PC1-2.pdf", plot = pca1,
                 width = 8, height = 6)


# PCA of RNA-seq - Nov 18, 2024
tmp_samps <-colnames(fsprwt.dwnld)[c(12:13,17:19,22:25)]
tmp_rna_cpm <- log2(fsprwt.dwnld[,tmp_samps] + 1)
tmp_rna_meta <- data.frame(sample = gsub("\\.FS.+WT", "", tmp_samps),
                           group = gsub("(.+)\\.FS.+", "\\1", tmp_samps),
                           rep = gsub(".+WT\\.(.+)", "\\1", tmp_samps))
row.names(tmp_rna_meta) <- tmp_samps
tmp_rna_meta$group <- factor(tmp_rna_meta$group, levels = c("P4", "P10", "P28"))

pca.rna <- PCAtools::pca(tmp_rna_cpm, metadata = tmp_rna_meta)

X='PC1'; Y='PC2'
pca2 <- biplot(pca.rna, x = X, y = Y,
       colby = "group", 
       legendPosition = 'right', 
       title = 'Peak Quant PCA',
       subtitle = paste0(X, ' versus ', Y))

cowplot::ggsave2(filename = "../data/processed/03_PCA_PC1-2_RNA.pdf", plot = pca2,
                 width = 8, height = 6)

```


################################################################################
### Differential binding analysis
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(edgeR)
library(rtracklayer)
library(qvalue)

y <- h3k18la.dge.filt

design <- model.matrix(~h3k18la.dge.filt$samples$group)
colnames(design) <- levels(h3k18la.dge.filt$samples$group)

y <- estimateDisp(y, design)
summary(y$trended.dispersion)

fit <- glmQLFit(y, design, robust=TRUE)
summary(fit$var.post)

par(mfrow=c(1,2))
o <- order(y$AveLogCPM)
plot(y$AveLogCPM[o], sqrt(y$trended.dispersion[o]), type="l", lwd=2,
ylim=c(0, 1), xlab=expression("Ave."~Log[2]~"CPM"),
ylab=("Biological coefficient of variation"))
plotQLDisp(fit)

summary(fit$df.prior)

results <- glmQLFTest(fit, coef=c(2, 3))
head(results$table)

results$table$qvalue <- qvalue(results$table$PValue)$qvalue
results$table$FDR <- p.adjust(results$table$PValue, method = "fdr")

h3k18la.db.results <- cbind(h3k18la.dge.filt$genes, results$table, h3k18la.dge.filt$FPKM_mean)
write.table(h3k18la.db.results, file = "../data/processed/03_H3k18la.db.results.tsv", sep = "\t", quote = F, row.names = F)

table(h3k18la.db.results$FDR < 0.01 & abs(h3k18la.db.results$logFC.P28) > 1)

h3k18la.db.results.sig.prom.pc <- h3k18la.db.results %>% dplyr::filter(FDR < 0.01, annotation == "Promoter", transcriptBiotype == "protein_coding")
```


################################################################################
#### Make heatmaps of DB C&T - Figure 4
################################################################################
```{r}

# Prom PC genes
###################
# Tidy table
dat.tidy <- log2(h3k18la.dge.filt$FPKM + 1) %>%
  as.data.frame() %>% 
  rownames_to_column(var = "peak_Id") %>% 
  dplyr::filter(peak_Id %in% (row.names(prom_pc_sig) %>% as.numeric())) %>% 
  gather(Sample, CPM, -peak_Id)

###################
#Factor sort sample names
dat.tidy$Sample <- gsub("_R", "_", dat.tidy$Sample)
dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample))


###################################################################
# Make Heatmap
###################################################################

# Sort by Heirarchical clustering using Ward's method with Euclidean distanced
distfunc <- function(x) dist(x,method="euclidean")
hclustfunc <- function(x) hclust(x, method="ward.D2")

#Get distance and heirarchical clustering
dat_2clst <- log2(h3k18la.dge.filt$FPKM + 1) %>%
  as.data.frame() %>% 
  rownames_to_column(var = "peak_Id") %>% 
  dplyr::filter(peak_Id %in% (row.names(prom_pc_sig) %>% as.numeric()))

d <- distfunc(as.matrix(dat_2clst[,4:7]))
hc <- hclustfunc(d)

dat.tidy$peak_Id <- factor(dat.tidy$peak_Id, 
                                      levels = dat.tidy$peak_Id[hc$order], ordered = T)


###############################################
# # Sort by PCA
pca <- prcomp(dat_2clst[,2:8], center = T, scale. = F)
pca.order <- order(pca$x[,1])

dat.tidy$peak_Id <- factor(dat.tidy$peak_Id,
                                      levels = dat.tidy$peak_Id[rev(pca.order)], ordered = T)

# Make heatmap
library(scales)
lim = c(0,8)
nm = "FPKM (log2)"
mb.colors <- colorRampPalette(c("#023858","#3690c0", "#e0f3f8","#fff7bc","#fec44f","#ec7014",
                                    "#fc4e2a","#bd0026","#800026"))(n=1000)
pp <- ggplot(dat.tidy, aes(Sample, peak_Id)) +
        geom_tile(aes(fill = CPM), colour = "transparent") + 
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        geom_vline(xintercept = c(2.5,4.5)) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT_hm-pca.pdf", plot = pp,
                 width = 6, height = 6)

```


################################################################################
### Correlate the C&T and RNA seq - Figure 4
################################################################################
```{r}
library(cowplot)

rna_tmp <- retina_cpm %>% 
  rownames_to_column(var = "ENSEMBL") %>% 
  dplyr::select(c("ENSEMBL", "external_gene_name", "P4", "P10", "P28")) %>% 
  group_by(external_gene_name) %>% 
  arrange(desc(P28)) %>% 
  slice_head(n=1)

prom_pc_sig_filt_rna <- merge(prom_pc_sig_filt, rna_tmp, by.x = "SYMBOL", by.y = "external_gene_name")

prom_pc_sig_filt_rna[,c(27:29)] <- log2(prom_pc_sig_filt_rna[,c(27:29)] + 1)

# Find Person correlation between the H3K28La peak quant and RNAseq expression
prom_pc_sig_filt_rna <- prom_pc_sig_filt_rna %>% 
  rowwise() %>% 
  mutate(pearson = cor(x=c(P4.x, P10.x, P28.x), y = c(P4.y, P10.y, P28.y), method = "pearson"))


# Split by correlated and anticorrelated
cor_sig = 0.8
posi_cor <- prom_pc_sig_filt_rna %>% 
  dplyr::filter(pearson > cor_sig)

anti_cor <- prom_pc_sig_filt_rna %>% 
  dplyr::filter(pearson < -cor_sig)


######################
# Plot the positive correlated peak and rna data
pca <- prcomp(posi_cor[,c(27:29, 31:33)], center = T, scale. = F)
pca.order <- order(pca$x[,1])

d <- distfunc(as.matrix(posi_cor[,c(27:29, 31:33)]))
hc <- hclustfunc(d)


dat.tidy <- posi_cor %>% 
  dplyr::select(SYMBOL, P4.x, P10.x, P28.x, P4.y, P10.y, P28.y) %>%
  gather(Sample, CPM, -SYMBOL)

dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample))

dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL, 
                                      levels = dat.tidy$SYMBOL[(pca.order)], ordered = T)

dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL, 
                                      levels = dat.tidy$SYMBOL[hc$order], ordered = T)

lim1 <- c(0,8)
p <- ggplot(dat.tidy %>% dplyr::filter(Sample %in% c("P4.x", "P10.x", "P28.x")), 
            aes(Sample, SYMBOL)) +
        geom_tile(aes(fill = CPM), colour = "transparent") + 
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim1, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))


lim2 <- c(0,16)
pp <- ggplot(dat.tidy %>% dplyr::filter(Sample %in% c("P4.y", "P10.y", "P28.y")), 
             aes(Sample, SYMBOL)) +
        geom_tile(aes(fill = CPM), colour = "transparent") + 
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim2, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))

pg <- plot_grid(p, pp)

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT-RNA_posi-cor_hm-hc2.pdf", plot = pg,
                 width = 8, height = 6)


pcacor <- biplot(PCAtools::pca(posi_cor[,c(27:29, 31:33)]), x = X, y = Y,
       legendPosition = 'right', 
       title = 'Peak Quant PCA',
       subtitle = paste0(X, ' versus ', Y))


######################
# Plot the negative correlated peak and rna data
pca <- prcomp(anti_cor[,c(27:29, 31:33)], center = T, scale. = F)
pca.order <- order(pca$x[,1])

dat.tidy <- anti_cor %>% 
  dplyr::select(SYMBOL, P4.x, P10.x, P28.x, P4.y, P10.y, P28.y) %>%
  gather(Sample, CPM, -SYMBOL)

dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample))
dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL, 
                                      levels = dat.tidy$SYMBOL[rev(pca.order)], ordered = T)

lim1 <- c(0,8)
p <- ggplot(dat.tidy %>% dplyr::filter(Sample %in% c("P4.x", "P10.x", "P28.x")), 
            aes(Sample, SYMBOL)) +
        geom_tile(aes(fill = CPM), colour = "transparent") + 
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim1, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))


lim2 <- c(0,16)
pp <- ggplot(dat.tidy %>% dplyr::filter(Sample %in% c("P4.y", "P10.y", "P28.y")), 
             aes(Sample, SYMBOL)) +
        geom_tile(aes(fill = CPM), colour = "transparent") + 
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim2, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))

pg <- plot_grid(p, pp)

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT-RNA_anti-cor_hm-pca2.pdf", plot = pg,
                 width = 8, height = 6)

```


################################################################################
#### FunctionalGeneEnrichment of cor and anti-cor genes - Figure S2
################################################################################
```{r}

library(clusterProfiler)
library(enrichplot)
library(org.Mm.eg.db)
source(file.path(dest_tools_dir, "GO_leaf.R"))
# source("../src/tools/Dotplot_GO.R")

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 10

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_enrich <- file.path(dir_export, paste0("03_cor-anticor_FuncGeneEnrich_", enrich_db, "_", GOont))
dir.create(dir_enrich)


# Get list of genes associated with the peaks for each subgroup
cor_gene_list <- posi_cor$SYMBOL %>% unique() %>% unlist() %>% as.vector()
anticor_gene_list <- anti_cor$SYMBOL %>% unique() %>% unlist() %>% as.vector()


# Prep the gene lists for the FGE loop
goi_list <- list(cor_gene_list, anticor_gene_list)
names(goi_list) <- c("Posi-cor", "Anti-cor")

univ <- h3k18la.db.results$SYMBOL %>% unique()

enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list, fun = "enrichGO",
                                            keyType = "SYMBOL", ont = GOont, OrgDb = org, universe = univ)
  
  
# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

tmp <- enrich_results_filt_fix[,-11]

# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_cor-anti_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_cor-anti_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>% 
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>% 
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("H3K18la Peaks in Cor anti-cor"))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("Correlation_FGE_top_", top, "_leaf_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)
```

