---
title: "H3K18la Impport & Analysis"
output: html_notebook

---


**NOTES**: The data were generated from nf-core cutandrun pipeline.

**UPDATE**: The definition of promoter (-1000 to +500) was used for peak annotation.
            Filter the peaks for FDR of 10E-6 prior to generating the concensus peaks.



### Get custom scripts and functions and outside data used for analysis
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

script_dir <- "https://raw.githubusercontent.com/MBrooks313/R_codes/master"
dest_tools_dir <- "../src/tools"
dest_vis_dir <- "../src/visualization"

download.file(url = file.path(script_dir, "GO_leaf.R"),
              destfile = file.path(dest_tools_dir, "GO_leaf.R"),
              method = "curl")

download.file(url = file.path(script_dir, "DFave.R"),
              destfile = file.path(dest_tools_dir, "DFave.R"),
              method = "curl")

download.file(url = file.path(script_dir, "Heatmap.R"),
              destfile = file.path(dest_tools_dir, "Heatmap.R"),
              method = "curl")

library(tidyverse)
options(java.parameters = "-Xmx8000m")
print("Completed...")
```

### Get Ensembl annotation database
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

ensDb <- ahdb[[1]]

```

################################################################################
## Import outside data
################################################################################
```{r}
source("../src/tools/DFave.R")

#############################################
# Import expression data for flow-sorted rods
fsprwt.dwnld <- read.csv("../data/external/Mouse_FSPR_WT_Development_Gene_nCPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)
# Get group info for averaging
grp <- gsub("(.+)\\.FS.+", "\\1", colnames(fsprwt.dwnld)[10:25])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
fsprwt_cpm <- data.frame(merge(fsprwt.dwnld[,1:9],
                               log2(DFave(fsprwt.dwnld[,10:25], exp_grp) + 1),
                               by = 0),
                         row.names = 1)

fsprwt_z <- data.frame(merge(fsprwt_cpm[,1:9],
                               t(scale(t(fsprwt_cpm[,10:15]), scale = T, center = T)),
                               by = 0),
                         row.names = 1)



#############################################
# Import expression data for developing retina
retina.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)

# Get group info for averaging
grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.dwnld)[10:31])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
retina_cpm <- data.frame(merge(retina.dwnld[,1:9],
                               log2(DFave(retina.dwnld[,10:31], exp_grp) + 1),
                               by = 0),
                         row.names = 1)

retina_z <- data.frame(merge(retina_cpm[,1:9],
                               t(scale(t(retina_cpm[,10:19]), scale = T, center = T)),
                               by = 0),
                         row.names = 1)



#############################################
# Import the ChromHMM from Aldiri data
chromhmm_beds <- list.files(path = "/Volumes/data-1/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/chromhmm", pattern = "*bed", recursive = T, full.names = T)
names(chromhmm_beds) <- gsub(".+\\/(.+)_16.+", "\\1", chromhmm_beds)

# Import the beds 
chromHMM <- sapply(chromhmm_beds, toGRanges, format="BED", skip=1)

# Add the HMM calls to the rows
for (i in 1:length(chromhmm_beds)){
  mcols(chromHMM[[i]])$HMM <- read.table(chromhmm_beds[i], skip = 1)[,4]
  seqlevelsStyle(chromHMM[[i]]) <- "NCBI"
}



#############################################
# Import the ATAC-seq from Aldiri
atac_beds <- list.files(path = "/Volumes/data-1/NGS/ATAC-seq/ENCODE_v1_9_0/Mouse/Aldiri_GSE102092/bed", pattern = "*narrowPeak.gz", recursive = T, full.names = T)
names(atac_beds) <- gsub(".+\\/(.+)_idr.+", "\\1", atac_beds)

# Import the beds 
atac <- sapply(atac_beds, toGRanges, format="narrowPeak")

for (i in 1:length(atac)){
  atac[[i]] <- reduce(atac[[i]])
  seqlevelsStyle(atac[[i]]) <- "NCBI"
}



#############################################
# Import the H3K27ac from Aldiri
k27_aldiri_beds <- list.files(path = "/Volumes/data-1/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bed/", pattern = "*H3K27Ac*", recursive = T, full.names = T)
names(k27_aldiri_beds) <- gsub(".+\\/(.+)_H3K27.+", "\\1", k27_aldiri_beds)

# Import the beds 
k27_aldiri <- sapply(k27_aldiri_beds, toGRanges, format="narrowPeak")

for (i in 1:length(k27_aldiri)){
  k27_aldiri[[i]] <- GenomicRanges::reduce(k27_aldiri[[i]])
  seqlevelsStyle(k27_aldiri[[i]]) <- "NCBI"
}


#############################################
# Import the H3K4me3 from Aldiri
k4me3_aldiri_beds <- list.files(path = "/Volumes/data-1/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bed/", pattern = "*H3K4me3_*", recursive = T, full.names = T)
names(k4me3_aldiri_beds) <- gsub(".+\\/(.+)_H3K4me3.+", "\\1", k4me3_aldiri_beds)

# Import the beds 
k4me3_aldiri <- sapply(k4me3_aldiri_beds, toGRanges, format="narrowPeak")

for (i in 1:length(k4me3_aldiri)){
  k4me3_aldiri[[i]] <- GenomicRanges::reduce(k4me3_aldiri[[i]])
  seqlevelsStyle(k4me3_aldiri[[i]]) <- "NCBI"
}



#############################################
# Import the Nrl peaks from Liang
nrl_beds <- list.files(path = "/Volumes/data-1/People/brooksma/PEGASUS/Projects/Xulong/CutNRun/Mouse/Analysis/20200309_16-29/200723/data/interim/beds/idr", pattern = "Nrl.+idr05.+narrowPeak", recursive = F, full.names = T)
names(nrl_beds) <- gsub(".+\\/Nrl_(.+)\\.Rep.+", "\\1", nrl_beds)

# Import the beds 
nrl <- sapply(nrl_beds, toGRanges, format="narrowPeak")

for (i in 1:length(nrl)){
  seqlevelsStyle(nrl[[i]]) <- "NCBI"
}


```


### Annotate outside data
```{r}

# Aldiri k27ac 
k27AnnoList_ens <- list()

for (i in 1:length(k27_aldiri)){
  
  peak_tmp <- k27_aldiri[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  
  k27AnnoList_ens[[i]] <- anno_tmp

}

names(k27AnnoList_ens) <- names(k27_aldiri)


# Aldiri k4me3 
k4me3AnnoList_ens <- list()

for (i in 1:length(k4me3_aldiri)){
  
  peak_tmp <- k4me3_aldiri[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  
  k4me3AnnoList_ens[[i]] <- anno_tmp

}

names(k4me3AnnoList_ens) <- names(k4me3_aldiri)


```


################################################################################
## Timecourse H3K18la samples
################################################################################

### Get metadata for samples
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}

# Macs2 beds from NF-core
macs_beds <- list.files(path = "../data/interim/bed/macs2/", pattern = "narrowPeak", recursive = T)

# Meta sheet
meta <- data.frame(Beds = macs_beds[1:7],
                   Samp = gsub("H3K18La_(P.+)\\.macs.+", "\\1", macs_beds[1:7]),
                   Group = gsub("H3K18La_(P\\d+)_R.+", "\\1", macs_beds[1:7]))

meta <- meta[c(6,7,1:5),]

meta$Samp <- factor(meta$Samp, levels = unique(meta$Samp))
meta$Group <- factor(meta$Group, levels = unique(meta$Group))

```


### Import the peaks into a merged list object and plot for number of peaks at each timepoint
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
library(ChIPpeakAnno)

macs_dir <- "../data/interim/bed/macs2/"

# Import the cons peaks
peakList <- sapply(file.path(macs_dir, meta$Beds), toGRanges, format="narrowPeak")
names(peakList) <- meta$Samp

peakList_1E6 <- GRangesList()
# Sort the cons peaks
for (i in 1:length(peakList)){
  
  # seqlevelsStyle(peakList[[i]]) <- "NCBI"
  tmp_peak <- sortSeqlevels(peakList[[i]])
  tmp_peak <- GenomicRanges::sort(tmp_peak)
  tmp_peak <- tmp_peak[(elementMetadata(tmp_peak)[,"qValue"] > 6) & (seqnames(tmp_peak) != "MT")]
  peakList_1E6[[i]] <- tmp_peak
}
names(peakList_1E6) <- meta$Samp

# Gen consensus peaks from each timepoint
ol_P4 <- findOverlapsOfPeaks(c(peakList_1E6[1],
                             peakList_1E6[2]), connectedPeaks="keepAll", ignore.strand = T)

ol_P10 <- findOverlapsOfPeaks(c(peakList_1E6[3],
                             peakList_1E6[4]), connectedPeaks="keepAll", ignore.strand = T)

ol_P28 <- findOverlapsOfPeaks(c(peakList_1E6[5],
                             peakList_1E6[6],
                             peakList_1E6[7]), connectedPeaks="keepAll", ignore.strand = T)

peakList_cons <- GRangesList(ol_P4$mergedPeaks, ol_P10$mergedPeaks, ol_P28$mergedPeaks)
names(peakList_cons) <- c("P4", "P10", "P28")

```


### Annotate peaks with ensembl db
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


peakAnnoList_ens <- list()

for (i in 1:length(peakList_cons)){
  
  peak_tmp <- peakList_cons[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  peakAnnoList_ens[[i]] <- anno_tmp

}

names(peakAnnoList_ens) <- names(peakList_cons)

```


### Evaluate location of peaks relative to annotation features - Figure 3
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}


#######################
# Consensus peak number
#######################

# Generate plot for peak number
df <- lapply(peakList_cons, length) %>% as.data.frame() %>% gather("Sample", "Peaks")
df$Sample <- gsub("K18La\\.", "", df$Sample)
df$Sample <- gsub("X", "", df$Sample)
df$Sample <- factor(df$Sample, levels = df$Sample)

uniq_genes <- lapply(peakAnnoList_ens, function(x){x@anno$SYMBOL %>% unique() %>% length()}) %>% unlist() %>% as.vector()
prom_genes <- lapply(peakAnnoList_ens, function(x){x@anno %>% 
    as.data.frame() %>% 
    dplyr::filter(annotation == "Promoter") %>% 
    dplyr::select(SYMBOL) %>% 
    as.vector() %>% unlist() %>% 
    unique() %>% length()}) %>% 
  unlist() %>% as.vector()

b <- ggplot(df, aes(Sample, Peaks))+
  geom_bar(stat = "identity", fill = "grey50") +
  geom_point(aes(x = c(1:3), y = uniq_genes), colour = "red", size = 2) +
  annotate("text", x = c(1:3), y = uniq_genes + 200, label = uniq_genes) +
  geom_point(aes(x = c(1:3), y = prom_genes), colour = "blue", size = 2) +
  annotate("text", x = c(1:3), y = prom_genes + 200, label = prom_genes) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 0.8, vjust = 0.8)) +
  labs(title = "Peaks") + xlab(NULL)

cowplot::ggsave2(plot = b, filename = "../data/processed/01_Peak_number_TP.pdf", height = 3, width = 2.5)


#######################
# Consensus peak evaluation
#######################

promoter <- getPromoters(TxDb = ensDb, upstream = 3000, downstream = 3000)
seqlevelsStyle(promoter) <- "NCBI"

tagMatrixList <- list()
for (i in names(peakList_cons)){
  tagMatrixList[[i]] <- getTagMatrix(peakList_cons[[i]], windows=promoter)
}


# Plot peak histogram around TSS
p <- plotAvgProf(tagMatrixList, xlim=c(-3000, 3000), resmple = 100,
            xlab="Genomic Region (5'->3')", ylab = "Read Count Frequency")
print(p)

pdf("../data/processed/01_Peak_TSS_hist_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


# Plot annotation bargraph
p <- plotAnnoBar(peakAnnoList_ens, title = "H3K18la peak annotation")
print(p)

pdf("../data/processed/01_Peak_annot_bar_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


# Plot disrtobution of peaks realtive to TSS
p <- plotDistToTSS(peakAnnoList_ens,
              title="Distribution of H3K18la peak to TSS")
print(p)

pdf("../data/processed/01_Peak_TSSdist_bar_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


```


### FunctionalGeneEnrichment - Figure S1
```{r}

library(clusterProfiler)
library(enrichplot)
library(org.Mm.eg.db)
source(file.path(dest_tools_dir, "GO_leaf.R"))
# source("../src/tools/Dotplot_GO.R")

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 5

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_enrich <- file.path(dir_export, paste0("01_FuncGeneEnrich_", enrich_db, "_", GOont))
dir.create(dir_enrich)



# Get subsets of annotation to run FGE based on last mtg's discussion
prom_list <- lapply(peakAnnoList_ens, function(x){x@anno[x@anno$annotation == "Promoter" & 
                                                           x@anno$transcriptBiotype == "protein_coding"]})

genebody_list <- lapply(peakAnnoList_ens, function(x){x@anno[!(x@anno$annotation %in% c("Promoter", "Downstream", "Distal Intergenic")) & 
                                                           x@anno$transcriptBiotype == "protein_coding"]})

enhancer_list <- lapply(peakAnnoList_ens, function(x){x@anno[x@anno$distanceToTSS > -10000 & 
                                                               x@anno$distanceToTSS < -1000 & 
                                                           x@anno$transcriptBiotype == "protein_coding"]})

# Get list of genes associated with the peaks for each subgroup
prom_gene_list <- lapply(prom_list, function(x){x@elementMetadata$SYMBOL %>% unique() %>% unlist() %>% as.vector()}) 

genebody_gene_list <- lapply(genebody_list, function(x){x@elementMetadata$SYMBOL %>% unique() %>% unlist() %>% as.vector()}) 

enhancer_gene_list <- lapply(enhancer_list, function(x){x@elementMetadata$SYMBOL %>% unique() %>% unlist() %>% as.vector()}) 


# Prep the gene lists for the FGE loop
goi_list <- list(prom_gene_list, genebody_gene_list, enhancer_gene_list)
names(goi_list) <- c("Promoter", "Genebody", "Enhancer")

# Perform FGE on the GOIs from each subset of peaks
for (i in names(goi_list)){
  
  enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list[[i]], fun = "enrichGO",
                                            keyType = "SYMBOL", ont = GOont, OrgDb = org)
  
  
  # Filter enrichment for significance
  enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)
  
  # Fix the Ratio
  enrich_results_filt_fix <- enrich_results_filt %>% 
    as.data.frame() %>% 
    mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
    rowwise() %>% 
    mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))
  
  # Export the FGE results table
  tmp <- enrich_results_filt_fix[,-11]
  # Export the enrichment analysis
  write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_full.tsv")),
                quote = F, row.names = F, sep = "\t")

  # Get leaf terms
  enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)
  
  # Fix the Ratio
  enrich_results_leaf_fix <- enrich_results_leaf %>% 
    as.data.frame() %>% 
    mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
    rowwise() %>% 
    mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))
  
  # Export the FGE results table
  tmp <- enrich_results_leaf_fix[,-11]
  # Export the enrichment analysis
  write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_leaf.tsv")),
                quote = F, row.names = F, sep = "\t")
  
  # Get top 5 pathways per cell type to get union set of pathways
  top5_union <- enrich_results_leaf %>% 
    group_by(Cluster) %>% 
    arrange(p.adjust) %>% 
    slice_head(n = top)
  
  
  # Filter the full table for the union of top5 pathways
  top5_enrich <- enrich_results_filt_fix %>% 
    dplyr::filter(ID %in% (top5_union$ID %>% unique()))
  
  # Prep the df for plotting
  top5_enrich <- top5_enrich %>% 
    arrange(Cluster, p.adjust)
  top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
  top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
  top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)
  
  # Make plot
  pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
      geom_point(aes(size = Ratio)) + 
      theme_bw() +
      scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
      guides(size=guide_legend(title="Ratio")) +
      theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1)) +
      ggtitle(paste0("H3K18la Peaks in ", i))
  
  #Open the pdf device
  cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("Timepoint_FGE_top_", top, "_", i, "_leaf_01_dotplot.pdf")),
                   height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                   width = 6)
    
}

```


################################################################################
## Correlate H3K18La peaks with ATAC, ChromHMM, NRL peaks - Figure 3 and S2
################################################################################
```{r}

########################
# Promoter peak overlaps

###########
# Nrl peaks 
nrl_ol_p4 <- findOverlapsOfPeaks(prom_list$`P4`, nrl$P4, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p10 <- findOverlapsOfPeaks(prom_list$`P10`, nrl$P10, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p28 <- findOverlapsOfPeaks(prom_list$`P28`, nrl$P28, connectedPeaks = "keepAll", maxgap = 500L)

nrl_df <- cbind(lapply(nrl_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Nrl = rep(c("Nrl", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Nrl != "Nrl")
colnames(nrl_df)[1] <- "Count"

nrl_df$Age <- factor(nrl_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

nb <- ggplot(nrl_df, aes(x = Age, y = Count, fill = Nrl)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Promoter Peak Overlap with Nrl")

cowplot::ggsave2(nb, filename = file.path(dir_export, paste0("01_Prom_H3K18la-NRL_OL.pdf")),
                 height = 2, width = 4)

############
# ATAC peaks 
atac_ol_p4 <- findOverlapsOfPeaks(prom_list$`P4`, atac$P3, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p10 <- findOverlapsOfPeaks(prom_list$`P10`, atac$P10, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p28 <- findOverlapsOfPeaks(prom_list$`P28`, atac$P21, connectedPeaks = "keepAll", maxgap = 500L)


atac_df <- cbind(lapply(atac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Atac = rep(c("ATAC", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Atac != "ATAC")
colnames(atac_df)[1] <- "Count"

atac_df$Age <- factor(atac_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(atac_df, aes(x = Age, y = Count, fill = Atac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Promoter Peak Overlap with ATAC")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("01_Prom_H3K18la-ATAC_OL.pdf")),
                 height = 2, width = 4)


############
# k27ac peaks Aldiri
k27ac_ol_p4 <- findOverlapsOfPeaks(prom_list$`P4`, k27_aldiri$P3, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p10 <- findOverlapsOfPeaks(prom_list$`P10`, k27_aldiri$P10, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p28 <- findOverlapsOfPeaks(prom_list$`P28`, k27_aldiri$P21, connectedPeaks = "keepAll", maxgap = 500L)


k27ac_df <- cbind(lapply(k27ac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         k27ac = rep(c("k27ac", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(k27ac != "k27ac")
colnames(k27ac_df)[1] <- "Count"

k27ac_df$Age <- factor(k27ac_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(k27ac_df, aes(x = Age, y = Count, fill = k27ac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Promoter Peak Overlap with H3K27ac")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("01_Prom_H3K18la-k27ac_aldiri_OL.pdf")),
                 height = 2, width = 4)


##########
# ChromHMM
chromHMM_ol_p4 <- findOverlapsOfPeaks(prom_list$`P4`, chromHMM$P3, connectedPeaks = "keepAll")
chromHMM_ol_p10 <- findOverlapsOfPeaks(prom_list$`P10`, chromHMM$P10, connectedPeaks = "keepAll")
chromHMM_ol_p28 <- findOverlapsOfPeaks(prom_list$`P28`, chromHMM$P21, connectedPeaks = "keepAll")

chromHMM_ol_p4_fix <- chromHMM_ol_p4$overlappingPeaks$`prom_list.P4///chromHMM.P3` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p10_fix <- chromHMM_ol_p10$overlappingPeaks$`prom_list.P10///chromHMM.P10` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p28_fix <- chromHMM_ol_p28$overlappingPeaks$`prom_list.P28///chromHMM.P21` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromhmm_df <- rbind(chromHMM_ol_p4_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P4"),
                     chromHMM_ol_p10_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P10"),
                     chromHMM_ol_p28_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P28"))
colnames(chromhmm_df)[1] <- "HMM"

chromhmm_df$HMM <- factor(chromhmm_df$HMM, levels = seq(16,1))
chromhmm_df$Age <- factor(chromhmm_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- RColorBrewer::brewer.pal(6, "Accent")

col_fill <- colorRampPalette(col_pal)(16)

cb <- ggplot(chromhmm_df, aes(x = Age, y = Freq, fill = HMM)) +
  geom_bar(position="fill", stat="identity") +
  viridis::scale_fill_viridis(discrete = T, option = "H", direction = 1, guide = guide_legend(reverse = TRUE)) +
  coord_flip() +
  labs(title = "H3K18la Promoter Peak Overlap with ChromHMM")

cowplot::ggsave2(cb, filename = file.path(dir_export, paste0("01_Prom_H3K18la-ChromHMM_OL.pdf")),
                 height = 2, width = 4)




########################
# Enhancer peak overlaps

###########
# Nrl peaks 
nrl_ol_p4 <- findOverlapsOfPeaks(enhancer_list$`P4`, nrl$P4, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p10 <- findOverlapsOfPeaks(enhancer_list$`P10`, nrl$P10, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p28 <- findOverlapsOfPeaks(enhancer_list$`P28`, nrl$P28, connectedPeaks = "keepAll", maxgap = 500L)

nrl_df <- cbind(lapply(nrl_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Nrl = rep(c("Nrl", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Nrl != "Nrl")
colnames(nrl_df)[1] <- "Count"

nrl_df$Age <- factor(nrl_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

nb <- ggplot(nrl_df, aes(x = Age, y = Count, fill = Nrl)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Enhancer Peak Overlap with Nrl")

cowplot::ggsave2(nb, filename = file.path(dir_export, paste0("01_Enhancer_H3NRL_OL.pdf")),
                 height = 2, width = 4)

############
# ATAC peaks 
atac_ol_p4 <- findOverlapsOfPeaks(enhancer_list$`P4`, atac$P3, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p10 <- findOverlapsOfPeaks(enhancer_list$`P10`, atac$P10, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p28 <- findOverlapsOfPeaks(enhancer_list$`P28`, atac$P21, connectedPeaks = "keepAll", maxgap = 500L)


atac_df <- cbind(lapply(atac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Atac = rep(c("ATAC", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Atac != "ATAC")
colnames(atac_df)[1] <- "Count"

atac_df$Age <- factor(atac_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(atac_df, aes(x = Age, y = Count, fill = Atac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Enhancer Peak Overlap with ATAC")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("01_Enhancer_H3ATAC_OL.pdf")),
                 height = 2, width = 4)


############
# k27ac peaks Aldiri
k27ac_ol_p4 <- findOverlapsOfPeaks(enhancer_list$`P4`, k27_aldiri$P3, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p10 <- findOverlapsOfPeaks(enhancer_list$`P10`, k27_aldiri$P10, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p28 <- findOverlapsOfPeaks(enhancer_list$`P28`, k27_aldiri$P21, connectedPeaks = "keepAll", maxgap = 500L)


k27ac_df <- cbind(lapply(k27ac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         k27ac = rep(c("k27ac", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(k27ac != "k27ac")
colnames(k27ac_df)[1] <- "Count"

k27ac_df$Age <- factor(k27ac_df$Age, levels = rev(c("P4", "P10", "P28")))

# col_pal <- RColorBrewer::brewer.pal(6, "Accent")
col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(k27ac_df, aes(x = Age, y = Count, fill = k27ac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Enhancer Peak Overlap with H3K27ac")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("01_Enhancer_H3k27ac_aldiri_OL.pdf")),
                 height = 2, width = 4)


##########
# ChromHMM
chromHMM_ol_p4 <- findOverlapsOfPeaks(enhancer_list$`P4`, chromHMM$P3, connectedPeaks = "keepAll")
chromHMM_ol_p10 <- findOverlapsOfPeaks(enhancer_list$`P10`, chromHMM$P10, connectedPeaks = "keepAll")
chromHMM_ol_p28 <- findOverlapsOfPeaks(enhancer_list$`P28`, chromHMM$P21, connectedPeaks = "keepAll")

chromHMM_ol_p4_fix <- chromHMM_ol_p4$overlappingPeaks$`enhancer_list.P4///chromHMM.P3` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p10_fix <- chromHMM_ol_p10$overlappingPeaks$`enhancer_list.P10///chromHMM.P10` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p28_fix <- chromHMM_ol_p28$overlappingPeaks$`enhancer_list.P28///chromHMM.P21` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromhmm_df <- rbind(chromHMM_ol_p4_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P4"),
                     chromHMM_ol_p10_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P10"),
                     chromHMM_ol_p28_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P28"))
colnames(chromhmm_df)[1] <- "HMM"

chromhmm_df$HMM <- factor(chromhmm_df$HMM, levels = seq(16,1))
chromhmm_df$Age <- factor(chromhmm_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- RColorBrewer::brewer.pal(6, "Accent")

col_fill <- colorRampPalette(col_pal)(16)

cb <- ggplot(chromhmm_df, aes(x = Age, y = Freq, fill = HMM)) +
  geom_bar(position="fill", stat="identity") +
  viridis::scale_fill_viridis(discrete = T, option = "H", direction = 1, guide = guide_legend(reverse = TRUE)) +
  coord_flip() +
  labs(title = "H3K18la Enhancer Peak Overlap with ChromHMM")

cowplot::ggsave2(cb, filename = file.path(dir_export, paste0("01_Enhancer_H3ChromHMM_OL.pdf")),
                 height = 2, width = 4)




########################
# Genebody peak overlaps

###########
# Nrl peaks 
nrl_ol_p4 <- findOverlapsOfPeaks(genebody_list$`P4`, nrl$P4, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p10 <- findOverlapsOfPeaks(genebody_list$`P10`, nrl$P10, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p28 <- findOverlapsOfPeaks(genebody_list$`P28`, nrl$P28, connectedPeaks = "keepAll", maxgap = 500L)

nrl_df <- cbind(lapply(nrl_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Nrl = rep(c("Nrl", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Nrl != "Nrl")
colnames(nrl_df)[1] <- "Count"

nrl_df$Age <- factor(nrl_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

nb <- ggplot(nrl_df, aes(x = Age, y = Count, fill = Nrl)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Genebody Peak Overlap with Nrl")

cowplot::ggsave2(nb, filename = file.path(dir_export, paste0("01_GeneBody_H3NRL_OL.pdf")),
                 height = 2, width = 4)

############
# ATAC peaks 
atac_ol_p4 <- findOverlapsOfPeaks(genebody_list$`P4`, atac$P3, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p10 <- findOverlapsOfPeaks(genebody_list$`P10`, atac$P10, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p28 <- findOverlapsOfPeaks(genebody_list$`P28`, atac$P21, connectedPeaks = "keepAll", maxgap = 500L)


atac_df <- cbind(lapply(atac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Atac = rep(c("ATAC", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Atac != "ATAC")
colnames(atac_df)[1] <- "Count"

atac_df$Age <- factor(atac_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(atac_df, aes(x = Age, y = Count, fill = Atac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Genebody Peak Overlap with ATAC")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("01_GeneBody_H3ATAC_OL.pdf")),
                 height = 2, width = 4)


############
# k27ac peaks Aldiri
k27ac_ol_p4 <- findOverlapsOfPeaks(genebody_list$`P4`, k27_aldiri$P3, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p10 <- findOverlapsOfPeaks(genebody_list$`P10`, k27_aldiri$P10, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p28 <- findOverlapsOfPeaks(genebody_list$`P28`, k27_aldiri$P21, connectedPeaks = "keepAll", maxgap = 500L)


k27ac_df <- cbind(lapply(k27ac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         k27ac = rep(c("k27ac", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(k27ac != "k27ac")
colnames(k27ac_df)[1] <- "Count"

k27ac_df$Age <- factor(k27ac_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(k27ac_df, aes(x = Age, y = Count, fill = k27ac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Genebody Peak Overlap with H3K27ac")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("01_GeneBody_H3k27ac_aldiri_OL.pdf")),
                 height = 2, width = 4)

##########
# ChromHMM
chromHMM_ol_p4 <- findOverlapsOfPeaks(genebody_list$`P4`, chromHMM$P3, connectedPeaks = "keepAll")
chromHMM_ol_p10 <- findOverlapsOfPeaks(genebody_list$`P10`, chromHMM$P10, connectedPeaks = "keepAll")
chromHMM_ol_p28 <- findOverlapsOfPeaks(genebody_list$`P28`, chromHMM$P21, connectedPeaks = "keepAll")

chromHMM_ol_p4_fix <- chromHMM_ol_p4$overlappingPeaks$`genebody_list.P4///chromHMM.P3` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p10_fix <- chromHMM_ol_p10$overlappingPeaks$`genebody_list.P10///chromHMM.P10` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p28_fix <- chromHMM_ol_p28$overlappingPeaks$`genebody_list.P28///chromHMM.P21` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromhmm_df <- rbind(chromHMM_ol_p4_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P4"),
                     chromHMM_ol_p10_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P10"),
                     chromHMM_ol_p28_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P28"))
colnames(chromhmm_df)[1] <- "HMM"

chromhmm_df$HMM <- factor(chromhmm_df$HMM, levels = seq(16,1))
chromhmm_df$Age <- factor(chromhmm_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- RColorBrewer::brewer.pal(6, "Accent")

col_fill <- colorRampPalette(col_pal)(16)

cb <- ggplot(chromhmm_df, aes(x = Age, y = Freq, fill = HMM)) +
  geom_bar(position="fill", stat="identity") +
  viridis::scale_fill_viridis(discrete = T, option = "H", direction = 1, guide = guide_legend(reverse = TRUE)) +
  coord_flip() +
  labs(title = "H3K18la Genebody Peak Overlap with ChromHMM")

cowplot::ggsave2(cb, filename = file.path(dir_export, paste0("01_GeneBody_H3ChromHMM_OL.pdf")),
                 height = 2, width = 4)


```


### Export the annotated peak file
```{r}

saveRDS(peakList_cons, file = "../data/interim/peakList_cons_240911.RDS")
saveRDS(peakAnnoList_ens, file = "../data/interim/peakAnnoList_cons_240911.RDS")

```

