---
title: "H3K18la Tissue Comparison Analysis"
output: html_notebook

---


### Get custom scripts and functions and outside data used for analysis
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

library(tidyverse)

```


### Get Ensembl annotation database
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

## Query for all available EnsDb databases
ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

ensDb <- ahdb[[1]]



ahdb_hs <- query(ah, pattern = c("EnsDb", "Homo sapiens", 109))

ensDb_hs <- ahdb_hs[[1]]

```


################################################################################
### Import the Galle peaks from nf-core pipeline, process them the same way as retina and then compare
################################################################################

### Import the peaks into a merged list object and plot for number of peaks at each timepoint
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
library(ChIPpeakAnno)

macs_dir <- "../data/external/Galle_nf-core/04_called_peaks/macs2/"
macs_beds <- list.files(path = macs_dir, pattern = "H3K18.+narrowPeak", recursive = T)

# Import the cons peaks
peakList <- sapply(file.path(macs_dir, macs_beds), toGRanges, format="narrowPeak")
names(peakList) <- gsub("(.+).macs.+", "\\1", macs_beds)

peakList_1E6 <- GRangesList()

for (i in 1:length(peakList)){
  
  # seqlevelsStyle(peakList[[i]]) <- "NCBI"
  tmp_peak <- sortSeqlevels(peakList[[i]])
  tmp_peak <- GenomicRanges::sort(tmp_peak)
  tmp_peak <- tmp_peak[(elementMetadata(tmp_peak)[,"qValue"] > 6) & (seqnames(tmp_peak) != "MT")]
  peakList_1E6[[i]] <- tmp_peak
}
names(peakList_1E6) <- gsub("(.+).macs.+", "\\1", macs_beds)


ol_ADIPO <- findOverlapsOfPeaks(c(peakList_1E6[1],
                             peakList_1E6[2]), connectedPeaks="keepAll", ignore.strand = T)

ol_ESC_2i <- findOverlapsOfPeaks(c(peakList_1E6[4],
                             peakList_1E6[5]), connectedPeaks="keepAll", ignore.strand = T)

ol_ESC_ser <- findOverlapsOfPeaks(c(peakList_1E6[6],
                             peakList_1E6[7]), connectedPeaks="keepAll", ignore.strand = T)

ol_GAS <- findOverlapsOfPeaks(c(peakList_1E6[8],
                             peakList_1E6[9]), connectedPeaks="keepAll", ignore.strand = T)

ol_MT <- findOverlapsOfPeaks(c(peakList_1E6[11],
                             peakList_1E6[12]), connectedPeaks="keepAll", ignore.strand = T)

ol_PIM <- findOverlapsOfPeaks(c(peakList_1E6[13],
                             peakList_1E6[14]), connectedPeaks="keepAll", ignore.strand = T)

peakList_cons <- GRangesList(ol_ADIPO$mergedPeaks, peakList_1E6$BMDM_H3K18la_R1, 
                             ol_ESC_2i$mergedPeaks, ol_ESC_ser$mergedPeaks,
                             ol_GAS$mergedPeaks, peakList_1E6$MB_H3K18la_R1, 
                             ol_MT$mergedPeaks, ol_PIM$mergedPeaks)
nm <- gsub("_H3K18la.+", "", names(peakList_1E6)) %>% unique()
nm <- gsub("-", "_", nm)
names(peakList_cons) <- nm

```


## Annotate outside peaks with ensembl db
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


peakAnnoList_galle_nf <- list()

for (i in 1:length(peakList_cons)){
  
  peak_tmp <- peakList_cons[[i]]
  names(peak_tmp) <- paste0("X", 1:length(peak_tmp))
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  
  peakAnnoList_galle_nf[[i]] <- anno_tmp

}

names(peakAnnoList_galle_nf) <- names(peakList_cons)

```


### Make upset and distance plot of promoter genes for the different tissues
```{r}

# Make upset plot based on presence of peak in promoter of gene
genes_list <- list(RETINA = (tc_peakAnno_cons$P28@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        ADIPO = (peakAnnoList_galle_nf$ADIPO@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        BMDM = (peakAnnoList_galle_nf$BMDM@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        GAS = (peakAnnoList_galle_nf$GAS@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        MB = (peakAnnoList_galle_nf$MB@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        MT = (peakAnnoList_galle_nf$MT@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()),
                        PIM = (peakAnnoList_galle_nf$PIM@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()))

u2 <- upset(fromList(genes_list), 
      nintersects = 14, 
      nsets = 7, 
      order.by = "freq", 
      decreasing = T, 
      mb.ratio = c(0.6, 0.4),
      number.angles = 0, 
      text.scale = 1.1, 
      point.size = 2.8, 
      line.size = 1
      )

pdf("../data/processed/P28-Galle_Prom-PC_nf_peak_upset_GeneBased.pdf", 
    width = 9, height = 6, onefile = F)
u2
dev.off()


# Make dendro plots based on presence of peak in promoter of gene
library(rafalib)
library(Hmisc)
library(corrplot)

# Get genes for matrix
genes_galle <- lapply(peakAnnoList_galle_nf, function(x){x@anno %>% 
    as.data.frame() %>% 
    dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
    dplyr::select(SYMBOL)}) %>% 
  unlist() %>% sort() %>% unique()

genes_P28 <- tc_peakAnno_cons$P28@anno %>% 
  as.data.frame() %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
  dplyr::select(SYMBOL) %>% 
  unlist() %>% sort() %>% unique()

genes_all <- c(genes_galle, genes_P28) %>% sort() %>% unique()

# Make matrix
genes_mat <- data.frame(RETINA = ifelse(genes_all %in% (tc_peakAnno_cons$P28@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        ADIPO = ifelse(genes_all %in% (peakAnnoList_galle_nf$ADIPO@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        BMDM = ifelse(genes_all %in% (peakAnnoList_galle_nf$BMDM@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        GAS = ifelse(genes_all %in% (peakAnnoList_galle_nf$GAS@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        MB = ifelse(genes_all %in% (peakAnnoList_galle_nf$MB@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        MT = ifelse(genes_all %in% (peakAnnoList_galle_nf$MT@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0),
                        PIM = ifelse(genes_all %in% (peakAnnoList_galle_nf$PIM@anno %>% 
                                          as.data.frame() %>% 
                                          dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
                                          dplyr::select(SYMBOL) %>% 
                                          unlist() %>% sort() %>% unique()), 1, 0), row.names = genes_all)


# Drop similar genes
genes_nonsim <- genes_mat[rowSums(genes_mat) %in% c(2:6),]

# Dendrogram of correlation
d2 <- dist(c2)
hc2 <- hclust(d2, method="ward.D2")

pdf(file="../data/processed/CorrDendro_sharedGenes.pdf", width=8, height=4, useDingbats = F)
par(mar=c(4,4,1,1))
myplclust(hclust = hc2)
dev.off()

```

