---
title: "H3K18la Analysis - May 28thth, 2024"
output: html_notebook

---

### Project Summary
This project investigates the H3K18la during development. 

**DESIGN**:  

**NOTES**: The data were generated from the ENCODE histone pipeline with duplicate 
removal with no controls. This notebool uses overlap of enhancer, promoter, and 
gene body to characterize QC and correlation with ATAC, Nrl and ChromHMM.

**UPDATE**: The definition of promoter (-1000 to +500) was used for peak annotation. 
Enhancer is defined as -10k to -1000 from TSS.

	• Figure 2
		○ # of genes with peaks
		○ FGE with gene numbers at each definition (GO:leaf)
			§ Promoter: -500 to +250
			§ Enhancer: -10K to -1kb
			§ Genebody
			§ Intergenic
		○ Heatmap
			§  -500 to 500
			§ -10kb to 10kb
		○ RNA-seq correlation (wt retina development)
	• Figure 3
		○ Corelate H3K18la
			§ ATAC
			§ Nrl
			§ RNA-seq (wt retina development
			§ ChromHMM
	• Figure 4
		○ Glycolytic genes show with H3K27ac and H3K18la
	• Figure 5
		○ Explant data CnR
			§ Use spike-in normalization to compare
		○ Correlate with RNA-seq (if data can be produced)


#### Get custom scripts and functions and outside data used for analysis
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

script_dir <- "https://raw.githubusercontent.com/MBrooks313/R_codes/master"
dest_tools_dir <- "../src/tools"
dest_vis_dir <- "../src/visualization"

download.file(url = file.path(script_dir, "GO_leaf.R"),
              destfile = file.path(dest_tools_dir, "GO_leaf.R"),
              method = "curl")

download.file(url = file.path(script_dir, "DFave.R"),
              destfile = file.path(dest_tools_dir, "DFave.R"),
              method = "curl")

download.file(url = file.path(script_dir, "Heatmap.R"),
              destfile = file.path(dest_tools_dir, "Heatmap.R"),
              method = "curl")
# system('cp ~/github/R_codes/DFave.R ../src/tools')
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/et_edgeR_v0.1.0.R ../src/tools")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/FuncGeneEnrich_v0.1.0.R ../src/tools")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/GO_leaf_v1.1.R ../src/tools")

#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/APk_cluster_BF_v0.1.0.R ../src/tools')
#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/HM_PCAsort_BF_v0.1.0.R ../src/visualization')
#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/parPlot_hc_v0.2.0.R ../src/visualization')
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/PCA_simple_v0.1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/VolcanoPlot_v0.1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_v1.7.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Dotplot_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Barplot_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_v1.7.R ../src/visualization")
# 
# system("cp /Volumes/NGS/RNA-seq/ENS_v98/Mouse/Counts/by_project/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv ../data/external")
# system("cp /Volumes/NGS/RNA-seq/ENS_v98/Mouse/Counts/by_project/Mouse_FSPR_WT_Development_Gene_nCPM_ENSv98.tsv ../data/external")

# Function to convert ENS ids to gene symbol 
# ens2sym <- function(x){paste(sort(unique(AnnotationDbi::mapIds(org.Mm.eg.db, keys = x, keytype = "ENSEMBL", column = "SYMBOL"))), sep = "", collapse = ",")}
library(tidyverse)
# options(java.parameters = "-Xmx8000m")
print("Completed...")
```

# Get Ensembl annotation database
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

## Query for all available EnsDb databases
ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

ensDb <- ahdb[[1]]

# supportedFilters(ensDb)

```



################################################################################
## Import the peak & peak anno list
################################################################################
```{r}
library(ChIPpeakAnno)
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


peakList_cons <- readRDS(file = "../data/interim/peakList_cons_250729.RDS")
# peakList_cons <- peakList_cons[c(1:3)]
peakAnnoList_ens <- readRDS(file = "../data/interim/peakAnnoList_cons_250729.RDS")
# peakAnnoList_ens <- peakAnnoList_ens[c(1:3)]

## Annotate peaks with ensembl db
# peakAnnoList_ens <- list()
# 
# for (i in 1:length(peakList_cons)){
# 
#   peak_tmp <- peakList_cons[[i]]
#   seqlevelsStyle(peak_tmp) <- "NCBI"
# 
#   peakAnnoList_ens[[i]] <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
#                          TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T,
#                          overlap = "all")
# 
# }
# 
# names(peakAnnoList_ens) <- names(peakList_cons)

```


################################################################################
## Import outside data
################################################################################
```{r}
source("../src/tools/DFave.R")

#############################################
# Import expression data for flow-sorted rods
fsprwt.dwnld <- read.csv("../data/external/Mouse_FSPR_WT_Development_Gene_nCPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)
# Get group info for averaging
grp <- gsub("(.+)\\.FS.+", "\\1", colnames(fsprwt.dwnld)[10:25])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
fsprwt_cpm <- data.frame(merge(fsprwt.dwnld[,1:9],
                               log2(DFave(fsprwt.dwnld[,10:25], exp_grp) + 1),
                               by = 0),
                         row.names = 1)

fsprwt_z <- data.frame(merge(fsprwt_cpm[,1:9],
                               t(scale(t(fsprwt_cpm[,10:15]), scale = T, center = T)),
                               by = 0),
                         row.names = 1)



#############################################
# Import expression data for developing retina
retina.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)

# Get group info for averaging
grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.dwnld)[10:31])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
retina_cpm <- data.frame(merge(retina.dwnld[,1:9],
                               log2(DFave(retina.dwnld[,10:31], exp_grp) + 1),
                               by = 0),
                         row.names = 1)

retina_z <- data.frame(merge(retina_cpm[,1:9],
                               t(scale(t(retina_cpm[,10:19]), scale = T, center = T)),
                               by = 0),
                         row.names = 1)



#############################################
# Import the ChromHMM from Aldiri data
chromhmm_beds <- list.files(path = "/Volumes/data-1/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/chromhmm", pattern = "*bed", recursive = T, full.names = T)
names(chromhmm_beds) <- gsub(".+\\/(.+)_16.+", "\\1", chromhmm_beds)

# Import the beds 
chromHMM <- sapply(chromhmm_beds, toGRanges, format="BED", skip=1)

# Add the HMM calls to the rows
for (i in 1:length(chromhmm_beds)){
  mcols(chromHMM[[i]])$HMM <- read.table(chromhmm_beds[i], skip = 1)[,4]
  seqlevelsStyle(chromHMM[[i]]) <- "NCBI"
}



#############################################
# Import the ATAC-seq from Aldiri
atac_beds <- list.files(path = "/Volumes/data-1/NGS/ATAC-seq/ENCODE_v1_9_0/Mouse/Aldiri_GSE102092/bed", pattern = "*narrowPeak.gz", recursive = T, full.names = T)
names(atac_beds) <- gsub(".+\\/(.+)_idr.+", "\\1", atac_beds)

# Import the beds 
atac <- sapply(atac_beds, toGRanges, format="narrowPeak")

for (i in 1:length(atac)){
  atac[[i]] <- reduce(atac[[i]])
  seqlevelsStyle(atac[[i]]) <- "NCBI"
}



#############################################
# Import the H3K27ac from Aldiri
k27_aldiri_beds <- list.files(path = "/Volumes/data-1/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bed/", pattern = "*H3K27Ac*", recursive = T, full.names = T)
names(k27_aldiri_beds) <- gsub(".+\\/(.+)_H3K27.+", "\\1", k27_aldiri_beds)

# Import the beds 
k27_aldiri <- sapply(k27_aldiri_beds, toGRanges, format="narrowPeak")

for (i in 1:length(k27_aldiri)){
  k27_aldiri[[i]] <- GenomicRanges::reduce(k27_aldiri[[i]])
  seqlevelsStyle(k27_aldiri[[i]]) <- "NCBI"
}


#############################################
# Import the H3K4me3 from Aldiri
k4me3_aldiri_beds <- list.files(path = "/Volumes/data-1/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Aldiri_GSE87037/bed/", pattern = "*H3K4me3_*", recursive = T, full.names = T)
names(k4me3_aldiri_beds) <- gsub(".+\\/(.+)_H3K4me3.+", "\\1", k4me3_aldiri_beds)

# Import the beds 
k4me3_aldiri <- sapply(k4me3_aldiri_beds, toGRanges, format="narrowPeak")

for (i in 1:length(k4me3_aldiri)){
  k4me3_aldiri[[i]] <- GenomicRanges::reduce(k4me3_aldiri[[i]])
  seqlevelsStyle(k4me3_aldiri[[i]]) <- "NCBI"
}



# #############################################
# # Import the H3K27ac from peaks from Gaur/Liang
# k27_gaur_beds <- list.files(path = "../data/interim/bed/", pattern = "K27ac.+optimal.+narrowPeak", recursive = F, full.names = T)
# names(k27_gaur_beds) <- gsub(".+\\/K27ac-(.+)_overlap.+", "\\1", k27_gaur_beds)
# 
# # Import the beds 
# k27_gaur <- sapply(k27_gaur_beds, toGRanges, format="narrowPeak")
# 
# for (i in 1:length(k27_gaur)){
#   seqlevelsStyle(k27_gaur[[i]]) <- "NCBI"
# }



#############################################
# Import the Nrl peaks from Liang
# nrl_beds <- list.files(path = "/Volumes/data/NGS/ChIP-seq/ENCODE_v1_4_0/Mouse/Liang_Crx-Nrl/bed", pattern = "Nrl.+narrowPeak.gz", recursive = T, full.names = T)
nrl_beds <- list.files(path = "/Volumes/data-1/People/brooksma/PEGASUS/Projects/Xulong/CutNRun/Mouse/Analysis/20200309_16-29/200723/data/interim/beds/idr", pattern = "Nrl.+idr05.+narrowPeak", recursive = F, full.names = T)
names(nrl_beds) <- gsub(".+\\/Nrl_(.+)\\.Rep.+", "\\1", nrl_beds)

# Import the beds 
nrl <- sapply(nrl_beds, toGRanges, format="narrowPeak")

for (i in 1:length(nrl)){
  seqlevelsStyle(nrl[[i]]) <- "NCBI"
}


save(list = c("peakList_cons", "peakAnnoList_ens", "ensDb", "fsprwt_cpm", "fsprwt_z", "retina_cpm", "retina_z", 
"chromHMM", "atac", "nrl"), file = "../data/interim/02_import_objects.RData")

# load(file = "../data/interim/02_import_objects.RData")
```


################################################################################
# Annotate outside data
################################################################################
```{r}

# Aldiri k27ac 
k27AnnoList_ens <- list()

for (i in 1:length(k27_aldiri)){
  
  peak_tmp <- k27_aldiri[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  # peakAnnoList_ens[[i]] <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
  #                        TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T,
  #                        overlap = "all")
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  anno_tmp@anno$GENENAME <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "GENENAME", 
                                         keytype = "GENEID")
  k27AnnoList_ens[[i]] <- anno_tmp

}

names(k27AnnoList_ens) <- names(k27_aldiri)


# Aldiri k4me3 
k4me3AnnoList_ens <- list()

for (i in 1:length(k4me3_aldiri)){
  
  peak_tmp <- k4me3_aldiri[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  # peakAnnoList_ens[[i]] <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
  #                        TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T,
  #                        overlap = "all")
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  anno_tmp@anno$GENENAME <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "GENENAME", 
                                         keytype = "GENEID")
  k4me3AnnoList_ens[[i]] <- anno_tmp

}

names(k4me3AnnoList_ens) <- names(k4me3_aldiri)


```


################################################################################
## Figure 2 QC plots
################################################################################
```{r}
dir_export <- "../data/processed"

#######################
# Consensus peak number
#######################

# Generate plot for peak number
df <- lapply(peakList_cons, length) %>% as.data.frame() %>% gather("Sample", "Peaks")
df$Sample <- gsub("K18La\\.", "", df$Sample)
df$Sample <- gsub("X", "", df$Sample)
df$Sample <- factor(df$Sample, levels = df$Sample)

uniq_genes <- lapply(peakAnnoList_ens, function(x){x@anno$SYMBOL %>% unique() %>% length()}) %>% unlist() %>% as.vector()
prom_genes <- lapply(peakAnnoList_ens, function(x){x@anno %>% 
    as.data.frame() %>% 
    dplyr::filter(annotation == "Promoter") %>% 
    dplyr::select(SYMBOL) %>% 
    as.vector() %>% unlist() %>% 
    unique() %>% length()}) %>% 
  unlist() %>% as.vector()

b <- ggplot(df, aes(Sample, Peaks))+
  geom_bar(stat = "identity", fill = "grey50") +
  geom_point(aes(x = c(1:6), y = uniq_genes), colour = "red", size = 2) +
  annotate("text", x = c(1:6), y = uniq_genes + 200, label = uniq_genes) +
  geom_point(aes(x = c(1:6), y = prom_genes), colour = "blue", size = 2) +
  annotate("text", x = c(1:6), y = prom_genes + 200, label = prom_genes) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 0.8, vjust = 0.8)) +
  labs(title = "Peaks") + xlab(NULL)

cowplot::ggsave2(plot = b, filename = "../data/processed/02_Peak_number_TP.pdf", height = 3, width = 2.5)



# # Generate plot for gene number
# df <- lapply(peakList_cons, length) %>% as.data.frame() %>% gather("Sample", "Peaks")
# df$Sample <- gsub("K18La\\.", "", df$Sample)
# df$Sample <- gsub("X", "", df$Sample)
# df$Sample <- factor(df$Sample, levels = df$Sample)
# 
# b <- ggplot(df, aes(Sample, Peaks))+
#   geom_bar(stat = "identity") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 30, hjust = 0.8, vjust = 0.8)) +
#   labs(title = "Peaks") + xlab(NULL)
# 
# cowplot::ggsave2(plot = b, filename = "../data/processed/02_Peak_number_TP.pdf", height = 3, width = 2)



##########################
# Peak annotation barplots
##########################

flank <- 1000
promoter <- getPromoters(TxDb = ensDb, upstream = flank, downstream = flank)
seqlevelsStyle(promoter) <- "NCBI"

tagMatrixList <- list()
for (i in names(peakList_cons)){
  tagMatrixList[[i]] <- getTagMatrix(peakList_cons[[i]], windows=promoter)
}


# Plot peak histogram around TSS
p <- plotAvgProf(tagMatrixList, xlim=c(-flank, flank), resmple = 100,
            xlab="Genomic Region (5'->3')", ylab = "Read Count Frequency")
print(p)

pdf("../data/processed/02_Peak_TSS_hist_TP.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


# Plot annotation bargraph
p <- plotAnnoBar(peakAnnoList_ens, title = "H3K18la peak annotation")
print(p)

pdf("../data/processed/02_Peak_annot_bar_TP.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


#########
# Heatmap
#########

##############
# Heatmap 5000
flank <- c(500, 1000, 3000, 10000)

for (i in flank[c(2,3)]){
  
  promoter <- getPromoters(TxDb = ensDb, upstream = i, downstream = i)
  seqlevelsStyle(promoter) <- "NCBI"
  
  tagMatrixList2 <- list()
  for (ii in names(peakList_cons)){
    tagMatrixList2[[ii]] <- getTagMatrix(peakList_cons[[ii]], windows=promoter)
  }
  
  # Heatmap plot
  pdf(file = file.path(dir_export, paste0("02_Peak_TSSdist_", i,"_heatmap_TP.pdf")), height = 4, width = 5)
  tagHeatmap(tagMatrixList2, xlim=c(-i, i), color = "grey25")
  dev.off()

}


```


################################################################################
## Figure 2 FGE
################################################################################
```{r}

library(clusterProfiler)
library(enrichplot)
library(org.Mm.eg.db)
source(file.path(dest_tools_dir, "GO_leaf.R"))
# source("../src/tools/Dotplot_GO.R")

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 5

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_enrich <- file.path(dir_export, paste0("02_FuncGeneEnrich_", enrich_db, "_", GOont))
dir.create(dir_enrich)



# Get subsets of annotation to run FGE based on last mtg's discussion
prom_list <- lapply(peakAnnoList_ens, function(x){x@anno[x@anno$annotation == "Promoter" & 
                                                           x@anno$transcriptBiotype == "protein_coding"]})

genebody_list <- lapply(peakAnnoList_ens, function(x){x@anno[!(x@anno$annotation %in% c("Promoter", "Downstream", "Distal Intergenic")) & 
                                                           x@anno$transcriptBiotype == "protein_coding"]})

enhancer_list <- lapply(peakAnnoList_ens, function(x){x@anno[x@anno$distanceToTSS > -10000 & 
                                                               x@anno$distanceToTSS < -1000 & 
                                                           x@anno$transcriptBiotype == "protein_coding"]})

# Get list of genes associated with the peaks for each subgroup
prom_gene_list <- lapply(prom_list, function(x){x@elementMetadata$SYMBOL %>% unique() %>% unlist() %>% as.vector()}) 

genebody_gene_list <- lapply(genebody_list, function(x){x@elementMetadata$SYMBOL %>% unique() %>% unlist() %>% as.vector()}) 

enhancer_gene_list <- lapply(enhancer_list, function(x){x@elementMetadata$SYMBOL %>% unique() %>% unlist() %>% as.vector()}) 


# Prep the gene lists for the FGE loop
goi_list <- list(prom_gene_list, genebody_gene_list, enhancer_gene_list)
names(goi_list) <- c("Promoter", "Genebody", "Enhancer")

# Perform FGE on the GOIs from each subset of peaks
for (i in names(goi_list)){
  
  enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list[[i]], fun = "enrichGO",
                                            keyType = "SYMBOL", ont = GOont, OrgDb = org)
  
  
  # Filter enrichment for significance
  enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)
  
  # Fix the Ratio
  enrich_results_filt_fix <- enrich_results_filt %>% 
    as.data.frame() %>% 
    mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
    rowwise() %>% 
    mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))
  
  # Export the FGE results table
  tmp <- enrich_results_filt_fix[,-11]
  # Export the enrichment analysis
  write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_full.tsv")),
                quote = F, row.names = F, sep = "\t")

  # Get leaf terms
  enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)
  
  # Fix the Ratio
  enrich_results_leaf_fix <- enrich_results_leaf %>% 
    as.data.frame() %>% 
    mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
    rowwise() %>% 
    mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))
  
  # Export the FGE results table
  tmp <- enrich_results_leaf_fix[,-11]
  # Export the enrichment analysis
  write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_leaf.tsv")),
                quote = F, row.names = F, sep = "\t")
  
  # Get top 5 pathways per cell type to get union set of pathways
  top5_union <- enrich_results_leaf %>% 
    group_by(Cluster) %>% 
    arrange(p.adjust) %>% 
    slice_head(n = top)
  
  
  # Filter the full table for the union of top5 pathways
  top5_enrich <- enrich_results_filt_fix %>% 
    dplyr::filter(ID %in% (top5_union$ID %>% unique()))
  
  # Prep the df for plotting
  top5_enrich <- top5_enrich %>% 
    arrange(Cluster, p.adjust)
  top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
  top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
  top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)
  
  # Make plot
  pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
      geom_point(aes(size = Ratio)) + 
      theme_bw() +
      scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
      guides(size=guide_legend(title="Ratio")) +
      theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1)) +
      ggtitle(paste0("H3K18la Peaks in ", i))
  
  #Open the pdf device
  cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("Timepoint_FGE_top_", top, "_", i, "_leaf_01_dotplot.pdf")),
                   height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                   width = 6)
    
}


# Make a barplot of the number of enhancer, promoter, and genebody peaks
region_df <- data.frame(Count = c(lapply(enhancer_list, length),
                          lapply(prom_list, length),
                          lapply(genebody_list, length)) %>% unlist() %>% as.vector(),
                         Region = rep(c("Enhancer", "Promoter", "Genebody"), each = 6),
                         Group = rep(c(names(enhancer_list)), 3))


region_df$Group <- factor(region_df$Group, levels = names(enhancer_list))
region_df$Region <- factor(region_df$Region, levels = rev(c("Genebody", "Enhancer", "Promoter")))

rb <- ggplot(region_df, aes(x = Group, y = Count, fill = Region)) +
  geom_bar(position="fill", stat="identity") + 
  # scale_fill_manual(values = rev(col_fill)) +
  labs(title = "H3K18la Peak Annotation Region with Age")

cowplot::ggsave2(rb, filename = file.path(dir_export, paste0("02_Peak_H3K18la-Region_OL.pdf")),
                 height = 3, width = 3)

```


################################################################################
## Figure 2 RNA heatmap of chipseq peaks
################################################################################
```{r}
source(file.path(dest_tools_dir, "Heatmap.R"))
library(scales)
library(RColorBrewer)

mb.colors <- colorRampPalette(c("#023858","#3690c0", "#e0f3f8","#fff7bc","#fec44f","#ec7014",
                                    "#fc4e2a","#bd0026","#800026"))(n=1000)
jet <- c("#7F0000","red","#FF7F00","yellow","white","cyan", "#007FFF", "blue","#00007F")
mb.colors <- colorRampPalette(rev(jet))(1000)
 
lim <- c(-3,3)


# Pheamap using row annotation

region_peaks <- list(enhancer_gene_list, prom_gene_list, genebody_gene_list)
names(region_peaks) <- c("Enhancer", "Promoter", "Genebody")

for (i in 1:length(region_peaks)){
  tmp_gene_list <- region_peaks[[i]]
  region_nm <- names(region_peaks)[i]
  
  # Gene union vector of genes
  tmp_union <- tmp_gene_list %>% unlist(use.names = F) %>% as.vector() %>% unique() 
  tmp_union <- tmp_union[!is.na(tmp_union)]
  
  # Make peak boolean DF
  tmp_peak_df <- data.frame(P28_ChIP = tmp_union %in% tmp_gene_list$`P28`,
                               P10_ChIP = tmp_union %in% tmp_gene_list$`P10`,
                               P4_ChIP = tmp_union %in% tmp_gene_list$`P4`)
  row.names(tmp_peak_df) <- tmp_union
  tmp_peak_df <- tmp_peak_df * 1
  
  # Subsett the expression DF with the GOIs
  tmp_chip_rna <- retina_z %>% 
    rownames_to_column() %>% 
    dplyr::filter(external_gene_name %in% tmp_union) %>% 
    dplyr::select(c(external_gene_name, P4, P10, P28)) %>% 
    dplyr::filter(!duplicated(external_gene_name), !is.na(P4)) %>% 
    column_to_rownames(var = "external_gene_name")
    
  # Oreder the peak DF with same order as RNA DF
  tmp_peak_df <- tmp_peak_df[row.names(tmp_chip_rna),]
  
  # CLuster by chip_peak
  distfunc <- function(x) dist(x,method="euclidean")
  hclustfunc <- function(x) hclust(x, method="ward.D2")
  d <- distfunc(as.matrix(tmp_peak_df))
  hc <- hclustfunc(d)
  
  # re-order the data.frames
  tmp_chip_rna <- tmp_chip_rna[hc$order,]
  tmp_peak_df <- tmp_peak_df[hc$order,]
  
  
  # Plot heatmaps
  h <- pheatmap::pheatmap(tmp_chip_rna, scale = "none", 
                          cluster_rows = T, cluster_cols = F, show_rownames = F, 
                          clustering_method = "ward.D2", 
                          color = colorRampPalette(rev(brewer.pal(n = 9, "RdBu")))(50),
                          main = paste0(region_nm, " Genes"),
                          annotation_row = tmp_peak_df)
  
  hh <- pheatmap::pheatmap(tmp_chip_rna, scale = "none", 
                          cluster_rows = F, cluster_cols = F, show_rownames = F, 
                          clustering_method = "ward.D2", 
                          color = colorRampPalette(rev(brewer.pal(n = 9, "RdBu")))(50),
                          main = paste0(region_nm, " Genes"),
                          annotation_row = tmp_peak_df)

  cowplot::ggsave2(filename = file.path(dir_export, paste0("02_ChIP_retRNA_HM_" , region_nm, "_v1.pdf")),
                   h, height = 6, width = 4)
  
  cowplot::ggsave2(filename = file.path(dir_export, paste0("02_ChIP_retRNA_HM_" , region_nm, "_v2.pdf")),
                   hh, height = 6, width = 4)
  
}







```


################################################################################
## Figure 3 Correlate peaks with ATAC, ChromHMM, NRL peaks
################################################################################
```{r}

########################
# Promoter peak overlaps

###########
# Nrl peaks 
nrl_ol_p4 <- findOverlapsOfPeaks(prom_list$`P4`, nrl$P4, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p10 <- findOverlapsOfPeaks(prom_list$`P10`, nrl$P10, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p28 <- findOverlapsOfPeaks(prom_list$`P28`, nrl$P28, connectedPeaks = "keepAll", maxgap = 500L)

nrl_df <- cbind(lapply(nrl_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Nrl = rep(c("Nrl", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Nrl != "Nrl")
colnames(nrl_df)[1] <- "Count"

nrl_df$Age <- factor(nrl_df$Age, levels = rev(c("P4", "P10", "P28")))

# col_pal <- RColorBrewer::brewer.pal(6, "Accent")
col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

nb <- ggplot(nrl_df, aes(x = Age, y = Count, fill = Nrl)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Promoter Peak Overlap with Nrl")

cowplot::ggsave2(nb, filename = file.path(dir_export, paste0("02_Prom_H3K18la-NRL_OL.pdf")),
                 height = 2, width = 4)

############
# ATAC peaks 
atac_ol_p4 <- findOverlapsOfPeaks(prom_list$`P4`, atac$P3, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p10 <- findOverlapsOfPeaks(prom_list$`P10`, atac$P10, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p28 <- findOverlapsOfPeaks(prom_list$`P28`, atac$P21, connectedPeaks = "keepAll", maxgap = 500L)


atac_df <- cbind(lapply(atac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Atac = rep(c("ATAC", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Atac != "ATAC")
colnames(atac_df)[1] <- "Count"

atac_df$Age <- factor(atac_df$Age, levels = rev(c("P4", "P10", "P28")))

# col_pal <- RColorBrewer::brewer.pal(6, "Accent")
col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(atac_df, aes(x = Age, y = Count, fill = Atac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Promoter Peak Overlap with ATAC")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("02_Prom_H3K18la-ATAC_OL.pdf")),
                 height = 2, width = 4)


############
# k27ac peaks Aldiri
k27ac_ol_p4 <- findOverlapsOfPeaks(prom_list$`P4`, k27_aldiri$P3, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p10 <- findOverlapsOfPeaks(prom_list$`P10`, k27_aldiri$P10, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p28 <- findOverlapsOfPeaks(prom_list$`P28`, k27_aldiri$P21, connectedPeaks = "keepAll", maxgap = 500L)


k27ac_df <- cbind(lapply(k27ac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         k27ac = rep(c("k27ac", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(k27ac != "k27ac")
colnames(k27ac_df)[1] <- "Count"

k27ac_df$Age <- factor(k27ac_df$Age, levels = rev(c("P4", "P10", "P28")))

# col_pal <- RColorBrewer::brewer.pal(6, "Accent")
col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(k27ac_df, aes(x = Age, y = Count, fill = k27ac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Promoter Peak Overlap with H3K27ac")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("02_Prom_H3K18la-k27ac_aldiri_OL.pdf")),
                 height = 2, width = 4)


# ############
# # k27ac peaks Gaur
# k27ac_ol_p4 <- findOverlapsOfPeaks(prom_list$`P4`, k27_gaur$P4, connectedPeaks = "keepAll", maxgap = 500L)
# k27ac_ol_p10 <- findOverlapsOfPeaks(prom_list$`P10`, k27_gaur$P10, connectedPeaks = "keepAll", maxgap = 500L)
# k27ac_ol_p28 <- findOverlapsOfPeaks(prom_list$`P28`, k27_gaur$P28, connectedPeaks = "keepAll", maxgap = 500L)
# 
# 
# k27ac_df <- cbind(lapply(k27ac_ol_p4$peaklist, length) %>% as.data.frame(),
#                  lapply(k27ac_ol_p10$peaklist, length) %>% as.data.frame(),
#                  lapply(k27ac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
#   as.data.frame() %>% 
#   mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
#          k27ac = rep(c("k27ac", "non-OL", "OL"), 3)) %>% 
#   dplyr::filter(k27ac != "k27ac")
# colnames(k27ac_df)[1] <- "Count"
# 
# k27ac_df$Age <- factor(k27ac_df$Age, levels = rev(c("P4", "P10", "P28")))
# 
# # col_pal <- RColorBrewer::brewer.pal(6, "Accent")
# col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
# col_fill <- colorRampPalette(col_pal)(2)
# 
# ab <- ggplot(k27ac_df, aes(x = Age, y = Count, fill = k27ac)) +
#   geom_bar(position="fill", stat="identity") + 
#   scale_fill_manual(values = rev(col_fill)) +
#   coord_flip() +
#   labs(title = "H3K18la Promoter Peak Overlap with H3K27ac")
# 
# cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("02_Prom_H3K18la-k27ac_gaur_OL.pdf")),
#                  height = 2, width = 4)



##########
# ChromHMM
chromHMM_ol_p4 <- findOverlapsOfPeaks(prom_list$`P4`, chromHMM$P3, connectedPeaks = "keepAll")
chromHMM_ol_p10 <- findOverlapsOfPeaks(prom_list$`P10`, chromHMM$P10, connectedPeaks = "keepAll")
chromHMM_ol_p28 <- findOverlapsOfPeaks(prom_list$`P28`, chromHMM$P21, connectedPeaks = "keepAll")

chromHMM_ol_p4_fix <- chromHMM_ol_p4$overlappingPeaks$`prom_list.P4///chromHMM.P3` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p10_fix <- chromHMM_ol_p10$overlappingPeaks$`prom_list.P10///chromHMM.P10` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p28_fix <- chromHMM_ol_p28$overlappingPeaks$`prom_list.P28///chromHMM.P21` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromhmm_df <- rbind(chromHMM_ol_p4_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P4"),
                     chromHMM_ol_p10_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P10"),
                     chromHMM_ol_p28_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P28"))
colnames(chromhmm_df)[1] <- "HMM"

chromhmm_df$HMM <- factor(chromhmm_df$HMM, levels = seq(16,1))
chromhmm_df$Age <- factor(chromhmm_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- RColorBrewer::brewer.pal(6, "Accent")
# col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(16)

cb <- ggplot(chromhmm_df, aes(x = Age, y = Freq, fill = HMM)) +
  geom_bar(position="fill", stat="identity") +
  viridis::scale_fill_viridis(discrete = T, option = "H", direction = 1, guide = guide_legend(reverse = TRUE)) +
  coord_flip() +
  labs(title = "H3K18la Promoter Peak Overlap with ChromHMM")

cowplot::ggsave2(cb, filename = file.path(dir_export, paste0("02_Prom_H3K18la-ChromHMM_OL.pdf")),
                 height = 2, width = 4)




########################
# Enhancer peak overlaps

###########
# Nrl peaks 
nrl_ol_p4 <- findOverlapsOfPeaks(enhancer_list$`P4`, nrl$P4, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p10 <- findOverlapsOfPeaks(enhancer_list$`P10`, nrl$P10, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p28 <- findOverlapsOfPeaks(enhancer_list$`P28`, nrl$P28, connectedPeaks = "keepAll", maxgap = 500L)

nrl_df <- cbind(lapply(nrl_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Nrl = rep(c("Nrl", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Nrl != "Nrl")
colnames(nrl_df)[1] <- "Count"

nrl_df$Age <- factor(nrl_df$Age, levels = rev(c("P4", "P10", "P28")))

# col_pal <- RColorBrewer::brewer.pal(6, "Accent")
col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

nb <- ggplot(nrl_df, aes(x = Age, y = Count, fill = Nrl)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Enhancer Peak Overlap with Nrl")

cowplot::ggsave2(nb, filename = file.path(dir_export, paste0("02_Enhancer_H3NRL_OL.pdf")),
                 height = 2, width = 4)

############
# ATAC peaks 
atac_ol_p4 <- findOverlapsOfPeaks(enhancer_list$`P4`, atac$P3, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p10 <- findOverlapsOfPeaks(enhancer_list$`P10`, atac$P10, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p28 <- findOverlapsOfPeaks(enhancer_list$`P28`, atac$P21, connectedPeaks = "keepAll", maxgap = 500L)


atac_df <- cbind(lapply(atac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Atac = rep(c("ATAC", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Atac != "ATAC")
colnames(atac_df)[1] <- "Count"

atac_df$Age <- factor(atac_df$Age, levels = rev(c("P4", "P10", "P28")))

# col_pal <- RColorBrewer::brewer.pal(6, "Accent")
col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(atac_df, aes(x = Age, y = Count, fill = Atac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Enhancer Peak Overlap with ATAC")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("02_Enhancer_H3ATAC_OL.pdf")),
                 height = 2, width = 4)


############
# k27ac peaks Aldiri
k27ac_ol_p4 <- findOverlapsOfPeaks(enhancer_list$`P4`, k27_aldiri$P3, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p10 <- findOverlapsOfPeaks(enhancer_list$`P10`, k27_aldiri$P10, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p28 <- findOverlapsOfPeaks(enhancer_list$`P28`, k27_aldiri$P21, connectedPeaks = "keepAll", maxgap = 500L)


k27ac_df <- cbind(lapply(k27ac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         k27ac = rep(c("k27ac", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(k27ac != "k27ac")
colnames(k27ac_df)[1] <- "Count"

k27ac_df$Age <- factor(k27ac_df$Age, levels = rev(c("P4", "P10", "P28")))

# col_pal <- RColorBrewer::brewer.pal(6, "Accent")
col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(k27ac_df, aes(x = Age, y = Count, fill = k27ac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Enhancer Peak Overlap with H3K27ac")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("02_Enhancer_H3k27ac_aldiri_OL.pdf")),
                 height = 2, width = 4)

# 
# ############
# # k27ac peaks Gaur
# k27ac_ol_p4 <- findOverlapsOfPeaks(enhancer_list$`P4`, k27_gaur$P4, connectedPeaks = "keepAll", maxgap = 500L)
# k27ac_ol_p10 <- findOverlapsOfPeaks(enhancer_list$`P10`, k27_gaur$P10, connectedPeaks = "keepAll", maxgap = 500L)
# k27ac_ol_p28 <- findOverlapsOfPeaks(enhancer_list$`P28`, k27_gaur$P28, connectedPeaks = "keepAll", maxgap = 500L)
# 
# 
# k27ac_df <- cbind(lapply(k27ac_ol_p4$peaklist, length) %>% as.data.frame(),
#                  lapply(k27ac_ol_p10$peaklist, length) %>% as.data.frame(),
#                  lapply(k27ac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
#   as.data.frame() %>% 
#   mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
#          k27ac = rep(c("k27ac", "non-OL", "OL"), 3)) %>% 
#   dplyr::filter(k27ac != "k27ac")
# colnames(k27ac_df)[1] <- "Count"
# 
# k27ac_df$Age <- factor(k27ac_df$Age, levels = rev(c("P4", "P10", "P28")))
# 
# # col_pal <- RColorBrewer::brewer.pal(6, "Accent")
# col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
# col_fill <- colorRampPalette(col_pal)(2)
# 
# ab <- ggplot(k27ac_df, aes(x = Age, y = Count, fill = k27ac)) +
#   geom_bar(position="fill", stat="identity") + 
#   scale_fill_manual(values = rev(col_fill)) +
#   coord_flip() +
#   labs(title = "H3K18la Enhancer Peak Overlap with H3K27ac")
# 
# cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("02_Enhancer_H3k27ac_gaur_OL.pdf")),
#                  height = 2, width = 4)


##########
# ChromHMM
chromHMM_ol_p4 <- findOverlapsOfPeaks(enhancer_list$`P4`, chromHMM$P3, connectedPeaks = "keepAll")
chromHMM_ol_p10 <- findOverlapsOfPeaks(enhancer_list$`P10`, chromHMM$P10, connectedPeaks = "keepAll")
chromHMM_ol_p28 <- findOverlapsOfPeaks(enhancer_list$`P28`, chromHMM$P21, connectedPeaks = "keepAll")

chromHMM_ol_p4_fix <- chromHMM_ol_p4$overlappingPeaks$`enhancer_list.P4///chromHMM.P3` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p10_fix <- chromHMM_ol_p10$overlappingPeaks$`enhancer_list.P10///chromHMM.P10` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p28_fix <- chromHMM_ol_p28$overlappingPeaks$`enhancer_list.P28///chromHMM.P21` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromhmm_df <- rbind(chromHMM_ol_p4_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P4"),
                     chromHMM_ol_p10_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P10"),
                     chromHMM_ol_p28_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P28"))
colnames(chromhmm_df)[1] <- "HMM"

chromhmm_df$HMM <- factor(chromhmm_df$HMM, levels = seq(16,1))
chromhmm_df$Age <- factor(chromhmm_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- RColorBrewer::brewer.pal(6, "Accent")
# col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(16)

cb <- ggplot(chromhmm_df, aes(x = Age, y = Freq, fill = HMM)) +
  geom_bar(position="fill", stat="identity") +
  viridis::scale_fill_viridis(discrete = T, option = "H", direction = 1, guide = guide_legend(reverse = TRUE)) +
  coord_flip() +
  labs(title = "H3K18la Enhancer Peak Overlap with ChromHMM")

cowplot::ggsave2(cb, filename = file.path(dir_export, paste0("02_Enhancer_H3ChromHMM_OL.pdf")),
                 height = 2, width = 4)




########################
# Genebody peak overlaps

###########
# Nrl peaks 
nrl_ol_p4 <- findOverlapsOfPeaks(genebody_list$`P4`, nrl$P4, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p10 <- findOverlapsOfPeaks(genebody_list$`P10`, nrl$P10, connectedPeaks = "keepAll", maxgap = 500L)
nrl_ol_p28 <- findOverlapsOfPeaks(genebody_list$`P28`, nrl$P28, connectedPeaks = "keepAll", maxgap = 500L)

nrl_df <- cbind(lapply(nrl_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(nrl_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Nrl = rep(c("Nrl", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Nrl != "Nrl")
colnames(nrl_df)[1] <- "Count"

nrl_df$Age <- factor(nrl_df$Age, levels = rev(c("P4", "P10", "P28")))

# col_pal <- RColorBrewer::brewer.pal(6, "Accent")
col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

nb <- ggplot(nrl_df, aes(x = Age, y = Count, fill = Nrl)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Genebody Peak Overlap with Nrl")

cowplot::ggsave2(nb, filename = file.path(dir_export, paste0("02_GeneBody_H3NRL_OL.pdf")),
                 height = 2, width = 4)

############
# ATAC peaks 
atac_ol_p4 <- findOverlapsOfPeaks(genebody_list$`P4`, atac$P3, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p10 <- findOverlapsOfPeaks(genebody_list$`P10`, atac$P10, connectedPeaks = "keepAll", maxgap = 500L)
atac_ol_p28 <- findOverlapsOfPeaks(genebody_list$`P28`, atac$P21, connectedPeaks = "keepAll", maxgap = 500L)


atac_df <- cbind(lapply(atac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(atac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         Atac = rep(c("ATAC", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(Atac != "ATAC")
colnames(atac_df)[1] <- "Count"

atac_df$Age <- factor(atac_df$Age, levels = rev(c("P4", "P10", "P28")))

# col_pal <- RColorBrewer::brewer.pal(6, "Accent")
col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(atac_df, aes(x = Age, y = Count, fill = Atac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Genebody Peak Overlap with ATAC")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("02_GeneBody_H3ATAC_OL.pdf")),
                 height = 2, width = 4)


############
# k27ac peaks Aldiri
k27ac_ol_p4 <- findOverlapsOfPeaks(genebody_list$`P4`, k27_aldiri$P3, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p10 <- findOverlapsOfPeaks(genebody_list$`P10`, k27_aldiri$P10, connectedPeaks = "keepAll", maxgap = 500L)
k27ac_ol_p28 <- findOverlapsOfPeaks(genebody_list$`P28`, k27_aldiri$P21, connectedPeaks = "keepAll", maxgap = 500L)


k27ac_df <- cbind(lapply(k27ac_ol_p4$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p10$peaklist, length) %>% as.data.frame(),
                 lapply(k27ac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
  as.data.frame() %>% 
  mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
         k27ac = rep(c("k27ac", "non-OL", "OL"), 3)) %>% 
  dplyr::filter(k27ac != "k27ac")
colnames(k27ac_df)[1] <- "Count"

k27ac_df$Age <- factor(k27ac_df$Age, levels = rev(c("P4", "P10", "P28")))

# col_pal <- RColorBrewer::brewer.pal(6, "Accent")
col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(2)

ab <- ggplot(k27ac_df, aes(x = Age, y = Count, fill = k27ac)) +
  geom_bar(position="fill", stat="identity") + 
  scale_fill_manual(values = rev(col_fill)) +
  coord_flip() +
  labs(title = "H3K18la Genebody Peak Overlap with H3K27ac")

cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("02_GeneBody_H3k27ac_aldiri_OL.pdf")),
                 height = 2, width = 4)


# ############
# # k27ac peaks Gaur
# k27ac_ol_p4 <- findOverlapsOfPeaks(genebody_list$`P4`, k27_gaur$P4, connectedPeaks = "keepAll", maxgap = 500L)
# k27ac_ol_p10 <- findOverlapsOfPeaks(genebody_list$`P10`, k27_gaur$P10, connectedPeaks = "keepAll", maxgap = 500L)
# k27ac_ol_p28 <- findOverlapsOfPeaks(genebody_list$`P28`, k27_gaur$P28, connectedPeaks = "keepAll", maxgap = 500L)
# 
# 
# k27ac_df <- cbind(lapply(k27ac_ol_p4$peaklist, length) %>% as.data.frame(),
#                  lapply(k27ac_ol_p10$peaklist, length) %>% as.data.frame(),
#                  lapply(k27ac_ol_p28$peaklist, length) %>% as.data.frame()) %>% t() %>% 
#   as.data.frame() %>% 
#   mutate(Age = rep(c("P4", "P10", "P28"), each = 3),
#          k27ac = rep(c("k27ac", "non-OL", "OL"), 3)) %>% 
#   dplyr::filter(k27ac != "k27ac")
# colnames(k27ac_df)[1] <- "Count"
# 
# k27ac_df$Age <- factor(k27ac_df$Age, levels = rev(c("P4", "P10", "P28")))
# 
# # col_pal <- RColorBrewer::brewer.pal(6, "Accent")
# col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
# col_fill <- colorRampPalette(col_pal)(2)
# 
# ab <- ggplot(k27ac_df, aes(x = Age, y = Count, fill = k27ac)) +
#   geom_bar(position="fill", stat="identity") + 
#   scale_fill_manual(values = rev(col_fill)) +
#   coord_flip() +
#   labs(title = "H3K18la Genebody Peak Overlap with H3K27ac")
# 
# cowplot::ggsave2(ab, filename = file.path(dir_export, paste0("02_GeneBody_H3k27ac_gaur_OL.pdf")),
#                  height = 2, width = 4)


##########
# ChromHMM
chromHMM_ol_p4 <- findOverlapsOfPeaks(genebody_list$`P4`, chromHMM$P3, connectedPeaks = "keepAll")
chromHMM_ol_p10 <- findOverlapsOfPeaks(genebody_list$`P10`, chromHMM$P10, connectedPeaks = "keepAll")
chromHMM_ol_p28 <- findOverlapsOfPeaks(genebody_list$`P28`, chromHMM$P21, connectedPeaks = "keepAll")

chromHMM_ol_p4_fix <- chromHMM_ol_p4$overlappingPeaks$`genebody_list.P4///chromHMM.P3` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p10_fix <- chromHMM_ol_p10$overlappingPeaks$`genebody_list.P10///chromHMM.P10` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromHMM_ol_p28_fix <- chromHMM_ol_p28$overlappingPeaks$`genebody_list.P28///chromHMM.P21` %>% 
  rownames_to_column(var = "peakMerge") %>% 
  dplyr::select(peakMerge, peaks1, peaks2, HMM, shortestDistance) %>% 
  group_by(peaks1) %>% 
  slice_max(shortestDistance)

chromhmm_df <- rbind(chromHMM_ol_p4_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P4"),
                     chromHMM_ol_p10_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P10"),
                     chromHMM_ol_p28_fix$HMM %>% table() %>% as.data.frame() %>% mutate(Age = "P28"))
colnames(chromhmm_df)[1] <- "HMM"

chromhmm_df$HMM <- factor(chromhmm_df$HMM, levels = seq(16,1))
chromhmm_df$Age <- factor(chromhmm_df$Age, levels = rev(c("P4", "P10", "P28")))

col_pal <- RColorBrewer::brewer.pal(6, "Accent")
# col_pal <- wesanderson::wes_palette("Zissou1", 5, "discrete")
col_fill <- colorRampPalette(col_pal)(16)

cb <- ggplot(chromhmm_df, aes(x = Age, y = Freq, fill = HMM)) +
  geom_bar(position="fill", stat="identity") +
  viridis::scale_fill_viridis(discrete = T, option = "H", direction = 1, guide = guide_legend(reverse = TRUE)) +
  coord_flip() +
  labs(title = "H3K18la Genebody Peak Overlap with ChromHMM")

cowplot::ggsave2(cb, filename = file.path(dir_export, paste0("02_GeneBody_H3ChromHMM_OL.pdf")),
                 height = 2, width = 4)


```


################################################################################
# Figure 4 - Venn/Euler/Boxplot of promoter genes from H3K18la, H3K27ac, H3K4me3
```{r}
library(ggVennDiagram)

###############
# Venn diagram

# P4
gene_list_P4 <- list(H3K27ac = k27AnnoList_ens$P3@anno %>% 
                       as.data.frame() %>% 
                       dplyr::filter(annotation == "Promoter",
                                     transcriptBiotype == "protein_coding") %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F) %>% unique(),
                  H3K4me3 = k4me3AnnoList_ens$P3@anno %>% 
                       as.data.frame() %>% 
                       dplyr::filter(annotation == "Promoter",
                                     transcriptBiotype == "protein_coding") %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F) %>% unique(),
                  H3K18la = peakAnnoList_ens$P4 %>% 
                       as.data.frame() %>% 
                       dplyr::filter(annotation == "Promoter",
                                     transcriptBiotype == "protein_coding") %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F) %>% unique())

venn <- Venn(gene_list_P4)
data <- process_data(venn)

gv <- ggplot() +
  # change mapping of color filling
  geom_sf(aes(fill = log10(count)), data = venn_region(data), show.legend = TRUE)+
  # adjust edge size and color
  geom_sf(aes(color=id), size = 3, data = venn_setedge(data), show.legend = FALSE) + 
  # geom_sf_label(aes(label=count), size = 3, data = venn_region(data)) +
  geom_sf_label(aes(label=(count / sum(count) * 100) %>% round(digits = 0)), size = 3, data = venn_region(data)) +
  # show set label in bold
  geom_sf_text(aes(label = name), size = 5, fontface = "bold", data = venn_setlabel(data)) +  
  # add a alternative region name
  # geom_sf_label(aes(label = name), data = venn_region(data), alpha = 0.5) +  
  theme_void()

gv <- gv + scale_fill_distiller(palette = "Purples", direction = 1)

cowplot::ggsave2(gv, filename = file.path(dir_export, paste0("02_Venn_P4_Prom_18la-27ac-4me3.pdf")),
                 height = 4, width = 4)


# P10
gene_list_P10 <- list(H3K27ac = k27AnnoList_ens$P10@anno %>% 
                       as.data.frame() %>% 
                       dplyr::filter(annotation == "Promoter",
                                     transcriptBiotype == "protein_coding") %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F) %>% unique(),
                  H3K4me3 = k4me3AnnoList_ens$P10@anno %>% 
                       as.data.frame() %>% 
                       dplyr::filter(annotation == "Promoter",
                                     transcriptBiotype == "protein_coding") %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F) %>% unique(),
                  H3K18la = peakAnnoList_ens$P10 %>% 
                       as.data.frame() %>% 
                       dplyr::filter(annotation == "Promoter",
                                     transcriptBiotype == "protein_coding") %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F) %>% unique())

venn <- Venn(gene_list_P10)
data <- process_data(venn)

gv <- ggplot() +
  # change mapping of color filling
  geom_sf(aes(fill = log10(count)), data = venn_region(data), show.legend = TRUE)+
  # adjust edge size and color
  geom_sf(aes(color=id), size = 3, data = venn_setedge(data), show.legend = FALSE) + 
  # geom_sf_label(aes(label=count), size = 3, data = venn_region(data)) +
  geom_sf_label(aes(label=(count / sum(count) * 100) %>% round(digits = 0)), size = 3, data = venn_region(data)) +
  # show set label in bold
  geom_sf_text(aes(label = name), size = 5, fontface = "bold", data = venn_setlabel(data)) +  
  # add a alternative region name
  # geom_sf_label(aes(label = name), data = venn_region(data), alpha = 0.5) +  
  theme_void()

gv <- gv + scale_fill_distiller(palette = "Purples", direction = 1)

cowplot::ggsave2(gv, filename = file.path(dir_export, paste0("02_Venn_P10_Prom_18la-27ac-4me3.pdf")),
                 height = 4, width = 4)


# P28
gene_list_P28 <- list(H3K27ac = k27AnnoList_ens$adult@anno %>% 
                       as.data.frame() %>% 
                       dplyr::filter(annotation == "Promoter",
                                     transcriptBiotype == "protein_coding") %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F) %>% unique(),
                  H3K4me3 = k4me3AnnoList_ens$adult@anno %>% 
                       as.data.frame() %>% 
                       dplyr::filter(annotation == "Promoter",
                                     transcriptBiotype == "protein_coding") %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F) %>% unique(),
                  H3K18la = peakAnnoList_ens$P28 %>% 
                       as.data.frame() %>% 
                       dplyr::filter(annotation == "Promoter",
                                     transcriptBiotype == "protein_coding") %>% 
                       dplyr::select(SYMBOL) %>% 
                       as.vector() %>% unlist(use.names = F) %>% unique())

venn <- Venn(gene_list_P28)
data <- process_data(venn)

gv <- ggplot() +
  # change mapping of color filling
  geom_sf(aes(fill = log10(count)), data = venn_region(data), show.legend = TRUE)+
  # adjust edge size and color
  geom_sf(aes(color=id), size = 3, data = venn_setedge(data), show.legend = FALSE) + 
  # geom_sf_label(aes(label=count), size = 3, data = venn_region(data)) +
  geom_sf_label(aes(label=(count / sum(count) * 100) %>% round(digits = 0)), size = 3, data = venn_region(data)) +
  # show set label in bold
  geom_sf_text(aes(label = name), size = 5, fontface = "bold", data = venn_setlabel(data)) +  
  # add a alternative region name
  # geom_sf_label(aes(label = name), data = venn_region(data), alpha = 0.5) +  
  theme_void()
gv <- gv + scale_fill_distiller(palette = "Purples", direction = -1)

cowplot::ggsave2(gv, filename = file.path(dir_export, paste0("02_Venn_P28_Prom_18la-27ac-4me3.pdf")),
                 height = 4, width = 4)



#####################
# Boxplots
#####################

# P4
venn <- Venn(gene_list_P4)
data <- process_data(venn)
 
data_item <- data@region$item
names(data_item) <- data@region$name

df <- data.frame()
tp = "P4"

for (i in names(data_item)){
  tmp_data <- retina_cpm %>% dplyr::filter(external_gene_name %in% data_item[[i]])
  tmp_tidy <- data.frame(Section = i, 
                         Gene = tmp_data$external_gene_name, 
                         CPM = tmp_data[tp], row.names = NULL)
  df <- rbind(df, tmp_tidy)
}

no_mark <- data_item %>% unlist(use.names = F)
tmp_data <- retina_cpm %>% 
  dplyr::filter(!(external_gene_name %in% no_mark) & gene_biotype == "protein_coding")
tmp_tidy <-  data.frame(Section = "no_mark", 
                         Gene = tmp_data$external_gene_name, 
                         CPM = tmp_data[tp], row.names = NULL)

df <- rbind(df, tmp_tidy)

df$Section <- factor(df$Section, levels = c(names(data_item)[c(7,4:6,1:3)], "no_mark") %>% rev())

tmp_count <- df %>% count(Section) %>% dplyr::select(n)
tmp_count <- paste0("n=", tmp_count$n)

gb <- ggplot(df, aes(x = Section, y = P4, color = Section)) +
  geom_boxplot() + coord_flip() + theme_classic() + theme(legend.position = "none") +
  annotate("text", label = tmp_count, x = seq(1:8), y = 16) +
  guides(color = guide_legend(reverse=T)) + ylab(label = "CPM (log2)") + ggtitle(paste0(tp, " gene expression with presence of histone marks in promoter"))

cowplot::ggsave2(gb, filename = file.path(dir_export, paste0("02_Box_P4_Prom_18la-27ac-4me3.pdf")),
                 height = 6, width = 8)


# P10
venn <- Venn(gene_list_P10)
data <- process_data(venn)
 
data_item <- data@region$item
names(data_item) <- data@region$name

df <- data.frame()
tp = "P10"

for (i in names(data_item)){
  tmp_data <- retina_cpm %>% dplyr::filter(external_gene_name %in% data_item[[i]])
  tmp_tidy <- data.frame(Section = i, 
                         Gene = tmp_data$external_gene_name, 
                         CPM = tmp_data[tp], row.names = NULL)
  df <- rbind(df, tmp_tidy)
}

no_mark <- data_item %>% unlist(use.names = F)
tmp_data <- retina_cpm %>% dplyr::filter(!(external_gene_name %in% no_mark) & gene_biotype == "protein_coding")
tmp_tidy <-  data.frame(Section = "no_mark", 
                         Gene = tmp_data$external_gene_name, 
                         CPM = tmp_data[tp], row.names = NULL)

df <- rbind(df, tmp_tidy)

df$Section <- factor(df$Section, levels = c(names(data_item)[c(7,4:6,1:3)], "no_mark") %>% rev())

tmp_count <- df %>% count(Section) %>% dplyr::select(n)
tmp_count <- paste0("n=", tmp_count$n)

gb <- ggplot(df, aes(x = Section, y = P10, color = Section)) +
  geom_boxplot() + coord_flip() + theme_classic() + theme(legend.position = "none") +
  annotate("text", label = tmp_count, x = seq(1:8), y = 16) +
  guides(color = guide_legend(reverse=T)) + ylab(label = "CPM (log2)") + ggtitle(paste0(tp, " gene expression with presence of histone marks in promoter"))

cowplot::ggsave2(gb, filename = file.path(dir_export, paste0("02_Box_P10_Prom_18la-27ac-4me3.pdf")),
                 height = 6, width = 8)


# P28
venn <- Venn(gene_list_P28)
data <- process_data(venn)
 
data_item <- data@region$item
names(data_item) <- data@region$name

df <- data.frame()
tp = "P28"

for (i in names(data_item)){
  tmp_data <- retina_cpm %>% dplyr::filter(external_gene_name %in% data_item[[i]])
  tmp_tidy <- data.frame(Section = i, 
                         Gene = tmp_data$external_gene_name, 
                         CPM = tmp_data[tp], row.names = NULL)
  df <- rbind(df, tmp_tidy)
}

no_mark <- data_item %>% unlist(use.names = F)
tmp_data <- retina_cpm %>% dplyr::filter(!(external_gene_name %in% no_mark) & gene_biotype == "protein_coding")
tmp_tidy <-  data.frame(Section = "no_mark", 
                         Gene = tmp_data$external_gene_name, 
                         CPM = tmp_data[tp], row.names = NULL)

df <- rbind(df, tmp_tidy)

df$Section <- factor(df$Section, levels = c(names(data_item)[c(7,4:6,1:3)], "no_mark") %>% rev())

tmp_count <- df %>% count(Section) %>% dplyr::select(n)
tmp_count <- paste0("n=", tmp_count$n)

gb <- ggplot(df, aes(x = Section, y = P28, color = Section)) +
  geom_boxplot() + coord_flip() + theme_classic() + theme(legend.position = "none") +
  annotate("text", label = tmp_count, x = seq(1:8), y = 16) +
  guides(color = guide_legend(reverse=T)) + ylab(label = "CPM (log2)") + ggtitle(paste0(tp, " gene expression with presence of histone marks in promoter"))

cowplot::ggsave2(gb, filename = file.path(dir_export, paste0("02_Box_P28_Prom_18la-27ac-4me3.pdf")),
                 height = 6, width = 6)


###################
# Euler plot
###################
library(eulerr)

# P4
eu <- euler(gene_list_P4)
eup <- plot(eu, quantities = list(type = "percent"))

cowplot::ggsave2(eup, filename = file.path(dir_export, paste0("02_Euler_P4_Prom_18la-27ac-4me3.pdf")),
                 height = 4, width = 4)

# P10
eu <- euler(gene_list_P10)
eup <- plot(eu, quantities = list(type = "percent"))

cowplot::ggsave2(eup, filename = file.path(dir_export, paste0("02_Euler_P10_Prom_18la-27ac-4me3.pdf")),
                 height = 4, width = 4)

# P28
eu <- euler(gene_list_P28)
eup <- plot(eu, quantities = list(type = "percent"))

cowplot::ggsave2(eup, filename = file.path(dir_export, paste0("02_Euler_P28_Prom_18la-27ac-4me3.pdf")),
                 height = 4, width = 4)

```


################################################################################
# Figure 4 - FGE of venn sections of promoter genes from H3K18la, H3K27ac, H3K4me3
```{r}
library(clusterProfiler)
library(enrichplot)
library(org.Mm.eg.db)
source(file.path(dest_tools_dir, "GO_leaf.R"))
# source("../src/tools/Dotplot_GO.R")

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 5

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_enrich <- file.path(dir_export, paste0("02_FGE_VennRegions_", enrich_db, "_", GOont))
dir.create(dir_enrich)

univ <- retina_cpm %>% dplyr::filter(gene_biotype == "protein_coding") %>% 
  dplyr::select(external_gene_name) %>% 
  as.vector() %>% unlist(use.names = F) %>% unique()

# Prep the gene lists for the FGE loop
goi_list <- list(gene_list_P4, gene_list_P10, gene_list_P28)
names(goi_list) <- c("P4", "P10", "P28")

# Perform FGE on the GOIs from each subset of peaks
for (i in names(goi_list)){
  
  tmp_venn <- Venn(goi_list[[i]])
  tmp_data <- process_data(tmp_venn)
  tmp_regions <- tmp_data@region$item
  names(tmp_regions) <- tmp_data@region$name
  
  enrich_results <- clusterProfiler::compareCluster(geneClusters = tmp_regions, fun = "enrichGO",
                                            keyType = "SYMBOL", ont = GOont, OrgDb = org, univ = univ)
  
  
  # Filter enrichment for significance
  enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)
  
  # Fix the Ratio
  enrich_results_filt_fix <- enrich_results_filt %>% 
    as.data.frame() %>% 
    mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
    rowwise() %>% 
    mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))
  
  # Export the FGE results table
  tmp <- enrich_results_filt_fix[,-11]
  # Export the enrichment analysis
  write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_full.tsv")),
                quote = F, row.names = F, sep = "\t")

  # Get leaf terms
  enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)
  
  # Fix the Ratio
  enrich_results_leaf_fix <- enrich_results_leaf %>% 
    as.data.frame() %>% 
    mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
    rowwise() %>% 
    mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))
  
  # Export the FGE results table
  tmp <- enrich_results_leaf_fix[,-11]
  # Export the enrichment analysis
  write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_leaf.tsv")),
                quote = F, row.names = F, sep = "\t")
  
  # Get top 5 pathways per cell type to get union set of pathways
  top5_union <- enrich_results_leaf %>% 
    group_by(Cluster) %>% 
    arrange(p.adjust) %>% 
    slice_head(n = top)
  
  
  # Filter the full table for the union of top5 pathways
  top5_enrich <- enrich_results_filt_fix %>% 
    dplyr::filter(ID %in% (top5_union$ID %>% unique()))
  
  # Prep the df for plotting
  top5_enrich <- top5_enrich %>% 
    arrange(Cluster, p.adjust)
  top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
  top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
  top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)
  
  # Make plot
  pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
      geom_point(aes(size = Ratio)) + 
      theme_bw() +
      scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
      guides(size=guide_legend(title="Ratio")) +
      theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
            axis.text.x = element_text(angle = 30, hjust = 1)) +
      ggtitle(paste0("H3K18la Peaks in ", i))
  
  #Open the pdf device
  cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0(i, "_VennSection_FGE_top_", top, "_", i, "_leaf_01_dotplot.pdf")),
                   height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                   width = 6)
    
}


```


##################
# Sankey plot of promoter peak histone presence
##################
```{r}
library(ggalluvial)

# Prep the gene lists for the FGE loop
goi_list <- list(gene_list_P4, gene_list_P10, gene_list_P28)
names(goi_list) <- c("P4", "P10", "P28")

total_gois <- c(gene_list_P4 %>% unlist(use.names = F) %>% unique(),
                gene_list_P10 %>% unlist(use.names = F) %>% unique(),
                gene_list_P28 %>% unlist(use.names = F) %>% unique()) %>% unique()


# Prep DF for alluvial plot
df_alluv <- data.frame()

# 
for (i in names(goi_list)){
  
  tmp_venn <- Venn(goi_list[[i]])
  tmp_data <- process_data(tmp_venn)
  tmp_regions <- tmp_data@region$item
  names(tmp_regions) <- tmp_data@region$name

  for (ii in names(tmp_regions)){
    
    tmp_genes <- tmp_regions[[ii]]
    tmp_df <- data.frame(Symbol = tmp_genes,
                         TP = i,
                         Region = ii)
   
    
    df_alluv <- rbind(df_alluv, tmp_df)
      
  }
  
   tmp_df_missing <- data.frame(Symbol = total_gois[!(total_gois %in% (tmp_regions %>% unlist(use.names = F) %>% unique()))],
                                TP = i,
                                Region = NA)
   df_alluv <- rbind(df_alluv, tmp_df_missing)
}

df_alluv$TP <- factor(df_alluv$TP, levels = c("P4", "P10", "P28"))
df_alluv <- df_alluv %>% as.tibble() %>% replace_na(list(Region = "no_mark"))

hist_barplot_levels <- gb$data$Section %>% levels()
# hist_barplot_levels[1] <- NA
df_alluv$Region <- factor(df_alluv$Region, levels = hist_barplot_levels)

gga <- ggplot(df_alluv, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) +
  # scale_fill_brewer(type = "qual", palette = "Set2") +
  # scale_fill_viridis_b() +
  geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) +
  geom_stratum(alpha = 0.75) +
  theme(legend.position = "bottom") +
  ggtitle("Histone occupancy in promoter region during development") 
cowplot::ggsave2(gga, filename = file.path(dir_export, paste0("02_Alluvial_Prom_18la-27ac-4me3.pdf")),
                 height = 8, width = 12)

saveRDS(object = df_alluv, file = "../data/interim/02_df_alluvial.RDS")

```


################################################################################
## Perform FGE on Alluvial segments
################################################################################
```{r}

df_alluv_spread <- df_alluv %>% spread(TP, Region)


goi_list_alluv <- list(all3_allTP = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3..H3K18la", 
                                       P10 == "H3K27ac..H3K4me3..H3K18la", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P4.all3_P10.K4..K27_P28.all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3..H3K18la", 
                                       P10 == "H3K27ac..H3K4me3", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P4.K4..K27_P10.all3_P28.all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3", 
                                       P10 == "H3K27ac..H3K4me3..H3K18la", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P4.K4..K27_P10.K4..K27_P28.all3 = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3", 
                                       P10 == "H3K27ac..H3K4me3", 
                                       P28 == "H3K27ac..H3K4me3..H3K18la") %>% 
                         select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       K4..K27_allTP = df_alluv_spread %>% 
                         dplyr::filter(P4 == "H3K27ac..H3K4me3", 
                                       P10 == "H3K27ac..H3K4me3", 
                                       P28 == "H3K27ac..H3K4me3") %>% 
                         select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28.K27..K18 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K27ac..H3K18la") %>% 
                         select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28.K4..K18 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K4me3..H3K18la") %>% 
                         select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F),
                       P28.K18 = df_alluv_spread %>% 
                         dplyr::filter(P28 == "H3K18la") %>% 
                         select(Symbol) %>% 
                         as.vector() %>% unlist(use.names = F))


# Perform FGE on the GOIs from each subset of peaks

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 12

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_enrich <- file.path(dir_export, paste0("02_Alluvial_Regions_", enrich_db, "_", GOont))
dir.create(dir_enrich)

univ <- df_alluv_spread$Symbol

  
enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list_alluv, fun = "enrichGO",
                                          keyType = "SYMBOL", ont = GOont, OrgDb = org, univ = univ)


# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>% 
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>% 
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("H3K18la Peaks in ", i))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0(i, "_Alluvial-Region_FGE_top_", top, "_", i, "_leaf_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)
    
```





<!-- ################################################################################ -->
<!-- ## Figure 3 Correlate peaks with RNA-seq -->
<!-- ################################################################################ -->
<!-- ```{r} -->
<!-- library(csaw) -->
<!-- library(limma) -->
<!-- library(edgeR) -->

<!-- # Make master peak domain -->
<!-- masterPeak <- GenomicRanges::reduce(c(peakList_cons[[1]], -->
<!--                                  peakList_cons[[2]], -->
<!--                                  peakList_cons[[3]])) -->

<!-- tmp_gr_list <- list() -->
<!-- for (i in 1:3){ -->
<!--   tmp_gr <- peakList_cons[[i]] -->
<!--   seqlevelsStyle(tmp_gr) <- "NCBI" -->
<!--   tmp_gr_list[[i]] <- tmp_gr  -->
<!-- } -->

<!-- masterPeak_anno <- GenomicRanges::reduce(c(tmp_gr_list[[1]], -->
<!--                                  tmp_gr_list[[2]], -->
<!--                                  tmp_gr_list[[3]])) -->

<!-- masterPeak_anno <- annotatePeak(masterPeak_anno, tssRegion=c(-1000, 500), -->
<!--                          TxDb=ensDb, annoDb=NULL, addFlankGeneInfo = T, -->
<!--                          overlap = "all") -->

<!-- # Get fragment counts for domains -->
<!-- bams <- list.files(path = "../data/interim/", pattern = "*bam$", recursive = T, full.names = T) -->
<!-- bams <- bams[c(23:26, 13:17, 18:22)] -->
<!-- names(bams) <- paste(gsub(".+\\/K18La-(.+)_cr.+", "\\1", bams), c(seq(1:4), seq(1:5), seq(1:5)), sep = "_" ) -->

<!-- pe.param <- readParam(max.frag=400, pe="both", minq = 20, dedup = F) -->

<!-- k18la.counts <- regionCounts(bams, masterPeak, ext=150, param=pe.param) -->
<!-- colnames(k18la.counts) <- paste(gsub(".+\\/K18La-(.+)_cr.+", "\\1", bams), c(seq(1:4), seq(1:5), seq(1:5)), sep = "_" ) -->


<!-- # Seq depth normalization -->
<!-- k18la.dge <- asDGEList(k18la.counts, -->
<!--                # norm.factors = filtered.data$norm.factors, -->
<!--                group = gsub(".+\\/K18La-(.+)_cr.+", "\\1", bams), -->
<!--                genes = as.data.frame(masterPeak_anno@anno)) -->
<!-- colnames(k18la.dge) <- names(bams) -->
<!-- k18la.dge <- edgeR::calcNormFactors(k18la.dge) -->
<!-- k18la.dge$rpkm <- edgeR::rpkm(k18la.dge, gene.length = "width") -->

<!-- k18la.dge$logRPKM <- log2(k18la.dge$rpkm + 1) -->

<!-- color.grp.pal <- wesanderson::wes_palette("Darjeeling1", 5, type = "discrete")[c(1,3,5)] -->

<!-- #Prep the PCA data -->
<!-- pca.k18la <- PCAtools::pca(k18la.dge$logRPKM, metadata = data.frame(k18la.dge$samples)) -->
<!-- pca.k18la$metadata$group <- factor(pca.k18la$metadata$group, levels = c("P4", "P10", "P28")) -->
<!-- sampkey <- color.grp.pal -->
<!-- names(sampkey) <- unique(pca.k18la$metadata$group) -->

<!-- #Plot all the samples -->
<!-- X='PC1'; Y='PC2' -->
<!-- p <- PCAtools::biplot(pca.k18la, x = X, y = Y, -->
<!--        colby = "group", colkey = sampkey, -->
<!--        #lab = "replicate", labSize = 2, -->
<!--        legendPosition = 'right',  -->
<!--        title = 'PCA All Samples', -->
<!--        subtitle = paste0(X, ' versus ', Y)) -->

<!-- cowplot::ggsave2(p, filename = file.path(dir_export, paste0("02_PCA_H3K18la_PC1vPC2.pdf")), -->
<!--                  height = 5, width = 6) -->


<!-- # Correlation with RNA -->
<!-- ``` -->



<!-- # Test EnrichedHeatmaps -->
<!-- https://bioconductor.org/packages/release/bioc/vignettes/EnrichedHeatmap/inst/doc/EnrichedHeatmap.html -->

<!-- ```{r} -->
<!-- library(EnrichedHeatmap) -->

<!-- load(system.file("extdata", "chr21_test_data.RData", package = "EnrichedHeatmap")) -->
<!-- ls() -->

<!-- tss = promoters(genes, upstream = 0, downstream = 1) -->
<!-- mat1 = normalizeToMatrix(H3K4me3, tss, value_column = "coverage",  -->
<!--     extend = 5000, mean_mode = "w0", w = 50) -->

<!-- EnrichedHeatmap(mat1, name = "H3K4me3") -->

<!-- tss2 <- getPromoters(TxDb = ensDb, upstream = 0, downstream = 1) -->
<!-- seqlevelsStyle(tss2) <- "UCSC" -->

<!-- mat_P28 <- normalizeToMatrix(peakList_cons$`K18La-P28`, tss2, value_column = "score",  -->
<!--     extend = 1000, mean_mode = "w0", w = 50) -->

<!-- EnrichedHeatmap(mat_P28, name = "P28-H3K18la", ylim = c(0,100)) -->

<!-- ``` -->






<!-- ## Get metadata for samples -->
<!-- ```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE} -->
<!-- # ord <- c(1:4) -->

<!-- # Macs2 beds -->
<!-- macs_beds <- list.files(path = "../data/interim/bed/", pattern = "optimal.+narrowPeak.gz", recursive = T) -->

<!-- # Meta sheet -->
<!-- meta <- data.frame(Beds = macs_beds, -->
<!--                    Samp = gsub("(.+)_overlap.+", "\\1", macs_beds)) -->
<!-- meta <- meta[c(5,3,4,2,1,8,6,7),] -->
<!-- # meta$Beds <- factor(meta$Beds, levels = levels(meta$Beds)) -->
<!-- meta$Samp <- factor(meta$Samp, levels = unique(meta$Samp)) -->


<!-- print("Meta data acquired...") -->
<!-- ``` -->


<!-- # Import the peaks into a merged list object and plot for number of peaks at each timepoint -->
<!-- ```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE} -->
<!-- library(ChIPpeakAnno) -->

<!-- cons_dir <- "../data/interim/bed/" -->

<!-- # Import the cons peaks -->
<!-- peakList_cons <- sapply(file.path(cons_dir, meta$Beds), toGRanges, format="narrowPeak") -->
<!-- names(peakList_cons) <- meta$Samp -->

<!-- # Sort the cons peaks -->
<!-- for (i in 1:length(peakList_cons)){ -->

<!--   # seqlevelsStyle(peakList_cons[[i]]) <- "NCBI" -->
<!--   peakList_cons[i] <- sortSeqlevels(peakList_cons[[i]]) -->
<!--   peakList_cons[i] <- GenomicRanges::sort(peakList_cons[[i]]) -->
<!-- } -->

<!-- # Generate plot for peak number -->
<!-- df <- lapply(peakList_cons, length) %>% as.data.frame() %>% gather("Sample", "Peaks") -->
<!-- df$Sample <- gsub("X", "", df$Sample) -->
<!-- df$Sample <- factor(df$Sample, levels = df$Sample) -->

<!-- b <- ggplot(df, aes(Sample, Peaks))+ -->
<!--   geom_bar(stat = "identity") + -->
<!--   theme_minimal() + -->
<!--   theme(axis.text.x = element_text(angle = 30, hjust = 0.8, vjust = 0.8)) + -->
<!--   labs(title = "Peaks") + xlab(NULL) -->

<!-- cowplot::ggsave2(plot = b, filename = "../data/processed/Peak_number_allPeaks.pdf", height = 4, width = 3) -->

<!-- print("Peak data acquired...") -->
<!-- ``` -->


<!-- ## Annotate peaks with ensembl db -->
<!-- ```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE} -->
<!-- library(ChIPseeker) -->
<!-- library(org.Mm.eg.db) -->
<!-- library(biomaRt) -->
<!-- library(GenomicFeatures) -->


<!-- peakAnnoList_ens <- list() -->

<!-- for (i in 1:length(peakList_cons)){ -->

<!--   peak_tmp <- peakList_cons[[i]] -->
<!--   seqlevelsStyle(peak_tmp) <- "NCBI" -->

<!--   peakAnnoList_ens[[i]] <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500), -->
<!--                          TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T, -->
<!--                          overlap = "all") -->

<!-- } -->

<!-- names(peakAnnoList_ens) <- levels(meta$Samp) -->

<!-- ``` -->




<!-- ################################################################################ -->
<!-- # Prepare plot for figures - Oct 30th, 2023  -->
<!-- ################################################################################ -->

<!-- #################### -->
<!-- # Timecourse K18La -->
<!-- #################### -->



<!-- # FGE of annotated genes -->
<!-- ```{r} -->

<!-- library(clusterProfiler) -->
<!-- library(enrichplot) -->
<!-- library(org.Mm.eg.db) -->
<!-- source(file.path(dest_tools_dir, "GO_leaf.R")) -->
<!-- # source("../src/tools/Dotplot_GO.R") -->

<!-- org = "org.Mm.eg.db" -->

<!-- enrich_db <- "GO_220310" -->
<!-- GOont <- "BP" -->

<!-- # dir_enrich <- file.path(dir_res, "FuncGeneEnrich") -->
<!-- dir_enrich <- file.path(dir_export, paste0("FuncGeneEnrich_", enrich_db, "_", GOont)) -->
<!-- dir.create(dir_enrich) -->


<!-- # Get list of genes for enrichment analysis -->

<!-- goi.list <- list(P4 = peakAnnoList_ens[[1]]@anno$SYMBOL %>% unique(), -->
<!--                  P10 = peakAnnoList_ens[[2]]@anno$SYMBOL %>% unique(), -->
<!--                  P28 = peakAnnoList_ens[[3]]@anno$SYMBOL %>% unique()) -->

<!-- # # Get universe -->
<!-- # univ <- row.names(seu_rodbp@assays$decontXcounts@data) -->

<!-- # Get enrichment -->
<!-- enrich_results <- clusterProfiler::compareCluster(geneClusters = goi.list, fun = "enrichGO", -->
<!--                                             keyType = "SYMBOL", ont = GOont, OrgDb = org) -->





<!-- ########################### -->
<!-- # Look at GO Level 6 -->
<!-- ########################### -->

<!-- go_level <- 8 -->
<!-- go_ids <- GOxploreR::Level2GOTermBP(go_level, organism = "mouse") -->
<!-- top <- 10 -->

<!-- # Filter enrichment for significance -->
<!-- enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.0001, Count > 2) -->

<!-- enrich_results_filt_fix <- enrich_results_filt %>%  -->
<!--   as.data.frame() %>%  -->
<!--   mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>%  -->
<!--   rowwise() %>%  -->
<!--   mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2])) -->

<!-- # Prep the enrichment DF -->
<!-- enrich <- enrich_results_filt_fix %>%  -->
<!--   as.data.frame() %>%  -->
<!--   dplyr::filter(ID %in% go_ids) -->
<!-- enrich$Description <- factor(enrich$Description, levels = rev(unique(enrich$Description))) -->


<!-- # Get top 5 pathways per cell type to get union set of pathways -->
<!-- top5_union <- enrich %>%  -->
<!--   group_by(Cluster) %>%  -->
<!--   arrange(p.adjust) %>%  -->
<!--   slice_head(n = top) -->


<!-- # Filter the full table for the union of top5 pathways -->
<!-- top5_enrich <- enrich %>%  -->
<!--   dplyr::filter(ID %in% (top5_union$ID %>% unique())) -->

<!-- # Prep the df for plotting -->
<!-- top5_enrich <- top5_enrich %>%  -->
<!--   arrange(Cluster, p.adjust) -->
<!-- top5_enrich$Description <- substring(top5_enrich$Description, 1, 60) -->
<!-- top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description))) -->
<!-- top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust) -->

<!-- # Make plot -->
<!-- p <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) +  -->
<!--     geom_point(aes(size = Ratio)) +  -->
<!--     theme_bw() + -->
<!--     scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) + -->
<!--     guides(size=guide_legend(title="Ratio")) + -->
<!--     theme(axis.title.x = element_blank(), axis.title.y = element_blank(), -->
<!--           axis.text.x = element_text(angle = 30, hjust = 1)) -->


<!-- #Open the pdf device -->
<!-- cowplot::ggsave2(plot = p, filename = file.path(dir_enrich, paste0("Timepoint_FGE_top_", top, "_level_", go_level, "_0001_dotplot.pdf")), -->
<!--                  height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5,  -->
<!--                  width = 6) -->



<!-- ########################### -->
<!-- # Look at leaf terms -->
<!-- ########################### -->


<!-- # # Filter enrichment for significance -->
<!-- # enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.0001, Count > 2) -->

<!-- # Get leaf terms -->
<!-- enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont) -->

<!-- # # Fix the Ratio -->
<!-- # enrich_results_leaf_fix <- enrich_results_leaf %>% -->
<!-- #   as.data.frame() %>% -->
<!-- #   mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% -->
<!-- #   rowwise() %>% -->
<!-- #   mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2])) -->

<!-- # Get top 5 pathways per cell type to get union set of pathways -->
<!-- top5_union <- enrich_results_leaf %>%  -->
<!--   group_by(Cluster) %>%  -->
<!--   arrange(p.adjust) %>%  -->
<!--   slice_head(n = top) -->


<!-- # Filter the full table for the union of top5 pathways -->
<!-- top5_enrich <- enrich_results_filt_fix %>%  -->
<!--   dplyr::filter(ID %in% (top5_union$ID %>% unique())) -->

<!-- # Prep the df for plotting -->
<!-- top5_enrich <- top5_enrich %>%  -->
<!--   arrange(Cluster, p.adjust) -->
<!-- top5_enrich$Description <- substring(top5_enrich$Description, 1, 60) -->
<!-- top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description))) -->
<!-- top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust) -->

<!-- # Make plot -->
<!-- pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) +  -->
<!--     geom_point(aes(size = Ratio)) +  -->
<!--     theme_bw() + -->
<!--     scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) + -->
<!--     guides(size=guide_legend(title="Ratio")) + -->
<!--     theme(axis.title.x = element_blank(), axis.title.y = element_blank(), -->
<!--           axis.text.x = element_text(angle = 30, hjust = 1)) -->

<!-- #Open the pdf device -->
<!-- cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("Timepoint_FGE_top_", top, "_leaf_0001_dotplot.pdf")), -->
<!--                  height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5,  -->
<!--                  width = 6) -->


<!-- #  -->
<!-- # # Make dotplot of results -->
<!-- # dot_go(enrich_results_leaf, top = 10, grp = "Cluster", desc = "Description", p.val = "p.adjust", ratio = "Ratio", filenm = file.path(dir_enrich, "plots", paste0(i, "_cluster_dotplot_05"))) -->
<!-- #  -->
<!-- # dot_go(enrich_results_leaf, top = 5, grp = "Cluster", desc = "Description", p.val = "p.adjust", ratio = "Ratio", filenm = file.path(dir_enrich, "plots", paste0(i, "_cluster_dotplot_05_5"))) -->

<!-- # Export the FGE results table -->
<!-- tmp <- enrich_results_filt_fix[,-11] -->
<!-- # Export the enrichment analysis -->
<!-- write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", "timepoint", "_filt.tsv")), -->
<!--               quote = F, row.names = F, sep = "\t") -->

<!-- # Export the FGE results table -->
<!-- tmp <- enrich_results_leaf_fix[,-11] -->
<!-- # Export the enrichment analysis -->
<!-- write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", "timepoint", "_leaf.tsv")), -->
<!--               quote = F, row.names = F, sep = "\t") -->


<!-- ``` -->


#
### Session Information
```{r, echo=FALSE}
sessionInfo()
```

#
## Explanation of Output Files

#### File structure of the data

* data
    + external
    + interim
    + processed   
    + raw
        - bams
        - beds
        - bigwig
* notebooks

#
#### Output File Description

Output files from analysis contain both tab-separated values table (.tsv), as well as, plots in pdf form. The tsv tables can be opened in Excel.

#
##### 'data/external'
* DevRet_Gene38v84_nCPM_MSTR.tsv
    + Gene expression values from deveolping mouse retina. Values are normalized CPM values. File is tab-separated value file.

#
##### 'data/interim'
Output from combining replicate peaks for the concensus bed file used for further analysis

* beds
    + Consensus bed files for each timepoint grouping using overlap method
    + cons
        - Consensus bed files for each timepoint using the cons method (*narrowPeak)
* bigwig
    + Merged bigwig files using samples from cons samples
        
#
##### 'data/processed'

* 01_PeakGeneAnno_*.txt
    + Peak annotation at each tf-timepoint
* gProf_AllPeaks
    + Directory containing the export table and visualization results from the functional gene enrichment analysis from genes containing peaks from the consensus bed files for the timepoint groupings
        - gProfileR_all.tsv - output from gProfileR
        - Heatmaps : heatmaps of all the top enriched categories using log2 CPM values 
* gProf_AllPeaks_Nrl_PN
    + Replot of the dotplots using all timepoints except Nrl E14
* gProf_[Crx|Nrl]
    + Replot of the dotplots separarting Crx and Nrl
* gProf_Nrl_PN
    + Replot of the dotplots separating Nrl and excluding the E14 timepoint

#
##### 'data/raw

* bams
    + Bam files generated from the pipeline for each sample.
* beds
    + The 1% FDR Macs peak bed files for each sample used in the analysis
* bigwig
    + Bigwig file exports for each sample. The bigwig file is a bin per million normalization. 