---
title: "03 - H3K18la DE Peak Analysis- Sept 13th, 2024"
output: html_notebook
---

### Project Summary
This project investigates the H3K18la during inhibition of glycolysis. 

**DESIGN**:  

**NOTES**: The data were generated from the ENCODE histone pipeline with duplicate 
removal with no controls. This notebool uses peak quant based on CSAW normalization. 
Differential binding and correlation with RNA changes.

**UPDATE**: The definition of promoter (-1000 to +500) was used for peak annotation. 
Enhancer is defined as -10k to -1000 from TSS.

	<!-- • Figure 2 -->
	<!-- 	○ # of genes with peaks -->
	<!-- 	○ FGE with gene numbers at each definition (GO:leaf) -->
	<!-- 		§ Promoter: -500 to +250 -->
	<!-- 		§ Enhancer: -10K to -1kb -->
	<!-- 		§ Genebody -->
	<!-- 		§ Intergenic -->
	<!-- 	○ Heatmap -->
	<!-- 		§  -500 to 500 -->
	<!-- 		§ -10kb to 10kb -->
	<!-- 	○ RNA-seq correlation (wt retina development) -->
	<!-- • Figure 3 -->
	<!-- 	○ Corelate H3K18la -->
	<!-- 		§ ATAC -->
	<!-- 		§ Nrl -->
	<!-- 		§ RNA-seq (wt retina development -->
	<!-- 		§ ChromHMM -->
	<!-- • Figure 4 -->
	<!-- 	○ Glycolytic genes show with H3K27ac and H3K18la -->
	<!-- • Figure 5 -->
	<!-- 	○ Explant data CnR -->
	<!-- 		§ Use spike-in normalization to compare -->
	<!-- 	○ Correlate with RNA-seq (if data can be produced) -->
	
	
################################################################################
#### Get custom scripts and functions and outside data used for analysis
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

script_dir <- "https://raw.githubusercontent.com/MBrooks313/R_codes/master"
dest_tools_dir <- "../src/tools"
dest_vis_dir <- "../src/visualization"

download.file(url = file.path(script_dir, "GO_leaf.R"),
              destfile = file.path(dest_tools_dir, "GO_leaf.R"),
              method = "curl")

download.file(url = file.path(script_dir, "DFave.R"),
              destfile = file.path(dest_tools_dir, "DFave.R"),
              method = "curl")

download.file(url = file.path(script_dir, "Heatmap.R"),
              destfile = file.path(dest_tools_dir, "Heatmap.R"),
              method = "curl")

download.file(url = file.path(script_dir, "Expr_Filter.R"),
              destfile = file.path(dest_tools_dir, "Expr_Filter.R"),
              method = "curl")

download.file(url = file.path(script_dir, "VolcanoPlot.R"),
              destfile = file.path(dest_tools_dir, "VolcanoPlot.R"),
              method = "curl")
# system('cp ~/github/R_codes/DFave.R ../src/tools')
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/et_edgeR_v0.1.0.R ../src/tools")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/FuncGeneEnrich_v0.1.0.R ../src/tools")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/GO_leaf_v1.1.R ../src/tools")

#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/APk_cluster_BF_v0.1.0.R ../src/tools')
#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/HM_PCAsort_BF_v0.1.0.R ../src/visualization')
#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/parPlot_hc_v0.2.0.R ../src/visualization')
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/PCA_simple_v0.1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/VolcanoPlot_v0.1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_v1.7.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Dotplot_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Barplot_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_v1.7.R ../src/visualization")
# 
# system("cp /Volumes/NGS/RNA-seq/ENS_v98/Mouse/Counts/by_project/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv ../data/external")
# system("cp /Volumes/NGS/RNA-seq/ENS_v98/Mouse/Counts/by_project/Mouse_FSPR_WT_Development_Gene_nCPM_ENSv98.tsv ../data/external")

# Function to convert ENS ids to gene symbol 
# ens2sym <- function(x){paste(sort(unique(AnnotationDbi::mapIds(org.Mm.eg.db, keys = x, keytype = "ENSEMBL", column = "SYMBOL"))), sep = "", collapse = ",")}
library(tidyverse)

dir_export <- "../data/processed"

# options(java.parameters = "-Xmx8000m")
print("Completed...")
```


################################################################################
#### Import outside data
################################################################################
```{r}
source("../src/tools/DFave.R")


#############################################
# Import db data for time-course H3K18La

db_tc <- read.csv("/Volumes/data/People/brooksma/PEGASUS/Projects/Gaur/Analysis/cutntag_240522/240911/data/processed/03_H3k18la.db.results.tsv", sep = "\t")

db_tc_sig_PromPC <- db_tc %>% dplyr::filter(annotation == "Promoter", 
                                            transcriptBiotype == "protein_coding",
                                            (abs(logFC.P10) > 1 | abs(logFC.P28)), 
                                            FDR < 0.01)





# #############################################
# # Import expression data for flow-sorted rods
# fsprwt.dwnld <- read.csv("../data/external/Mouse_FSPR_WT_Development_Gene_nCPM_ENSv98.tsv",
#                          sep = "\t", row.names = 1)
# # Get group info for averaging
# grp <- gsub("(.+)\\.FS.+", "\\1", colnames(fsprwt.dwnld)[10:25])
# exp_grp <- factor(grp, levels = unique(grp))
# 
# # Get average CPM values
# fsprwt_cpm <- data.frame(merge(fsprwt.dwnld[,1:9],
#                                log2(DFave(fsprwt.dwnld[,10:25], exp_grp) + 1),
#                                by = 0),
#                          row.names = 1)
# 
# fsprwt_z <- data.frame(merge(fsprwt_cpm[,1:9],
#                                t(scale(t(fsprwt_cpm[,10:15]), scale = T, center = T)),
#                                by = 0),
#                          row.names = 1)
# 
# 
# 
# #############################################
# # Import expression data for developing retina
# retina.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv",
#                          sep = "\t", row.names = 1)
# 
# # Get group info for averaging
# grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.dwnld)[10:31])
# exp_grp <- factor(grp, levels = unique(grp))
# 
# # Get average CPM values
# retina_cpm <- data.frame(merge(retina.dwnld[,1:9],
#                                log2(DFave(retina.dwnld[,10:31], exp_grp) + 1),
#                                by = 0),
#                          row.names = 1)
# 
# retina_z <- data.frame(merge(retina_cpm[,1:9],
#                                t(scale(t(retina_cpm[,10:19]), scale = T, center = T)),
#                                by = 0),
#                          row.names = 1)
# 
# 
# #############################################
# # Import transcript expression data for developing retina
# retina.trans.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Trans_nRKPM_ENSv98.tsv",
#                          sep = "\t", row.names = 1)
# 
# # Get group info for averaging
# grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.trans.dwnld)[12:33])
# exp_grp <- factor(grp, levels = unique(grp))
# 
# # Get average CPM values
# retina_trans_cpm <- data.frame(merge(retina.trans.dwnld[,1:11],
#                                log2(DFave(retina.trans.dwnld[,12:33], exp_grp) + 1),
#                                by = 0),
#                          row.names = 1)
# 
# retina_trans_z <- data.frame(merge(retina_trans_cpm[,1:11],
#                                t(scale(t(retina_trans_cpm[,12:21]), scale = T, center = T)),
#                                by = 0),
#                          row.names = 1)


```


################################################################################
#### Get Ensembl annotation database
################################################################################
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

## Query for all available EnsDb databases
ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

ensDb <- ahdb[[1]]
# ensDb_pc <- addFilter(ensDb, GenebiotypeFilter("protein_coding", condition = "=="))
ensDb_pc <- addFilter(ensDb, TxBiotypeFilter("protein_coding", condition = "=="))
ensDb_pc_hard <- genes(ensDb_pc)
ensDb_pc_hard <- ensembldb::filter(ensDb, filter = ~ tx_biotype == "protein_coding")

# supportedFilters(ensDb)

```


################################################################################
#### Get list of bam files and make master bed feature for quantitation
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(tidyverse)
library(ChIPpeakAnno)
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


# Macs2 beds
bam_dir <- "../data/raw/bams/"
macs_bams <- list.files(path = bam_dir, pattern = ".+.bam$", full.names = T, recursive = T)
# macs_bams <- macs_bams[c(6:7, 1:2, 3:5)]
macs_bams <- macs_bams[c(4:6, 1:3, 14,7:8,
                         11:12, 9:10, 13, 15:16)]
# Meta sheet
meta <- data.frame(bams = macs_bams,
                   Mark = gsub(".+\\/(.+)_.+_R.+\\.target.+", "\\1", macs_bams),
                   Treat = gsub(".+\\/H3.+_(.+)_R.+\\.target.+", "\\1", macs_bams),
                   Group = gsub(".+\\/(.+)_R.+", "\\1", macs_bams),
                   Rep = paste0("R", c(rep(1:3, times = 3), rep(1:2, times = 2), 1:3)),
                   Run = gsub(".+\\/(25.+)\\/H.+", "\\1", macs_bams))

row.names(meta) <- paste(meta$Group, meta$Rep, sep = "_")

names(macs_bams) <- paste(meta$Group, meta$Rep, sep = "_")
meta$Group <- factor(meta$Group, levels = unique(meta$Group))


# Import consensus peaks
peakList_cons <- readRDS(file = "../data/interim/peakList_cons_250729.RDS")
# peakList_cons <- peakList_cons

mstr_features <- IRanges::reduce(c(peakList_cons[[1]],
                         peakList_cons[[2]],
                         peakList_cons[[3]],
                         peakList_cons[[4]],
                         peakList_cons[[5]],
                         peakList_cons[[6]]))

# Annotate the features
mstr_features_anno <- annotatePeak(mstr_features, tssRegion=c(-1000, 500),
                         # TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T,
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
mstr_features_anno@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = mstr_features_anno@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")

# Export bed file of the regions
mstr_feature_bed <- data.frame(chrom = seqnames(mstr_features_anno@anno),
                             start = start(mstr_features_anno@anno),
                             end = end(mstr_features_anno@anno),
                             name = mstr_features_anno@anno$SYMBOL,
                             score = width(mstr_features_anno@anno),
                             strand = strand(mstr_features_anno@anno))

write.table(mstr_feature_bed, file = "../data/processed/03_mstr_feature.bed", sep = "\t", quote = F, col.names = F, row.names = F)

```


################################################################################
#### Get bam counts over mstr_feature
################################################################################
```{r}
library(csaw)

# # Determine optimal fragment length using cross-correlation
# wind.param <- readParam(pe="both", minq = 20, dedup = T)
# max.delay <- 500
# bam.files <- paste0(bam_dir, macs_bams)
# x <- correlateReads(bam.files, max.delay, param=wind.param)
# plot(0:max.delay, x, type="l", ylab="CCF", xlab="Delay (bp)")
# print(paste0("Optimal fragment size (bp): ", maximizeCcf(x)))



# Perform region counts
pe.param <- readParam(max.frag=600, pe="both", minq = 20, dedup = T)
# bam.files <- paste0(bam_dir, macs_bams)

h3k18la.counts <- regionCounts(meta$bams, mstr_features_anno@anno, ext=150, param=pe.param)


# Normalization
binned <- windowCounts(meta$bams, bin=TRUE, width=10000, param=pe.param)
h3k18la_se <- normFactors(binned, se.out=T)

```


################################################################################
#### Prep the DGE list objects
################################################################################
```{r}
library(edgeR)

# source(file.path(dest_tools_dir, "DFave.R"))
source(file.path(dest_tools_dir, "Expr_Filter.R"))


##########
## h3k18la
# samp_nm <- gsub(".+H3K18La_(.+)\\.target.+", "\\1", h3k18la.counts@colData[[1]])

h3k18la.dge <- asDGEList(h3k18la.counts,
               norm.factors = h3k18la_se$norm.factors,
               # group = gsub("_.+", "", samp_nm),
               group = meta$Group,
               genes = as.data.frame(mstr_features_anno))
colnames(h3k18la.dge) <- paste(meta$Group, meta$Rep, sep = "_")
# h3k18la.dge$samples$group <- factor(h3k18la.dge$samples$group, levels = unique(h3k18la.dge$samples$group))

h3k18la.dge$samples$Run <- meta$Run

h3k18la.dge$rpkm <- rpkm(h3k18la.dge, gene.length = "width")
ExpFilt(h3k18la.dge$rpkm, exp = 1, group = h3k18la.dge$samples$group)

h3k18la.dge.filt <- h3k18la.dge[idx.filt, keep.lib.sizes = T]

h3k18la.dge.filt$logCPM <- log2(cpm(h3k18la.dge.filt) + 1)
h3k18la.dge.filt$logFPKM <- rpkm(h3k18la.dge.filt, gene.length = "width", normalized.lib.sizes = T, log = T, prior.count = 1)
h3k18la.dge.filt$FPKM <- rpkm(h3k18la.dge.filt, gene.length = "width", normalized.lib.sizes = T, log = F)
h3k18la.dge.filt$FPKM_mean <- h3k18la.dge.filt$FPKM %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(H3K18LA_CNTRL = mean(c(H3K18LA_CNTRL_R1, H3K18LA_CNTRL_R2, H3K18LA_CNTRL_R3)),
         H3K18LA_2DG = mean(c(H3K18LA_2DG_R1, H3K18LA_2DG_R2, H3K18LA_2DG_R3)),
         H3K18LA_SO = mean(c(H3K18LA_SO_R1, H3K18LA_SO_R2, H3K18LA_SO_R3)),
         H3K27AC_CNTRL = mean(c(H3K27AC_CNTRL_R1, H3K27AC_CNTRL_R2)),
         H3K27AC_2DG = mean(c(H3K27AC_2DG_R1, H3K27AC_2DG_R2)),
         H3K27AC_SO = mean(c(H3K27AC_SO_R1, H3K27AC_SO_R2, H3K27AC_SO_R3)))

# Export RPKM quant
tmp <- cbind(as.data.frame(mstr_features_anno), h3k18la.dge$rpkm)
write.table(tmp, file = "../data/processed/03_tc_quant_rpkm.tsv", sep = "\t", quote = F, col.names = F, row.names = F)


# Export bed file of the filtered regions
mstr_feature_bed_filt <- data.frame(chrom = h3k18la.dge.filt$genes$seqnames,
                             start = h3k18la.dge.filt$genes$start,
                             end = h3k18la.dge.filt$genes$end,
                             name = h3k18la.dge.filt$genes$SYMBOL,
                             score = h3k18la.dge.filt$genes$width,
                             strand = h3k18la.dge.filt$genes$strand)

write.table(mstr_feature_bed_filt, file = "../data/processed/03_mstr_feature_expfilt.bed", sep = "\t", quote = F, col.names = F, row.names = F)


```


################################################################################
#### PCA analysis
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
# library(plot3D)
library(PCAtools)
library(wesanderson)

# wind.dge <- asDGEList(filtered.data,
#                group = paste(meta_used$stage, meta_used$treatment, sep = "_"))
# colnames(wind.dge) <- meta_used$sample_name
# 
# wind.dge$logCPM <- log2(cpm(wind.dge) + 1)
# 
# color.grp.pal <- wes_palette("Zissou1", 5, "discrete")[c(5,3)]
# color.grp.pal <- c(color.grp.pal, "plum")

#
#Prep the PCA data
pca.h3k18la <- PCAtools::pca(h3k18la.dge.filt$logFPKM, metadata = meta)
# pca.h3k18la$metadata$group <- factor(pca.h3k18la$metadata$group, levels = unique(pca.h3k18la$metadata$group))

scree <- screeplot(pca.h3k18la)
cowplot::ggsave2(filename = "../data/processed/03_PCA_screeplot.pdf", plot = scree,
                 width = 8, height = 6)


eigcorp <- eigencorplot(pca.h3k18la, metavars = colnames(pca.h3k18la$metadata)[2:6])
pdf(file = "../data/processed/03_PCA_eigenplot.pdf",
                 width = 8, height = 6, useDingbats = F)
print(eigcorp)
dev.off()
# sampkey <- color.grp.pal
# names(sampkey) <- unique(pca.wind$metadata$Group)[c(2:3,1)]
# 
# 
# #Plot rb samples
# rb.dge <- wind.dge[,grep("normal", wind.dge$samples$group, invert = T)]
# rb.dge$logCPM <- rb.dge$logCPM[,grep("normal", wind.dge$samples$group, invert = T)]
# pca.rb <- PCAtools::pca(rb.dge$logCPM, metadata = data.frame(Group = rb.dge$samples$group, row.names = row.names(rb.dge$samples)))

X='PC1'; Y='PC2'
pca1 <- biplot(pca.h3k18la, x = X, y = Y,
       colby = "Group", #colkey = sampkey,
       #lab = "replicate", labSize = 2,
       legendPosition = 'right', 
       title = 'Peak Quant PCA',
       subtitle = paste0(X, ' versus ', Y))

cowplot::ggsave2(filename = "../data/processed/03_PCA_PC1-2_group.pdf", plot = pca1,
                 width = 8, height = 6)

X='PC1'; Y='PC3'
pca1 <- biplot(pca.h3k18la, x = X, y = Y,
       colby = "Group", #colkey = sampkey,
       #lab = "replicate", labSize = 2,
       legendPosition = 'right', 
       title = 'Peak Quant PCA',
       subtitle = paste0(X, ' versus ', Y))

cowplot::ggsave2(filename = "../data/processed/03_PCA_PC1-2_group.pdf", plot = pca1,
                 width = 8, height = 6)

X='PC1'; Y='PC2'
pca1 <- biplot(pca.h3k18la, x = X, y = Y,
       colby = "Run", #colkey = sampkey,
       #lab = "replicate", labSize = 2,
       legendPosition = 'right', 
       title = 'Peak Quant PCA',
       subtitle = paste0(X, ' versus ', Y))

cowplot::ggsave2(filename = "../data/processed/03_PCA_PC1-2_run.pdf", plot = pca1,
                 width = 8, height = 6)

# # PCA of RNA-seq - Nov 18, 2024
# tmp_samps <-colnames(fsprwt.dwnld)[c(12:13,17:19,22:25)]
# tmp_rna_cpm <- log2(fsprwt.dwnld[,tmp_samps] + 1)
# tmp_rna_meta <- data.frame(sample = gsub("\\.FS.+WT", "", tmp_samps),
#                            group = gsub("(.+)\\.FS.+", "\\1", tmp_samps),
#                            rep = gsub(".+WT\\.(.+)", "\\1", tmp_samps))
# row.names(tmp_rna_meta) <- tmp_samps
# tmp_rna_meta$group <- factor(tmp_rna_meta$group, levels = c("P4", "P10", "P28"))
# 
# pca.rna <- PCAtools::pca(tmp_rna_cpm, metadata = tmp_rna_meta)
# 
# X='PC1'; Y='PC2'
# pca2 <- biplot(pca.rna, x = X, y = Y,
#        colby = "group", #colkey = sampkey,
#        #lab = "replicate", labSize = 2,
#        legendPosition = 'right', 
#        title = 'Peak Quant PCA',
#        subtitle = paste0(X, ' versus ', Y))
# 
# cowplot::ggsave2(filename = "../data/processed/03_PCA_PC1-2_RNA.pdf", plot = pca2,
#                  width = 8, height = 6)

```


################################################################################
#### Differential binding analysis
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(edgeR)
library(rtracklayer)
library(qvalue)

source(file.path(dest_tools_dir, "VolcanoPlot.R"))
# y <- asDGEList(filtered.data)
# y <- y[,which(meta_used$stage == "blastoma")]
# colnames(y) <- meta_used$sample_name[which(meta_used$stage == "blastoma")]
# 
# stage_treat <- factor(paste(meta_used$stage[which(meta_used$stage == "blastoma")],
#                      meta_used$treatment[which(meta_used$stage == "blastoma")], sep = "_"))
# stage_treat[8] <- "blastoma_N"
y <- h3k18la.dge.filt

design <- model.matrix(~0 + h3k18la.dge.filt$samples$group)
colnames(design) <- levels(h3k18la.dge.filt$samples$group)

y <- estimateDisp(y, design)
summary(y$trended.dispersion)

fit <- glmQLFit(y, design, robust=TRUE)
summary(fit$var.post)

par(mfrow=c(1,2))
o <- order(y$AveLogCPM)
plot(y$AveLogCPM[o], sqrt(y$trended.dispersion[o]), type="l", lwd=2,
ylim=c(0, 1), xlab=expression("Ave."~Log[2]~"CPM"),
ylab=("Biological coefficient of variation"))
plotQLDisp(fit)

summary(fit$df.prior)

LA_2DG <- glmQLFTest(fit, contrast=c(-1,1,0,0,0,0))
LA_SO <- glmQLFTest(fit, contrast=c(-1,0,1,0,0,0))
AC_2DG <- glmQLFTest(fit, contrast=c(0,0,0,-1,1,0))
AC_SO <- glmQLFTest(fit, contrast=c(0,0,0,-1,0,1))

# Export Results

# LA_2DG
LA_2DG$table$qvalue <- qvalue(LA_2DG$table$PValue)$qvalue
LA_2DG$table$FDR <- p.adjust(LA_2DG$table$PValue, method = "fdr")
# crx.db$table$lfdr <- qvalue::qvalue(crx.db$table$PValue, fdr.level = 0.05)$significant
expt <- cbind(LA_2DG$genes[,-c(7:11)], LA_2DG$table, y$FPKM_mean)
write.table(expt, file = "../data/processed/03_H3k18la.db.LA_2DG.tsv", sep = "\t", quote = F, row.names = F)











# LA_2DG
res_LA_2DG <- topTags(LA_2DG, n = Inf, sort.by = "logFC")$table %>% 
  as.tibble() %>% 
  rowwise() %>% 
  mutate(label_gene = ifelse(FDR < 0.01, SYMBOL, NA))
 
res_LA_2DG_promPC <- res_LA_2DG %>% dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")

vp <- volPlot(res_LA_2DG_promPC, FC.nm = "logFC", FDR.nm = "FDR", fc = 1, fdr = 0.05, sym.lab = FALSE,
                     expt.nm = "03_H3K18LA_2DG", res.path = "../data/processed/", ret = T)

vp <- vp +
   geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) 

cowplot::ggsave2(filename = "../data/processed/03_H3K18LA_2DG_vol.pdf", plot = vp,
                 width = 8, height = 6)


# LA_SO
res_LA_SO <- topTags(LA_SO, n = Inf, sort.by = "logFC")$table %>% 
  as.tibble() %>% 
  rowwise() %>% 
  mutate(label_gene = ifelse(FDR < 0.01, SYMBOL, NA))
 
res_LA_SO_promPC <- res_LA_SO %>% dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")

vp <- volPlot(res_LA_SO_promPC, FC.nm = "logFC", FDR.nm = "FDR", fc = 1, fdr = 0.05, sym.lab = FALSE,
                     expt.nm = "03_H3K18LA_SO", res.path = "../data/processed/", ret = T)

vp <- vp +
   geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) 

cowplot::ggsave2(filename = "../data/processed/03_H3K18LA_SO_vol.pdf", plot = vp,
                 width = 8, height = 6)


# AC_2DG
res_AC_2DG <- topTags(AC_2DG, n = Inf, sort.by = "logFC")$table %>% 
  as.tibble() %>% 
  rowwise() %>% 
  mutate(label_gene = ifelse(FDR < 0.002, SYMBOL, NA))
 
res_AC_2DG_promPC <- res_AC_2DG %>% dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")

vp <- volPlot(res_AC_2DG_promPC, FC.nm = "logFC", FDR.nm = "FDR", fc = 1, fdr = 0.05, sym.lab = FALSE,
                     expt.nm = "03_H3K27AC_2DG", res.path = "../data/processed/", ret = T)

vp <- vp +
   geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) 

cowplot::ggsave2(filename = "../data/processed/03_H3K27AC_2DG_vol.pdf", plot = vp,
                 width = 8, height = 6)


# AC_SO
res_AC_SO <- topTags(AC_SO, n = Inf, sort.by = "logFC")$table %>% 
  as.tibble() %>% 
  rowwise() %>% 
  mutate(label_gene = ifelse(FDR < 0.01, SYMBOL, NA))
 
res_AC_SO_promPC <- res_AC_SO %>% dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")

vp <- volPlot(res_AC_SO_promPC, FC.nm = "logFC", FDR.nm = "FDR", fc = 1, fdr = 0.05, sym.lab = FALSE,
                     expt.nm = "03_H3K27AC_SO", res.path = "../data/processed/", ret = T)

vp <- vp +
   geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) 

cowplot::ggsave2(filename = "../data/processed/03_H3K27AC_SO_vol.pdf", plot = vp,
                 width = 8, height = 6)

```




################################################################################
#### Make Heatmap of DB Prom PC genes
################################################################################
```{r}

db_goi <- rbind(res_LA_2DG_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame(),
            res_LA_SO_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame(),
            res_AC_2DG_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame(),
            res_AC_SO_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame())

LA_db_goi <- rbind(res_LA_2DG_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame(),
            res_LA_SO_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05,  annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame())

AC_db_goi <- rbind(res_AC_2DG_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame(),
            res_AC_SO_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame())

# Testing for sanity
AC_db_goi <- rbind(res_AC_SO_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame())


db_goi <- db_goi %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(!duplicated(peakID))

LA_db_goi <- LA_db_goi %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(!duplicated(peakID))

AC_db_goi <- AC_db_goi %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(!duplicated(peakID))

# Filter quant binding for sig DB peaks
dat <- cbind(h3k18la.dge.filt$genes, h3k18la.dge.filt$FPKM_mean) %>%
  as.data.frame() %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(peakID %in% db_goi$peakID)

LA_dat <- cbind(h3k18la.dge.filt$genes, h3k18la.dge.filt$FPKM_mean) %>%
  as.data.frame() %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(peakID %in% LA_db_goi$peakID)

AC_dat <- cbind(h3k18la.dge.filt$genes, h3k18la.dge.filt$FPKM_mean) %>%
  as.data.frame() %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(peakID %in% AC_db_goi$peakID)

# Tidy the table
dat.tidy <- dat %>%
  dplyr::select(SYMBOL, peakID, H3K18LA_CNTRL, H3K18LA_2DG, H3K18LA_SO, H3K27AC_CNTRL, H3K27AC_2DG, H3K27AC_SO) %>%
  gather(Sample, RPKM, -SYMBOL, -peakID)

dat.tidy$RPKM <- log2(dat.tidy$RPKM + 1)
dat.tidy$Sample <- factor(dat.tidy$Sample, levels = colnames(dat)[20:25])

# LA
LA_dat.tidy <- LA_dat %>%
  dplyr::select(SYMBOL, peakID, H3K18LA_CNTRL, H3K18LA_2DG, H3K18LA_SO) %>%
  gather(Sample, RPKM, -SYMBOL, -peakID)

LA_dat.tidy$RPKM <- log2(LA_dat.tidy$RPKM + 1)
LA_dat.tidy$Sample <- factor(LA_dat.tidy$Sample, levels = colnames(dat)[20:22])

#AC
AC_dat.tidy <- AC_dat %>%
  dplyr::select(SYMBOL, peakID, H3K27AC_CNTRL, H3K27AC_2DG, H3K27AC_SO) %>%
  gather(Sample, RPKM, -SYMBOL, -peakID)

AC_dat.tidy$RPKM <- log2(AC_dat.tidy$RPKM + 1)
AC_dat.tidy$Sample <- factor(AC_dat.tidy$Sample, levels = colnames(dat)[23:25])

###################
# Sort by Heirarchical clustering using Ward's method with Euclidean distanced
distfunc <- function(x) dist(x,method="euclidean")
hclustfunc <- function(x) hclust(x, method="ward.D2")

#Get distance and heirarchical clustering
d <- distfunc(as.matrix(dat[,20:22]))
hc <- hclustfunc(d)

d <- distfunc(as.matrix(AC_dat[,23:25]))
hc <- hclustfunc(d)

###################
# Sort by PCA
pca <- prcomp(dat[,20:22], center = T, scale. = F)
pca.order <- order(pca$x[,1])

LA_pca <- prcomp(LA_dat[,20:22], center = T, scale. = F)
LA_pca.order <- order(LA_pca$x[,1])

AC_pca <- prcomp(AC_dat[,23:25], center = T, scale. = F)
AC_pca.order <- order(AC_pca$x[,1])


###################################################################
# Make Heatmap
###################################################################



# Make heatmap
library(scales)
lim = c(0,6)
nm = "RPKM (log2)"
mb.colors <- colorRampPalette(c("#023858","#3690c0", "#e0f3f8","#fff7bc","#fec44f","#ec7014",
                                    "#fc4e2a","#bd0026","#800026"))(n=1000)

# Sort by HC clustering
tmp_dat <- dat.tidy
tmp_dat$peakID <- factor(tmp_dat$peakID,
                                      levels = tmp_dat$peakID[hc$order], ordered = T)

pp <- ggplot(tmp_dat, aes(Sample, peakID)) +
        geom_tile(aes(fill = RPKM), colour = "transparent") +
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT_hm-HC.pdf", plot = pp,
                 width = 8, height = 6)


# Sort by PCA
tmp_dat <- dat.tidy
tmp_dat$peakID <- factor(tmp_dat$peakID,
                                      levels = tmp_dat$peakID[rev(pca.order)], ordered = T)

pp <- ggplot(tmp_dat, aes(Sample, peakID)) +
        geom_tile(aes(fill = RPKM), colour = "transparent") +
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT_hm-PCA.pdf", plot = pp,
                 width = 8, height = 6)

# LA PCA
tmp_dat <- LA_dat.tidy
tmp_dat$peakID <- factor(tmp_dat$peakID,
                                      levels = tmp_dat$peakID[rev(LA_pca.order)], ordered = T)

pp <- ggplot(tmp_dat, aes(Sample, peakID)) +
        geom_tile(aes(fill = RPKM), colour = "transparent") +
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT_LA_hm-PCA.pdf", plot = pp,
                 width = 8, height = 6)

# AC PCA
tmp_dat <- AC_dat.tidy
tmp_dat$peakID <- factor(tmp_dat$peakID,
                                      levels = tmp_dat$peakID[rev(AC_pca.order)], ordered = T)

pp <- ggplot(tmp_dat, aes(Sample, peakID)) +
        geom_tile(aes(fill = RPKM), colour = "transparent") +
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT_AC_hm-PCA.pdf", plot = pp,
                 width = 8, height = 6)


```



```{r}
library(ggVennDiagram)

################
# 2DG and SO in LA data

venn_goi <- list(DG = res_LA_2DG_promPC %>% 
                   dplyr::filter(abs(logFC) > 1, FDR < 0.05) %>% 
                   dplyr::select(SYMBOL) %>% 
                   unique() %>% unlist() %>% as.vector(),
                 TC = db_tc_sig_PromPC$SYMBOL %>% unique(),
                 SO = res_LA_SO_promPC %>% 
                   dplyr::filter(abs(logFC) > 1, FDR < 0.05) %>% 
                   dplyr::select(SYMBOL) %>% 
                   unique() %>% unlist() %>% as.vector())

# v <- ggVennDiagram(venn_goi, label = "count") +
#   ggtitle(label = "H3K18La DB Analysis of Promoter Peaks")

# ggvenn package
library(ggvenn)

v2 <- ggplot(venn_goi %>% list_to_data_frame(), aes(A = DG, B = SO, C = TC)) +
      geom_venn(show_percentage = F) +
      theme_void() + # Removes background and axes
      coord_fixed() # Ensures circles are circular

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_DG-SO-TC_venn.pdf", plot = v2,
                 width = 4, height = 3)
```


################################################################################
#### FGE of 2DG and SO expts overlap with Timecourse data
################################################################################
```{r}

library(clusterProfiler)
library(enrichplot)
library(org.Mm.eg.db)
source(file.path(dest_tools_dir, "GO_leaf.R"))
# source("../src/tools/Dotplot_GO.R")

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 10

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_enrich <- file.path(dir_export, paste0("03_DG-SO-TC_overlap_FuncGeneEnrich_", enrich_db, "_", GOont))
dir.create(dir_enrich)


# Prep the gene lists for the FGE loop
goi_list <- list(DG = vg %>% dplyr::filter(DG & TC) %>% 
                   dplyr::select(key) %>% 
                   unlist() %>% as.vector(),
                 SO = vg %>% dplyr::filter(SO & TC) %>% 
                   dplyr::select(key) %>% 
                   unlist() %>% as.vector())

univ <- h3k18la.dge.filt$genes %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
  dplyr::select(SYMBOL) %>% 
  unique() %>% unlist() %>% as.vector()

enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list, fun = "enrichGO",
                                            keyType = "SYMBOL", ont = GOont, OrgDb = org, universe = univ)


# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>%
  as.data.frame() %>%
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>%
  rowwise() %>%
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_LA_DG-SO-TC_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>%
  as.data.frame() %>%
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>%
  rowwise() %>%
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_LA_DG-SO-TC_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>%
  group_by(Cluster) %>%
  arrange(p.adjust) %>%
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>%
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>%
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) +
    geom_point(aes(size = Ratio)) +
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("2DG and SO DB Peaks in Timecourse "))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("LA_DG-TC_FGE_top_", top, "_leaf_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5,
                 width = 6)
```



<!-- ################################################################################ -->
<!-- #### Merge CUT&RUN and RNAseq -->
<!-- ################################################################################ -->
<!-- ```{r} -->

<!-- ################################################################### -->
<!-- # Prep the peak DB data -->
<!-- ################################################################### -->

<!-- # Get highest expressing transcript at P28 -->
<!-- ret_trans_high_dups <- retina_trans_cpm %>%  -->
<!--   rownames_to_column("ensembl_transcript_id") %>%  -->
<!--   group_by(ensembl_gene_id) %>%  -->
<!--   arrange(desc(P28)) %>%  -->
<!--   top_n(1) -->

<!-- # Get sig DB promoter peaks from protein coding genes -->
<!-- prom_pc_sig <- h3k18la.db.results %>% dplyr::filter(annotation == "Promoter", -->
<!--                                                      transcriptBiotype == "protein_coding", -->
<!--                                                      FDR < 0.01) -->

<!-- # Get gene with multiple promoter peaks -->
<!-- peak_dup_geneID <-  prom_pc_sig[duplicated(prom_pc_sig$geneId), 'geneId'] %>% unique() -->

<!-- # Subset promoter peaks for genes with a single promoter peak -->
<!-- prom_pc_uniq <- prom_pc_sig %>% dplyr::filter(!(geneId %in% peak_dup_geneID)) -->

<!-- # Subset promoter peaks for genes with multiple promoter peaks -->
<!-- prom_pc_dup <- prom_pc_sig %>%  -->
<!--   dplyr::filter(geneId %in% peak_dup_geneID) -->

<!-- # Select promoter peaks for highest transcript expression at P28 and secondarily by highest F score -->
<!-- prom_pc_dup_trans <- prom_pc_dup %>%  -->
<!--   dplyr::filter(transcriptId %in% ret_trans_high_dups$ensembl_transcript_id) %>%  -->
<!--   group_by(transcriptId) %>%  -->
<!--   arrange(desc(F)) %>%  -->
<!--   top_n(1) -->

<!-- # Merge the single promoter peak with redundancy removed peaks -->
<!-- prom_pc_sig_filt <- rbind(prom_pc_uniq, prom_pc_dup_trans) %>%  -->
<!--   arrange(seqnames, start) -->


<!-- ################################################################### -->
<!-- # Prep tidy table and add RNAseq -->
<!-- ################################################################### -->

<!-- ################### -->
<!-- # Tidy table -->
<!-- dat.tidy <- prom_pc_sig_filt %>%  -->
<!--   dplyr::select(SYMBOL, P4, P10, P28) %>% -->
<!--   gather(Sample, CPM, -SYMBOL) -->
<!-- # dat.tidy$CPM <- log2(dat.tidy$CPM + 1) -->


<!-- ################### -->
<!-- # Get RNA data for same genes -->
<!-- dat.tidy.rna <- retina_cpm %>%  -->
<!--   dplyr::select(external_gene_name, P4, P10, P28) %>% -->
<!--   dplyr::rename(SYMBOL = external_gene_name, -->
<!--                 P4_RNA = P4, -->
<!--                 P10_RNA = P10, -->
<!--                 P28_RNA = P28) %>%  -->
<!--   gather(Sample, CPM, -SYMBOL) -->

<!-- ################### -->
<!-- #Factor sort sample names -->
<!-- dat.tidy <- rbind(dat.tidy, dat.tidy.rna) -->
<!-- dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample)) -->


<!-- ################################################################### -->
<!-- # Cluster the peak DB -->
<!-- ################################################################### -->

<!-- ################### -->
<!-- # Sort by Heirarchical clustering using Ward's method with Euclidean distanced -->
<!-- distfunc <- function(x) dist(x,method="euclidean") -->
<!-- hclustfunc <- function(x) hclust(x, method="ward.D2") -->

<!-- #Get distance and heirarchical clustering -->
<!-- d <- distfunc(as.matrix(prom_pc_sig_filt[,27:29])) -->
<!-- hc <- hclustfunc(d) -->


<!-- ################### -->
<!-- # Sort by PCA -->
<!-- pca <- prcomp(prom_pc_sig_filt[,27:29], center = T, scale. = F) -->
<!-- pca.order <- order(pca$x[,1]) -->


<!-- ################################################################### -->
<!-- # Make Heatmap -->
<!-- ################################################################### -->

<!-- # Sort by HC clustering -->
<!-- dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL,  -->
<!--                                       levels = dat.tidy$SYMBOL[hc$order], ordered = T) -->

<!-- # Sort by PCA -->
<!-- dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL,  -->
<!--                                       levels = dat.tidy$SYMBOL[rev(pca.order)], ordered = T) -->

<!-- # Make heatmap -->
<!-- library(scales) -->
<!-- lim = c(0,16) -->
<!-- nm = "CPM (log2)" -->
<!-- mb.colors <- colorRampPalette(c("#023858","#3690c0", "#e0f3f8","#fff7bc","#fec44f","#ec7014", -->
<!--                                     "#fc4e2a","#bd0026","#800026"))(n=1000) -->
<!-- pp <- ggplot(dat.tidy, aes(Sample, SYMBOL)) + -->
<!--         geom_tile(aes(fill = CPM), colour = "transparent") +  -->
<!--         #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") + -->
<!--         scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) + -->
<!--         theme(axis.text.x = element_text(angle = 45, hjust = 1), -->
<!--               axis.title.x = element_blank(), axis.title.y = element_blank(), -->
<!--               panel.background = element_rect(fill = NA)) -->

<!-- cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT-RNA_hm-pca.pdf", plot = pp, -->
<!--                  width = 8, height = 6) -->
<!-- ``` -->


<!-- # Correlate the C&T and RNA seq -->
<!-- ```{r} -->
<!-- library(cowplot) -->

<!-- rna_tmp <- retina_cpm %>%  -->
<!--   rownames_to_column(var = "ENSEMBL") %>%  -->
<!--   dplyr::select(c("ENSEMBL", "external_gene_name", "P4", "P10", "P28")) %>%  -->
<!--   group_by(external_gene_name) %>%  -->
<!--   arrange(desc(P28)) %>%  -->
<!--   slice_head(n=1) -->

<!-- prom_pc_sig_filt_rna <- merge(prom_pc_sig_filt, rna_tmp, by.x = "SYMBOL", by.y = "external_gene_name") -->

<!-- prom_pc_sig_filt_rna[,c(27:29)] <- log2(prom_pc_sig_filt_rna[,c(27:29)] + 1) -->

<!-- # Find Person correlation between the H3K28La peak quant and RNAseq expression -->



<!-- prom_pc_sig_filt_rna <- prom_pc_sig_filt_rna %>%  -->
<!--   rowwise() %>%  -->
<!--   mutate(pearson = cor(x=c(P4.x, P10.x, P28.x), y = c(P4.y, P10.y, P28.y), method = "pearson")) -->

<!-- a <- ggplot(prom_pc_sig_filt_rna, aes(x=pearson)) + -->
<!--   geom_histogram(binwidth = 0.01) -->

<!-- cowplot::ggsave2(filename = paste0("../data/processed/03_Prom_pc_sigDE-RNA_corr-dist.pdf"), -->
<!--                  plot = a, -->
<!--                  width = 6, height = 4) -->

<!-- # Split by correlated and anticorrelated -->
<!-- cor_sig = 0.8 -->
<!-- posi_cor <- prom_pc_sig_filt_rna %>%  -->
<!--   dplyr::filter(pearson > cor_sig) -->

<!-- anti_cor <- prom_pc_sig_filt_rna %>%  -->
<!--   dplyr::filter(pearson < -cor_sig) -->


<!-- ###################### -->
<!-- # Plot the positive correlated peak and rna data -->
<!-- pca <- prcomp(posi_cor[,c(27:29, 31:33)], center = T, scale. = F) -->
<!-- pca.order <- order(pca$x[,1]) -->
<!-- # pca.order <- order(pca$x[,1] * pca$x[,2]) -->

<!-- d <- distfunc(as.matrix(posi_cor[,c(27:29, 31:33)])) -->
<!-- hc <- hclustfunc(d) -->


<!-- dat.tidy <- posi_cor %>%  -->
<!--   dplyr::select(SYMBOL, P4.x, P10.x, P28.x, P4.y, P10.y, P28.y) %>% -->
<!--   gather(Sample, CPM, -SYMBOL) -->

<!-- dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample)) -->

<!-- dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL,  -->
<!--                                       levels = dat.tidy$SYMBOL[(pca.order)], ordered = T) -->

<!-- dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL,  -->
<!--                                       levels = dat.tidy$SYMBOL[hc$order], ordered = T) -->

<!-- lim1 <- c(0,8) -->
<!-- p <- ggplot(dat.tidy %>% dplyr::filter(Sample %in% c("P4.x", "P10.x", "P28.x")),  -->
<!--             aes(Sample, SYMBOL)) + -->
<!--         geom_tile(aes(fill = CPM), colour = "transparent") +  -->
<!--         scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim1, oob=squish) + -->
<!--         # geom_vline(xintercept = 3.5) + -->
<!--         theme(axis.text.x = element_text(angle = 45, hjust = 1), -->
<!--               axis.title.x = element_blank(), axis.title.y = element_blank(), -->
<!--               panel.background = element_rect(fill = NA)) -->


<!-- lim2 <- c(0,16) -->
<!-- pp <- ggplot(dat.tidy %>% dplyr::filter(Sample %in% c("P4.y", "P10.y", "P28.y")),  -->
<!--              aes(Sample, SYMBOL)) + -->
<!--         geom_tile(aes(fill = CPM), colour = "transparent") +  -->
<!--         scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim2, oob=squish) + -->
<!--         # geom_vline(xintercept = 3.5) + -->
<!--         theme(axis.text.x = element_text(angle = 45, hjust = 1), -->
<!--               axis.title.x = element_blank(), axis.title.y = element_blank(), -->
<!--               panel.background = element_rect(fill = NA)) -->

<!-- pg <- plot_grid(p, pp) -->

<!-- cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT-RNA_posi-cor_hm-hc2.pdf", plot = pg, -->
<!--                  width = 8, height = 6) -->


<!-- pcacor <- biplot(PCAtools::pca(posi_cor[,c(27:29, 31:33)]), x = X, y = Y, -->
<!--        # colby = "group", #colkey = sampkey, -->
<!--        #lab = "replicate", labSize = 2, -->
<!--        legendPosition = 'right',  -->
<!--        title = 'Peak Quant PCA', -->
<!--        subtitle = paste0(X, ' versus ', Y)) -->








<!-- ###################### -->
<!-- # Plot the negative correlated peak and rna data -->
<!-- pca <- prcomp(anti_cor[,c(27:29, 31:33)], center = T, scale. = F) -->
<!-- pca.order <- order(pca$x[,1]) -->

<!-- dat.tidy <- anti_cor %>%  -->
<!--   dplyr::select(SYMBOL, P4.x, P10.x, P28.x, P4.y, P10.y, P28.y) %>% -->
<!--   gather(Sample, CPM, -SYMBOL) -->

<!-- dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample)) -->
<!-- dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL,  -->
<!--                                       levels = dat.tidy$SYMBOL[rev(pca.order)], ordered = T) -->

<!-- lim1 <- c(0,8) -->
<!-- p <- ggplot(dat.tidy %>% dplyr::filter(Sample %in% c("P4.x", "P10.x", "P28.x")),  -->
<!--             aes(Sample, SYMBOL)) + -->
<!--         geom_tile(aes(fill = CPM), colour = "transparent") +  -->
<!--         scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim1, oob=squish) + -->
<!--         # geom_vline(xintercept = 3.5) + -->
<!--         theme(axis.text.x = element_text(angle = 45, hjust = 1), -->
<!--               axis.title.x = element_blank(), axis.title.y = element_blank(), -->
<!--               panel.background = element_rect(fill = NA)) -->


<!-- lim2 <- c(0,16) -->
<!-- pp <- ggplot(dat.tidy %>% dplyr::filter(Sample %in% c("P4.y", "P10.y", "P28.y")),  -->
<!--              aes(Sample, SYMBOL)) + -->
<!--         geom_tile(aes(fill = CPM), colour = "transparent") +  -->
<!--         scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim2, oob=squish) + -->
<!--         # geom_vline(xintercept = 3.5) + -->
<!--         theme(axis.text.x = element_text(angle = 45, hjust = 1), -->
<!--               axis.title.x = element_blank(), axis.title.y = element_blank(), -->
<!--               panel.background = element_rect(fill = NA)) -->

<!-- pg <- plot_grid(p, pp) -->

<!-- cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT-RNA_anti-cor_hm-pca2.pdf", plot = pg, -->
<!--                  width = 8, height = 6) -->

<!-- ``` -->


################################################################################
#### FGE of 2DG and SO expts
################################################################################
```{r}

library(clusterProfiler)
library(enrichplot)
library(org.Mm.eg.db)
source(file.path(dest_tools_dir, "GO_leaf.R"))
# source("../src/tools/Dotplot_GO.R")

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 10

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_enrich <- file.path(dir_export, paste0("03_PromPC_v2_FuncGeneEnrich_", enrich_db, "_", GOont))
dir.create(dir_enrich)


# Prep the gene lists for the FGE loop
goi_list <- list(LA_2DG = res_LA_2DG_promPC %>% 
                   dplyr::filter(abs(logFC) > 1, FDR < 0.05) %>% 
                   dplyr::select(SYMBOL) %>% 
                   unique() %>% unlist() %>% as.vector(),
                 LA_SO = res_LA_SO_promPC %>% 
                   dplyr::filter(abs(logFC) > 1, FDR < 0.05) %>% 
                   dplyr::select(SYMBOL) %>% 
                   unique() %>% unlist() %>% as.vector(),
                 AC_2DG = res_AC_2DG_promPC %>% 
                   dplyr::filter(abs(logFC) > 1, FDR < 0.05) %>% 
                   dplyr::select(SYMBOL) %>% 
                   unique() %>% unlist() %>% as.vector(),
                 AC_SO = res_AC_SO_promPC %>% 
                   dplyr::filter(abs(logFC) > 1, FDR < 0.05) %>% 
                   dplyr::select(SYMBOL) %>% 
                   unique() %>% unlist() %>% as.vector())

univ <- h3k18la.dge.filt$genes %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
  dplyr::select(SYMBOL) %>% 
  unique() %>% unlist() %>% as.vector()

enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list, fun = "enrichGO",
                                            keyType = "SYMBOL", ont = GOont, OrgDb = org, universe = univ)


# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>%
  as.data.frame() %>%
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>%
  rowwise() %>%
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_LA-AC_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>%
  as.data.frame() %>%
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>%
  rowwise() %>%
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_LA-AC_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>%
  group_by(Cluster) %>%
  arrange(p.adjust) %>%
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>%
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>%
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) +
    geom_point(aes(size = Ratio)) +
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("H3K18la and H3K27ac Peaks in 2DG and SO explants"))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("LA-AC_FGE_top_", top, "_leaf_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5,
                 width = 6)
```



################################################################################
#### Make heatmaps of DB C&T
################################################################################
```{r}

# Prom PC genes
###################
# Tidy table
dat.tidy <- log2(h3k18la.dge.filt$FPKM + 1) %>%
  as.data.frame() %>%
  rownames_to_column(var = "peak_Id") %>%
  dplyr::filter(peak_Id %in% (row.names(prom_pc_sig) %>% as.numeric())) %>%
  # dplyr::select(peak_Id, P4, P10, P28) %>%
  gather(Sample, CPM, -peak_Id)

###################
#Factor sort sample names
dat.tidy$Sample <- gsub("_R", "_", dat.tidy$Sample)
dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample))


###################################################################
# Make Heatmap
###################################################################

# Sort by Heirarchical clustering using Ward's method with Euclidean distanced
distfunc <- function(x) dist(x,method="euclidean")
hclustfunc <- function(x) hclust(x, method="ward.D2")

#Get distance and heirarchical clustering
dat_2clst <- log2(h3k18la.dge.filt$FPKM + 1) %>%
  as.data.frame() %>%
  rownames_to_column(var = "peak_Id") %>%
  dplyr::filter(peak_Id %in% (row.names(prom_pc_sig) %>% as.numeric()))

d <- distfunc(as.matrix(dat_2clst[,4:7]))
hc <- hclustfunc(d)

dat.tidy$peak_Id <- factor(dat.tidy$peak_Id,
                                      levels = dat.tidy$peak_Id[hc$order], ordered = T)

# Make heatmap
library(scales)
lim = c(0,8)
nm = "FPKM (log2)"
mb.colors <- colorRampPalette(c("#023858","#3690c0", "#e0f3f8","#fff7bc","#fec44f","#ec7014",
                                    "#fc4e2a","#bd0026","#800026"))(n=1000)
pp <- ggplot(dat.tidy, aes(Sample, peak_Id)) +
        geom_tile(aes(fill = CPM), colour = "transparent") +
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        geom_vline(xintercept = c(2.5,4.5)) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT_hm-hc.pdf", plot = pp,
                 width = 6, height = 6)


###############################################
# # Sort by PCA
pca <- prcomp(dat_2clst[,2:8], center = T, scale. = F)
pca.order <- order(pca$x[,1])

dat.tidy$peak_Id <- factor(dat.tidy$peak_Id,
                                      levels = dat.tidy$peak_Id[rev(pca.order)], ordered = T)

# Make heatmap
library(scales)
lim = c(0,8)
nm = "FPKM (log2)"
mb.colors <- colorRampPalette(c("#023858","#3690c0", "#e0f3f8","#fff7bc","#fec44f","#ec7014",
                                    "#fc4e2a","#bd0026","#800026"))(n=1000)
pp <- ggplot(dat.tidy, aes(Sample, peak_Id)) +
        geom_tile(aes(fill = CPM), colour = "transparent") +
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        geom_vline(xintercept = c(2.5,4.5)) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT_hm-pca.pdf", plot = pp,
                 width = 6, height = 6)

```


<!-- ################################################################################ -->
<!-- #### GSEA enrichment of targeted pathways  -->
<!-- ################################################################################ -->
<!-- ```{r} -->
<!-- library(KEGGREST) -->
<!-- library(org.Mm.eg.db) -->

<!-- orgdb <- org.Mm.eg.db::org.Mm.eg.db -->


<!-- ############################### -->
<!-- # Prep DB results -->

<!-- # Get sig DB promoter peaks from protein coding genes -->
<!-- prom_pc <- h3k18la.db.results %>% dplyr::filter(annotation == "Promoter", -->
<!--                                                      transcriptBiotype == "protein_coding") -->

<!-- # Get gene with multiple promoter peaks -->
<!-- peak_dup_geneID <-  prom_pc[duplicated(prom_pc$geneId), 'geneId'] %>% unique() -->

<!-- # Subset promoter peaks for genes with a single promoter peak -->
<!-- prom_pc_uniq <- prom_pc %>% dplyr::filter(!(geneId %in% peak_dup_geneID)) -->

<!-- # Subset promoter peaks for genes with multiple promoter peaks -->
<!-- prom_pc_dup <- prom_pc %>%  -->
<!--   dplyr::filter(geneId %in% peak_dup_geneID) -->

<!-- # Select promoter peaks for highest transcript expression at P28 and secondarily by highest F score -->
<!-- prom_pc_dup_trans <- prom_pc_dup %>%  -->
<!--   dplyr::filter(transcriptId %in% ret_trans_high_dups$ensembl_transcript_id) %>%  -->
<!--   group_by(SYMBOL) %>%  -->
<!--   arrange(desc(F)) %>%  -->
<!--   top_n(1) -->

<!-- # Merge the single promoter peak with redundancy removed peaks -->
<!-- prom_pc_filt <- rbind(prom_pc_uniq, prom_pc_dup_trans) %>%  -->
<!--   arrange(seqnames, start) -->

<!-- prom_pc_filt <- prom_pc_filt[!duplicated(prom_pc_filt$SYMBOL),] -->

<!-- # Make stat vector -->
<!-- tmp_stat <- prom_pc_filt %>% arrange(desc(logFC.P28)) -->
<!-- stat_vect <- tmp_stat$logFC.P28 -->
<!-- names(stat_vect) <- tmp_stat$SYMBOL -->

<!-- library(fgsea) -->
<!-- # FGSEA -->

<!-- ###################################################### -->
<!-- # Prep FGSEA -->

<!-- # List of KEGG IDs and their pathway names -->
<!-- kegg.path <- keggList("pathway") -->
<!-- kegg.ids <- gsub("map", "mmu", names(kegg.path)) -->
<!-- kegg.path <- unlist(lapply(1:length(kegg.path), function(x){kegg.path[[x]]})) -->

<!-- # Use specific pathway ID to get name -->
<!-- kegg.i <- c("mmu00010", "mmu00020", "mmu00030", "mmu00190", -->
<!--             "mmu04979", "mmu03410", "mmu03420", -->
<!--             "mmu04115", "mmu04064", "mmu04150", "mmu04310", -->
<!--             "mmu04744") -->
<!-- # kegg.p <- kegg.path[match(kegg.i, kegg.ids)] -->


<!-- path_goi <- list() -->

<!-- for (i in kegg.i){ -->

<!--   # Pathway name -->
<!--   kegg.p <- kegg.path[match(i, kegg.ids)] -->

<!--   # Get pathway gene symbols -->
<!--   kegg <- KEGGREST::keggGet(i)[[1]]$GENE -->
<!--   kegg.ez <- kegg[seq(1,length(kegg), by = 2)] -->
<!--   kegg.sym <- AnnotationDbi::select(orgdb, keys = kegg.ez, keytype = "ENTREZID", columns = "SYMBOL")[,2] -->

<!--   tmp_goi <- list(kegg.sym) -->
<!--   names(tmp_goi) <- kegg.p -->
<!--   path_goi <- c(path_goi, tmp_goi) -->

<!-- } -->

<!-- # # Pathway name -->
<!-- # kegg.p <- kegg.path[match(kegg.i, kegg.ids)] -->
<!-- #  -->
<!-- # # Get pathway gene symbols -->
<!-- # kegg <- KEGGREST::keggGet(kegg.i)[[1]]$GENE -->
<!-- # kegg.ez <- kegg[seq(1,length(kegg), by = 2)] -->
<!-- # kegg.sym <- AnnotationDbi::select(orgdb, keys = kegg.ez, keytype = "ENTREZID", columns = "SYMBOL")[,2] -->
<!-- #  -->
<!-- # path_goi <- list(kegg.sym) -->
<!-- # names(path_goi) <- kegg.p -->


<!-- ##################################################### -->
<!-- # Perform the enrichment analysis -->

<!-- res <- fgsea(pathways = path_goi, stats = stat_vect) -->


<!-- ##################################################### -->
<!-- # Plot the results -->
<!-- res_plot <- list() -->

<!-- for (i in 1:length(path_goi)){ -->
<!--   res_plot[[i]] <- plotEnrichment(pathway = path_goi[[i]], stats = stat_vect) + -->
<!--   ggtitle(label = res[i]$pathway, subtitle = paste0("FDR = ", round(res[i]$padj, 3), ", NES = ", round(res[i]$NES, 3))) + -->
<!--   labs(caption = paste0("Leading edge = ", res[i]$leadingEdge)) -->
<!-- } -->

<!-- # Export plot -->
<!-- for (i in 1:length(res_plot)){ -->
<!--   nm <- gsub(" ","-", res[i,]$pathway) -->
<!--   nm <- gsub("/", "", nm) -->
<!--   cowplot::ggsave2(filename = paste0("../data/processed/03_Prom_pc_GSEA_",  -->
<!--                                      nm, -->
<!--                                      ".pdf"),  -->
<!--                    plot = res_plot[[i]], -->
<!--                    width = 6, height = 4) -->
<!-- } -->





<!-- # getGSEAdata <- function (pathway, stats, gseaParam = 1, ticksSize = 0.2) -->
<!-- # { -->
<!-- #   rnk <- rank(-stats) -->
<!-- #     ord <- order(rnk) -->
<!-- #     statsAdj <- stats[ord] -->
<!-- #     statsAdj <- sign(statsAdj) * (abs(statsAdj)^gseaParam) -->
<!-- #     statsAdj <- statsAdj/max(abs(statsAdj)) -->
<!-- #     pathway <- unname(as.vector(na.omit(match(pathway, names(statsAdj))))) -->
<!-- #     pathway <- sort(pathway) -->
<!-- #     gseaRes <- calcGseaStat(statsAdj, selectedStats = pathway, -->
<!-- #         returnAllExtremes = TRUE) -->
<!-- #     bottoms <- gseaRes$bottoms -->
<!-- #     tops <- gseaRes$tops -->
<!-- #     n <- length(statsAdj) -->
<!-- #     xs <- as.vector(rbind(pathway - 1, pathway)) -->
<!-- #     ys <- as.vector(rbind(bottoms, tops)) -->
<!-- #     toPlot <- data.frame(x = c(0, xs, n + 1), y = c(0, ys, 0)) -->
<!-- #     diff <- (max(tops) - min(bottoms))/8 -->
<!-- #     x = y = NULL -->
<!-- #     return(toPlot) -->
<!-- # } -->
<!-- #  -->
<!-- # fgsea.lfc <- getGSEAdata(path_goi, stats = stat_vect) -->
<!-- #  -->
<!-- # enrich.fgsea.lfc <- ggplot(fgsea.lfc, aes(x = x, ymin = 0, ymax = y, y = y)) + -->
<!-- #   geom_ribbon(fill = "salmon", size = 1, alpha = 0.25) + -->
<!-- #   geom_line(color = "salmon", size = 1) + -->
<!-- #   geom_hline(yintercept = 0, size = 0.25, alpha = 0.8) + -->
<!-- #   theme_classic() + -->
<!-- #   theme(panel.border = element_blank(), panel.grid.minor = element_blank()) + -->
<!-- #   labs(x = "rank", y = "enrichment score") #+ -->
<!-- #   # ggtitle(paste0("FGSEA enrichment lfc\nProteostasis Network\nNES = ", -->
<!-- #   #                round(fg.lfc.path.pn$NES,2), -->
<!-- #   #                " ; Adj. p-value = ", -->
<!-- #                  # fg.lfc.path.pn$padj)) -->
<!-- ``` -->



<!-- ################################################################################ -->
<!-- ## RNAseq correlated Alluvial plot -->
<!-- ################################################################################ -->
<!-- ```{r} -->
<!-- library(ggalluvial) -->

<!-- # Load alluvial data.frame from 02 -->
<!-- df_alluv <- readRDS("../data/interim/02_df_alluvial.RDS") -->

<!-- # Correlated genes -->
<!-- df_alluv_corr <- df_alluv %>%  -->
<!--   dplyr::filter(Symbol %in% posi_cor$SYMBOL) -->

<!-- gga <- ggplot(df_alluv_corr, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) + -->
<!--   # scale_fill_brewer(type = "qual", palette = "Set2") + -->
<!--   # scale_fill_viridis_b() + -->
<!--   geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) + -->
<!--   geom_stratum(alpha = 0.75) + -->
<!--   theme(legend.position = "bottom") + -->
<!--   ggtitle("Histone occupancy in correlated promoter region during development")  -->

<!-- cowplot::ggsave2(gga, filename = file.path(dir_export, paste0("03_Alluvial_Corr_Prom_18la-27ac-4me3.pdf")), -->
<!--                  height = 8, width = 12) -->


<!-- # Anti-Correlated genes -->
<!-- df_alluv_anticorr <- df_alluv %>%  -->
<!--   dplyr::filter(Symbol %in% anti_cor$SYMBOL) -->

<!-- gga <- ggplot(df_alluv_anticorr, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) + -->
<!--   # scale_fill_brewer(type = "qual", palette = "Set2") + -->
<!--   # scale_fill_viridis_b() + -->
<!--   geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) + -->
<!--   geom_stratum(alpha = 0.75) + -->
<!--   theme(legend.position = "bottom") + -->
<!--   ggtitle("Histone occupancy in anti-correlated promoter region during development")  -->

<!-- cowplot::ggsave2(gga, filename = file.path(dir_export, paste0("03_Alluvial_AntiCorr_Prom_18la-27ac-4me3.pdf")), -->
<!--                  height = 8, width = 12) -->


<!-- # Un-Correlated genes -->
<!-- df_alluv_uncorr <- df_alluv %>%  -->
<!--   dplyr::filter(!(Symbol %in% c(posi_cor$SYMBOL, anti_cor$SYMBOL))) -->

<!-- gga <- ggplot(df_alluv_uncorr, aes(x = TP, stratum = Region, alluvium = Symbol, fill = Region, label = Region)) + -->
<!--   # scale_fill_brewer(type = "qual", palette = "Set2") + -->
<!--   # scale_fill_viridis_b() + -->
<!--   geom_flow(stat = "alluvium", lode.guidance = "rightleft", alpha = 0.2) + -->
<!--   geom_stratum(alpha = 0.75) + -->
<!--   theme(legend.position = "bottom") + -->
<!--   ggtitle("Histone occupancy in uncorrelated promoter region during development")  -->

<!-- cowplot::ggsave2(gga, filename = file.path(dir_export, paste0("03_Alluvial_UnCorr_Prom_18la-27ac-4me3.pdf")), -->
<!--                  height = 8, width = 12) -->






<!-- ``` -->



# ################################################################################
# ## Perform FGE on Alluvial segments
# ################################################################################
# ```{r}
# source("../src/tools/GO_leaf.R")
# 
# org = "org.Mm.eg.db"
# 
# enrich_db <- "GO_220310"
# GOont <- "BP"
# top <- 10
# 
# # dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
# dir_enrich <- file.path(dir_export, paste0("03_PromPC_FuncGeneEnrich", enrich_db, "_", GOont))
# dir.create(dir_enrich)
# 
# univ <- df_alluv_spread$Symbol
# 
# 
# enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list_alluv, fun = "enrichGO",
#                                           keyType = "SYMBOL", ont = GOont, OrgDb = org)
# 
# 
# # Filter enrichment for significance
# enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)
# 
# # Fix the Ratio
# enrich_results_filt_fix <- enrich_results_filt %>%
#   as.data.frame() %>%
#   mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>%
#   rowwise() %>%
#   mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))
# 
# # Export the FGE results table
# tmp <- enrich_results_filt_fix[,-11]
# # Export the enrichment analysis
# write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_full.tsv")),
#               quote = F, row.names = F, sep = "\t")
# 
# # Get leaf terms
# enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)
# 
# # Fix the Ratio
# enrich_results_leaf_fix <- enrich_results_leaf %>%
#   as.data.frame() %>%
#   mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>%
#   rowwise() %>%
#   mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))
# 
# # Export the FGE results table
# tmp <- enrich_results_leaf_fix[,-11]
# # Export the enrichment analysis
# write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_", i, "_leaf.tsv")),
#               quote = F, row.names = F, sep = "\t")
# 
# # Get top 5 pathways per cell type to get union set of pathways
# top_union <- enrich_results_leaf %>%
#   group_by(Cluster) %>%
#   arrange(p.adjust) %>%
#   slice_head(n = top)
# 
# 
# # Filter the full table for the union of top5 pathways
# top_enrich <- enrich_results_filt_fix %>%
#   dplyr::filter(ID %in% (top_union$ID %>% unique()))
# 
# # Prep the df for plotting
# top_enrich <- top_enrich %>%
#   arrange(Cluster, p.adjust)
# top_enrich$Description <- substring(top_enrich$Description, 1, 60)
# top_enrich$Description <- factor(top_enrich$Description, levels = rev(unique(top_enrich$Description)))
# top_enrich$p.adjust <- -log10(top_enrich$p.adjust)
# 
# # Make plot
# pp <- ggplot(top_enrich, aes(Cluster, Description, colour = p.adjust)) +
#     geom_point(aes(size = Ratio)) +
#     theme_bw() +
#     scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
#     guides(size=guide_legend(title="Ratio")) +
#     theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
#           axis.text.x = element_text(angle = 30, hjust = 1)) +
#     ggtitle(paste0("GO:BP in Histone Occupancy in Correlated peaks"))
# 
# #Export plot
# cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0(i, "_Alluvial-Region_FGE_top_", top, "_", i, "_leaf_01_dotplot.pdf")),
#                  height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5,
#                  width = 6)
# 
# # Make barplots for each section
# 
# for (i in top_union$Cluster){
#   tmp_dat <- top_union %>% dplyr::filter(Cluster == i)
#   tmp_dat$Description <- factor(tmp_dat$Description, levels = rev(tmp_dat$Description))
# 
#   bp <- ggplot(tmp_dat) +
#       geom_bar(stat = 'identity', aes(x=Description, y = Count, fill = -log10(p.adjust))) +
#       coord_flip() +
#       scale_fill_gradient(low="#3B9AB2", high="#F21A00", name="P Value \n(-log10)",limits=c(0,10), oob = scales::squish) +
#       #scale_fill_gradientn(colors = c("#3B9AB2", "#F21A00"), name="P Value (log10)", trans = 'reverse', limits=c(-50,0), oob=squish) +
#       labs(y="Number of Significant Genes", x=NULL) +
#       ggtitle(paste0("GO:BP - ", i)) +
#       theme_bw()
#   cowplot::ggsave2(plot = bp, filename = file.path(dir_enrich, paste0("Alluvial-Region_FGE_top_", top, "_", i, "_leaf_01_barplot.pdf")),
#                  height = 4,
#                  width = 8)
# }
# 










<!-- ``` -->


<!-- ################################################################################ -->
<!-- # Make expression plots for several genes - Oct 11, 2024 -->
<!-- ################################################################################ -->
<!-- ```{r} -->

<!-- goi <- c("Akt3", "Rorb", "Reep6", "Vamp2") -->

<!-- for (i in goi){ -->

<!--   # RNAseq -->
<!--   tmp_dat <- retina.dwnld %>% dplyr::filter(external_gene_name == i) %>% -->
<!--     dplyr::select(c(5, 22:23, 26:27, 30:31)) %>%  -->
<!--     gather("Sample", "CPM", -external_gene_name) %>%  -->
<!--     mutate(Group = gsub("(.+)\\.Retina.+", "\\1", Sample)) %>%  -->
<!--     rename(c(Symbol = "external_gene_name")) -->

<!--   tmp_dat$Group <- factor(tmp_dat$Group, levels = tmp_dat$Group %>% unique()) -->

<!--   p <- ggplot(tmp_dat, aes(x = Group, y = CPM)) + -->
<!--     geom_boxplot(fill = c("#fdbe85", "#fd8d3c", "#d94701")) + -->
<!--     ggtitle("RNAseq") -->

<!--   # CUT&Tag -->
<!--   tmp_dat2 <- h3k18la.dge$rpkm %>% as_tibble()  %>%  -->
<!--     mutate(annotation = h3k18la.dge$genes$annotation,  -->
<!--            transcriptBiotype = h3k18la.dge$genes$transcriptBiotype, -->
<!--            Symbol = h3k18la.dge$genes$SYMBOL) %>%  -->
<!--     dplyr::filter(Symbol == i,  -->
<!--                   annotation == "Promoter", -->
<!--                   transcriptBiotype == "protein_coding") -->
<!--   tmp_dat2 <- tmp_dat2 %>% -->
<!--     gather("Sample", "FPKM", -Symbol, -annotation, -transcriptBiotype) %>%  -->
<!--     mutate(Group = gsub("(.+)_R\\d", "\\1", Sample)) -->

<!--   tmp_dat2$Group <- factor(tmp_dat2$Group, levels = tmp_dat2$Group %>% unique()) -->

<!--   pp <- ggplot(tmp_dat2, aes(x = Group, y = FPKM)) + -->
<!--     geom_boxplot(fill = c("#fdbe85", "#fd8d3c", "#d94701")) + -->
<!--     ggtitle("CUT&Tag") -->

<!--   pg <- cowplot::plot_grid(pp, p) -->

<!--   cowplot::ggsave2(plot = pg, filename = file.path(dir_export, paste0("03_exp_boxplot_", i , ".pdf")), -->
<!--                    height = 4, width = 4) -->

<!-- } -->


<!-- ``` -->


# Export objects
```{r}

saveRDS(ensDb, file = "../data/interim/ensDb.RDS")
saveRDS(mstr_features_anno, file = "../data/interim/03_mstr_features_anno_240913.RDS")
saveRDS(peakList_cons, file = "../data/interim/03_peakList_cons_240913.RDS")
saveRDS(prom_pc_sig, file = "../data/interim/03_prom_pc_sig_240913.RDS")
saveRDS(h3k18la.db.results, file = "../data/interim/03_h3k18la_db_results_240913.RDS")
saveRDS(posi_cor, file = "../data/interim/03_h3k18la_db_dge_positive_corr_241001.RDS")
saveRDS(anti_cor, file = "../data/interim/03_h3k18la_db_dge_anti_corr_241001.RDS")


```