---
title: "05 - H3K18la Explant Analysis - Sept 27th, 2024"
output: html_notebook
---

### Project Summary
This project investigates the H3K18la during development. 

**DESIGN**:  

**NOTES**: The data were generated from the ENCODE histone pipeline with duplicate 
removal with no controls. This notebool uses peak quant based on CSAW normalization. 
Differential binding and correlation with RNA changes.

**UPDATE**: The definition of promoter (-1000 to +500) was used for peak annotation. 
Enhancer is defined as -10k to -1000 from TSS.

	<!-- • Figure 2 -->
	<!-- 	○ # of genes with peaks -->
	<!-- 	○ FGE with gene numbers at each definition (GO:leaf) -->
	<!-- 		§ Promoter: -500 to +250 -->
	<!-- 		§ Enhancer: -10K to -1kb -->
	<!-- 		§ Genebody -->
	<!-- 		§ Intergenic -->
	<!-- 	○ Heatmap -->
	<!-- 		§  -500 to 500 -->
	<!-- 		§ -10kb to 10kb -->
	<!-- 	○ RNA-seq correlation (wt retina development) -->
	<!-- • Figure 3 -->
	<!-- 	○ Corelate H3K18la -->
	<!-- 		§ ATAC -->
	<!-- 		§ Nrl -->
	<!-- 		§ RNA-seq (wt retina development -->
	<!-- 		§ ChromHMM -->
	<!-- • Figure 4 -->
	<!-- 	○ Glycolytic genes show with H3K27ac and H3K18la -->
	<!-- • Figure 5 -->
	<!-- 	○ Explant data CnR -->
	<!-- 		§ Use spike-in normalization to compare -->
	<!-- 	○ Correlate with RNA-seq (if data can be produced) -->
	
	
################################################################################
#### Get custom scripts and functions and outside data used for analysis
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

script_dir <- "https://raw.githubusercontent.com/MBrooks313/R_codes/master"
dest_tools_dir <- "../src/tools"
dest_vis_dir <- "../src/visualization"

# download.file(url = file.path(script_dir, "GO_leaf.R"),
#               destfile = file.path(dest_tools_dir, "GO_leaf.R"),
#               method = "curl")
# 
# download.file(url = file.path(script_dir, "DFave.R"),
#               destfile = file.path(dest_tools_dir, "DFave.R"),
#               method = "curl")
# 
# download.file(url = file.path(script_dir, "Heatmap.R"),
#               destfile = file.path(dest_tools_dir, "Heatmap.R"),
#               method = "curl")
# 
# download.file(url = file.path(script_dir, "Expr_Filter.R"),
#               destfile = file.path(dest_tools_dir, "Expr_Filter.R"),
#               method = "curl")
# system('cp ~/github/R_codes/DFave.R ../src/tools')
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/et_edgeR_v0.1.0.R ../src/tools")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/FuncGeneEnrich_v0.1.0.R ../src/tools")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/GO_leaf_v1.1.R ../src/tools")

#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/APk_cluster_BF_v0.1.0.R ../src/tools')
#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/HM_PCAsort_BF_v0.1.0.R ../src/visualization')
#system('cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/parPlot_hc_v0.2.0.R ../src/visualization')
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/PCA_simple_v0.1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/VolcanoPlot_v0.1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_v1.7.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Dotplot_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Barplot_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_GO_v1.0.R ../src/visualization")
# system("cp /Volumes/PEGASUS/Projects/BioInf_Scripts/R_functions/ggplots/Heatmap_v1.7.R ../src/visualization")
# 
# system("cp /Volumes/NGS/RNA-seq/ENS_v98/Mouse/Counts/by_project/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv ../data/external")
# system("cp /Volumes/NGS/RNA-seq/ENS_v98/Mouse/Counts/by_project/Mouse_FSPR_WT_Development_Gene_nCPM_ENSv98.tsv ../data/external")

# Function to convert ENS ids to gene symbol 
# ens2sym <- function(x){paste(sort(unique(AnnotationDbi::mapIds(org.Mm.eg.db, keys = x, keytype = "ENSEMBL", column = "SYMBOL"))), sep = "", collapse = ",")}
library(tidyverse)
dir_export <- "../data/processed/"
options(java.parameters = "-Xmx8000m")
print("Completed...")
```


################################################################################
#### Import outside data
################################################################################
```{r}
source("../src/tools/DFave.R")

invivo_peakAnnoList_ens <- readRDS("../data/interim/peakAnnoList_cons_240911.RDS")



# #############################################
# # Import expression data for flow-sorted rods
# fsprwt.dwnld <- read.csv("../data/external/Mouse_FSPR_WT_Development_Gene_nCPM_ENSv98.tsv",
#                          sep = "\t", row.names = 1)
# # Get group info for averaging
# grp <- gsub("(.+)\\.FS.+", "\\1", colnames(fsprwt.dwnld)[10:25])
# exp_grp <- factor(grp, levels = unique(grp))
#
# # Get average CPM values
# fsprwt_cpm <- data.frame(merge(fsprwt.dwnld[,1:9],
#                                log2(DFave(fsprwt.dwnld[,10:25], exp_grp) + 1),
#                                by = 0),
#                          row.names = 1)
#
# fsprwt_z <- data.frame(merge(fsprwt_cpm[,1:9],
#                                t(scale(t(fsprwt_cpm[,10:15]), scale = T, center = T)),
#                                by = 0),
#                          row.names = 1)
#
#
#
# #############################################
# # Import expression data for developing retina
retina.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)

# Get group info for averaging
grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.dwnld)[10:31])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
retina_cpm <- data.frame(merge(retina.dwnld[,1:9],
                               log2(DFave(retina.dwnld[,10:31], exp_grp) + 1),
                               by = 0),
                         row.names = 1)
#
# retina_z <- data.frame(merge(retina_cpm[,1:9],
#                                t(scale(t(retina_cpm[,10:19]), scale = T, center = T)),
#                                by = 0),
#                          row.names = 1)
#
#
#############################################
# Import transcript expression data for developing retina
retina.trans.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Trans_nRKPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)

# Get group info for averaging
grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.trans.dwnld)[12:33])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
retina_trans_cpm <- data.frame(merge(retina.trans.dwnld[,1:11],
                               log2(DFave(retina.trans.dwnld[,12:33], exp_grp) + 1),
                               by = 0),
                         row.names = 1)

# retina_trans_z <- data.frame(merge(retina_trans_cpm[,1:11],
#                                t(scale(t(retina_trans_cpm[,12:21]), scale = T, center = T)),
#                                by = 0),
#                          row.names = 1)


```


################################################################################
#### Get Ensembl annotation database
################################################################################
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

## Query for all available EnsDb databases
ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

# ensDb <- ahdb[[1, force=TRUE, verbose=FALSE]]
ensDb <- ahdb[[1]]
ensDb
# ensDb_pc <- addFilter(ensDb, GenebiotypeFilter("protein_coding", condition = "=="))
# ensDb_pc <- addFilter(ensDb, TxBiotypeFilter("protein_coding", condition = "=="))
# ensDb_pc_hard <- genes(ensDb_pc)
# ensDb_pc_hard <- ensembldb::filter(ensDb, filter = ~ tx_biotype == "protein_coding")

# supportedFilters(ensDb)
# ensDb <- readRDS("../data/interim/ensDb.RDS")
```


################################################################################
#### Get list of bam files and make master bed feature for quantitation
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(tidyverse)
library(ChIPpeakAnno)
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


# Macs2 beds
bam_dir <- "../data/raw/bams/"
macs_bams <- list.files(path = bam_dir, pattern = ".+5mM.+.bam$")
macs_bams <- macs_bams[c(4:6, 1:3, 9:10, 7:8)]

# Import consensus peaks from timecourse data
peakList_cons_tc <- readRDS(file = "../data/interim/03_peakList_cons_240913.RDS")
# peakList_cons <- peakList_cons[c(1:3)]

# Import master consensus
mstr_features_anno <- readRDS(file = "../data/interim/03_mstr_features_anno_240913.RDS")

# Import the timecourse DB results
h3k18la.db.results <- readRDS("../data/interim/03_h3k18la_db_results_240913.RDS")

# Import the timecourse prom pc filtered DB results
prom_pc_sig <- readRDS("../data/interim/03_prom_pc_sig_240913.RDS")

time_peakAnnoList_ens <- readRDS("../data/interim/peakAnnoList_cons_240911.RDS")
# mstr_features <- IRanges::reduce(c(peakList_cons[[1]],
#                          peakList_cons[[2]],
#                          peakList_cons[[3]]))

# # Annotate the features
# mstr_features_anno <- annotatePeak(mstr_features, tssRegion=c(-1000, 500),
#                          # TxDb=ensDb, annoDb="org.Mm.eg.db", addFlankGeneInfo = T,
#                          TxDb=ensDb, addFlankGeneInfo = T,
#                          overlap = "TSS")
# mstr_features_anno@anno$SYMBOL <- mapIds(x = ensDb, 
#                                          keys = mstr_features_anno@anno$geneId, 
#                                          column = "SYMBOL", 
#                                          keytype = "GENEID")
# 
# # Export bed file of the regions
# mstr_feature_bed <- data.frame(chrom = seqnames(mstr_features_anno@anno),
#                              start = start(mstr_features_anno@anno),
#                              end = end(mstr_features_anno@anno),
#                              name = mstr_features_anno@anno$SYMBOL,
#                              score = width(mstr_features_anno@anno),
#                              strand = strand(mstr_features_anno@anno))
# 
# write.table(mstr_feature_bed, file = "../data/processed/03_mstr_feature.bed", sep = "\t", quote = F, col.names = F, row.names = F)

```


################################################################################
## All explant samples
################################################################################

## Get metadata for samples
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
# ord <- c(1:4)

# Macs2 beds
# macs_beds <- list.files(path = "../data/interim/bed/explant/consensus_nf-core/", pattern = ".+MB.+bed", recursive = T)
macs_beds <- list.files(path = "../data/interim/bed/explant/macs2/", pattern = "narrowPeak", recursive = T)

meta <- data.frame(Beds = macs_beds,
                   Samp = gsub("(.+)\\.macs.+", "\\1", macs_beds),
                   Group = gsub("(.+)_R.+", "\\1", macs_beds))
meta <- meta[c(4:6, 1:3, 9:10, 7:8),]
# meta$Beds <- factor(meta$Beds, levels = levels(meta$Beds))
meta$Samp <- factor(meta$Samp, levels = unique(meta$Samp))
meta$Group <- factor(meta$Group, levels = unique(meta$Group))

# # Meta sheet
# meta <- data.frame(Beds = macs_beds,
#                    Samp = gsub("(.+)\\.consensus.+", "\\1", macs_beds))
# meta <- meta[c(2,1,4,3),]
# # meta$Beds <- factor(meta$Beds, levels = levels(meta$Beds))
# meta$Samp <- factor(meta$Samp, levels = unique(meta$Samp))


print("Meta data acquired...")
```


# Import the peaks into a merged list object and plot for number of peaks at each timepoint
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
library(ChIPpeakAnno)

macs_dir <- "../data/interim/bed/explant/macs2//"

# Import the cons peaks
peakList <- sapply(file.path(macs_dir, meta$Beds), toGRanges, format="narrowPeak")
names(peakList) <- meta$Samp

peakList_1E6 <- GRangesList()
# Sort the cons peaks
for (i in 1:length(peakList)){
  
  # seqlevelsStyle(peakList[[i]]) <- "NCBI"
  tmp_peak <- sortSeqlevels(peakList[[i]])
  tmp_peak <- GenomicRanges::sort(tmp_peak)
  tmp_peak <- tmp_peak[(elementMetadata(tmp_peak)[,"qValue"] > 6) & (seqnames(tmp_peak) != "MT")]
  peakList_1E6[[i]] <- tmp_peak
}
names(peakList_1E6) <- meta$Samp

# Overlap peaks per timepoint
# Get overlaps
ol_k18_5 <- findOverlapsOfPeaks(c(peakList_1E6[1],
                             peakList_1E6[2],
                             peakList_1E6[3]), connectedPeaks="keepAll", ignore.strand = T)

ol_k18_25 <- findOverlapsOfPeaks(c(peakList_1E6[4],
                             peakList_1E6[5],
                             peakList_1E6[6]), connectedPeaks="keepAll", ignore.strand = T)

ol_k27_5 <- findOverlapsOfPeaks(c(peakList_1E6[7],
                             peakList_1E6[8]), connectedPeaks="keepAll", ignore.strand = T)

ol_k27_25 <- findOverlapsOfPeaks(c(peakList_1E6[9],
                             peakList_1E6[10]), connectedPeaks="keepAll", ignore.strand = T)

peakList_cons <- GRangesList(ol_k18_5$mergedPeaks, ol_k18_25$mergedPeaks, ol_k27_5$mergedPeaks, ol_k27_25$mergedPeaks)
names(peakList_cons) <- c("K18_5", "K18_25", "K27_5", "K27_25")

# cons_dir <- "../data/interim/bed/explant/consensus_nf-core"
# 
# # Import the cons peaks
# peakList_cons <- sapply(file.path(cons_dir, meta$Beds), toGRanges, format="BED")
# names(peakList_cons) <- meta$Samp
# 
# # Sort the cons peaks
# for (i in 1:length(peakList_cons)){
#   
#   # seqlevelsStyle(peakList_cons[[i]]) <- "NCBI"
#   peakList_cons[i] <- sortSeqlevels(peakList_cons[[i]])
#   peakList_cons[i] <- GenomicRanges::sort(peakList_cons[[i]])
# }

# # Generate plot for peak number
# df <- lapply(peakList_cons, length) %>% as.data.frame() %>% gather("Sample", "Peaks")
# df$Sample <- gsub("X", "", df$Sample)
# df$Sample <- factor(df$Sample, levels = df$Sample)
# 
# b <- ggplot(df, aes(Sample, Peaks))+
#   geom_bar(stat = "identity") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 30, hjust = 0.8, vjust = 0.8)) +
#   labs(title = "Peaks") + xlab(NULL)
# 
# cowplot::ggsave2(plot = b, filename = "../data/processed/05_Peak_number_allPeaks.pdf", height = 4, width = 3)

print("Peak data acquired...")
```


## Annotate peaks with ensembl db
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


peakAnnoList_ens <- list()

for (i in 1:length(peakList_cons)){
  
  peak_tmp <- peakList_cons[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  peakAnnoList_ens[[i]] <- anno_tmp
}

names(peakAnnoList_ens) <- names(peakList_cons)

```


## Export peak annotation as bed and xlsx files
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(rtracklayer)
library(xlsx)


for (i in 1:length(peakAnnoList_ens)){
  
  gr <- peakAnnoList_ens[[i]]@anno
  name_gr <- names(peakAnnoList_ens[i])
  
  df <- data.frame(seqnames=seqnames(gr),
    starts=start(gr)-1,
    ends=end(gr),
    SYMBOL=gr$SYMBOL,
    # names=names(gr),
    #scores=gr$score,
    scores=".",
    strands=strand(gr),
    #signalValue=gr$signalValue,
    #qValue=gr$qValue,
    annotation=gr$annotation,
    # names=names(gr),
    # SYMBOL=gr$SYMBOL,
    # ENTREZID=gr$ENTREZID,
    geneId=gr$geneId,
    transcriptId=gr$transcriptId,
    distanceToTSS=gr$distanceToTSS)
  
  # write.table(df, file=paste0("../data/processed/05_", name_gr, "_annotated.bed"), 
  #             quote=F, sep="\t", row.names=F, col.names=F)

  openxlsx::write.xlsx(df, file=paste0("../data/processed/05_", name_gr, "_annotated.xlsx"),
             row.names=F, col.names=T)

}


```

# Evaluate location of peaks relative to annotation features
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
#library(clusterProfiler)
promoter <- getPromoters(TxDb = ensDb, upstream = 3000, downstream = 3000)
seqlevelsStyle(promoter) <- "NCBI"

tagMatrixList <- list()
for (i in names(peakList_cons)){
  tagMatrixList[[i]] <- getTagMatrix(peakList_cons[[i]], windows=promoter)
}


# # tagMatrix <- getTagMatrix(peakList_cons, windows=promoter)
# tagHeatmap(tagMatrixList, xlim=c(-3000, 3000))

# Plot peak histogram around TSS
p <- plotAvgProf(tagMatrixList, xlim=c(-3000, 3000), resmple = 100,
            xlab="Genomic Region (5'->3')", ylab = "Read Count Frequency")
print(p)

pdf("../data/processed/05_Peak_TSS_hist_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


# Plot annotation bargraph
p <- plotAnnoBar(peakAnnoList_ens, title = "H3K18la peak annotation")
print(p)

pdf("../data/processed/05_Peak_annot_bar_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


# Plot disrtobution of peaks realtive to TSS
p <- plotDistToTSS(peakAnnoList_ens,
              title="Distribution of H3K18la peak to TSS")
print(p)

pdf("../data/processed/05_Peak_TSSdist_bar_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


#######################
# Consensus peak number
#######################

# Generate plot for peak number
df <- lapply(peakList_cons, length) %>% as.data.frame() %>% gather("Sample", "Peaks")
df$Sample <- gsub("K18La\\.", "", df$Sample)
df$Sample <- gsub("X", "", df$Sample)
df$Sample <- factor(df$Sample, levels = df$Sample)

uniq_genes <- lapply(peakAnnoList_ens, function(x){x@anno$SYMBOL %>% unique() %>% length()}) %>% unlist() %>% as.vector()
prom_genes <- lapply(peakAnnoList_ens, function(x){x@anno %>% 
    as.data.frame() %>% 
    dplyr::filter(annotation == "Promoter") %>% 
    dplyr::select(SYMBOL) %>% 
    as.vector() %>% unlist() %>% 
    unique() %>% length()}) %>% 
  unlist() %>% as.vector()

b <- ggplot(df, aes(Sample, Peaks))+
  geom_bar(stat = "identity", fill = "grey50") +
  geom_point(aes(x = c(1:4), y = uniq_genes), colour = "red", size = 2) +
  annotate("text", x = c(1:4), y = uniq_genes + 200, label = uniq_genes) +
  geom_point(aes(x = c(1:4), y = prom_genes), colour = "blue", size = 2) +
  annotate("text", x = c(1:4), y = prom_genes + 200, label = prom_genes) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 0.8, vjust = 0.8)) +
  labs(title = "Peaks") + xlab(NULL)

cowplot::ggsave2(plot = b, filename = "../data/processed/05_Peak_number_explant.pdf", height = 3, width = 2.5)

```


# Import the stat files
```{r}

total_import <- read.csv("../data/interim/explant/H3k18la_total-counts.txt", header = F)
# total <- data.frame(sample = gsub(".+_(P.+)_rm.+", "\\1", total_import[c(1, 3, 5, 7, 9, 11),]),
#                     total_reads = gsub("(.+)\\s\\+.+", "\\1", total_import[c(2,4,6,8,10,12),]))

ecoli_import <- read.csv("../data/interim/explant/H3k18la_ecoli-counts.txt", header = F)


df <- data.frame(sample = gsub(".+_(P.+)_rm.+", "\\1", total_import[c(1, 3, 5, 7, 9, 11),]),
                 total_reads = gsub("(.+)\\s\\+.+", "\\1", total_import[c(2,4,6,8,10,12),]) %>%  as.numeric(),
                 ecoli_reads = strsplit(ecoli_import[c(2,4,6,8,10,12),], "\t") %>% 
                      lapply(function(x){x[[3]]}) %>% 
                      unlist() %>% as.numeric)

df$pct_ecoli <- (df$ecoli_reads / df$total_reads * 100) %>% round(1)


write.table(df , file = "../data/processed/05_ecoli_counts.tsv", sep = "\t", quote = F, row.names = F)
```




################################################################################
#### Get bam counts over mstr_feature
################################################################################
```{r}
library(csaw)

# # Determine optimal fragment length using cross-correlation
# wind.param <- readParam(pe="both", minq = 20, dedup = T)
# max.delay <- 500
# bam.files <- paste0(bam_dir, macs_bams)
# x <- correlateReads(bam.files, max.delay, param=wind.param)
# plot(0:max.delay, x, type="l", ylab="CCF", xlab="Delay (bp)")
# print(paste0("Optimal fragment size (bp): ", maximizeCcf(x)))



# Perform region counts
pe.param <- readParam(max.frag=600, pe="both", minq = 20, dedup = T)
bam.files <- paste0(bam_dir, macs_bams)

h3k18la.counts <- regionCounts(bam.files, mstr_features_anno@anno, ext=150, param=pe.param)


# Normalization
binned <- windowCounts(bam.files, bin=TRUE, width=10000, param=pe.param)
h3k18la_se <- normFactors(binned, se.out=T)

```


################################################################################
#### Prep the DGE list objects
################################################################################
```{r}
library(edgeR)

# source(file.path(dest_tools_dir, "DFave.R"))
source(file.path(dest_tools_dir, "Expr_Filter.R"))


##########
## h3k18la
samp_nm <- gsub(".+(H.+)\\.target.+", "\\1", h3k18la.counts@colData[[1]])

h3k18la.dge <- asDGEList(h3k18la.counts,
               norm.factors = h3k18la_se$norm.factors,
               group = gsub("_.+", "", samp_nm),
               genes = as.data.frame(mstr_features_anno))
colnames(h3k18la.dge) <- samp_nm
h3k18la.dge$samples$group <- factor(h3k18la.dge$samples$group, levels = unique(h3k18la.dge$samples$group))

# Filter for peak quant
h3k18la.dge$rpkm <- rpkm(h3k18la.dge, gene.length = "width")
ExpFilt(h3k18la.dge$rpkm, exp = 1, group = h3k18la.dge$samples$group)

h3k18la.dge.filt <- h3k18la.dge[idx.filt, keep.lib.sizes = T]

# Add expression tables to unfiltered object
h3k18la.dge$logCPM <- log2(cpm(h3k18la.dge) + 1)
h3k18la.dge$logFPKM <- rpkm(h3k18la.dge, gene.length = "width", normalized.lib.sizes = T, log = T, prior.count = 1)
h3k18la.dge$FPKM <- rpkm(h3k18la.dge, gene.length = "width", normalized.lib.sizes = T, log = F)
h3k18la.dge$FPKM_mean <- h3k18la.dge$FPKM %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(H3K18La_5mM= mean(c(H3K18La_5mM_R1, H3K18La_5mM_R2, H3K18La_5mM_R3)),
         H3K18La_25mM = mean(c(H3K18La_25mM_R1, H3K18La_25mM_R2, H3K18La_25mM_R3)),
         H3K27ac_5mM = mean(c(H3K27ac_5mM_R1, H3K27ac_5mM_R2)),
         H3K27ac_25mM = mean(c(H3K27ac_25mM_R1, H3K27ac_25mM_R2)))

```


################################################################################
#### PCA analysis
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(plot3D)
library(PCAtools)
library(wesanderson)

# wind.dge <- asDGEList(filtered.data,
#                group = paste(meta_used$stage, meta_used$treatment, sep = "_"))
# colnames(wind.dge) <- meta_used$sample_name
# 
# wind.dge$logCPM <- log2(cpm(wind.dge) + 1)
# 
# color.grp.pal <- wes_palette("Zissou1", 5, "discrete")[c(5,3)]
# color.grp.pal <- c(color.grp.pal, "plum")

#
#Prep the PCA data
pca.h3k18la <- PCAtools::pca(h3k18la.dge$logFPKM, metadata = data.frame(h3k18la.dge$samples))
# pca.h3k18la$metadata$group <- factor(pca.h3k18la$metadata$group, levels = unique(pca.h3k18la$metadata$group))

# sampkey <- color.grp.pal
# names(sampkey) <- unique(pca.wind$metadata$Group)[c(2:3,1)]
# 
# 
# #Plot rb samples
# rb.dge <- wind.dge[,grep("normal", wind.dge$samples$group, invert = T)]
# rb.dge$logCPM <- rb.dge$logCPM[,grep("normal", wind.dge$samples$group, invert = T)]
# pca.rb <- PCAtools::pca(rb.dge$logCPM, metadata = data.frame(Group = rb.dge$samples$group, row.names = row.names(rb.dge$samples)))

X='PC1'; Y='PC2'
pca1 <- biplot(pca.h3k18la, x = X, y = Y,
       colby = "group", #colkey = sampkey,
       #lab = "replicate", labSize = 2,
       legendPosition = 'right', 
       title = 'Peak Quant PCA',
       subtitle = paste0(X, ' versus ', Y))

cowplot::ggsave2(filename = "../data/processed/05_PCA_PC1-2.pdf", plot = pca1,
                 width = 8, height = 6)

```


################################################################################
#### Get the DB peaks of H3k18la from 5 to 25mM glucose 
################################################################################
```{r}

library(edgeR)
library(rtracklayer)
library(qvalue)
# y <- asDGEList(filtered.data)
# y <- y[,which(meta_used$stage == "blastoma")]
# colnames(y) <- meta_used$sample_name[which(meta_used$stage == "blastoma")]
# 
# stage_treat <- factor(paste(meta_used$stage[which(meta_used$stage == "blastoma")],
#                      meta_used$treatment[which(meta_used$stage == "blastoma")], sep = "_"))
# stage_treat[8] <- "blastoma_N"

# Subset for only the lactylation samples 
y <- h3k18la.dge[,h3k18la.dge$samples$group == "H3K18La"]
y$samples$group <- gsub(".+_(.+)_.+", "\\1", row.names(y$samples)) %>% as.factor()
y$samples$group <- relevel(y$samples$group, ref = "5mM")

y$FPKM <- rpkm(y, gene.length = "width", normalized.lib.sizes = T, log = F)
y$FPKM_mean <- y$FPKM %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(H3K18La_5mM= mean(c(H3K18La_5mM_R1, H3K18La_5mM_R2, H3K18La_5mM_R3)),
         H3K18La_25mM = mean(c(H3K18La_25mM_R1, H3K18La_25mM_R2, H3K18La_25mM_R3)))

ExpFilt(y$rpkm, exp = 1, group = y$samples$group)

y <- y[idx.filt, keep.lib.sizes = T]
y$FPKM_mean <- y$FPKM_mean[idx.filt,]
# design <- model.matrix(~h3k18la.dge.filt$samples$group)
# colnames(design) <- levels(h3k18la.dge.filt$samples$group)

y <- estimateDisp(y)
summary(y$trended.dispersion)

et <- exactTest(y)
et$table$FDR <- p.adjust(et$table$PValue, method = "fdr")

fdr <- 0.01
fc <- 2
# Filter for significance
et.sig <- et$table %>%
  # rownames_to_column('gene') %>% 
  dplyr::filter(FDR < fdr, abs(logFC) >= fc)# %>% 
  # column_to_rownames('gene')

results <- decideTests(et, p.value=fdr, lfc = fc)
d <- summary(results)
  
# Merge results tables
et_results <- list(all = data.frame(merge(et$genes, et$table, by = 0), row.names = 1),
               sig = data.frame(merge(et$genes[row.names(et.sig),], et.sig, by = 0), row.names = 1))
# fit <- glmQLFit(y, design, robust=TRUE)
# summary(fit$var.post)

# par(mfrow=c(1,2))
# o <- order(y$AveLogCPM)
# plot(y$AveLogCPM[o], sqrt(y$trended.dispersion[o]), type="l", lwd=2,
# ylim=c(0, 1), xlab=expression("Ave."~Log[2]~"CPM"),
# ylab=("Biological coefficient of variation"))
# plotQLDisp(fit)


explant.db.results <- cbind(et$genes, et$table, y$FPKM_mean)
write.table(explant.db.results, file = "../data/processed/05_H3k18la.explant.db.results.tsv", sep = "\t", quote = F, row.names = F)

table(h3k18la.db.results$FDR < 0.01)

explant.db.results.sig.prom.pc <- h3k18la.db.results %>% dplyr::filter(FDR < 0.01, annotation == "Promoter", transcriptBiotype == "protein_coding")

#Export the bed foile of DB regions
df <- data.frame(seqnames=explant.db.results$seqnames,
    starts=explant.db.results$start-1,
    ends=explant.db.results$end,
    SYMBOL=explant.db.results$SYMBOL,
    # names=names(gr),
    #scores=gr$score,
    scores=".",
    strands=explant.db.results$strand,
    #signalValue=gr$signalValue,
    #qValue=gr$qValue,
    annotation=explant.db.results$annotation,
    # names=explant.db.results$,
    # SYMBOL=gr$SYMBOL,
    # ENTREZID=gr$ENTREZID,
    geneId=explant.db.results$geneId,
    transcriptId=explant.db.results$transcriptId,
    distanceToTSS=explant.db.results$distanceToTSS,
    logfc=explant.db.results$logFC,
    fdr=explant.db.results$FDR)
  
write.table(df, file=paste0("../data/processed/05_explant_H3K18la-sigDB.bed"), 
            quote=F, sep="\t", row.names=F, col.names=F)


```



################################################################################
#### Look at DB of timecourse and explant for prom_pc_sig DB peaks
################################################################################

```{r}

# Get the explant mean FPKM for the prom pc sig peaks
rnames <- row.names(prom_pc_sig) %>% as.numeric()
ex_prom_pc_RPKM <- h3k18la.dge$FPKM_mean[rnames,]

# Merge the timcourse and explant data
prom_pc_sig_ex <- cbind(prom_pc_sig, ex_prom_pc_RPKM)



###################################################################
# Prep tidy table and add RNAseq
###################################################################

###################
# Tidy table
dat.tidy <- prom_pc_sig_ex %>% 
  rownames_to_column(var = "peakId") %>% 
  dplyr::select(peakId, P4, P10, P28, H3K18La_5mM, H3K18La_25mM, H3K27ac_5mM, H3K27ac_25mM) %>%
  gather(Sample, CPM, -peakId)


###################
#Factor sort sample names

dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample))


###################################################################
# Cluster the peak DB
###################################################################

# ###################
# # Sort by Heirarchical clustering using Ward's method with Euclidean distanced
distfunc <- function(x) dist(x,method="euclidean")
hclustfunc <- function(x) hclust(x, method="ward.D2")

#Get distance and heirarchical clustering
d <- distfunc(as.matrix(prom_pc_sig_ex[,27:29]))
hc <- hclustfunc(d)
ct <-- cutree(hc, k=3)

###################
# Sort by PCA of 
pca <- prcomp(prom_pc_sig_ex[,27:29], center = T, scale. = F)
pca.order <- order(pca$x[,1])


###################################################################
# Make Heatmap
###################################################################

# # Sort by HC clustering
dat.tidy$peakId <- factor(dat.tidy$peakId,
                                      levels = dat.tidy$peakId[hc$order], ordered = T)

# Sort by PCA
dat.tidy$peakId <- factor(dat.tidy$peakId, 
                                      levels = dat.tidy$peakId[rev(pca.order)], ordered = T)

# Make heatmap
library(scales)
# lim = c(0,16)
lim = c(0,8)
# nm = "CPM (log2)"
nm = "FPKM (log2)"
mb.colors <- colorRampPalette(c("#023858","#3690c0", "#e0f3f8","#fff7bc","#fec44f","#ec7014",
                                    "#fc4e2a","#bd0026","#800026"))(n=1000)
pp <- ggplot(dat.tidy, aes(Sample, peakId)) +
        geom_tile(aes(fill = log2(CPM)), colour = "transparent") + 
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        geom_vline(xintercept = c(3.5,5.5)) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.text.y = element_blank(),
              axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/05_Prom_pc_CT-Ex_hm-hc_v2.pdf", plot = pp,
                 width = 8, height = 6)


```


################################################################################
#### Make Euler plot of timecourse and explant peaks
################################################################################
```{r}
library(eulerr)

time_ex_ol <- findOverlapsOfPeaks(time_peakAnnoList_ens$P28@anno, 
                                  peakAnnoList_ens$K18_5@anno, 
                                  peakAnnoList_ens$K18_25@anno, connectedPeaks = "keepAll")

eul_vect <- lapply(time_ex_ol$peaklist, length) %>% as.vector() %>% unlist()

names(eul_vect) <- gsub(".anno", "", names(eul_vect))
names(eul_vect) <- gsub("peakAnnoList_ens.", "", names(eul_vect))
names(eul_vect) <- gsub("time_", "", names(eul_vect))
names(eul_vect) <- gsub("///", "&", names(eul_vect))
eul_vect <- eul_vect[c(2,1,3:7)]
eul <- euler(eul_vect) 

eup <- plot(eul, quantities = list(type = c("counts", "percent")))

cowplot::ggsave2(eup, filename = file.path(dir_export, paste0("05_Euler_P28_Explant_peakOL.pdf")),
                 height = 4, width = 4)



```


################################################################################
#### Look at promoter peaks from DGE from RNA analysis (231220)
################################################################################
```{r}
library(ggrepel)

# Get highest expressing transcript at P28
ret_trans_high_dups <- retina_trans_cpm %>% 
  rownames_to_column("ensembl_transcript_id") %>% 
  group_by(ensembl_gene_id) %>% 
  arrange(desc(P28)) %>% 
  top_n(1)

# Import explant RNAseq
rna_ex_cpm <- read.csv("../data/external/Gene_CPM_MSTR_231220explant.tsv", sep = "\t", row.names = 1)
rna_ex_dge <- read.csv("../data/external/DEG_HighGlucose_v_LowGlucose_231220explant.csv")


# Box plot of explant DB in explant RNAseq
dat <- rna_ex_cpm %>% 
  rowwise() %>% 
  mutate(mM_5 = mean(mM_5_rep1, mM_5_rep2, mM_5_rep3),
         mM_25 = mean(mM_25_rep1, mM_25_rep2, mM_25_rep3))

dat <- dat %>% 
  dplyr::filter(external_gene_name %in% explant.db.results.sig.prom.pc$SYMBOL) %>% 
  dplyr::select(external_gene_name, mM_5, mM_25) %>% 
  gather(Sample, CPM, -external_gene_name)

bp <- dat %>% ggplot(aes(Sample, log2(CPM+1))) +
  geom_boxplot()

cowplot::ggsave2(bp, filename = file.path(dir_export, paste0("05_DB_Explant-Timecourse_euler.pdf")),
                 height = 4, width = 4)

dat2 <- merge(rna_ex_dge, explant.db.results, by.x = "ensembl_ID", by.y = "geneId") %>% 
  dplyr::filter(transcriptBiotype == "protein_coding", annotation == "Promoter")

bp <- dat2 %>% ggplot(aes(logFC.y, logFC.x)) +
  geom_point() +
  geom_text_repel(data = dat2,
                    aes(label = external_gene_name),
                    nudge_y = 0.25, nudge_x = 0.25, 
                    min.segment.length = 0.25, 
                    direction = "both", 
                    segment.color = "grey70")

cowplot::ggsave2(bp, filename = file.path(dir_export, paste0("05_Explant_DBvDGE_biplot.pdf")),
                 height = 6, width = 6)


write.table(dat2, file = "../data/processed/05_RNA-H3k18la.explant.db.results.tsv", row.names = F, sep = "\t", quote = F)



########################################################
# Filter the DB table for one promoter peak rep per gene

# Filter explant peaks quant for promoter of protein coding genes
# h3k18la.dge.prom.pc <- h3k18la.dge$genes %>% 
#   dplyr::filter(annotation == "Promoter",
#                 transcriptBiotype == "protein_coding")

# Get gene with multiple promoter peaks
peak_dup_geneID <-  prom_pc_sig[duplicated(explant.db.results.sig.prom.pc$geneId), 'geneId'] %>% unique()

# Subset promoter peaks for genes with a single promoter peak
prom_pc_uniq <- explant.db.results.sig.prom.pc %>% dplyr::filter(!(geneId %in% peak_dup_geneID))

# Subset promoter peaks for genes with multiple promoter peaks
prom_pc_dup <- explant.db.results.sig.prom.pc %>% 
  dplyr::filter(geneId %in% peak_dup_geneID)

# Select promoter peaks for highest transcript expression at P28 and secondarily by highest F score
prom_pc_dup_trans <- prom_pc_dup %>% 
  dplyr::filter(transcriptId %in% ret_trans_high_dups$ensembl_transcript_id) %>% 
  group_by(transcriptId) %>% 
  arrange(desc(F)) %>% 
  top_n(1)

# Merge the single promoter peak with redundancy removed peaks
prom_pc_sig_filt <- rbind(prom_pc_uniq, prom_pc_dup_trans) %>% 
  arrange(seqnames, start)





```


################################################################################
#### Make Euler plot of timecourse DB and explant DB peaks
################################################################################
```{r}
library(eulerr)


eul_vect <- list(in_vivo=prom_pc_sig$SYMBOL %>% unique(),
                 explant=explant.db.results.sig.prom.pc$SYMBOL %>% unique())
eul <- euler(eul_vect) 

eup <- plot(eul, quantities = list(type = c("counts", "percent")))

cowplot::ggsave2(eup, filename = file.path(dir_export, paste0("05_Euler_invivo_Explant_DB.pdf")),
                 height = 4, width = 4)



```



















################################################################################
#### DO NOT RUN BELOW HERE ... EXAMPLE CODE FOR POACHING
################################################################################



################################################################################
#### Differential binding analysis
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(edgeR)
library(rtracklayer)
library(qvalue)
# y <- asDGEList(filtered.data)
# y <- y[,which(meta_used$stage == "blastoma")]
# colnames(y) <- meta_used$sample_name[which(meta_used$stage == "blastoma")]
# 
# stage_treat <- factor(paste(meta_used$stage[which(meta_used$stage == "blastoma")],
#                      meta_used$treatment[which(meta_used$stage == "blastoma")], sep = "_"))
# stage_treat[8] <- "blastoma_N"
y <- h3k18la.dge.filt

design <- model.matrix(~h3k18la.dge.filt$samples$group)
colnames(design) <- levels(h3k18la.dge.filt$samples$group)

y <- estimateDisp(y, design)
summary(y$trended.dispersion)

fit <- glmQLFit(y, design, robust=TRUE)
summary(fit$var.post)

par(mfrow=c(1,2))
o <- order(y$AveLogCPM)
plot(y$AveLogCPM[o], sqrt(y$trended.dispersion[o]), type="l", lwd=2,
ylim=c(0, 1), xlab=expression("Ave."~Log[2]~"CPM"),
ylab=("Biological coefficient of variation"))
plotQLDisp(fit)

summary(fit$df.prior)

results <- glmQLFTest(fit, coef=c(2, 3))
head(results$table)

results$table$qvalue <- qvalue(results$table$PValue)$qvalue
results$table$FDR <- p.adjust(results$table$PValue, method = "fdr")
# crx.db$table$lfdr <- qvalue::qvalue(crx.db$table$PValue, fdr.level = 0.05)$significant
h3k18la.db.results <- cbind(h3k18la.dge.filt$genes, results$table, h3k18la.dge.filt$FPKM_mean)
write.table(h3k18la.db.results, file = "../data/processed/03_H3k18la.db.results.tsv", sep = "\t", quote = F, row.names = F)

table(h3k18la.db.results$FDR < 0.01)

```


################################################################################
#### Merge CUT&RUN and RNAseq
################################################################################
```{r}

###################################################################
# Prep the peak DB data
###################################################################

# Get highest expressing transcript at P28
ret_trans_high_dups <- retina_trans_cpm %>% 
  rownames_to_column("ensembl_transcript_id") %>% 
  group_by(ensembl_gene_id) %>% 
  arrange(desc(P28)) %>% 
  top_n(1)

# Get sig DB promoter peaks from protein coding genes
prom_pc_sig <- h3k18la.db.results %>% dplyr::filter(annotation == "Promoter",
                                                     transcriptBiotype == "protein_coding",
                                                     FDR < 0.01)

# Get gene with multiple promoter peaks
peak_dup_geneID <-  prom_pc_sig[duplicated(prom_pc_sig$geneId), 'geneId'] %>% unique()

# Subset promoter peaks for genes with a single promoter peak
prom_pc_uniq <- prom_pc_sig %>% dplyr::filter(!(geneId %in% peak_dup_geneID))

# Subset promoter peaks for genes with multiple promoter peaks
prom_pc_dup <- prom_pc_sig %>% 
  dplyr::filter(geneId %in% peak_dup_geneID)

# Select promoter peaks for highest transcript expression at P28 and secondarily by highest F score
prom_pc_dup_trans <- prom_pc_dup %>% 
  dplyr::filter(transcriptId %in% ret_trans_high_dups$ensembl_transcript_id) %>% 
  group_by(transcriptId) %>% 
  arrange(desc(F)) %>% 
  top_n(1)

# Merge the single promoter peak with redundancy removed peaks
prom_pc_sig_filt <- rbind(prom_pc_uniq, prom_pc_dup_trans) %>% 
  arrange(seqnames, start)


###################################################################
# Prep tidy table and add RNAseq
###################################################################

###################
# Tidy table
dat.tidy <- prom_pc_sig_filt %>% 
  dplyr::select(SYMBOL, P4, P10, P28) %>%
  gather(Sample, CPM, -SYMBOL)

###################
# Get RNA data for same genes
dat.tidy.rna <- retina_cpm %>% 
  dplyr::select(external_gene_name, P4, P10, P28) %>%
  dplyr::rename(SYMBOL = external_gene_name,
                P4_RNA = P4,
                P10_RNA = P10,
                P28_RNA = P28) %>% 
  gather(Sample, CPM, -SYMBOL)

###################
#Factor sort sample names
dat.tidy <- rbind(dat.tidy, dat.tidy.rna)
dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample))


###################################################################
# Cluster the peak DB
###################################################################

###################
# Sort by Heirarchical clustering using Ward's method with Euclidean distanced
distfunc <- function(x) dist(x,method="euclidean")
hclustfunc <- function(x) hclust(x, method="ward.D2")

#Get distance and heirarchical clustering
d <- distfunc(as.matrix(prom_pc_sig_filt[,27:29]))
hc <- hclustfunc(d)


###################
# Sort by PCA
pca <- prcomp(prom_pc_sig_filt[,27:29], center = T, scale. = F)
pca.order <- order(pca$x[,1])


###################################################################
# Make Heatmap
###################################################################

# Sort by HC clustering
dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL, 
                                      levels = dat.tidy$SYMBOL[hc$order], ordered = T)

# Sort by PCA
dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL, 
                                      levels = dat.tidy$SYMBOL[rev(pca.order)], ordered = T)

# Make heatmap
library(scales)
lim = c(0,16)
nm = "CPM (log2)"
mb.colors <- colorRampPalette(c("#023858","#3690c0", "#e0f3f8","#fff7bc","#fec44f","#ec7014",
                                    "#fc4e2a","#bd0026","#800026"))(n=1000)
pp <- ggplot(dat.tidy, aes(Sample, SYMBOL)) +
        geom_tile(aes(fill = CPM), colour = "transparent") + 
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT-RNA_hm-pca.pdf", plot = pp,
                 width = 8, height = 6)
```


# Correlate the C&T and RNA seq
```{r}

rna_tmp <- retina_cpm %>% 
  rownames_to_column(var = "ENSEMBL") %>% 
  dplyr::select(c("ENSEMBL", "external_gene_name", "P4", "P10", "P28")) %>% 
  group_by(external_gene_name) %>% 
  arrange(desc(P28)) %>% 
  slice_head(n=1)

prom_pc_sig_filt_rna <- merge(prom_pc_sig_filt, rna_tmp, by.x = "SYMBOL", by.y = "external_gene_name")



# Find Person correlation between the H3K28La peak quant and RNAseq expression
prom_pc_sig_filt_rna <- prom_pc_sig_filt_rna %>% 
  rowwise() %>% 
  mutate(pearson = cor(x=c(P4.x, P10.x, P28.x), y = c(P4.y, P10.y, P28.y), method = "pearson"))

a <- ggplot(prom_pc_sig_filt_rna, aes(x=pearson)) +
  geom_histogram(binwidth = 0.01)

cowplot::ggsave2(filename = paste0("../data/processed/03_Prom_pc_sigDE-RNA_corr-dist.pdf"),
                 plot = a,
                 width = 6, height = 4)

# Split by correlated and anticorrelated
cor_sig = 0.8
posi_cor <- prom_pc_sig_filt_rna %>% 
  dplyr::filter(pearson > cor_sig)

anti_cor <- prom_pc_sig_filt_rna %>% 
  dplyr::filter(pearson < cor_sig)


######################
# Plot the positive correlated peak and rna data
pca <- prcomp(posi_cor[,c(27:29, 31:33)], center = T, scale. = F)
pca.order <- order(pca$x[,1])

dat.tidy <- posi_cor %>% 
  dplyr::select(SYMBOL, P4.x, P10.x, P28.x, P4.y, P10.y, P28.y) %>%
  gather(Sample, CPM, -SYMBOL)

dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample))
dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL, 
                                      levels = dat.tidy$SYMBOL[rev(pca.order)], ordered = T)

pp <- ggplot(dat.tidy, aes(Sample, SYMBOL)) +
        geom_tile(aes(fill = CPM), colour = "transparent") + 
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        geom_vline(xintercept = 3.5) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT-RNA_posi-cor_hm-pca.pdf", plot = pp,
                 width = 8, height = 6)

######################
# Plot the negative correlated peak and rna data
pca <- prcomp(anti_cor[,c(27:29, 31:33)], center = T, scale. = F)
pca.order <- order(pca$x[,1])

dat.tidy <- anti_cor %>% 
  dplyr::select(SYMBOL, P4.x, P10.x, P28.x, P4.y, P10.y, P28.y) %>%
  gather(Sample, CPM, -SYMBOL)

dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample))
dat.tidy$SYMBOL <- factor(dat.tidy$SYMBOL, 
                                      levels = dat.tidy$SYMBOL[rev(pca.order)], ordered = T)

pq <- ggplot(dat.tidy, aes(Sample, SYMBOL)) +
        geom_tile(aes(fill = CPM), colour = "transparent") + 
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        geom_vline(xintercept = 3.5) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03_Prom_pc_CT-RNA_anti-cor_hm-pca.pdf", plot = pq,
                 width = 8, height = 6)

```


################################################################################
# Make expression plots for several genes - Oct 11, 2024
################################################################################
```{r}

goi <- c("Akt3", "Rorb", "Reep6", "Vamp2")

for (i in goi){
  
  # RNAseq
  tmp_dat <- retina.dwnld %>% dplyr::filter(external_gene_name == i) %>%
    dplyr::select(c(5, 22:23, 26:27, 30:31)) %>% 
    gather("Sample", "CPM", -external_gene_name) %>% 
    mutate(Group = gsub("(.+)\\.Retina.+", "\\1", Sample)) %>% 
    rename(c(Symbol = "external_gene_name"))
  
  tmp_dat$Group <- factor(tmp_dat$Group, levels = tmp_dat$Group %>% unique())
  
  p <- ggplot(tmp_dat, aes(x = Group, y = CPM)) +
    geom_boxplot(fill = c("#fdbe85", "#fd8d3c", "#d94701")) +
    ggtitle(i)
   
  # CUT&Tag
  tmp_dat2 <- h3k18la.dge$FPKM %>% as_tibble()  %>% 
    dplyr::select(c(1:6)) %>% 
    mutate(annotation = h3k18la.dge$genes$annotation, 
           transcriptBiotype = h3k18la.dge$genes$transcriptBiotype,
           Symbol = h3k18la.dge$genes$SYMBOL) %>% 
    dplyr::filter(Symbol == i, 
                  annotation == "Promoter",
                  transcriptBiotype == "protein_coding")
  tmp_dat2 <- tmp_dat2 %>%
    gather("Sample", "FPKM", -Symbol, -annotation, -transcriptBiotype) %>% 
    mutate(Group = gsub("(.+)_R\\d", "\\1", Sample))
  
  pp <- p <- ggplot(tmp_dat2, aes(x = Group, y = CPM)) +
    geom_boxplot(fill = c("#fdbe85", "#fd8d3c", "#d94701")) +
    ggtitle(i)
  
  cowplot::ggsave2(plot = p, filename = file.path(dir_export, paste0("08_", i, "_exp_boxplot.pdf")),
                   height = 4, width = 4)

}


```





# Export objects
```{r}

saveRDS(mstr_features_anno, file = "../data/interim/03_mstr_features_anno_240529.RDS")
saveRDS(peakList_cons, file = "../data/interim/03_peakList_cons_240529.RDS")

```