---
title: "H3K18la Explant Glucose Analysis"
output: html_notebook
---

	
################################################################################
### Get custom scripts and functions and outside data used for analysis
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

script_dir <- "https://raw.githubusercontent.com/MBrooks313/R_codes/master"
dest_tools_dir <- "../src/tools"
dest_vis_dir <- "../src/visualization"


library(tidyverse)
dir_export <- "../data/processed/"
options(java.parameters = "-Xmx8000m")

```


################################################################################
### Import outside data
################################################################################
```{r}
source("../src/tools/DFave.R")

invivo_peakAnnoList_ens <- readRDS("../data/interim/peakAnnoList_cons_240911.RDS")

# Import expression data for developing retina
retina.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Gene_nCPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)

# Get group info for averaging
grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.dwnld)[10:31])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
retina_cpm <- data.frame(merge(retina.dwnld[,1:9],
                               log2(DFave(retina.dwnld[,10:31], exp_grp) + 1),
                               by = 0),
                         row.names = 1)

# Import transcript expression data for developing retina
retina.trans.dwnld <- read.csv("../data/external/Mouse_Retina_Development_Trans_nRKPM_ENSv98.tsv",
                         sep = "\t", row.names = 1)

# Get group info for averaging
grp <- gsub("(.+)\\.Retina.+", "\\1", colnames(retina.trans.dwnld)[12:33])
exp_grp <- factor(grp, levels = unique(grp))

# Get average CPM values
retina_trans_cpm <- data.frame(merge(retina.trans.dwnld[,1:11],
                               log2(DFave(retina.trans.dwnld[,12:33], exp_grp) + 1),
                               by = 0),
                         row.names = 1)

```


################################################################################
### Get Ensembl annotation database
################################################################################
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

ensDb <- ahdb[[1]]
ensDb

```


################################################################################
### Get list of bam files and make master bed feature for quantitation
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(tidyverse)
library(ChIPpeakAnno)
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


# Macs2 beds
bam_dir <- "../data/raw/bams/"
macs_bams <- list.files(path = bam_dir, pattern = ".+5mM.+.bam$")
macs_bams <- macs_bams[c(4:6, 1:3, 9:10, 7:8)]

# Import consensus peaks from timecourse data
peakList_cons_tc <- readRDS(file = "../data/interim/03_peakList_cons_240913.RDS")

# Import master consensus
mstr_features_anno <- readRDS(file = "../data/interim/03_mstr_features_anno_240913.RDS")

# Import the timecourse DB results
h3k18la.db.results <- readRDS("../data/interim/03_h3k18la_db_results_240913.RDS")

# Import the timecourse prom pc filtered DB results
prom_pc_sig <- readRDS("../data/interim/03_prom_pc_sig_240913.RDS")

time_peakAnnoList_ens <- readRDS("../data/interim/peakAnnoList_cons_240911.RDS")

```


################################################################################
### All explant samples meta
################################################################################

### Get metadata for samples
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}

# Macs2 beds
macs_beds <- list.files(path = "../data/interim/bed/explant/macs2/", pattern = "narrowPeak", recursive = T)

meta <- data.frame(Beds = macs_beds,
                   Samp = gsub("(.+)\\.macs.+", "\\1", macs_beds),
                   Group = gsub("(.+)_R.+", "\\1", macs_beds))
meta <- meta[c(4:6, 1:3, 9:10, 7:8),]

meta$Samp <- factor(meta$Samp, levels = unique(meta$Samp))
meta$Group <- factor(meta$Group, levels = unique(meta$Group))

```


### Import the peaks into a merged list object and plot for number of peaks at each timepoint
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
library(ChIPpeakAnno)

macs_dir <- "../data/interim/bed/explant/macs2//"

# Import the cons peaks
peakList <- sapply(file.path(macs_dir, meta$Beds), toGRanges, format="narrowPeak")
names(peakList) <- meta$Samp

peakList_1E6 <- GRangesList()
# Sort the cons peaks
for (i in 1:length(peakList)){
  
  # seqlevelsStyle(peakList[[i]]) <- "NCBI"
  tmp_peak <- sortSeqlevels(peakList[[i]])
  tmp_peak <- GenomicRanges::sort(tmp_peak)
  tmp_peak <- tmp_peak[(elementMetadata(tmp_peak)[,"qValue"] > 6) & (seqnames(tmp_peak) != "MT")]
  peakList_1E6[[i]] <- tmp_peak
}
names(peakList_1E6) <- meta$Samp

# Overlap peaks per timepoint
ol_k18_5 <- findOverlapsOfPeaks(c(peakList_1E6[1],
                             peakList_1E6[2],
                             peakList_1E6[3]), connectedPeaks="keepAll", ignore.strand = T)

ol_k18_25 <- findOverlapsOfPeaks(c(peakList_1E6[4],
                             peakList_1E6[5],
                             peakList_1E6[6]), connectedPeaks="keepAll", ignore.strand = T)

ol_k27_5 <- findOverlapsOfPeaks(c(peakList_1E6[7],
                             peakList_1E6[8]), connectedPeaks="keepAll", ignore.strand = T)

ol_k27_25 <- findOverlapsOfPeaks(c(peakList_1E6[9],
                             peakList_1E6[10]), connectedPeaks="keepAll", ignore.strand = T)

peakList_cons <- GRangesList(ol_k18_5$mergedPeaks, ol_k18_25$mergedPeaks, ol_k27_5$mergedPeaks, ol_k27_25$mergedPeaks)
names(peakList_cons) <- c("K18_5", "K18_25", "K27_5", "K27_25")

```


### Annotate peaks with ensembl db
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


peakAnnoList_ens <- list()

for (i in 1:length(peakList_cons)){
  
  peak_tmp <- peakList_cons[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")
  peakAnnoList_ens[[i]] <- anno_tmp
}

names(peakAnnoList_ens) <- names(peakList_cons)

```


### Evaluate location of peaks relative to annotation features - Figure 6 & S3
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
#library(clusterProfiler)
promoter <- getPromoters(TxDb = ensDb, upstream = 3000, downstream = 3000)
seqlevelsStyle(promoter) <- "NCBI"

tagMatrixList <- list()
for (i in names(peakList_cons)){
  tagMatrixList[[i]] <- getTagMatrix(peakList_cons[[i]], windows=promoter)
}


# Plot peak histogram around TSS
p <- plotAvgProf(tagMatrixList, xlim=c(-3000, 3000), resmple = 100,
            xlab="Genomic Region (5'->3')", ylab = "Read Count Frequency")
print(p)

pdf("../data/processed/05_Peak_TSS_hist_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


# Plot annotation bargraph
p <- plotAnnoBar(peakAnnoList_ens, title = "H3K18la peak annotation")
print(p)

pdf("../data/processed/05_Peak_annot_bar_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


# Plot disrtobution of peaks realtive to TSS
p <- plotDistToTSS(peakAnnoList_ens,
              title="Distribution of H3K18la peak to TSS")
print(p)

pdf("../data/processed/05_Peak_TSSdist_bar_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


#######################
# Consensus peak number
#######################

# Generate plot for peak number
df <- lapply(peakList_cons, length) %>% as.data.frame() %>% gather("Sample", "Peaks")
df$Sample <- gsub("K18La\\.", "", df$Sample)
df$Sample <- gsub("X", "", df$Sample)
df$Sample <- factor(df$Sample, levels = df$Sample)

uniq_genes <- lapply(peakAnnoList_ens, function(x){x@anno$SYMBOL %>% unique() %>% length()}) %>% unlist() %>% as.vector()
prom_genes <- lapply(peakAnnoList_ens, function(x){x@anno %>% 
    as.data.frame() %>% 
    dplyr::filter(annotation == "Promoter") %>% 
    dplyr::select(SYMBOL) %>% 
    as.vector() %>% unlist() %>% 
    unique() %>% length()}) %>% 
  unlist() %>% as.vector()

b <- ggplot(df, aes(Sample, Peaks))+
  geom_bar(stat = "identity", fill = "grey50") +
  geom_point(aes(x = c(1:4), y = uniq_genes), colour = "red", size = 2) +
  annotate("text", x = c(1:4), y = uniq_genes + 200, label = uniq_genes) +
  geom_point(aes(x = c(1:4), y = prom_genes), colour = "blue", size = 2) +
  annotate("text", x = c(1:4), y = prom_genes + 200, label = prom_genes) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 0.8, vjust = 0.8)) +
  labs(title = "Peaks") + xlab(NULL)

cowplot::ggsave2(plot = b, filename = "../data/processed/05_Peak_number_explant.pdf", height = 3, width = 2.5)

```


################################################################################
### Get bam counts over mstr_feature
################################################################################
```{r}
library(csaw)

# Perform region counts
pe.param <- readParam(max.frag=600, pe="both", minq = 20, dedup = T)
bam.files <- paste0(bam_dir, macs_bams)

h3k18la.counts <- regionCounts(bam.files, mstr_features_anno@anno, ext=150, param=pe.param)


# Normalization
binned <- windowCounts(bam.files, bin=TRUE, width=10000, param=pe.param)
h3k18la_se <- normFactors(binned, se.out=T)

```


### Prep the DGE list objects
```{r}
library(edgeR)

# source(file.path(dest_tools_dir, "DFave.R"))
source(file.path(dest_tools_dir, "Expr_Filter.R"))


##########
## h3k18la
samp_nm <- gsub(".+(H.+)\\.target.+", "\\1", h3k18la.counts@colData[[1]])

h3k18la.dge <- asDGEList(h3k18la.counts,
               norm.factors = h3k18la_se$norm.factors,
               group = gsub("_.+", "", samp_nm),
               genes = as.data.frame(mstr_features_anno))
colnames(h3k18la.dge) <- samp_nm
h3k18la.dge$samples$group <- factor(h3k18la.dge$samples$group, levels = unique(h3k18la.dge$samples$group))

# Filter for peak quant
h3k18la.dge$rpkm <- rpkm(h3k18la.dge, gene.length = "width")
ExpFilt(h3k18la.dge$rpkm, exp = 1, group = h3k18la.dge$samples$group)

h3k18la.dge.filt <- h3k18la.dge[idx.filt, keep.lib.sizes = T]

# Add expression tables to unfiltered object
h3k18la.dge$logCPM <- log2(cpm(h3k18la.dge) + 1)
h3k18la.dge$logFPKM <- rpkm(h3k18la.dge, gene.length = "width", normalized.lib.sizes = T, log = T, prior.count = 1)
h3k18la.dge$FPKM <- rpkm(h3k18la.dge, gene.length = "width", normalized.lib.sizes = T, log = F)
h3k18la.dge$FPKM_mean <- h3k18la.dge$FPKM %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(H3K18La_5mM= mean(c(H3K18La_5mM_R1, H3K18La_5mM_R2, H3K18La_5mM_R3)),
         H3K18La_25mM = mean(c(H3K18La_25mM_R1, H3K18La_25mM_R2, H3K18La_25mM_R3)),
         H3K27ac_5mM = mean(c(H3K27ac_5mM_R1, H3K27ac_5mM_R2)),
         H3K27ac_25mM = mean(c(H3K27ac_25mM_R1, H3K27ac_25mM_R2)))

```



### PCA analysis - Figure S3
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(plot3D)
library(PCAtools)
library(wesanderson)

#Prep the PCA data
pca.h3k18la <- PCAtools::pca(h3k18la.dge$logFPKM, metadata = data.frame(h3k18la.dge$samples))

X='PC1'; Y='PC2'
pca1 <- biplot(pca.h3k18la, x = X, y = Y,
       colby = "group", #colkey = sampkey,
       #lab = "replicate", labSize = 2,
       legendPosition = 'right', 
       title = 'Peak Quant PCA',
       subtitle = paste0(X, ' versus ', Y))

cowplot::ggsave2(filename = "../data/processed/05_PCA_PC1-2.pdf", plot = pca1,
                 width = 8, height = 6)

```


################################################################################
### Get the DB peaks of H3k18la from 5 to 25mM glucose 
################################################################################
```{r}

library(edgeR)
library(rtracklayer)
library(qvalue)


# Subset for only the lactylation samples 
y <- h3k18la.dge[,h3k18la.dge$samples$group == "H3K18La"]
y$samples$group <- gsub(".+_(.+)_.+", "\\1", row.names(y$samples)) %>% as.factor()
y$samples$group <- relevel(y$samples$group, ref = "5mM")

y$FPKM <- rpkm(y, gene.length = "width", normalized.lib.sizes = T, log = F)
y$FPKM_mean <- y$FPKM %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(H3K18La_5mM= mean(c(H3K18La_5mM_R1, H3K18La_5mM_R2, H3K18La_5mM_R3)),
         H3K18La_25mM = mean(c(H3K18La_25mM_R1, H3K18La_25mM_R2, H3K18La_25mM_R3)))

ExpFilt(y$rpkm, exp = 1, group = y$samples$group)

y <- y[idx.filt, keep.lib.sizes = T]
y$FPKM_mean <- y$FPKM_mean[idx.filt,]


y <- estimateDisp(y)
summary(y$trended.dispersion)

et <- exactTest(y)
et$table$FDR <- p.adjust(et$table$PValue, method = "fdr")

fdr <- 0.01
fc <- 2

# Filter for significance
et.sig <- et$table %>%
  dplyr::filter(FDR < fdr, abs(logFC) >= fc)# %>% 

results <- decideTests(et, p.value=fdr, lfc = fc)
d <- summary(results)
  
# Merge results tables
et_results <- list(all = data.frame(merge(et$genes, et$table, by = 0), row.names = 1),
               sig = data.frame(merge(et$genes[row.names(et.sig),], et.sig, by = 0), row.names = 1))


explant.db.results <- cbind(et$genes, et$table, y$FPKM_mean)
write.table(explant.db.results, file = "../data/processed/05_H3k18la.explant.db.results.tsv", sep = "\t", quote = F, row.names = F)

table(h3k18la.db.results$FDR < 0.01)

explant.db.results.sig.prom.pc <- h3k18la.db.results %>% dplyr::filter(FDR < 0.01, annotation == "Promoter", transcriptBiotype == "protein_coding")


```



################################################################################
### Look at DB of timecourse and explant for prom_pc_sig DB peaks - Figure 6
################################################################################

```{r}

# Get the explant mean FPKM for the prom pc sig peaks
rnames <- row.names(prom_pc_sig) %>% as.numeric()
ex_prom_pc_RPKM <- h3k18la.dge$FPKM_mean[rnames,]

# Merge the timcourse and explant data
prom_pc_sig_ex <- cbind(prom_pc_sig, ex_prom_pc_RPKM)



###################################################################
# Prep tidy table and add RNAseq
###################################################################

###################
# Tidy table
dat.tidy <- prom_pc_sig_ex %>% 
  rownames_to_column(var = "peakId") %>% 
  dplyr::select(peakId, P4, P10, P28, H3K18La_5mM, H3K18La_25mM, H3K27ac_5mM, H3K27ac_25mM) %>%
  gather(Sample, CPM, -peakId)


###################
#Factor sort sample names

dat.tidy$Sample <- factor(dat.tidy$Sample, levels = unique(dat.tidy$Sample))


###################################################################
# Cluster the peak DB
###################################################################

# ###################
# # Sort by Heirarchical clustering using Ward's method with Euclidean distanced
distfunc <- function(x) dist(x,method="euclidean")
hclustfunc <- function(x) hclust(x, method="ward.D2")

#Get distance and heirarchical clustering
d <- distfunc(as.matrix(prom_pc_sig_ex[,27:29]))
hc <- hclustfunc(d)
ct <-- cutree(hc, k=3)

###################
# Sort by PCA of 
pca <- prcomp(prom_pc_sig_ex[,27:29], center = T, scale. = F)
pca.order <- order(pca$x[,1])


###################################################################
# Make Heatmap
###################################################################

# # Sort by HC clustering
dat.tidy$peakId <- factor(dat.tidy$peakId,
                                      levels = dat.tidy$peakId[hc$order], ordered = T)

# Sort by PCA
dat.tidy$peakId <- factor(dat.tidy$peakId, 
                                      levels = dat.tidy$peakId[rev(pca.order)], ordered = T)

# Make heatmap
library(scales)
# lim = c(0,16)
lim = c(0,8)
# nm = "CPM (log2)"
nm = "FPKM (log2)"
mb.colors <- colorRampPalette(c("#023858","#3690c0", "#e0f3f8","#fff7bc","#fec44f","#ec7014",
                                    "#fc4e2a","#bd0026","#800026"))(n=1000)
pp <- ggplot(dat.tidy, aes(Sample, peakId)) +
        geom_tile(aes(fill = log2(CPM)), colour = "transparent") + 
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        geom_vline(xintercept = c(3.5,5.5)) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(),
              axis.text.y = element_blank(),
              axis.title.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/05_Prom_pc_CT-Ex_hm-hc_v2.pdf", plot = pp,
                 width = 8, height = 6)


```



################################################################################
### Look at promoter peaks from DGE from RNA analysis - Figure 6
################################################################################
```{r}
library(ggrepel)

# Get highest expressing transcript at P28
ret_trans_high_dups <- retina_trans_cpm %>% 
  rownames_to_column("ensembl_transcript_id") %>% 
  group_by(ensembl_gene_id) %>% 
  arrange(desc(P28)) %>% 
  top_n(1)

# Import explant RNAseq
rna_ex_cpm <- read.csv("../data/external/Gene_CPM_MSTR_231220explant.tsv", sep = "\t", row.names = 1)
rna_ex_dge <- read.csv("../data/external/DEG_HighGlucose_v_LowGlucose_231220explant.csv")


dat2 <- merge(rna_ex_dge, explant.db.results, by.x = "ensembl_ID", by.y = "geneId") %>% 
  dplyr::filter(transcriptBiotype == "protein_coding", annotation == "Promoter")

bp <- dat2 %>% ggplot(aes(logFC.y, logFC.x)) +
  geom_point() +
  geom_text_repel(data = dat2,
                    aes(label = external_gene_name),
                    nudge_y = 0.25, nudge_x = 0.25, 
                    min.segment.length = 0.25, 
                    direction = "both", 
                    segment.color = "grey70")

cowplot::ggsave2(bp, filename = file.path(dir_export, paste0("05_Explant_DBvDGE_biplot.pdf")),
                 height = 6, width = 6)


write.table(dat2, file = "../data/processed/05_RNA-H3k18la.explant.db.results.tsv", row.names = F, sep = "\t", quote = F)


```

