---
title: "H3K18la Explant 2-DG Analysis"
output: html_notebook

---


### Get custom scripts and functions and outside data used for analysis
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

script_dir <- "https://raw.githubusercontent.com/MBrooks313/R_codes/master"
dest_tools_dir <- "../src/tools"
dest_vis_dir <- "../src/visualization"

download.file(url = file.path(script_dir, "GO_leaf.R"),
              destfile = file.path(dest_tools_dir, "GO_leaf.R"),
              method = "curl")

download.file(url = file.path(script_dir, "Expr_Filter.R"),
              destfile = file.path(dest_tools_dir, "Expr_Filter.R"),
              method = "curl")

download.file(url = file.path(script_dir, "VolcanoPlot.R"),
              destfile = file.path(dest_tools_dir, "VolcanoPlot.R"),
              method = "curl")

library(tidyverse)
options(java.parameters = "-Xmx8000m")

```

# Get Ensembl annotation database
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(AnnotationHub)
library(ensembldb)

ah <- AnnotationHub()

ahdb <- query(ah, pattern = c("EnsDb", "Mus musculus", 102))

ensDb <- ahdb[[1]]

```


################################################################################
### Import outside data
################################################################################
```{r}
source("../src/tools/DFave.R")

#############################################
# Import db data for time-course H3K18La

db_tc <- read.csv("/Volumes/data/People/brooksma/PEGASUS/Projects/Gaur/Analysis/cutntag_240522/240911/data/processed/H3k18la.db.results.tsv", sep = "\t")

db_tc_sig_PromPC <- db_tc %>% dplyr::filter(annotation == "Promoter", 
                                            transcriptBiotype == "protein_coding",
                                            (abs(logFC.P10) > 1 | abs(logFC.P28)), 
                                            FDR < 0.01)

#############################################
# Import data for 2-DG RNA-seq experiment

rna.dge <- readRDS("../data/interim/gene.dge.filt.rds")
rna_de_res <- read.csv("../data/processed/RNA_DE_2DG_results.tsv", sep = "\t")
```


################################################################################
### All Explant 2-DG Samples
################################################################################

### Get metadata for samples
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
# ord <- c(1:4)

# Macs2 beds
macs_beds <- list.files(path = "../data/interim/bed/", pattern = "narrowPeak", recursive = T, full.names = T)
macs_beds <- macs_beds[c(4:6, 1:3,
                         11:12, 9:10)]
# Meta sheet
meta <- data.frame(Beds = macs_beds,
                   Mark = gsub(".+\\/(.+)_.+_R.+\\.macs.+", "\\1", macs_beds),
                   Treat = gsub(".+\\/H3.+_(.+)_R.+\\.macs.+", "\\1", macs_beds),
                   Group = gsub(".+\\/(.+)_R.+", "\\1", macs_beds),
                   Rep = paste0("R", c(rep(1:3, times = 3), rep(1:2, times = 2))),
                   Run = gsub(".+\\/(25.+)\\/H.+", "\\1", macs_beds))

names(macs_beds) <- paste(meta$Group, meta$Rep, sep = "_")

meta$Group <- factor(meta$Group, levels = unique(meta$Group))

```


### Import the peaks into a merged list object and plot for number of peaks at each timepoint
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
library(ChIPpeakAnno)

# Import the cons peaks
peakList <- sapply(file.path(meta$Beds), toGRanges, format="narrowPeak")
names(peakList) <- names(macs_beds)

peakList_1E6 <- GRangesList()
# Sort the cons peaks
for (i in 1:length(peakList)){
  
  tmp_peak <- sortSeqlevels(peakList[[i]])
  tmp_peak <- GenomicRanges::sort(tmp_peak)
  tmp_peak <- tmp_peak[(elementMetadata(tmp_peak)[,"qValue"] > 6) & (seqnames(tmp_peak) != "MT")]
  peakList_1E6[[i]] <- tmp_peak
}
names(peakList_1E6) <- names(macs_beds)

# Get overlaps
ol_La_cntl <- findOverlapsOfPeaks(c(peakList_1E6[1],
                             peakList_1E6[2], 
                             peakList_1E6[3]), connectedPeaks="keepAll", ignore.strand = T)

ol_La_2DG <- findOverlapsOfPeaks(c(peakList_1E6[4],
                             peakList_1E6[5], 
                             peakList_1E6[6]), connectedPeaks="keepAll", ignore.strand = T)

ol_Ac_cntl <- findOverlapsOfPeaks(c(peakList_1E6[10],
                             peakList_1E6[11]), 
                             connectedPeaks="keepAll", ignore.strand = T)

ol_Ac_2DG <- findOverlapsOfPeaks(c(peakList_1E6[12],
                             peakList_1E6[13]), connectedPeaks="keepAll", ignore.strand = T)

peakList_cons <- GRangesList(ol_La_cntl$mergedPeaks, ol_La_2DG$mergedPeaks,
                             ol_Ac_cntl$mergedPeaks, ol_Ac_2DG$mergedPeaks)
names(peakList_cons) <- c("La_cntl", "La_2DG", "Ac_cntl", "Ac_2DG")


# Get master features for quantitation
mstr_features <- IRanges::reduce(c(peakList_cons[[1]],
                         peakList_cons[[2]],
                         peakList_cons[[4]],
                         peakList_cons[[5]]))

# Annotate the features
mstr_features_anno <- annotatePeak(mstr_features, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
mstr_features_anno@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = mstr_features_anno@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")

```


### Import the peaks into a merged list object and plot for number of peaks at each timepoint. -Figure S4
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}
library(ChIPpeakAnno)


# Generate plot for peak number
df <- lapply(peakList_cons, length) %>% as.data.frame() %>% gather("Sample", "Peaks")
df$Sample <- gsub("K18La\\.", "", df$Sample)
df$Sample <- gsub("X", "", df$Sample)
df$Sample <- factor(df$Sample, levels = df$Sample)

uniq_genes <- lapply(peakAnnoList_ens, function(x){x@anno$SYMBOL %>% unique() %>% length()}) %>% unlist() %>% as.vector()
prom_genes <- lapply(peakAnnoList_ens, function(x){x@anno %>% 
    as.data.frame() %>% 
    dplyr::filter(annotation == "Promoter") %>% 
    dplyr::select(SYMBOL) %>% 
    as.vector() %>% unlist() %>% 
    unique() %>% length()}) %>% 
  unlist() %>% as.vector()

b <- ggplot(df, aes(Sample, Peaks))+
  geom_bar(stat = "identity", fill = "grey50") +
  geom_point(aes(x = c(1:6), y = uniq_genes), colour = "red", size = 2) +
  annotate("text", x = c(1:6), y = uniq_genes + 200, label = uniq_genes) +
  geom_point(aes(x = c(1:6), y = prom_genes), colour = "blue", size = 2) +
  annotate("text", x = c(1:6), y = prom_genes + 200, label = prom_genes) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 0.8, vjust = 0.8)) +
  labs(title = "Peaks") + xlab(NULL)

cowplot::ggsave2(plot = b, filename = "../data/processed/Peak_number_TP.pdf", height = 3, width = 2.5)

```


### Annotate peaks with ensembl db
```{r, echo = FALSE, results = 'hide', warning=FALSE, message=FALSE}
library(ChIPseeker)
library(org.Mm.eg.db)
library(biomaRt)
library(GenomicFeatures)


peakAnnoList_ens <- list()

for (i in 1:length(peakList_cons)){
  
  peak_tmp <- peakList_cons[[i]]
  seqlevelsStyle(peak_tmp) <- "NCBI"
  
  anno_tmp <- annotatePeak(peak_tmp, tssRegion=c(-1000, 500),
                         TxDb=ensDb, addFlankGeneInfo = T,
                         overlap = "all")
  anno_tmp@anno$SYMBOL <- mapIds(x = ensDb, 
                                         keys = anno_tmp@anno$geneId, 
                                         column = "SYMBOL", 
                                         keytype = "GENEID")

  peakAnnoList_ens[[i]] <- anno_tmp

}

names(peakAnnoList_ens) <- names(peakList_cons)

```


### Evaluate location of peaks relative to annotation features - Figure 6
```{r, echo = FALSE, results = 'asis', warning=FALSE, message=FALSE}

promoter <- getPromoters(TxDb = ensDb, upstream = 3000, downstream = 3000)
seqlevelsStyle(promoter) <- "NCBI"

tagMatrixList <- list()
for (i in names(peakList_cons)){
  tagMatrixList[[i]] <- getTagMatrix(peakList_cons[[i]], windows=promoter)
}


# Plot annotation bargraph
p <- plotAnnoBar(peakAnnoList_ens, title = "H3K18la peak annotation")
print(p)

pdf("../data/processed/Peak_annot_bar_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


# Plot disrtobution of peaks realtive to TSS
p <- plotDistToTSS(peakAnnoList_ens,
              title="Distribution of H3K18la peak to TSS")
print(p)

pdf("../data/processed/Peak_TSSdist_bar_allPeaks.pdf",
    height = 3, width = 5, useDingbats = F)
print(p)
dev.off()


```


################################################################################
### Get bam counts over mstr_feature
################################################################################
```{r}
library(csaw)

# Perform region counts
pe.param <- readParam(max.frag=600, pe="both", minq = 20, dedup = T)

h3k18la.counts <- regionCounts(meta$bams, mstr_features_anno@anno, ext=150, param=pe.param)


# Normalization
binned <- windowCounts(meta$bams, bin=TRUE, width=10000, param=pe.param)
h3k18la_se <- normFactors(binned, se.out=T)

```


################################################################################
### Prep the DGE list objects
################################################################################
```{r}
library(edgeR)

source(file.path(dest_tools_dir, "Expr_Filter.R"))


##########
## h3k18la

h3k18la.dge <- asDGEList(h3k18la.counts,
               norm.factors = h3k18la_se$norm.factors,
               group = meta$Group,
               genes = as.data.frame(mstr_features_anno))
colnames(h3k18la.dge) <- paste(meta$Group, meta$Rep, sep = "_")

h3k18la.dge$samples$Run <- meta$Run

h3k18la.dge$rpkm <- rpkm(h3k18la.dge, gene.length = "width")
ExpFilt(h3k18la.dge$rpkm, exp = 1, group = h3k18la.dge$samples$group)

h3k18la.dge.filt <- h3k18la.dge[idx.filt, keep.lib.sizes = T]

h3k18la.dge.filt$logCPM <- log2(cpm(h3k18la.dge.filt) + 1)
h3k18la.dge.filt$logFPKM <- rpkm(h3k18la.dge.filt, gene.length = "width", normalized.lib.sizes = T, log = T, prior.count = 1)
h3k18la.dge.filt$FPKM <- rpkm(h3k18la.dge.filt, gene.length = "width", normalized.lib.sizes = T, log = F)
h3k18la.dge.filt$FPKM_mean <- h3k18la.dge.filt$FPKM %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(H3K18LA_CNTRL = mean(c(H3K18LA_CNTRL_R1, H3K18LA_CNTRL_R2, H3K18LA_CNTRL_R3)),
         H3K18LA_2DG = mean(c(H3K18LA_2DG_R1, H3K18LA_2DG_R2, H3K18LA_2DG_R3)),
         H3K27AC_CNTRL = mean(c(H3K27AC_CNTRL_R1, H3K27AC_CNTRL_R2)),
         H3K27AC_2DG = mean(c(H3K27AC_2DG_R1, H3K27AC_2DG_R2)))

```


################################################################################
### PCA analysis - Figure S4
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}

library(PCAtools)
library(wesanderson)

#Prep the PCA data
pca.h3k18la <- PCAtools::pca(h3k18la.dge.filt$logFPKM, metadata = meta)


X='PC1'; Y='PC2'
pca1 <- biplot(pca.h3k18la, x = X, y = Y,
       colby = "Group", 
       legendPosition = 'right', 
       title = 'Peak Quant PCA',
       subtitle = paste0(X, ' versus ', Y))

cowplot::ggsave2(filename = "../data/processed/03b_PCA_PC1-2_group.pdf", plot = pca1,
                 width = 8, height = 6)

```


################################################################################
### Differential binding analysis - Figure 6
################################################################################
```{r,echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library(edgeR)
library(rtracklayer)
library(qvalue)

source(file.path(dest_tools_dir, "VolcanoPlot.R"))

y <- h3k18la.dge.filt

design <- model.matrix(~0 + h3k18la.dge.filt$samples$group)
colnames(design) <- levels(h3k18la.dge.filt$samples$group)

y <- estimateDisp(y, design)
summary(y$trended.dispersion)

fit <- glmQLFit(y, design, robust=TRUE)
summary(fit$var.post)

par(mfrow=c(1,2))
o <- order(y$AveLogCPM)
plot(y$AveLogCPM[o], sqrt(y$trended.dispersion[o]), type="l", lwd=2,
ylim=c(0, 1), xlab=expression("Ave."~Log[2]~"CPM"),
ylab=("Biological coefficient of variation"))
plotQLDisp(fit)

summary(fit$df.prior)

LA_2DG <- glmQLFTest(fit, contrast=c(-1,1,0,0))
AC_2DG <- glmQLFTest(fit, contrast=c(0,0,-1,1))


# Export Results

# LA_2DG
LA_2DG$table$qvalue <- qvalue(LA_2DG$table$PValue)$qvalue
LA_2DG$table$FDR <- p.adjust(LA_2DG$table$PValue, method = "fdr")

expt <- cbind(LA_2DG$genes[,-c(7:11)], LA_2DG$table, y$FPKM_mean)
write.table(expt, file = "../data/processed/03b_H3k18la.db.LA_2DG.tsv", sep = "\t", quote = F, row.names = F)


# AC_2DG
AC_2DG$table$qvalue <- qvalue(AC_2DG$table$PValue)$qvalue
AC_2DG$table$FDR <- p.adjust(AC_2DG$table$PValue, method = "fdr")

expt <- cbind(AC_2DG$genes[,-c(7:11)], AC_2DG$table, y$FPKM_mean)
write.table(expt, file = "../data/processed/03b_H3k18la.db.AC_2DG.tsv", sep = "\t", quote = F, row.names = F)



###############
# Volcano plots
###############

# LA_2DG
res_LA_2DG <- topTags(LA_2DG, n = Inf, sort.by = "logFC")$table %>% 
  as.tibble() %>% 
  rowwise() %>% 
  mutate(label_gene = ifelse((FDR < 0.001 | logFC < -3) | (FDR < 0.05 & logFC > 1), SYMBOL, NA))
 
res_LA_2DG_promPC <- res_LA_2DG %>% dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")

vp <- volPlot(res_LA_2DG_promPC, FC.nm = "logFC", FDR.nm = "FDR", fc = 1, fdr = 0.05, sym.lab = FALSE,
                     expt.nm = "03b_H3K18LA_2DG", res.path = "../data/processed/", ret = T)

vp <- vp +
   geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) 

cowplot::ggsave2(filename = "../data/processed/03b_H3K18LA_2DG_vol.pdf", plot = vp,
                 width = 8, height = 6)




# AC_2DG
res_AC_2DG <- topTags(AC_2DG, n = Inf, sort.by = "logFC")$table %>% 
  as.tibble() %>% 
  rowwise() %>% 
  mutate(label_gene = ifelse((FDR < 0.002 | (FDR < 0.05 & logFC < -2) | (FDR < 0.05 & logFC > 1)), SYMBOL, NA))
 
res_AC_2DG_promPC <- res_AC_2DG %>% dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding")

vp <- volPlot(res_AC_2DG_promPC, FC.nm = "logFC", FDR.nm = "FDR", fc = 1, fdr = 0.05, sym.lab = FALSE,
                     expt.nm = "03b_H3K27AC_2DG", res.path = "../data/processed/", ret = T)

vp <- vp +
   geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) 

cowplot::ggsave2(filename = "../data/processed/03b_H3K27AC_2DG_vol.pdf", plot = vp,
                 width = 8, height = 6)


```


################################################################################
### Make Heatmap of DB Prom PC genes - Figure S4
################################################################################
```{r}

db_goi <- rbind(res_LA_2DG_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame(),
            res_AC_2DG_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame())

LA_db_goi <- rbind(res_LA_2DG_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame())

AC_db_goi <- rbind(res_AC_2DG_promPC %>%
              dplyr::filter(abs(logFC)>1, FDR < 0.05, annotation == "Promoter", transcriptBiotype == "protein_coding") %>% as.data.frame())


db_goi <- db_goi %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(!duplicated(peakID))

LA_db_goi <- LA_db_goi %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(!duplicated(peakID))

AC_db_goi <- AC_db_goi %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(!duplicated(peakID))

# Filter quant binding for sig DB peaks
dat <- cbind(h3k18la.dge.filt$genes, h3k18la.dge.filt$FPKM_mean) %>%
  as.data.frame() %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(peakID %in% db_goi$peakID)

LA_dat <- cbind(h3k18la.dge.filt$genes, h3k18la.dge.filt$FPKM_mean) %>%
  as.data.frame() %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(peakID %in% LA_db_goi$peakID)

AC_dat <- cbind(h3k18la.dge.filt$genes, h3k18la.dge.filt$FPKM_mean) %>%
  as.data.frame() %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>%
  dplyr::filter(peakID %in% AC_db_goi$peakID)

# Tidy the table
dat.tidy <- dat %>%
  dplyr::select(SYMBOL, peakID, H3K18LA_CNTRL, H3K18LA_2DG, H3K27AC_CNTRL, H3K27AC_2DG) %>%
  gather(Sample, RPKM, -SYMBOL, -peakID)

dat.tidy$RPKM <- log2(dat.tidy$RPKM + 1)
dat.tidy$Sample <- factor(dat.tidy$Sample, levels = colnames(dat)[20:23])

# LA
LA_dat.tidy <- LA_dat %>%
  dplyr::select(SYMBOL, peakID, H3K18LA_CNTRL, H3K18LA_2DG) %>%
  gather(Sample, RPKM, -SYMBOL, -peakID)

LA_dat.tidy$RPKM <- log2(LA_dat.tidy$RPKM + 1)
LA_dat.tidy$Sample <- factor(LA_dat.tidy$Sample, levels = colnames(dat)[20:21])

#AC
AC_dat.tidy <- AC_dat %>%
  dplyr::select(SYMBOL, peakID, H3K27AC_CNTRL, H3K27AC_2DG) %>%
  gather(Sample, RPKM, -SYMBOL, -peakID)

AC_dat.tidy$RPKM <- log2(AC_dat.tidy$RPKM + 1)
AC_dat.tidy$Sample <- factor(AC_dat.tidy$Sample, levels = colnames(dat)[22:23])

###################
# Sort by Heirarchical clustering using Ward's method with Euclidean distanced
distfunc <- function(x) dist(x,method="euclidean")
hclustfunc <- function(x) hclust(x, method="ward.D2")

#Get distance and heirarchical clustering
d <- distfunc(as.matrix(dat[,20:21]))
hc <- hclustfunc(d)

d <- distfunc(as.matrix(AC_dat[,22:23]))
hc <- hclustfunc(d)

###################
# Sort by PCA
pca <- prcomp(dat[,20:23], center = T, scale. = F)
pca.order <- order(pca$x[,1])

LA_pca <- prcomp(LA_dat[,20:21], center = T, scale. = F)
LA_pca.order <- order(LA_pca$x[,1])

AC_pca <- prcomp(AC_dat[,22:23], center = T, scale. = F)
AC_pca.order <- order(AC_pca$x[,1])


###################################################################
# Make Heatmap
###################################################################

# Make heatmap
library(scales)
lim = c(0,6)
nm = "RPKM (log2)"
mb.colors <- colorRampPalette(c("#023858","#3690c0", "#e0f3f8","#fff7bc","#fec44f","#ec7014",
                                    "#fc4e2a","#bd0026","#800026"))(n=1000)

# LA PCA
tmp_dat <- LA_dat.tidy
tmp_dat$peakID <- factor(tmp_dat$peakID,
                                      levels = tmp_dat$peakID[(LA_pca.order)], ordered = T)

pp <- ggplot(tmp_dat, aes(Sample, peakID)) +
        geom_tile(aes(fill = RPKM), colour = "transparent") +
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03b_Prom_pc_CT_LA_hm-PCA.pdf", plot = pp,
                 width = 8, height = 6)

# AC PCA
tmp_dat <- AC_dat.tidy
tmp_dat$peakID <- factor(tmp_dat$peakID,
                                      levels = tmp_dat$peakID[rev(AC_pca.order)], ordered = T)

pp <- ggplot(tmp_dat, aes(Sample, peakID)) +
        geom_tile(aes(fill = RPKM), colour = "transparent") +
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03b_Prom_pc_CT_AC_hm-PCA.pdf", plot = pp,
                 width = 8, height = 6)


```


### Euler of DB with Timecourse - Figure S4
```{r}
library(ggVennDiagram)

################
# 2DG and SO in LA data

venn_goi <- list(DG = res_LA_2DG_promPC %>% 
                   dplyr::filter(abs(logFC) > 1, FDR < 0.05) %>% 
                   dplyr::select(SYMBOL) %>% 
                   unique() %>% unlist() %>% as.vector(),
                 TC = db_tc_sig_PromPC$SYMBOL %>% unique())


library(eulerr)

fit1 <- euler(c(
  DG = 833,
  TC = 4156,
  "DG&TC" = 1014
))

p <-plot(fit1, quantities = TRUE)
cowplot::ggsave2(filename = "../data/processed/03b_Prom_pc_DG-TC_euler.pdf", plot = p,
                 width = 4, height = 3)

```

################################################################################
### FGE of 2DG expts - Figure 6
################################################################################
```{r}

library(clusterProfiler)
library(enrichplot)
library(org.Mm.eg.db)
source(file.path(dest_tools_dir, "GO_leaf.R"))

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 10


dir_enrich <- file.path(dir_export, paste0("03b_PromPC_v2_FuncGeneEnrich_", enrich_db, "_", GOont))
dir.create(dir_enrich)


# Prep the gene lists for the FGE loop
goi_list <- list(LA_2DG = res_LA_2DG_promPC %>% 
                   dplyr::filter(abs(logFC) > 1, FDR < 0.05) %>% 
                   dplyr::select(SYMBOL) %>% 
                   unique() %>% unlist() %>% as.vector(),
                 AC_2DG = res_AC_2DG_promPC %>% 
                   dplyr::filter(abs(logFC) > 1, FDR < 0.05) %>% 
                   dplyr::select(SYMBOL) %>% 
                   unique() %>% unlist() %>% as.vector())

univ <- h3k18la.dge.filt$genes %>% 
  dplyr::filter(annotation == "Promoter", transcriptBiotype == "protein_coding") %>% 
  dplyr::select(SYMBOL) %>% 
  unique() %>% unlist() %>% as.vector()

enrich_results <- clusterProfiler::compareCluster(geneClusters = goi_list, fun = "enrichGO",
                                            keyType = "SYMBOL", ont = GOont, OrgDb = org, universe = univ)


# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>%
  as.data.frame() %>%
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>%
  rowwise() %>%
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]

# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_LA-AC_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>%
  as.data.frame() %>%
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>%
  rowwise() %>%
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_LA-AC_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>%
  group_by(Cluster) %>%
  arrange(p.adjust) %>%
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>%
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>%
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) +
    geom_point(aes(size = Ratio)) +
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("H3K18la and H3K27ac Peaks in 2DG explants"))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("LA-AC_FGE_top_", top, "_leaf_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5,
                 width = 6)
```


################################################################################
### RNA-seq and H3K18La DB binding integration - Figure 6
################################################################################
```{r}

tmp_db <- res_LA_2DG_promPC %>% 
  as.data.frame() %>% 
  arrange(desc(F)) %>% 
  dplyr::filter(!duplicated(geneId))

rna_db_merge <- inner_join(rna_de_res, tmp_db, by = c("ens_id" = "geneId"))

rna_db_merge <- rna_db_merge %>% mutate(sig_de = ifelse((abs(logFC.x) > 1 & FDR.x < 0.05) & (abs(logFC.y) > 1 & FDR.y < 0.05), "both",
                                                        ifelse(abs(logFC.x) > 1 & FDR.x < 0.05, "RNA",
                                                               ifelse(abs(logFC.y) > 1 & FDR.y < 0.05, "CUT&Tag", "NA"))))
rna_db_merge <- rna_db_merge %>% mutate(label_gene = ifelse(sig_de == "both", SYMBOL, NA))

dat <- rna_db_merge %>%
  mutate(rna_sig = ifelse((abs(logFC.x) >= 1 & FDR.x < 0.05), "sig", NA),
         cuttag_sig = ifelse((abs(logFC.y) >= 1 & FDR.y < 0.05), "sig", NA)) 
  

bp_all <- ggplot(dat, aes(x = logFC.y, y = logFC.x, color = sig_de, alpha = sig_de)) +
  geom_point(size = 0.5) +
  scale_color_manual(values = c("both" = "#F8766D", "CUT&Tag" = "#00BA38", "RNA" = "#619CFF", "NA" = "grey80"), name = "Significant") +
  scale_alpha_manual(values=c("both" = 1, "CUT&Tag" = 1, "RNA" = 1, "NA" = 0.1)) +
  geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) +
  xlim(-4,2) +
  ggtitle("2-DG RNAseq vs H3K18La CUT&Tag") +
  xlab("CUT&Tag 2-DG vs Control (logFC)") +
  ylab("RNAseq 2-DG vs Control (logFC)")

cowplot::ggsave2(filename = file.path(dir_export, paste0("03b", "_2DG_RNA-CUTtag-LA_point_all.pdf")), plot = bp_all,
                   width = 5, height = 4)

  
bp_sig <- ggplot(dat %>% dplyr::filter(rna_sig == "sig" | cuttag_sig == "sig"), 
                 aes(x = logFC.y, y = logFC.x, color = sig_de)) +
  geom_point() +
  scale_color_manual(values = c("both" = "#F8766D", "CUT&Tag" = "#00BA38", "RNA" = "#619CFF", "NA" = "grey95"), name = "Significant") +
  scale_alpha_manual(values=c(1, 1, 1, 0.1)) +
  geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) +
  xlim(-4,2) +
  ggtitle("2-DG RNAseq vs H3K18La CUT&Tag") +
  xlab("CUT&Tag 2-DG vs Control (logFC)") +
  ylab("RNAseq 2-DG vs Control (logFC)")

cowplot::ggsave2(filename = file.path(dir_export, paste0("03b", "_2DG_RNA-CUTtag-LA_point_sig.pdf")), plot = bp_sig,
                   width = 4, height = 3)

```

################################################################################
### RNA-seq and H3K27ac DB binding integration - Figure 6
################################################################################
```{r}

tmp_db <- res_AC_2DG_promPC %>% 
  as.data.frame() %>% 
  arrange(desc(F)) %>% 
  dplyr::filter(!duplicated(geneId))

rna_db_merge <- inner_join(rna_de_res, tmp_db, by = c("ens_id" = "geneId"))

rna_db_merge <- rna_db_merge %>% mutate(sig_de = ifelse((abs(logFC.x) > 1 & FDR.x < 0.05) & (abs(logFC.y) > 1 & FDR.y < 0.05), "both",
                                                        ifelse(abs(logFC.x) > 1 & FDR.x < 0.05, "RNA",
                                                               ifelse(abs(logFC.y) > 1 & FDR.y < 0.05, "CUT&Tag", "NA"))))
rna_db_merge <- rna_db_merge %>% mutate(label_gene = ifelse(sig_de == "both", SYMBOL, NA))

dat <- rna_db_merge %>%
  mutate(rna_sig = ifelse((abs(logFC.x) >= 1 & FDR.x < 0.05), "sig", NA),
         cuttag_sig = ifelse((abs(logFC.y) >= 1 & FDR.y < 0.05), "sig", NA)) 
  

bp_all <- ggplot(dat, aes(x = logFC.y, y = logFC.x, color = sig_de, alpha = sig_de)) +
  geom_point(size = 0.5) +
  scale_color_manual(values = c("both" = "#F8766D", "CUT&Tag" = "#00BA38", "RNA" = "#619CFF", "NA" = "grey80"), name = "Significant") +
  scale_alpha_manual(values=c("both" = 1, "CUT&Tag" = 1, "RNA" = 1, "NA" = 0.1)) +
  geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) +
  xlim(-4,2) +
  ggtitle("2-DG RNAseq vs H3K27Ac CUT&Tag") +
  xlab("CUT&Tag 2-DG vs Control (logFC)") +
  ylab("RNAseq 2-DG vs Control (logFC)")

cowplot::ggsave2(filename = file.path(dir_export, paste0("03b", "_2DG_RNA-CUTtag-AC_point_all.pdf")), plot = bp_all,
                   width = 5, height = 4)
```



################################################################################
### Perform quant of 2-DG analysis using TimeCourse DB peaks
################################################################################


```{r}
library(csaw)

gr <- makeGRangesFromDataFrame(db_tc_sig_PromPC, keep.extra.columns = TRUE)

# Perform region counts
pe.param <- readParam(max.frag=600, pe="both", minq = 20, dedup = T)

h3k18la_tc.counts <- regionCounts(meta$bams, gr, ext=150, param=pe.param)


```


### Make a heatmap of 2-DG projected on TimeCourse DB genes
```{r}

# Make DGE list object

h3k18la_tc.dge <- asDGEList(h3k18la_tc.counts,
               norm.factors = h3k18la_se$norm.factors,
               group = meta$Group,
               genes = as.data.frame(gr))
colnames(h3k18la_tc.dge) <- paste(meta$Group, meta$Rep, sep = "_")


h3k18la_tc.dge$samples$Run <- meta$Run

h3k18la_tc.dge$logCPM <- log2(cpm(h3k18la_tc.dge) + 1)
h3k18la_tc.dge$logFPKM <- rpkm(h3k18la_tc.dge, gene.length = "width", normalized.lib.sizes = T, log = T, prior.count = 1)
h3k18la_tc.dge$FPKM <- rpkm(h3k18la_tc.dge, gene.length = "width", normalized.lib.sizes = T, log = F)
h3k18la_tc.dge$FPKM_mean <- h3k18la_tc.dge$FPKM %>% 
  as_tibble() %>% 
  rowwise() %>% 
  transmute(H3K18LA_CNTRL = mean(c(H3K18LA_CNTRL_R1, H3K18LA_CNTRL_R2, H3K18LA_CNTRL_R3)),
         H3K18LA_2DG = mean(c(H3K18LA_2DG_R1, H3K18LA_2DG_R2, H3K18LA_2DG_R3)),
         H3K27AC_CNTRL = mean(c(H3K27AC_CNTRL_R1, H3K27AC_CNTRL_R2)),
         H3K27AC_2DG = mean(c(H3K27AC_2DG_R1, H3K27AC_2DG_R2)))

# Export RPKM quant
tmp <- cbind(h3k18la_tc.dge$genes, h3k18la_tc.dge$FPKM_mean)
write.table(tmp, file = "../data/processed/03b_DG_tc_db_quant_rpkm.tsv", sep = "\t", quote = F, col.names = F, row.names = F)


# Make a heatmap
dat.tmp <- cbind(h3k18la_tc.dge$genes, h3k18la_tc.dge$FPKM_mean) %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_"))


dat.tidy <- dat.tmp %>%
  mutate(peakID = paste(seqnames, start, end, sep = "_")) %>% 
  dplyr::select(SYMBOL, peakID, P28, H3K18LA_CNTRL, H3K18LA_2DG) %>%
  gather(Sample, RPKM, -SYMBOL, -peakID) %>% 
  mutate(RPKM = log2(RPKM + 1))

dat.tidy$Sample <- factor(dat.tidy$Sample, levels = dat.tidy$Sample %>% unique())

# P28 order
ord_p28 <- dat.tmp %>% 
  arrange(P28) %>% 
  dplyr::select(peakID) %>% 
  unlist() %>% as.vector()

# HC clustering
d <- distfunc(as.matrix(dat.tmp %>% dplyr::select(P28, H3K18LA_CNTRL, H3K18LA_2DG)))
hc <- hclustfunc(d)
ord_hc <- dat.tmp$peakID[hc$order]

# PCA ordering
pca.tmp <- prcomp(dat.tmp %>% dplyr::select(P28, H3K18LA_CNTRL, H3K18LA_2DG), center = T, scale. = T)
ord_pca <- dat.tmp$peakID[order(pca.tmp$x[,1])]


library(scales)
lim = c(0,8)
nm = "RPKM (log2)"
mb.colors <- colorRampPalette(c("#023858","#3690c0", "#e0f3f8","#fff7bc","#fec44f","#ec7014",
                                    "#fc4e2a","#bd0026","#800026"))(n=1000)

# Sort by order in Fig 3
peakID_ord <- readRDS("/Volumes/data/People/brooksma/PEGASUS/Projects/Gaur/Analysis/cutntag_240522/240911/data/interim/DB_Prom-PC_peakID_PCA_ord.Rds")

tmp_dat <- dat.tidy
tmp_dat$peakID <- factor(tmp_dat$peakID,
                                      levels = peakID_ord$peakID, ordered = T)

pp <- ggplot(tmp_dat, aes(Sample, peakID)) +
        geom_tile(aes(fill = RPKM), colour = "transparent") +
        #scale_fill_gradient(high = "white", low = "steelblue", name = "CPM (log2)") +
        scale_fill_gradientn(colours = mb.colors, name = nm, limits = lim, oob=squish) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text.y = element_blank(),
              panel.background = element_rect(fill = NA))

cowplot::ggsave2(filename = "../data/processed/03b_Prom_pc_TC-DG_hm-Fig5Border.pdf", plot = pp,
                 width = 8, height = 6)

```
