---
title: "Mouse glycolysis inhibitor - Differential Expression"
output: html_notebook
---

#
##### R Version
Built with `r getRversion()`

#
## Project Essentials

#
#### Load data
```{r, echo=FALSE, results='asis'}
# load("RNAseq_Analysis_v0.3.1.RData")

meta <- readRDS("../data/interim/meta.rds")
gene.dge <- readRDS("../data/interim/gene.dge.rds")
gene.dge.filt = readRDS("../data/interim/gene.dge.filt.rds")


print("Completed...")
```

#
#### Get custom scripts used for analysis
```{r,echo=FALSE, results='asis'}

script_dir <- "~/github/R_codes/"
dest_dir <- "../src/"

system(paste0("cp ", script_dir, "APk_cluster_BF.R ", dest_dir, "tools"))
system(paste0("cp ", script_dir, "Expr_Filter.R ", dest_dir, "tools"))
system(paste0("cp ", script_dir, "GO_leaf.R ", dest_dir, "tools"))

system(paste0("cp ", script_dir, "Heatmap.R ", dest_dir, "visualization"))
system(paste0("cp ", script_dir, "VolcanoPlot.R ", dest_dir, "visualization"))

print("Completed...")
```

#### Set color scheme and breaks
```{r, echo=FALSE, fig.height = 1.5, fig.width = 1.5}
library(wesanderson)
library(RColorBrewer)
library(viridis)

#
#Cluster palette
clust.pal <- brewer.pal(8, "Spectral")
#
#Heatmap colors
jet <- c("#7F0000","red","#FF7F00","yellow","white","cyan", "#007FFF", "blue","#00007F")
colorsZ <- colorRampPalette(rev(jet))(1000)
#nnrl.colors <- colorRampPalette(c("#023858","#3690c0","#fff7bc","#fec44f","#ec7014",
#                                  "#fc4e2a","#e31a1c","#bd0026","#800026"))(n=1000)
mb.col <- c("#023858","#3690c0","#e0f3f8","#fff7bc","#fec44f","#ec7014","#fc4e2a","#bd0026","#800026")
colorsCPM <- colorRampPalette(mb.col)(n=1000)
#
#colorsVir <- viridis_pal(option = "C")(1000)
#
#Generate the breaks for the color scale
col.breaksZ <- seq(-3, 3, length.out = 1001)
col.breaksCPM <- seq(0,16,length = 1001)
#

```


################################################################################
### DE of 2DG and SO experiments
################################################################################
*Fold change is set for 1.5 and FDR is set at 5%.*

```{r,echo=FALSE, results='asis', message=FALSE, warning=FALSE}
library(tidyverse)
library(edgeR)

# source("../src/tools/et_edgeR_v0.1.0.R")
# source("../src/tools/MBave.R")
source("../src/visualization/VolcanoPlot.R")
source("../src/visualization/Heatmap.R")


########################
# Prepare dgelist object
########################

# Remove Cntl_rep2 from analysis, low RIN/outlier
gene.dge.filt.ro <- gene.dge.filt[,-2]

# Get mean values for each group
gene.dge.filt.ro$cpm_mean <- gene.dge.filt.ro$cpm %>% 
  as_tibble() %>%
  rowwise() %>% 
  transmute(CNTRL = mean(c(Ctrl_rep1, Ctrl_rep3, Ctrl_rep4, Ctrl_rep5)),
         DG = mean(c(`2DG_rep1`,`2DG_rep2`, `2DG_rep3`, `2DG_rep4`)),
         SO = mean(c(SO_rep1, SO_rep2, SO_rep3)))


######################
# Perform DGE analysis
######################

y <- gene.dge.filt.ro

design <- model.matrix(~0 + y$samples$group)
colnames(design) <- levels(y$samples$group)

y <- estimateDisp(y, design)
summary(y$trended.dispersion)

fit <- glmQLFit(y, design, robust=TRUE)
summary(fit$var.post)

par(mfrow=c(1,2))
o <- order(y$AveLogCPM)
plot(y$AveLogCPM[o], sqrt(y$trended.dispersion[o]), type="l", lwd=2,
ylim=c(0, 1), xlab=expression("Ave."~Log[2]~"CPM"),
ylab=("Biological coefficient of variation"))
plotQLDisp(fit)

summary(fit$df.prior)

testDG <- glmQLFTest(fit, contrast=c(-1,1,0))
testSO <- glmQLFTest(fit, contrast=c(-1,0,1))


################
# Get Results and make volcano and heatmap
################

#####
# 2DG
res_DG <- topTags(testDG, n = Inf, sort.by = "logFC")$table %>% 
  rownames_to_column("ens_id") %>% 
  as.tibble() %>% 
  rowwise() %>% 
  mutate(label_gene = ifelse(FDR < 0.01 | (FDR < 0.05 & logFC < -2) | (FDR < 0.05 & logFC > 1), external_gene_name, NA))

write.table(res_DG, file = "../data/processed/RNA_DE_2DG_results.tsv", sep = "\t", quote = F, row.names = F)

# Violin Plot
vp <- volPlot(res_DG, FC.nm = "logFC", FDR.nm = "FDR", fc = 1, fdr = 0.05, sym.lab = FALSE,
                     expt.nm = "RNA_DE_2DG", res.path = "../data/processed/", ret = T)

vp <- vp +
   geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) 

cowplot::ggsave2(filename = "../data/processed/RNA_DE_2DG_vol.pdf", plot = vp,
                 width = 8, height = 6)

# Heatmap data
lcpm <- gene.dge.filt.ro$lcpm[,-2] %>% 
  as.data.frame() %>% 
  rownames_to_column("ens_id") %>% 
  as.tibble() 

dat <- inner_join(res_DG, lcpm, by = "ens_id") %>% 
  dplyr::filter(FDR < 0.05, abs(logFC) >= 1)

# Heatmap plot
heatmap_MB(expdat = dat, sym_col = 6, dat_col = 17:27, sort_col = 17:20, filenm = "../data/processed/RNA_DE_2DG")


#######
# LA_SO
res_SO <- topTags(testSO, n = Inf, sort.by = "logFC")$table %>% 
  rownames_to_column("ens_id") %>% 
  as.tibble() %>% 
  # arrange(logFC) %>% 
  rowwise() %>% 
  # mutate(label_gene = ifelse(FDR < 0.01 | (FDR < 0.05 & logFC < -4), external_gene_name, NA))
  mutate(rank_logFC = rank(-logFC),              # Top 20 (most positive)
         rank_logFC_asc = rank(logFC),           # Bottom 20 (most negative)
         label_gene = case_when(
           (rank_logFC <= 40 & FDR < 0.05) ~ external_gene_name,
           (rank_logFC_asc <= 50 & FDR < 0.05) ~ external_gene_name,
           TRUE ~ NA_character_
         ))
 
write.table(res_SO, file = "../data/processed/RNA_DE_SO_results.tsv", sep = "\t", quote = F, row.names = F)

# Volcano plot
vp <- volPlot(res_SO, FC.nm = "logFC", FDR.nm = "FDR", fc = 1, fdr = 0.05, sym.lab = FALSE,
                     expt.nm = "RNA_DE_SO", res.path = "../data/processed/", ret = T)

vp <- vp +
   geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) 

cowplot::ggsave2(filename = "../data/processed/RNA_DE_SO_vol.pdf", plot = vp,
                 width = 8, height = 6)

# Heatmap data
lcpm <- gene.dge.filt.ro$lcpm[,-2] %>% 
  as.data.frame() %>% 
  rownames_to_column("ens_id") %>% 
  as.tibble() 

dat <- inner_join(res_SO, lcpm, by = "ens_id") %>% 
  dplyr::filter(FDR < 0.05, abs(logFC) >= 1) %>% 
  distinct(external_gene_name, .keep_all = T)

# Heatmap plot
heatmap_MB(expdat = dat, sym_col = 6, dat_col = 17:27, sort_col = c(17:27), cluster = "HC", filenm = "../data/processed/RNA_DE_SO")



```


################################################################################
#### FGE of 2DG and SO genes
################################################################################
```{r}

library(clusterProfiler)
library(enrichplot)
library(org.Mm.eg.db)
source("../src/tools/GO_leaf.R")
# source("../src/tools/Dotplot_GO.R")

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 10

fc <- 1; fdr <- 0.05

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_export <- "../data/processed/"
dir_enrich <- file.path(dir_export, paste0("RNA_FuncGeneEnrich_", enrich_db, "_", GOont))
dir.create(dir_enrich)


# Prep the gene lists for the FGE loop
dg.list <- list(Down = res_DG %>% 
                  dplyr::filter(logFC < -fc, FDR < fdr) %>% 
                  dplyr::select(external_gene_name) %>% 
                  unlist(use.names = F) %>% as.vector(),
                Up = res_DG %>% 
                  dplyr::filter(logFC > fc, FDR < fdr) %>% 
                  dplyr::select(external_gene_name) %>% 
                  unlist(use.names = F) %>% as.vector()) 

so.list <- list(Down = res_SO %>% 
                  dplyr::filter(logFC < -fc, FDR < fdr) %>% 
                  dplyr::select(external_gene_name) %>% 
                  unlist(use.names = F) %>% as.vector(),
                Up = res_SO %>% 
                  dplyr::filter(logFC > fc, FDR < fdr) %>% 
                  dplyr::select(external_gene_name) %>% 
                  unlist(use.names = F) %>% as.vector()) 


univ <- gene.dge.filt.ro$genes$external_gene_name %>% unique()


###################
# DG FGE
###################
enrich_results <- clusterProfiler::compareCluster(geneClusters = dg.list, fun = "enrichGO",
                                            keyType = "SYMBOL", ont = GOont, OrgDb = org, universe = univ)
  
  
# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_2DG_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_2DG_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")


# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>% 
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>% 
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("2DG DGE FGE"))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("2DG_FGE_top_", top, "_leaf_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)

# Assign names
dg_enrich <- enrich_results_filt_fix
dg_enrich_leaf <- enrich_results_leaf_fix



###################
# SO FGE
###################
enrich_results <- clusterProfiler::compareCluster(geneClusters = so.list, fun = "enrichGO",
                                            keyType = "SYMBOL", ont = GOont, OrgDb = org, universe = univ)
  
  
# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_SO_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_SO_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")


# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>% 
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>% 
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("SO DGE FGE"))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("SO_FGE_top_", top, "_leaf_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)

# Assign names
so_enrich <- enrich_results_filt_fix
so_enrich_leaf <- enrich_results_leaf_fix

```


#
## Session Information
```{r, echo=FALSE}
sessionInfo()
```

#
## Explanation of Output Files

#### File structure of the data

* data
    + external
    + interim
    + processed   
    + raw
        - kallisto
        - star
* notebooks

#
#### Output File Description

Output files from analysis contain both tab-separated values table (.tsv), as well as, plots in pdf form. The tsv tables can be opened in Excel.

#
##### 'data/external'
Not used in this analysis.

#
##### 'data/interim'
Output from kallisto and tximport. This data is NOT normalized and shouldn't be used 'as is'.

* Trans_Counts_raw.tsv
    + Kallisto transcript-level counts
* Trans_EffLength_raw.tsv
    + Bias corrected effective transcript length
* Trans_TPM_raw.tsv
    + Transcript-level transcript-per-million (TPM) expression values
* Gene_Counts_raw.tsv
    + Tximport gene-level counts
* Gene_EffLength_raw.tsv
    + Bias corrected effective gene length
* Gene_TPM_raw.tsv
    + Gene-level transcript-per-million (TPM) expression values

#
##### 'data/processed'

* Gene_CPM_MSTR.tsv
    + Full TMM normalized gene-level results for all the genes in the annotation. Expression values are in counts-per-million (CPM). The Gene_log2CPM_MSTR.tsv file is the same except values are expressed as log2 with 1 offset.
* Gene_filtCPM_MSTR_5CPM.tsv
    + CPM filtered gene-level results of the full data table
* Trans_RPKM_MSTR.tsv
    + Full TMM normalized transcript-level results for all the transcripts in the annotation. Expression values are in reads-per-kilobase-per-million reads (RPKM). The Trans_log2RPKM_MSTR.tsv file is the same except values are expressed as log2 with 1 offset. 
* DE.et.all.tsv
    + The differential expression results output from edgeR. "All" is the full DE results, whereas the "sig" is only the significance filtered.
* gProfileR_ET.tsv
    + The full functional gene enrichment analysis output
* gProf_Expt_*
    + Directory containing the visualization results from the functional gene enrichment analysis
        - Heatmaps : heatmaps of all the top enriched categories using log2 CPM values 

#
##### 'data/raw
The raw kallisto data and star alignment summaries. If you need the actual BAM files from the star alignment the transfer can be arranged.
