---
title: "Glucose Explant RNA-seq"
output: html_notebook
---


### Get meta file
```{r}
meta <- read.csv("meta.csv", stringsAsFactors = F)
meta

```


### Experiment sample details - Eliminate samples/levels if needed
```{r, results="hide"}
## Number of samples and levels
nm <- dim(meta)[1]
expt.pheno <- factor(meta$Glucose)
nm.lvl <- length(levels(expt.pheno))

## Order of samples
ord <- c(1:nm)
ord.lvl <- unique(as.vector(expt.pheno))

## Remove outlier samples, if needed
remove.samp <- c() #Remove a sample(s)

## Make grp and export suffixes
if (!is.null(remove.samp)){
  grp <- factor(expt.pheno[ord[-remove.samp]], levels = ord.lvl)
  ord <- ord[-remove.samp]
  meta <- meta[-remove.samp,]
} else{
  grp <- factor(expt.pheno[ord], levels = ord.lvl)
}


meta$Group <- meta$Glucose_Rep

```


################################################################################
### Project Essentials
################################################################################

#### Get transcript and gene annotation
```{r}

# Import annotation
annot.gene <- read.csv(file = paste0("external/",short_sp,"_ENS",ens_ver,"_GeneAnnot.tsv"), sep="\t", row.names = 1)
annot.trans <- read.csv(file = paste0("external/",short_sp,"_ENS",ens_ver,"_TransAnnot.tsv"), sep="\t", row.names = 1)

# Create tx2gene needed for tximport
tx2gene <- data.frame(cbind(as.character(row.names(annot.trans)), as.character(annot.trans$ensembl_gene_id)))

```


## Set color scheme and breaks
```{r, results="hide"}
library(RColorBrewer)
## Sample colors
label.col <- c("red","blue", "black")

## Sample color palette
colorsSampPal <- colorRampPalette(label.col)(nm.lvl)

## Assign colors to samples based on grp
colorsSamp <- colorsSampPal[grp]

## Alignment plot colors
colorsAlign <- brewer.pal(3, "Spectral")

## Look at color assignments
pie(x = rep(1,nm.lvl), col = colorsSampPal, main = "Sample Group", labels = levels(grp))

# Cluster palette
clust.pal <- brewer.pal(8, "Spectral")

# Heatmap colors
jet <- c("#7F0000","red","#FF7F00","yellow","white","cyan", "#007FFF", "blue","#00007F")
colorsZ <- colorRampPalette(rev(jet))(1000)

mb.col <- c("#023858","#3690c0","#e0f3f8","#fff7bc","#fec44f","#ec7014","#fc4e2a","#bd0026","#800026")
colorsCPM <- colorRampPalette(mb.col)(n=1000)

# Generate the breaks for the color scale
col.breaksZ <- seq(-3, 3, length.out = 1001)
col.breaksCPM <- seq(0,16,length = 1001)
```


################################################################################
### Import Data
################################################################################

### Import kallisto data and summarize to gene-level
```{r, results="hide"}
library(tximport)
library(readr)

# kallisto results location
files <- file.path(sample_dirs, "abundance.h5")
names(files) <- paste0(meta$Glucose_Rep,"_",meta$Tech_Rep)


## Remove files, if necessary
files <- files[]
#meta <- meta[-18,]

## Import trans data
trans <- tximport(files, txOut=T, type = "kallisto", importer = read_tsv, tx2gene = NULL)
rnames <- gsub("(.+)\\..+", "\\1", row.names(trans$abundance), perl=T)
row.names(trans$abundance) <- rnames
row.names(trans$counts) <- rnames
row.names(trans$length) <- rnames


## Summarize to gene
gene <- summarizeToGene(trans, tx2gene, countsFromAbundance = "lengthScaledTPM")
```


### TMM Normalize the data
```{r}
library(edgeR)

#Transcript-level normalization
trans.dge <- DGEList(trans$counts[,ord], group = grp, genes = annot.trans[row.names(trans$counts),])
trans.dge <- sumTechReps(trans.dge, meta$Glucose_Rep)

trans.dge <- calcNormFactors(trans.dge)
trans.dge$rpkm <- rpkm(trans.dge, gene.length = "transcript_length")
trans.dge$lrpkm <- log2(trans.dge$rpkm + 1)

#Gene-level normalization
gene.dge <- DGEList(gene$counts[,ord], group = grp, genes = annot.gene[row.names(gene$counts),])
gene.dge <- sumTechReps(gene.dge, meta$Glucose_Rep)
gene.dge <- calcNormFactors(gene.dge)
gene.dge$cpm <- cpm(gene.dge)
gene.dge$lcpm <- log2(gene.dge$cpm + 1)
gene.dge$ltpm <- log2(gene$abundance[,ord] + 1)

#View the data frames
trans.dge$rpkm[1:6, 1:6]
gene.dge$cpm[1:6, 1:6]
```


### Export TMM transcript and gene level data using all genes
```{r, results="hide"}

#Export raw kallisto outputs
write.table(trans$counts, file="interim/Trans_Counts_raw.tsv", quote=F, sep="\t", col.names = NA)
write.table(trans$abundance, file="interim/Trans_TPM_raw.tsv", quote=F, sep="\t", col.names = NA)
write.table(trans$length, file="interim/Trans_EffLength_raw.tsv", quote=F, sep="\t", col.names = NA)

write.table(gene$counts, file="interim/Gene_Counts_raw.tsv", quote=F, sep="\t", col.names = NA)
write.table(gene$abundance, file="interim/Gene_TPM_raw.tsv", quote=F, sep="\t", col.names = NA)
write.table(gene$length, file="interim/Gene_EffLength_raw.tsv", quote=F, sep="\t", col.names = NA)


#Export transcript tables
write.table(data.frame(merge(trans.dge$genes, trans.dge$rpkm, by=0), row.names=1), 
            file="processed/Trans_RPKM_MSTR.tsv", quote=F, sep="\t", col.names = NA)
write.table(data.frame(merge(trans.dge$genes, trans.dge$lrpkm, by=0), row.names=1), 
            file="processed/Trans_log2RPKM_MSTR.tsv", quote=F, sep="\t", col.names = NA)


#Export gene tables
write.table(data.frame(merge(annot.gene, gene.dge$cpm, by=0), row.names=1), 
            file="processed/Gene_CPM_MSTR.tsv", quote=F, sep="\t", col.names = NA)
write.table(data.frame(merge(annot.gene, gene.dge$lcpm, by=0), row.names=1), 
            file="processed/Gene_log2CPM_MSTR.tsv", quote=F, sep="\t", col.names = NA)

```


### Perform expression filtered normalization and export final data
```{r}
source("src/filterFPKM_v1.R")

#Expression level for gene filtering
exprlvl = 1

## Clean up groups and meta sheet after merging technical replicates
small_meta <- meta %>% filter(Tech_Rep==1)
grp <- as.factor(small_meta$Glucose)


#Filter data for those expressed
filterFPKM(as.data.frame(gene.dge$counts), grp, exp = exprlvl)

#Perform normalization for DE
gene.dge.filt <- gene.dge[idx.filt, , keep.lib.sizes=FALSE]
gene.dge.filt <- calcNormFactors(gene.dge.filt)
gene.dge.filt$cpm <- cpm(gene.dge.filt, log=F)
gene.dge.filt$lcpm <- log2(gene.dge.filt$cpm + 1)

#Export the normalized filtered CPM values
write.table(data.frame(merge(gene.dge.filt$genes, gene.dge.filt$cpm, by = 0), row.names = 1), 
            file=paste0("processed/Gene_filtCPM_MSTR_", exprlvl, "CPM.tsv"), 
            quote=F, sep="\t", col.names = NA)

#Export the normalized filtered logCPM values
write.table(data.frame(merge(gene.dge.filt$genes, gene.dge.filt$lcpm, by = 0), row.names = 1), 
            file=paste0("processed/Gene_filtlog2CPM_MSTR_", exprlvl, "CPM.tsv"), 
            quote=F, sep="\t", col.names = NA)

print(paste0("Number of genes after filtering for ", exprlvl, " CPM: ", dim(gene.dge.filt)[1]))
```


### PCA
```{r}
library(PCAtools)

rownames(small_meta) <- c(small_meta$Glucose_Rep)
p <- pca(gene.dge.filt$lcpm, metadata = small_meta, removeVar = 0.1)

## Variance per PC (Scree Plot)
screeplot(p,
          hline=75,
          vline = 5)

## PCA Plot
biplot(p, 
       colby = "Glucose")+
  theme_classic()



## Eigencor plot - Pearson's correlation of factor to PCs with FDR corrected p-value
eigencorplot(p,
             col = colorRampPalette(c("white", "red3"))(20),
             components = getComponents(p, 1:5),
             metavars = c('Glucose','Bio_Rep'),
             plotRsquared = TRUE,
             corFUN = 'pearson',
             corUSE = 'pairwise.complete.obs',
             corMultipleTestCorrection = 'fdr')
```

################################################################################
### Differential Expression Analysis
################################################################################

### #Make the model matrix for analysis
```{r}
library(edgeR)

# Set up ANOVA-like difference
small_meta$Group <- as.factor(small_meta$Glucose)

mod <- model.matrix(~0+small_meta$Group)
colnames(mod) <- gsub("small_meta\\$", "", colnames(mod))
colnames(mod) <- gsub("Group", "", colnames(mod))
mod

# Estimate dispersion
gene.dge.filt <- estimateDisp(gene.dge.filt, mod)

# Fit
fit <- glmQLFit(gene.dge.filt, mod)
colnames(fit)

# Set desired individual contrasts
conMatrix <- makeContrasts(
  HighGlucose_v_LowGlucose = mM_25 - mM_5,
  levels=mod)
print("Matrix of examined contrasts")
conMatrix


```


### Perform differential expression 
```{r}

# Set chromosome list
chrList <- c(1:22,"Y","X")

# Set thresholds
library(dplyr)
fc <- 1
fdr <- 0.05

   
## Run Quasi-likelihood F Tests for all comparisons
qlf <- glmQLFTest(fit, contrast = conMatrix[,"HighGlucose_v_LowGlucose"])

## Store all genes 
DE <- c()
DE$HighGlucose_v_LowGlucose <- data.frame(topTags(qlf, n = Inf, p.value = 1, adjust.method = "BH", sort.by = "logFC")) %>% tibble::rownames_to_column("ensembl_ID")

## Store sig genes
DE$HighGlucose_v_LowGlucose <-data.frame(topTags(qlf, n = Inf, p.value = fdr, adjust.method = "BH", sort.by = "logFC")) %>% filter(abs(logFC) > fc) %>% filter(chromosome_name %in% chrList) %>% tibble::rownames_to_column("ensembl_ID")

## Sig genes merged with CPM data
cpm_table <- as.data.frame(gene.dge.filt$cpm)  %>% tibble::rownames_to_column("ensembl_ID")
DE$HighGlucose_v_LowGlucose_sig_cpm <- DE$HighGlucose_v_LowGlucose %>% left_join(cpm_table)

#Export Results
write.csv(DE$HighGlucose_v_LowGlucose_sig_cpm, "processed/DEG_HighGlucose_v_LowGlucose.csv", row.names = TRUE)
```

################################################################################
### Cluster Analysis and Functional Gene Enrichment
################################################################################

### Average the data and get Z-score
```{r}
source("src/MBave.R")


# Get the average log2 CPM expression per group for plotting
gene.dge$ave <- MBave(dat = gene.dge$lcpm, grp = grp)
gene.dge$cpm_ave <- MBave(dat = gene.dge$cpm, grp = grp)
```


```{r}
# Get the averaged Z-score for the DE sig genes
gene.dge$aveZ <- t(scale(t(gene.dge$ave[DE$HighGlucose_v_LowGlucose_sig_cpm$ensembl_ID,]), center = T, scale = T))
```


```{r}
library(gprofiler2)

DE_Up <- DE$HighGlucose_v_LowGlucose_sig_cpm %>% filter(logFC>0)
DE_Down <- DE$HighGlucose_v_LowGlucose_sig_cpm %>% filter(logFC<0)

gostres_up <- gost(query = DE_Up$ensembl_ID, 
                organism = "mmusculus", ordered_query = FALSE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = FALSE, 
                user_threshold = 0.05, correction_method = "g_SCS", 
                domain_scope = "annotated", custom_bg = NULL, 
                numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = TRUE)

gostres_down <- gost(query = DE_Down$ensembl_ID, 
                organism = "mmusculus", ordered_query = FALSE, 
                multi_query = FALSE, significant = TRUE, exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, evcodes = FALSE, 
                user_threshold = 0.05, correction_method = "g_SCS", 
                domain_scope = "annotated", custom_bg = NULL, 
                numeric_ns = "", sources = NULL, as_short_link = FALSE, highlight = TRUE)

gostplot(gostres_up, capped = TRUE, interactive = TRUE)
gostplot(gostres_down, capped = TRUE, interactive = TRUE)
```


