---
title: "Base Project - Data Import and QC"
output: html_notebook
---


### Meta file
```{r, echo = FALSE, results = 'asis'}
library(tidyverse)


kal_files <- list.files(path = "../data/raw/kallisto/", pattern = "abundance.tsv", recursive = T, full.names = T)
kal_files <- kal_files[c(5:9, 1:4, 10:12)]

# Meta sheet
meta <- data.frame(Beds = kal_files,
                   Treat = gsub(".+\\/(.+)_rep.+", "\\1", kal_files),
                   Rep = gsub(".+\\/.+_(rep.+)\\/abundance.+", "\\1", kal_files)) %>% 
  mutate(Rid = paste(Treat, Rep, sep = "_"))

names(macs_beds) <- paste(meta$Group, meta$Rep, sep = "_")

# RNA file
rna <- read.csv("../src/data/RNA_quant.csv")
rna <- rna %>% 
  dplyr::filter(ToUse ) %>% 
  arrange(Sample)


meta <- inner_join(meta, rna, by = join_by(Rid == Sample))
```


### Experiment sample details - Eliminate samples/levels if needed
```{r, echo = FALSE, results = 'asis'}

#Number of samples and levels
nm <- dim(meta)[1]
expt.pheno <- factor(meta$Treat)
nm.lvl <- length(levels(expt.pheno))

#Order of samples
ord <- c(1:nm)
ord.lvl <- unique(as.vector(expt.pheno))[c(1:nm.lvl)]

#Remove outlier samples, if needed
remove.samp <- NULL

#Make grp and export suffixes
if (!is.null(remove.samp)){
  grp <- factor(expt.pheno[ord[-remove.samp]], levels = ord.lvl)
  ord <- ord[-remove.samp]
} else{
  grp <- factor(expt.pheno[ord], levels = ord.lvl)
}

```

################################################################################
## Project Essentials
################################################################################

### Get custom scripts used for analysis
```{r, echo = FALSE, results = 'asis'}
script_dir <- "~/github/R_codes/"
dest_dir <- "../src/"

system(paste0("cp ", script_dir, "ENSannotDownload.R ", dest_dir, "tools"))
system(paste0("cp ", script_dir, "Expr_Filter.R ", dest_dir, "tools"))
system(paste0("cp ", script_dir, "GO_leaf.R ", dest_dir, "tools"))
system(paste0("cp ", script_dir, "Heatmap.R ", dest_dir, "visualization"))
system(paste0("cp ", script_dir, "VolcanoPlot.R ", dest_dir, "visualization"))

print("Completed...")
```


### Get transcript and gene annotation
```{r, echo = FALSE, results = 'asis'}

# Get annotation
source(paste0(dest_dir, "/tools/ENSannotDownload.R"))
annot <- ens_annot(species = "Mus musculus", release = "102")

#Create tx2gene needed for tximport
tx2gene <- data.frame(cbind(as.character(row.names(annot$trans)), as.character(annot$trans$ensembl_gene_id)))

```


### Set color scheme and breaks
```{r, echo = FALSE, results = 'asis'}
library(wesanderson)
library(RColorBrewer)

#Sample colors
label.col <- wes_palette("Zissou1", 5, "discrete")[c(2,3,1,5)]

#Sample color palette
colorsSampPal <- colorRampPalette(label.col)(nm.lvl)

colorsSamp <- colorsSampPal[grp]


#Alignment plot colors
colorsAlign <- wes_palette(n=5, name = "Zissou1")[c(1,3,5)]

#Look at color assignments
pie(x = rep(1,nm.lvl), col = colorsSampPal, main = "Sample Group", labels = levels(grp))

#Cluster palette
clust.pal <- brewer.pal(8, "Spectral")

#Heatmap colors
jet <- c("#7F0000","red","#FF7F00","yellow","white","cyan", "#007FFF", "blue","#00007F")
colorsZ <- colorRampPalette(rev(jet))(1000)
                                "#fc4e2a","#e31a1c","#bd0026","#800026"))(n=1000)
mb.col <- c("#023858","#3690c0","#e0f3f8","#fff7bc","#fec44f","#ec7014","#fc4e2a","#bd0026","#800026")
colorsCPM <- colorRampPalette(mb.col)(n=1000)

#Generate the breaks for the color scale
col.breaksZ <- seq(-3, 3, length.out = 1001)
col.breaksCPM <- seq(0,16,length = 1001)

```


################################################################################
### Import Data
################################################################################

#### Import kallisto data and summarize to gene-level
```{r, echo = F, message = F}
library(tximport)
library(readr)

#########
#kallisto results location
wd <- "../data/raw/kallisto"
samples <- dir(path=wd, pattern=".+")

files <- meta$Beds
names(files) <- meta$Rid 

#Import trans data
trans <- tximport(files, txOut=T, type = "kallisto", importer = read_tsv, tx2gene = NULL)
rnames <- gsub("(.+)\\..+", "\\1", row.names(trans$abundance), perl=T)
row.names(trans$abundance) <- rnames
row.names(trans$counts) <- rnames
row.names(trans$length) <- rnames

#Summarize to gene
gene <- summarizeToGene(trans, tx2gene, countsFromAbundance = "lengthScaledTPM")

```

#
#### TMM Normalize the data
```{r, echo = FALSE, results = 'asis', message=F}
library(edgeR)

#Transcript-level normalization
trans.dge <- DGEList(trans$counts[,ord], group = grp, genes = annot$trans[row.names(trans$counts),])
trans.dge <- calcNormFactors(trans.dge)
trans.dge$rpkm <- rpkm(trans.dge, gene.length = "transcript_length")
trans.dge$lrpkm <- log2(trans.dge$rpkm + 1)

#Gene-level normalization
gene.dge <- DGEList(gene$counts[,ord], group = grp, genes = annot$gene[row.names(gene$counts),])
gene.dge <- calcNormFactors(gene.dge)
gene.dge$cpm <- cpm(gene.dge)
gene.dge$lcpm <- log2(gene.dge$cpm + 1)
gene.dge$ltpm <- log2(gene$abundance[,ord] + 1)

```

################################################################################
### Export Interim and Final Data
################################################################################

### Export TMM transcript and gene level data using all genes
```{r, echo = FALSE, results = 'asis'}
#Export raw kallisto outputs
write.table(trans$counts, file="../data/interim/Trans_Counts_raw.tsv", quote=F, sep="\t", col.names = NA)
write.table(trans$abundance, file="../data/interim/Trans_TPM_raw.tsv", quote=F, sep="\t", col.names = NA)
write.table(trans$length, file="../data/interim/Trans_EffLength_raw.tsv", quote=F, sep="\t", col.names = NA)

write.table(gene$counts, file="../data/interim/Gene_Counts_raw.tsv", quote=F, sep="\t", col.names = NA)
write.table(gene$abundance, file="../data/interim/Gene_TPM_raw.tsv", quote=F, sep="\t", col.names = NA)
write.table(gene$length, file="../data/interim//Gene_EffLength_raw.tsv", quote=F, sep="\t", col.names = NA)

#Export transcript tables
write.table(data.frame(merge(trans.dge$genes, trans.dge$rpkm, by=0), row.names=1), 
            file="../data/processed/Trans_RPKM_MSTR.tsv", quote=F, sep="\t", col.names = NA)
write.table(data.frame(merge(trans.dge$genes, trans.dge$lrpkm, by=0), row.names=1), 
            file="../data/processed/Trans_log2RPKM_MSTR.tsv", quote=F, sep="\t", col.names = NA)

#Export gene tables
write.table(data.frame(merge(annot$gene, gene.dge$cpm, by=0), row.names=1), 
            file="../data/processed/Gene_CPM_MSTR.tsv", quote=F, sep="\t", col.names = NA)
write.table(data.frame(merge(annot$gene, gene.dge$lcpm, by=0), row.names=1), 
            file="../data/processed/Gene_log2CPM_MSTR.tsv", quote=F, sep="\t", col.names = NA)

```


### Perform expression filtered normalization and export final data
```{r, echo = FALSE, results = 'asis'}
source("../src/tools/Expr_Filter.R")

#Expression level for gene filtering
exprlvl = 5

#Filter data for those expressed
ExpFilt(as.data.frame(gene.dge$cpm), grp, exp = exprlvl)

#Perform normalization for DE
gene.dge.filt <- gene.dge[idx.filt, , keep.lib.sizes=FALSE]
gene.dge.filt <- calcNormFactors(gene.dge.filt)
gene.dge.filt$cpm <- cpm(gene.dge.filt, log=F)
gene.dge.filt$lcpm <- log2(gene.dge.filt$cpm + 1)

#Export the normalized filtered CPM values
write.table(data.frame(merge(gene.dge.filt$genes, gene.dge.filt$cpm, by = 0), row.names = 1), 
            file=paste0("../data/processed/Gene_filtCPM_MSTR_", exprlvl, "CPM.tsv"), 
            quote=F, sep="\t", col.names = NA)

print(paste0("Number of genes after filtering for ", exprlvl, " CPM: ", dim(gene.dge.filt)[1]))
```


### Render PCA on all samples
```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(pca3d)
library(PCAtools)


#Prep the PCA data
pca.all <- PCAtools::pca(gene.dge.filt$lcpm, metadata = data.frame(meta[ord,], row.names = 4), scale = T)
pca.ro <- PCAtools::pca(gene.dge.filt$lcpm[,-2], metadata = data.frame(meta[ord[-2],], row.names = 4), scale = T)

sampkey <- unique(colorsSamp)
names(sampkey) <- unique(pca.all$metadata$Treat)

s <- screeplot(pca.all)

cowplot::ggsave2(filename = "../data/processed/03_PCA.all_scree.pdf", plot = s,
                 width = 8, height = 6)

#Plot the eigencorplot
pdf(file = "../data/processed/03_PCA.all_eigen.pdf", width = 8, height = 6, useDingbats = T)
eigencorplot(pca.all, metavars = names(pca.all$metadata[c(2:6)]),
             main = "Correlation of PCs to Meta Features for all samples")
dev.off()


#Plot all the samples
X='PC1'; Y='PC2'
pca1 <- biplot(pca.all, x = X, y = Y,
       colby = "Treat", colkey = sampkey,
       legendPosition = 'right', 
       title = 'PCA All Samples',
       subtitle = paste0(X, ' versus ', Y))

cowplot::ggsave2(filename = "../data/processed/03_PCA.all_PC1-2_treat.pdf", plot = pca1,
                 width = 8, height = 6)

```


################################################################################
### DE of 2DG experiment
################################################################################


```{r,echo=FALSE, results='asis', message=FALSE, warning=FALSE}
library(tidyverse)
library(edgeR)

source("../src/visualization/VolcanoPlot.R")
source("../src/visualization/Heatmap.R")


########################
# Prepare dgelist object
########################

# Remove Cntl_rep2 from analysis, low RIN/outlier
gene.dge.filt.ro <- gene.dge.filt[,-2]

# Get mean values for each group
gene.dge.filt.ro$cpm_mean <- gene.dge.filt.ro$cpm %>% 
  as_tibble() %>%
  rowwise() %>% 
  transmute(CNTRL = mean(c(Ctrl_rep1, Ctrl_rep3, Ctrl_rep4, Ctrl_rep5)),
         DG = mean(c(`2DG_rep1`,`2DG_rep2`, `2DG_rep3`, `2DG_rep4`)))


######################
# Perform DGE analysis
######################

y <- gene.dge.filt.ro

design <- model.matrix(~0 + y$samples$group)
colnames(design) <- levels(y$samples$group)

y <- estimateDisp(y, design)
summary(y$trended.dispersion)

fit <- glmQLFit(y, design, robust=TRUE)
summary(fit$var.post)

par(mfrow=c(1,2))
o <- order(y$AveLogCPM)
plot(y$AveLogCPM[o], sqrt(y$trended.dispersion[o]), type="l", lwd=2,
ylim=c(0, 1), xlab=expression("Ave."~Log[2]~"CPM"),
ylab=("Biological coefficient of variation"))
plotQLDisp(fit)

summary(fit$df.prior)

testDG <- glmQLFTest(fit, contrast=c(-1,1))



################
# Get Results and make volcano and heatmap
################

#####
# 2DG
res_DG <- topTags(testDG, n = Inf, sort.by = "logFC")$table %>% 
  rownames_to_column("ens_id") %>% 
  as.tibble() %>% 
  rowwise() %>% 
  mutate(label_gene = ifelse(FDR < 0.01 | (FDR < 0.05 & logFC < -2) | (FDR < 0.05 & logFC > 1), external_gene_name, NA))

write.table(res_DG, file = "../data/processed/RNA_DE_2DG_results.tsv", sep = "\t", quote = F, row.names = F)

# Violin Plot
vp <- volPlot(res_DG, FC.nm = "logFC", FDR.nm = "FDR", fc = 1, fdr = 0.05, sym.lab = FALSE,
                     expt.nm = "RNA_DE_2DG", res.path = "../data/processed/", ret = T)

vp <- vp +
   geom_text_repel(aes(label = label_gene), size = 3, max.overlaps = Inf) 

cowplot::ggsave2(filename = "../data/processed/RNA_DE_2DG_vol.pdf", plot = vp,
                 width = 8, height = 6)

# Heatmap data
lcpm <- gene.dge.filt.ro$lcpm[,-2] %>% 
  as.data.frame() %>% 
  rownames_to_column("ens_id") %>% 
  as.tibble() 

dat <- inner_join(res_DG, lcpm, by = "ens_id") %>% 
  dplyr::filter(FDR < 0.05, abs(logFC) >= 1)

# Heatmap plot
heatmap_MB(expdat = dat, sym_col = 6, dat_col = 17:27, sort_col = 17:20, filenm = "../data/processed/RNA_DE_2DG")

```


################################################################################
#### FGE of 2DG and SO genes
################################################################################
```{r}

library(clusterProfiler)
library(enrichplot)
library(org.Mm.eg.db)
source("../src/tools/GO_leaf.R")
# source("../src/tools/Dotplot_GO.R")

org = "org.Mm.eg.db"

enrich_db <- "GO_220310"
GOont <- "BP"
top <- 10

fc <- 1; fdr <- 0.05

# dir_enrich <- file.path(dir_res, "FuncGeneEnrich")
dir_export <- "../data/processed/"
dir_enrich <- file.path(dir_export, paste0("RNA_FuncGeneEnrich_", enrich_db, "_", GOont))
dir.create(dir_enrich)


# Prep the gene lists for the FGE loop
dg.list <- list(Down = res_DG %>% 
                  dplyr::filter(logFC < -fc, FDR < fdr) %>% 
                  dplyr::select(external_gene_name) %>% 
                  unlist(use.names = F) %>% as.vector(),
                Up = res_DG %>% 
                  dplyr::filter(logFC > fc, FDR < fdr) %>% 
                  dplyr::select(external_gene_name) %>% 
                  unlist(use.names = F) %>% as.vector()) 


univ <- gene.dge.filt.ro$genes$external_gene_name %>% unique()


###################
# DG FGE
###################
enrich_results <- clusterProfiler::compareCluster(geneClusters = dg.list, fun = "enrichGO",
                                            keyType = "SYMBOL", ont = GOont, OrgDb = org, universe = univ)
  
  
# Filter enrichment for significance
enrich_results_filt <- enrich_results@compareClusterResult %>% dplyr::filter(p.adjust < 0.05, Count > 3)

# Fix the Ratio
enrich_results_filt_fix <- enrich_results_filt %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_filt_fix[,-11]

# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_2DG_full.tsv")),
              quote = F, row.names = F, sep = "\t")

# Get leaf terms
enrich_results_leaf <- GO_leaf(enrich_results_filt, query = 1, id.col = 2, ont = GOont)

# Fix the Ratio
enrich_results_leaf_fix <- enrich_results_leaf %>% 
  as.data.frame() %>% 
  mutate(tmp_col = str_split(GeneRatio, pattern = "/")) %>% 
  rowwise() %>% 
  mutate(Ratio = as.numeric(unlist(tmp_col)[1]) / as.numeric(unlist(tmp_col)[2]))

# Export the FGE results table
tmp <- enrich_results_leaf_fix[,-11]
# Export the enrichment analysis
write.table(tmp, file = file.path(dir_enrich, paste0(enrich_db, "_", GOont, "_2DG_leaf.tsv")),
              quote = F, row.names = F, sep = "\t")


# Get top 5 pathways per cell type to get union set of pathways
top5_union <- enrich_results_leaf %>% 
  group_by(Cluster) %>% 
  arrange(p.adjust) %>% 
  slice_head(n = top)


# Filter the full table for the union of top5 pathways
top5_enrich <- enrich_results_filt_fix %>% 
  dplyr::filter(ID %in% (top5_union$ID %>% unique()))

# Prep the df for plotting
top5_enrich <- top5_enrich %>% 
  arrange(Cluster, p.adjust)
top5_enrich$Description <- substring(top5_enrich$Description, 1, 60)
top5_enrich$Description <- factor(top5_enrich$Description, levels = rev(unique(top5_enrich$Description)))
top5_enrich$p.adjust <- -log10(top5_enrich$p.adjust)

# Make plot
pp <- ggplot(top5_enrich, aes(Cluster, Description, colour = p.adjust)) + 
    geom_point(aes(size = Ratio)) + 
    theme_bw() +
    scale_colour_gradient(low = "#3B9AB2", high = "#F21A00", name = "P Value\n-log10", limits=c(0,10), oob = scales::squish) +
    guides(size=guide_legend(title="Ratio")) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    ggtitle(paste0("2DG DGE FGE"))

#Open the pdf device
cowplot::ggsave2(plot = pp, filename = file.path(dir_enrich, paste0("2DG_FGE_top_", top, "_leaf_01_dotplot.pdf")),
                 height = top5_enrich$Description %>% unique() %>% length() / 10 + 2.5, 
                 width = 6)

# Assign names
dg_enrich <- enrich_results_filt_fix
dg_enrich_leaf <- enrich_results_leaf_fix

```